### 2022-08-10 Methods.Rmd
initial commit of exploratory figures and methods

---
title: "Aquatic Resource Trade in Species (ARTIS) Methods"
header-includes:
  - \usepackage{floatrow}
  - \floatplacement{figure}{H}
  - \hypersetup{linktocpage, colorlinks = true, linkcolor = blue}
  - \usepackage{draftwatermark}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \renewcommand{\headrulewidth}{0pt}
  - \rfoot{\thepage}
output: 
  bookdown::pdf_document2:
    fig_caption: yes
    toc: yes
    toc_depth: 2
---

# Overview

In order to estimate the seafood species trade network, we compiled and aligned diverse data on fishery and aquaculture production, live weight conversion factors, bilateral global trade, and nutrient content. While our data comes from open/published sources, its compilation was an immense task. Its comprehensive scale spans the globe and encompasses decades of changes in country and species names and seafood product forms. Over 4000 live weight conversion factors were compiled and matched to 2000+ farmed and wild capture aquatic species which in turn were matched to 900+ traded seafood product descriptions. This is then  aligned with the nutritional data that has been compiled across all aquatic seafood taxa for over 300 micronutrients. Our estimation of species trade flows occurs in two steps. First, we take a mass balance approach, where each country’s seafood exports must equal the domestic production, plus imports, minus domestic consumption, after accounting for processing losses. For each country, we estimate the proportion of seafood production going into each possible commodity, the proportion of each imported commodity processed and exported, and the domestic consumption of each commodity. We then use these estimates with bilateral trade data to solve for the global species flows. This approach substantially improves upon previous efforts by estimating species-level trade, covering all production environments (marine and freshwater) and production methods (farmed and wild caught), and including the processing and export of imported products.

# Data

## Production

Seafood production comes from the Food and Agriculture Organization (FAO), which provides national capture and aquaculture production (FAO 2020). The Food and Agriculture Organization provides annual capture and aquaculture production data for around 240 countries, territories or land areas from 1950 to 2019 in the FAO FishStatJ database. The FAO data reports production in tonnes (live weight equivalent) of around 550 farmed and 1600 wild capture species and species groups  classified by International Standard Statistical Classification of Aquatic Animals and Plants (ISSCAAP). FAO production data consists primarily of official national statistics, with some verifiable supplemental information from academic reviews, consultant reports, and other specialist literature. Data reported by nations are checked by the FAO for consistency and questionable values are verified with the reporting offices. When countries fail to report production, FAO uses past values to estimate production. For the purposes of this analysis, we do not distinguish between nationally-reported and FAO estimated values.

According to the Coordinating Working Party on Fishery Statistics, catch and landings should be assigned to the country of the flag flown by the fishing vessel irrespective of the location of the fishing. This means that production resulting from a country operating a fishing vessel in a foreign (host) country's territory should be recorded in the national statistics of the foreign fishing vessel. However, if the vessel is chartered by a company based in the home country or the vessel is fishing for the country under a joint venture contract or similar agreement and the operation is integral to the economy of the host country, this does not apply.

## Bilateral trade data

We use the CEPII BACI world trade database, which is a reconciled version of the UN Comtrade database (Gaulier and Zignago 2010). Trade data are reported to the UN by both importers and exporters following the Harmonized System (HS) codes. The HS trade code system organizes traded goods into a hierarchy, with the highest level represented by two digit codes (e.g., Chapter 03 covers "Fish and Crustaceans, Molluscs and Other Aquatic Invertebrates"), which are broken down into 4-digit headings (e.g., heading 0301 covers "Live fish"), which are then subdivided into 6-digit subheadings (e.g., subheading 030111 covers "Live ornamental freshwater fish"). National statistics offices may further subdivide HS codes into 7- to 12-digit codes but since these are not standard across countries, the HS 6-digit codes are the most highly resolved trade codes available globally. HS codes are administered by the World Customs Organization, which updates the codes every five years. HS versions can be used from their introduction through the present, meaning that the HS 2002 version provides a time series of trade from 2002 to the present whereas the HS 2017 version only provides a time series back to 2017. Notably, HS version 2012 included major revisions to the HS codes relevant to fisheries and aquaculture products. Changes in individual trade codes can be viewed [here](https://hstracker.wto.org/).

CEPII reconciles discrepancies by first removing transportation costs and using a weighting scheme based on each country's reporting reliability to average discrepancies in reported mirror flows, which occur in around 35% of observations (for all traded commodities). The BACI data focuses on trade flows between individual countries since 1994 and therefore drops flows within some groups of countries (e.g. Belgium-Luxembourg) to ensure consistent geographies. The resulting data set covers trade for over 200 countries and 5,000 products. Further details on the BACI data set are available in Gaulier and Zignago (2010).

## Live weight conversions

Global trade data is reported in terms of the product weight. In order to convert from product weight (i.e., net weight) to the live weight equivalent, a live weight conversion factor must be applied for each HS code. Live weight conversion factors are sourced from the European Market Observatory for Fisheries and Aquaculture Products (EUMOFA) (EUMOFA 2019), along with various national and international governmental report values. The EUMOFA data reports live weight conversion factors by CN-8 codes, so the mean of the live weight conversion factors falling within each HS 6-digit code are used here. The EUMOFA data assigns products primarily destined for industrial purposes (e.g., fish meal and fish oil), co-products (e.g., caviar) and live trade a value of zero. In this analysis, co-products retained a live weight conversion factor value of zero to avoid double counting, but live animal trade was assigned a live weight conversion factor of 1 and fish meal and fish oil was assigned an average value of 2.98 (Jackson 2009). Data compiled from reports were categorized into taxa types (mollusks, crustaceans, fishes, and other aquatic invertebrates), FAO ISSCAAP groups, species or taxon name, type of processing, and country of processing.

The live weight conversion factors applied to trade data introduce a potential source of uncertainty and error due to the fact that uncertainty in conversion factors is not reported and a single live weight conversion factor is presented per code, regardless of the species or region of origin. This is a limitation given that there is known geographical variation in live weight conversion factors due to differences in processing technology. Despite this limitation, the EUMOFA data offers better documentation and alignment with HS commodity codes than other live weight conversion factor data sources (Gephart et al. 2019) and is updated annually, providing documentation for changes in live weight conversion factors. Additionally, by supplementing the EUMOFA data  with the other reported values we can better capture specific species processing into various product forms and some regional variability.

All conversion factors were reported as live weight to product weight ratios.Estimated means and standard deviations for the live weight conversion factors are provided in XX. These conversion factors were mapped onto possible species to commodity or commodity to commodity conversions (described in section XX). For commodity to commodity conversions, we estimate the conversion factors (i.e., processing loss rate) as the additional mass lost when converting from the live weight to the original product form relative to converting from live weight to the processed product form. This can be calculated as the live weight conversion factor for the original product form divided by the live weight factor for the processed product form. We assume that mass cannot be gained through processing and set a minimum of 1% mass loss for processing conversion factors. Products which are not processed for export (i.e., are retained for domestic consumption) receive a value of -1 to account for them as consumption rather than export.

# Seafood production and commodity conversion

For each country, we estimate the proportion of each species going into each commodity and the proportion of each imported commodity processed into each other commodity for export. Each species can only be converted into a subset of the commodities. For example, Atlantic salmon, *Salmo salar*, can be converted into whole frozen salmon or frozen salmon fillets, but cannot be converted to a frozen tilapia fillet. Similarly, each commodity can only be converted to a subset of other commodities through processing. For example, whole frozen salmon can be processed into frozen salmon fillets, but not vice versa and neither salmon commodity can be converted to a tilapia commodity through processing. Defining possible conversions restricts the solution space to realistic results and improves estimation by substantially reducing the number of unknowns. We describe this assignment process below.

## Taxonomic group to commodity assignment

The species production to commodity assignment is a many-to-many matching problem, wherein one commodity can consist of multiple species and one species can be converted to multiple commodities. To identify predominantly marine, capture-sourced HS codes, it is necessary to identify the species falling within each code. All taxonomic groups reported in the FAO production data were therefore matched to the HS 6-digit codes based on the code descriptions and hierarchy. We use the HS code descriptions to identify possible species to commodity conversion.

The first matching step required dividing all taxonomic groups into the broad commodity groups at the 4-digit level (fish, crustaceans, molluscs and aquatic invertebrates). Within each of these groups, taxonomic groups were matched based on 6 types of matching categories:

1. Explicit taxa match - Matched based on taxonomic information provided in the code description
2. NEC match - All remaining unmatched species within the 4-digit level are assigned to the “not elsewhere considered” (NEC) code
3. NEC by taxa match - When a code description signifies an NEC group, but limits this based on a taxonomic category (e.g., Salmonidae, N.E.C.), the NEC grouping occurs at this level, rather than the broad NEC match
4. Broad commodity match - Only the broad taxonomic groups inform this assignment since no further taxonomic information is provided
5. Aquarium trade match - Assigned to ornamental species trade based on species found in the aquarium/ornamental trade (REF)
6. Fish meal and fish oil match - Assigned to fish meal and fish oil codes based on reported species used in fish oils and fish meals (REF)

All species are matched to at least one HS code.

## Commodity to commodity processing assignment

As with the species to commodity assignment, the commodity-to-commodity assignment is a many-to-many data problem. Here, one commodity can be processed into multiple other commodities (i.e., frozen salmon can be processed into salmon fillets or canned salmon), which also means one commodity could have come from multiple other commodities. To create these assignments, we established rules for which product transformations are technically possible. First, a product could not transfer outside of its broad commodity group (e.g., fish, crustaceans, mollusc, aquatic invertebrate). Second, where a more refined species or species group was given (e.g., tunas, salmons, etc.) a product cannot be transformed outside that group. Third, products are classified in terms of their state (e.g., alive, fresh, frozen, etc.) and presentation (e.g., e.g., whole, filleted, salted/dried/preserved meats, reductions such as fish meal and fish oil, etc.) and cannot be converted into less processed forms (e.g., frozen salmon fillets cannot turn into fresh salmon fillets, and frozen salmon fillets cannot turn into a frozen whole salmon). We ensure all commodities match to at least one taxon and all taxa match at least one commodity.

## Country standardization

Describe country standardization and point to country list, with territories included. Country list based on Comtrade documentation (https://comtrade.un.org/db/mr/rfReportersList.aspx)

# Network Estimation

Estimating species bilateral trade flows occurs in two steps: first, solving the national production-trade mass balance, and second, converting reported commodity trade flow estimates to species trade flow estimates based on the estimated species mix going into each commodity.

## National mass-balance

We start with the fact that exports must equal production and imports, minus consumption. Since exports are reported as commodities, we solve this mass balance problem in terms of commodities. Production data are reported for each species, so we estimate the elements of a matrix that represents the proportion of production going into each commodity. Since an imported commodity can be processed and exported as a different commodity, we also estimate the proportion of each import being converted into a different commodity (possible conversions described in section XX). Then for a given country,

$$e=V_1 \circ X \cdot p+V_2 \circ W \cdot g - c + \epsilon$$

If $n$ is the number of species and $m$ is the number of commodities, then: $V_1$ is a sparse ($m$ x $n$) matrix with product conversion factors corresponding to the unknowns in $X$; $X$ is a sparse ($m$ x $n$) matrix of the proportion of each species in each commodity; $p$ is a vector of domestic species production ($n$ x $1$); $V_2$ is a sparse ($m$ x $m$) matrix with product conversion factors corresponding to the entries of $W$; $W$ is a ($m$ x $m$) matrix of the processed imported commodities; g be a vector of imports ($m$ x $1$), $c$ is a vector of domestic consumption ($m$ x $1$), and; $\epsilon$ is a vector of error terms ($m$ x $1$).

We compiled reported values for $V_1$, $V_2$, $e$, $p$ and $g$, and estimate the entries of $X$, $W$, $c$, and $\epsilon$. We first converted this problem to a system of linear equations. Using the property that $vec(ABC) = (C^T \otimes A)vec(B)$, we can create $A_b = (y^T \otimes D_m)D_V$, where $D_m$ is a diagonal matrix of ones, with dimension $m$ and $D_V$ is a diagonal matrix with the elements of $vec(V)$. The vector of unknowns is then $x_b = vec(Z)$. We then solve this system of equations with a quadratic optimization solver such that the mass balance equalities are satisfied, trade codes with higher species resolution in $X$ are filled first, the elements of $X$ are otherwise as even as possible (i.e., we assume an even distribution of production among commodities unless the data suggests otherwise), the elements of $W$ are as uneven as possible (i.e., we assume imported products are for domestic consumption unless the data suggests otherwise), and that $\epsilon$ is as small as possible (i.e., minimize the error).

Positive error terms represent situations where reported production and imports cannot explain exports. This can occur due to under- or un-reported production or imports, over reporting of exports, errors in the live weight conversion factors, or inconsistencies in the year production and trade are attributed to.

We solve the mass-balance problem for each country-year-HS version combination using the *Python* package "solve_qp." The estimated species mixes in national production ($X$), proportion of imports processed for export (off-diagonal values of $W$) and the error term ($\epsilon$) are passed to the next stage of the analysis.

## Converting the product trade network to a species trade network

To convert reported bilateral trade of seafood products to species trade flows, we first identify the mix of species going into each trade code. To do this, we first must reweight $X$ so it represents the proportion of each species in a given code rather than the proportion of production of a species going into each product form. To do this, we take each country's estimated $X$, multiply by  production, and divide by the column sums.

Next, we calculate the proportion of exports sourced from national production, imported and processed products, and from an unknown source (i.e., the error term). Since the mix of these sources cannot be derived from the mass balance equation alone, we calculate a range for the sourcing following Asche et al. (2022). To calculate the maximum exports from domestic production,  We apply the national reweighted $X$ to the exports from national production and the reweighted $X$ of trade partners (weighted based on import volume of each trade code) to import-sourced exports. Since the exports are of an unknown origin, we assign the most resolved common taxonomic name for the species falling within the code to the exports derived from the error term. [ADD EXAMPLE]

## Analysis implementation

- Add details about implementation in R and on Zorro

## Network post-estimation processing

- Removal of small trade flows
- Dealing with different HS versions

# Estimated network checks

## Comparison of total trade volumes to other reported values

## Comparison of trade patterns with simplified versions of the model

## Agreement across HS versions

## Comparison of production and trade data

# Conversion to nutrient flows

Questions for Daniel: 
- I heard AFCD has a species ID number - what system is used? 
- Key questions will be:
  - What product forms should we use? 
    - Start with muscle tissue but need to talk to Chris 
  - How should we aggregate up for genus, family, etc. level observations? 
    - 
  - Handling missing values
    - Fill in with a hierarchical approach - if no direct match, go with genus, etc. 
- Note: Initially, we just need nutrient flows for the live weight equivalents, but we could incorporate product forms later

# Planned future improvements

- Add economic realism
- Potentially incorporate more detailed national trade statistics
- Consider constructing matrix analogous to Xq for W that would push product transformations toward staying within a species group before using NEC codes to explain mass
- Add in FAO 2020 data - BACI data is already available and FAO data is also available, but we would need to test for structural chances in the data release before incorporating

# Integration with SAU

- Double check that producer corresponds to fishing entity
  - Yep!
- Marine has capture and mariculture, freshwater won't be available until the end of the year (she can provide a species list)
- Fishing entity, species, year, catch, gear, reporting status (FAO or reconstructed), sector, end use
- Deng will provide taxon table in addition to the catch data
- Send species list 
- Use a lineage approach for species groupings (prompt to look at peneus as well when they click peneus)
- Create species and country profile pages that link with SAU and Sea Life and Fish Base

# Outstanding issues

- Resolve whether to allow imports staying the same code to count toward exports or not (i.e., technical definition). This needs to be consistent across solve_country and make_snet
- Add trade flow confidence score

# Preliminary plots to add

- Network metrics over time by capture vs aquaculture
  - Average degree
  - Measure of node and edge distribution
  - Community detection of cliches over time (how does membership change)
  - Centrality 
  - Edge density
-

### 2022-08-10 SI.Rmd
initial commit of exploratory figures and methods

---
title: "Seafood Globalization SI"
author: "Jessica Gephart"
date: '2022-06-24'
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(kableExtra)
```

# Optimization diagnostics

# ARTIS compared with BACI

```{r baci_artis_compare}
# Directory file paths
baci_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs"
artis_dir <- "/Volumes/jgephart/ARTIS/Outputs/S_net/20220614_ARTIS-timeseries/artis_ts"
artis_filename <- "mid_artis_ts.csv"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/S_net/20220614_ARTIS-timeseries"

# Read in ARTIS totals
artis <- read.csv(file.path(artis_dir, artis_filename))
artis_custom <- read.csv(file.path(outdir, "custom_artis_ts.csv"))

# Read in compiled BACI totals (output from compare_baci_snet.R)
baci <- read.csv("/Volumes/jgephart/ARTIS/Outputs/baci_totals_allhs.csv")

```

After filtering to include the same set of product codes, total estimated ARTIS trade flows follow BACI trade flows closely (SI Fig X). Since a threshold is imposed to remove small trade flows in the ARTIS model, a small volume loss is expected. To balance using the most recent trade code version with the fact that total volumes lag for several years compared to total volumes reported under the previous code, we created a composite time series consisting of select years and code versions (SI Table X).

```{r}
kable(artis_custom %>% 
        select(year, hs_version), col.names = c("Year", "HS version"))
```

```{r plot_baci_artis, fig.cap = "Comparison of total export volume by HS code version for BACI (solid lines) and ARTIS (dashed lines). The solid black line indicates the composite ARTIS time series."}
ggplot() +
  geom_line(baci, mapping = aes(x = year, y = product_weight_t, color = factor(as.numeric(hs_version)))) +
  geom_line(artis, mapping = aes(x = year, y = product_weight_t, color = factor(hs_version)), linetype = "dashed") +
  geom_line(artis_custom, mapping = aes(x = year, y = product_weight_t), color = "black") +
  labs(x = "", y = "Exports (t, product weight)", color = "HS version") +
  theme_bw()
```

FIX IT: the total weight differences are greater than we estimated previously, likely due to the differences in what is estimated with quadprog versus cvxopt

```{r compare_custom_ts}
baci_artis <- baci %>%
  mutate(hs_version = as.numeric(hs_version)) %>%
  rename("baci_product_weight_t" = "product_weight_t") %>%
  right_join(artis_custom %>% 
              rename("artis_product_weight_t" = "product_weight_t"), 
            by = c("year", "hs_version")) %>%
  mutate(percent_diff = 100*(baci_product_weight_t-artis_product_weight_t)/baci_product_weight_t)

ggplot(baci_artis, aes(x = year, y = percent_diff)) +
  geom_line() + 
  labs(x = "", y = "Percent difference") +
  theme_bw()
```

# Species presence in trade

# Estimated range for import processing

# References

### 2022-08-10 prelim_results.Rmd
initial commit of exploratory figures and methods

---
title: "Seafood globalization preliminary results"
author: "Jessica Gephart"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)

```

```{r load_database_data}
# Load ARTIS data from SQL local database
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("POSTGRES_DB"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("POSTGRES_USER"),
                 password=Sys.getenv("POSTGRES_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# Pull all ARTIS and production data
artis <- dbGetQuery(con, 
"SELECT * FROM snet 
WHERE (snet_est = \'mid\') AND
((\"year\" <= 2003 AND hs_version = \'HS96\') OR
((\"year\" > 2003 AND \"year\" <= 2009) AND hs_version = \'HS02\') OR
((\"year\" > 2009 AND \"year\" <= 2012) AND hs_version = \'HS07\') OR
((\"year\" > 2012 AND \"year\" <= 2020) AND hs_version = \'HS12\'))")

artis <- artis %>%
  select(-record_id, -snet_est)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id)

# Close database connection
dbDisconnect(con)

rm(list = c("con"))
```

```{r set_directories}
# Set time series dates
date_nodash <- "20220614"
date_dash <- "2022-06-14"

# Set directories 
# datadir <-  paste("/Volumes/jgephart/ARTIS/Outputs/S_net/", date_nodash, "_ARTIS-timeseries", sep="")
model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_archive_20220718"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"

# Set color palettes
# All colors from palette: #FFE6BE (light yellow), #F7AF75 (light orange), #F37542 (medium orange), 
# #E24027 (dark orange), #0F2D59 (navy), #1B708F (blue), #2E8D9A (teal), #86ADA7 (light blue)
method_colors <- c("#F7AF75", "#2E8D9A", "#86ADA7") # light orange, teal, light blue
source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue

```

```{r load_drive_data}
code_max_resolved_taxa <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/code_max_resolved_taxa.csv") %>%
              mutate(hs6 = as.character(hs6))

sciname_metadata <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/sciname_metadata.csv")

```

```{r join_data}
# Load cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method" = "prod_method", "environment", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, environment) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

# Join production data to ARTIS
artis <- artis %>%
  filter(live_weight_t >= 0.1) %>%
# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                           "sciname", "method", "environment")) %>%
  ungroup()

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  # FIX IT: Change metadata to HSXX format and then remove this line
  mutate(hs_version = as.integer(parse_number(hs_version))) %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = c("sciname" = "SciName")) %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  ))

# Remove data frames after merged with ARTIS
rm(list = c("code_max_resolved_taxa", "sciname_metadata"))

```

# Trend in global seafood trade

```{r global_total_trend}
artis %>% 
  group_by(year) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000)) +
  geom_line() +
  labs(x = "", y = "Seafood exports (million t, live weight)") +
  theme_minimal()

total_1996 <- round(sum(artis %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(artis %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)
```

Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.

```{r global_species_pattern}
n_species <- artis %>% 
  group_by(year, sciname_hs_modified) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  filter(live_weight_t > 1000) %>%
  group_by(year) %>% 
  tally()

n_species %>%
  ggplot(aes(x = year, y = n)) +
  geom_line() +
  labs(x = "", y = "Number of species/groups with >1000 t exported") +
  theme_minimal()

n_species_1996 <- sum(n_species %>% filter(year == 1996) %>% pull(n))
n_species_2019 <- sum(n_species %>% filter(year == 2019) %>% pull(n))

```

The number of species/species groups identified in exports increased from `r n_species_1996` m in 1996 to `r n_species_2019` in 2019, representing a `r round(100*((n_species_2019-n_species_1996)/n_species_1996),0)`% increase over that period. Increases in species/species groups exported can arise from either an increase in the species traded internationally or from an increase in the resolution of the trade and production data over time.

# Top traded species

```{r exports_by_species}
# Top species exported
top_species_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 1) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 species \nexported (1996-2000)", fill = "Production method") +
  theme_minimal()

top_species_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 1) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 species \nexported (2015-2019)", fill = "Production method") +
  theme_minimal()

# Top species groups exported
top_species_groups_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 0) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 groups \nexported (1996-2000)", fill = "Production method") +
  theme_minimal()

top_species_groups_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  filter(str_count(sciname_hs_modified, pattern = " ") == 0) %>% 
  group_by(sciname_hs_modified, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(sciname_hs_modified) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(sciname_hs_modified = fct_reorder(as.factor(sciname_hs_modified), total)) %>%
  slice_max(n = 31, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = sciname_hs_modified, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 groups \nexported (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_species_groups_1, top_species_groups_2, common.legend = TRUE)
ggarrange(top_species_1, top_species_2, common.legend = TRUE)

```

# How does capture versus aquaculture trade globalization compare in terms of quantity and connectedness and how has this changed over time?

```{r}
# Stacked lines, colored by method
artis %>% 
  group_by(year, method) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = method)) +
  geom_area() +
  scale_fill_manual(values = method_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", "Production method") +
  theme_minimal()

# Line color by method
artis %>% 
  group_by(year, method) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = method)) +
  geom_line() +
  scale_color_manual(values = method_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Production method") +
  theme_minimal()

# Summary stats
## Percent of exports by source
kable(artis %>% 
        group_by(year, method) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        mutate(percent = round(100*live_weight_t/sum(live_weight_t), 1)) %>%
        select(year, method, percent) %>% 
        pivot_wider(names_from = method, values_from = percent),
      caption = "Percent of exports by production source")

## Number of exporters by source
kable(artis %>% 
        select(year, exporter_iso3c, method) %>%
        distinct() %>%
        group_by(year, method) %>%
        tally() %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Number of exporters by production source")

## Number of importers by source
kable(artis %>% 
        select(year, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, method) %>%
        tally() %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Number of importers by production source")

# Average out degree by source
kable(artis %>% 
        select(year, exporter_iso3c, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, exporter_iso3c, method) %>%
        tally() %>%
        group_by(year, method) %>%
        summarise(n = round(mean(n), 1)) %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Average out degree by production source")

# Out degree distribution by year
artis %>% 
  select(year, exporter_iso3c, importer_iso3c, method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, method) %>%
  tally() %>%
  ggplot(aes(n, fill = method)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = method_colors) +
  facet_wrap(~year) +
  labs(title = "Out degree distribution by method", fill = "Production method")

# Average in degree by source
kable(artis %>% 
        select(year, exporter_iso3c, importer_iso3c, method) %>%
        distinct() %>%
        group_by(year, importer_iso3c, method) %>%
        tally() %>%
        group_by(year, method) %>%
        summarise(n = round(mean(n), 1)) %>%
        select(year, method, n) %>% 
        pivot_wider(names_from = method, values_from = n),
      caption = "Average in degree by production source")

# In degree distribution by year
artis %>% 
  select(year, exporter_iso3c, importer_iso3c, method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, method) %>%
  tally() %>%
  ggplot(aes(n, fill = method)) +
  geom_density(alpha = .5) +
  scale_fill_manual(values = method_colors) +
  facet_wrap(~year) +
  labs(title = "Out degree distribution by method", fill = "Production method")

```

# How have regional trade patterns changed over time?

```{r}
# Top exporters, averaged over beginning and end time periods
top_exp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_exp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_exp_1, top_exp_2, common.legend = TRUE)

# Top importers, averaged over beginning and end time periods
top_imp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_imp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_imp_1, top_imp_2, common.legend = TRUE)

```

```{r regional_bilateral_trends}
artis_region <- artis %>%
  mutate(importer_region = countrycode(importer_iso3c, origin = "iso3c", destination = "region"),
         exporter_region = countrycode(exporter_iso3c, origin = "iso3c", destination = "region")) %>%
  group_by(year, exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t))

artis_region %>%
  filter(!is.na(exporter_region), !is.na(importer_region)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000)) +
  geom_line() +
  labs(x = "", y = "Exports (million t)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_grid(exporter_region~importer_region, scales = "free",
             labeller = label_wrap_gen(width = 5))

```

```{r top_bilateral_flows}
top_flows_1 <- artis %>%
  filter(year %in% 1996:2000) %>% 
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  arrange(desc(live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  mutate(exporter_to_importer = paste(
    countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name"), 
    "to",
    countrycode(importer_iso3c, origin = "iso3c", destination = "country.name"),
    sep = " ")) %>%
  mutate(exporter_to_importer = fct_reorder(exporter_to_importer, live_weight_t)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_to_importer)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 25 bilateral \nflows (1996-2000)", x = "Average trade (million t)", y = "") +
  theme_bw()

top_flows_2 <- artis %>%
  filter(year %in% 2015:2019) %>% 
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  arrange(desc(live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  mutate(exporter_to_importer = paste(
    countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name"), 
    "to",
    countrycode(importer_iso3c, origin = "iso3c", destination = "country.name"),
    sep = " ")) %>%
  mutate(exporter_to_importer = fct_reorder(exporter_to_importer, live_weight_t)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_to_importer)) +
  geom_bar(stat = "identity") +
  labs(title = "Top 25 bilateral \nflows (2015-2019)", x = "Average trade (million t)", y = "") +
  theme_bw()

ggarrange(top_flows_1, top_flows_2, nrow = 1)

```

# Exports from domestic, imported and unknown sources

```{r exports_by_source}
# Stacked lines, colored by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = dom_source)) +
  geom_area() +
  scale_fill_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

# Line color by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

## Percent of exports by source
kable(artis %>% 
        group_by(year, dom_source) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        mutate(percent = round(100*live_weight_t/sum(live_weight_t), 1)) %>%
        select(year, dom_source, percent) %>% 
        pivot_wider(names_from = dom_source, values_from = percent),
      caption = "Percent of exports by export source")
```

# How has foreign dependence evolved over time?

```{r exports_percent_by_source}
annual_total_prod <- prod %>%
  select(year, iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(year, sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

# Stacked lines, colored by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(annual_total_prod, by = "year") %>%
  mutate(percent_export = 100*live_weight_t/production_t) %>%
  ggplot(aes(x = year, y = percent_export, fill = dom_source)) +
  geom_area() +
  scale_fill_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

# Line color by export source
artis %>% 
  group_by(year, dom_source) %>%
  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
  left_join(annual_total_prod, by = "year") %>%
  mutate(percent_export = 100*live_weight_t/production_t) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = dom_source)) +
  geom_line() +
  scale_color_manual(values = source_colors) +
  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "Export source") +
  theme_minimal()

## Percent of exports by source
kable(artis %>% 
        group_by(year, dom_source) %>%
        summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
        ungroup() %>%
        left_join(annual_total_prod, by = "year") %>%
        mutate(percent = round(100*live_weight_t/production_t, 1)) %>%
        select(year, dom_source, percent) %>% 
        pivot_wider(names_from = dom_source, values_from = percent),
      caption = "Percent of exported production by export source")
```

```{r exports_percent_by_species}
prod_1 <- artis %>%
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>%
  select(exporter_iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_1 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>% 
  group_by(sciname, CommonName) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_1, by = c("sciname")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(CommonName = fct_reorder(as.factor(CommonName), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  ggplot(aes(x = percent_exported, y = CommonName)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported as domestic export", y = "",
       title = "Top 25 species \n(1996-2000)") +
  theme_minimal()

prod_2 <- artis %>%
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>%
  select(exporter_iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(sciname) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_2 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>% 
  group_by(sciname, CommonName) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_2, by = c("sciname")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(CommonName = fct_reorder(as.factor(CommonName), percent_exported)) %>%
  ggplot(aes(x = percent_exported, y = CommonName)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported as domestic export", y = "",
       title = "Top 25 species \n(2015-2019)") +
  theme_minimal()

ggarrange(percent_export_1, percent_export_2, common.legend = TRUE)

```

```{r exports_percent_by_country}
prod_1 <- prod %>%
  filter(year %in% 1996:2000) %>%
  select(iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(iso3c) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_1 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 1996:2000) %>% 
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_1, by = c("exporter_iso3c" = "iso3c")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  ggplot(aes(x = percent_exported, y = exporter_iso3c)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported \nas domestic export", y = "",
       title = "Top 25 countries \n(1996-2000)") +
  theme_minimal()

prod_2 <- prod %>%
  filter(year %in% 2015:2019) %>%
  select(iso3c, sciname, production_t) %>%
  distinct() %>%
  group_by(iso3c) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE))

percent_export_2 <- artis %>% 
  filter(dom_source == "domestic export", 
         year %in% 2015:2019) %>% 
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_2, by = c("exporter_iso3c" = "iso3c")) %>%
  mutate(percent_exported = 100*live_weight_t/production_t) %>%
  ungroup() %>%
  slice_max(n = 25, order_by = percent_exported) %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), percent_exported)) %>%
    mutate(percent_exported = case_when(
    percent_exported > 100 ~ 100,
    TRUE ~ percent_exported
  )) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  ggplot(aes(x = percent_exported, y = exporter_iso3c)) +
  geom_bar(stat="identity", position = "stack") + 
  labs(x = "Percent of production exported \nas domestic export", y = "",
       title = "Top 25 countries \n(2015-2019)") +
  theme_minimal()

ggarrange(percent_export_1, percent_export_2, common.legend = TRUE)

```

```{r calculate_supply}
exports <- artis %>%
  group_by(year, exporter_iso3c, dom_source, sciname, CommonName, method, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
  pivot_wider(names_from = "dom_source", values_from = "live_weight_t")

imports <- artis %>%
  group_by(year, importer_iso3c, sciname, CommonName, method, environment) %>%
  summarise(imports_live_weight_t = sum(live_weight_t))

supply <- exports %>%
  full_join(imports, by = c("year", "exporter_iso3c" = "importer_iso3c", 
                            "sciname", "CommonName", "method", "environment")) %>%
  full_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                          "sciname", "method", "environment"))

supply$foreign_export[is.na(supply$foreign_export)] <- 0
supply$domestic_export[is.na(supply$domestic_export)] <- 0
supply$error_export[is.na(supply$error_export)] <- 0
supply$imports_live_weight_t[is.na(supply$imports_live_weight_t)] <- 0
supply$production_t[is.na(supply$production_t)] <- 0

supply <- supply %>%
  mutate(supply = production_t + imports_live_weight_t - domestic_export - foreign_export - error_export,
         supply_no_error = production_t + imports_live_weight_t - domestic_export - foreign_export, 
         supply_domestic = production_t - domestic_export,
         supply_foreign = imports_live_weight_t - foreign_export) %>%
  rename("iso3c" = "exporter_iso3c") %>%
  ungroup() %>%
  mutate(supply_no_error = replace_na(supply_no_error, 0),
         supply_domestic = replace_na(supply_domestic, 0),
         supply_foreign = replace_na(supply_foreign, 0))

```

```{r supply_source_by_country}
select_countries <- c("CHN", "USA", "CHL", "GHA", "AUS", "NOR", "SWE", "MEX", "THA")

# Line graph of total supply for select countries
supply %>% 
  filter(iso3c %in% select_countries) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(iso3c, year) %>%
  summarise(supply_no_error = sum(supply_no_error)) %>%
  ggplot(aes(x = year, y = supply_no_error)) +
  geom_line() +
  labs(x = "", y = "Supply (t)") +
  facet_wrap(~iso3c) +
  theme_bw()

# Percent of supply by source for select countries
supply %>% 
  filter(iso3c %in% select_countries) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(iso3c, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  ggplot() +
  geom_line(aes(x = year, y = supply_domestic), color = "#1B708F") +
  geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
  facet_wrap(~iso3c) +
  theme_bw()

supply %>% 
  mutate(region = countrycode(iso3c, origin = "iso3c", destination = "region")) %>%
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  group_by(region, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  filter(region != "NA") %>%
  ggplot() +
  geom_line(aes(x = year, y = supply_domestic), color = "#1B708F") +
  geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
  facet_wrap(~region) +
  theme_bw()

supply %>% 
  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error)) %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  #filter(sum(supply_no_error) > 100000) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  arrange(desc(supply_foreign))

```

```{r supply_source_by_species}
supply_species <- supply %>% 
  group_by(sciname, CommonName, year) %>%
  mutate(total_production_t = sum(production_t)) %>%
  summarise(percent_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            percent_foreign = 100*sum(supply_foreign)/sum(supply_no_error), 
            production_t = sum(production_t)) %>%
  arrange(desc(percent_foreign))

supply_species %>% 
  group_by(sciname, CommonName) %>% 
  mutate(total_production_t = sum(production_t)) %>%
  ungroup() %>%
  slice_max(n = 16*length(unique(supply_species$year)), order_by = total_production_t) %>%
  ggplot() +
  geom_line(aes(x = year, y = percent_domestic), color = "#1B708F") + 
  geom_line(aes(x = year, y = percent_foreign), color = "#E24027") +
  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
  facet_wrap(~CommonName) +
  theme_bw()

```

# Trends in estimated error

```{r explore_country_error}
# Plot time series of species with largest volume error exports
supply %>%
  group_by(year, iso3c) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  group_by(iso3c) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_percent)) +
  geom_line() +
  labs(x = "", y = "Percent exports from unknown source") +
  facet_wrap(~iso3c) +
  theme_bw()

supply %>%
  group_by(year, iso3c) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  group_by(iso3c) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_export)) +
  geom_line() +
  labs(x = "", y = "Exports from unknown source (t, live weight)") +
  facet_wrap(~iso3c) +
  theme_bw()

```

```{r explore_species_error}
# Plot time series of species with largest volume error exports
supply %>%
  group_by(year, sciname, CommonName) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  # Remove groupings only used by error 
  filter(!(sciname %in% c("animalia", "actinopteri"))) %>%
  group_by(sciname, CommonName) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_percent)) +
  geom_line() +
  labs(x = "", y = "Percent exports from unknown source") +
  facet_wrap(~CommonName) +
  theme_bw()

supply %>%
  group_by(year, sciname, CommonName) %>%
  summarise(error_export = sum(error_export, na.rm = TRUE), 
            domestic_export = sum(domestic_export, na.rm = TRUE),
            foreign_export = sum(foreign_export, na.rm = TRUE),
            error_percent = 100*sum(error_export, na.rm = TRUE)/(sum(domestic_export, na.rm = TRUE) + 
                                                     sum(foreign_export, na.rm = TRUE) + 
                                                     sum(error_export, na.rm = TRUE))) %>%
  # Remove groupings only used by error 
  filter(!(sciname %in% c("animalia", "actinopteri"))) %>%
  group_by(sciname, CommonName) %>%
  mutate(total_error = sum(error_export)) %>%
  ungroup() %>%
  slice_max(n = 12*length(unique(supply$year)), order_by = total_error) %>%
  ggplot(aes(x = year, y = error_export)) +
  geom_line() +
  labs(x = "", y = "Exports from unknown source (t, live weight)") +
  facet_wrap(~CommonName) +
  theme_bw()

```

# Processing hotspots

```{r}
artis %>% 
  filter(dom_source == "foreign export",
         year %in% 2015:2019) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)/5) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, live_weight_t)) %>%
  slice_max(n = 25, order_by = live_weight_t) %>%
  ggplot(aes(x = live_weight_t, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average exports of foreign origin (2015-2019)",
       x = "Foreign exports (t)", y = "") +
  theme_bw()

artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c) %>%
  mutate(total_export = sum(live_weight_t)) %>%
  ungroup() %>%
  filter(dom_source == "foreign export") %>%
  group_by(exporter_iso3c) %>%
  summarise(export_percent_foreign = 100*sum(live_weight_t)/mean(total_export)) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, export_percent_foreign)) %>%
  slice_max(n = 25, order_by = export_percent_foreign) %>%
  ggplot(aes(x = export_percent_foreign, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average percent exports of foreign origin (2015-2019)",
       x = "Foreign exports as percent of total exports", y = "") +
  theme_bw()

artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c) %>%
  mutate(total_export = sum(live_weight_t)) %>%
  ungroup() %>%
  filter(dom_source == "foreign export") %>%
  group_by(exporter_iso3c) %>%
  summarise(export_percent_foreign = 100*sum(live_weight_t)/mean(total_export)) %>%
  mutate(country_name = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
  mutate(country_name = fct_reorder(country_name, export_percent_foreign)) %>%
  slice_max(n = 25, order_by = export_percent_foreign) %>%
  ggplot(aes(x = export_percent_foreign, y = country_name)) +
  geom_bar(stat = "identity") +
  labs(title = "Average percent of imports re-exported (2015-2019)",
       x = "Foreign exports as percent of total imports", y = "") +
  theme_bw()

```

### 2022-08-10 Exploratory_figs.R

@@ -1,0 +1,167 @@
+# Preliminary analysis
+
+# Load packages
+library(tidyverse)
+library(countrycode)
+library(circlize)
+source("Oceana_report_functions/summarize_S_net_functions.R")
+
+# Set directories 
+datadir <-  "/Volumes/jgephart/ARTIS/Outputs"
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/Oceana Report"
+S_net_file <- "S_net/2021-03-18_S-net_2016-2018_mean-quantity_mean-source.csv"
+model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs"
+
+# Load data
+S_net <- load_clean_dat(file.path(datadir, S_net_file)) 
+S_net <- S_net %>%
+  select(-X) %>%
+  filter(quantity > 1)
+
+prod_data <- read.csv(file.path(model_inputs_dir, "clean_fao_prod.csv"))
+
+# Chord diagram for each production source/environment combination
+region.colors <- c("#142B58", # Navy 
+                   "#792560", # Purple
+                   "#F35D2D", #Other option: "#13808F" # Teal
+                   "#39A584", # Sea Green
+                   "#C78F0B", # Gold
+                   "#355936", # Forest Green
+                   "#0686E5") # Light Blue "#00FFFF" # Cyan 
+## All trade network
+plot_df <- S_net %>%
+  group_by(exporter_region, importer_region) %>%
+  summarise(trade = sum(quantity)) 
+
+plot_df <- abbrev_region(plot_df)
+
+chordDiagram(plot_df,
+             grid.col = region.colors,
+             annotationTrack = c("name", "grid"),
+             directional = 1,
+             direction.type = c("diffHeight", "arrows"),
+             link.arr.type = "big.arrow",
+             link.target.prop = FALSE,
+             diffHeight = 0.08,
+             link.sort = TRUE)
+title("All marine and inland trade")
+circos.clear()
+
+## Marine capture trade network
+plot_df <- S_net %>%
+  filter(prod_method == "Capture", habitat == "Marine") %>%
+  group_by(exporter_region, importer_region) %>%
+  summarise(trade = sum(quantity)) 
+
+plot_df <- abbrev_region(plot_df)
+
+chordDiagram(plot_df,
+             grid.col = region.colors,
+             annotationTrack = c("name", "grid"),
+             directional = 1,
+             direction.type = c("diffHeight", "arrows"),
[…]

### 2022-08-10 Exploratory_figs.R
move exploratory figs script to Archived_code folder

@@ -1,167 +1,0 @@
-# Preliminary analysis
-
-# Load packages
-library(tidyverse)
-library(countrycode)
-library(circlize)
-source("Oceana_report_functions/summarize_S_net_functions.R")
-
-# Set directories 
-datadir <-  "/Volumes/jgephart/ARTIS/Outputs"
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/Oceana Report"
-S_net_file <- "S_net/2021-03-18_S-net_2016-2018_mean-quantity_mean-source.csv"
-model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs"
-
-# Load data
-S_net <- load_clean_dat(file.path(datadir, S_net_file)) 
-S_net <- S_net %>%
-  select(-X) %>%
-  filter(quantity > 1)
-
-prod_data <- read.csv(file.path(model_inputs_dir, "clean_fao_prod.csv"))
-
-# Chord diagram for each production source/environment combination
-region.colors <- c("#142B58", # Navy 
-                   "#792560", # Purple
-                   "#F35D2D", #Other option: "#13808F" # Teal
-                   "#39A584", # Sea Green
-                   "#C78F0B", # Gold
-                   "#355936", # Forest Green
-                   "#0686E5") # Light Blue "#00FFFF" # Cyan 
-## All trade network
-plot_df <- S_net %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(trade = sum(quantity)) 
-
-plot_df <- abbrev_region(plot_df)
-
-chordDiagram(plot_df,
-             grid.col = region.colors,
-             annotationTrack = c("name", "grid"),
-             directional = 1,
-             direction.type = c("diffHeight", "arrows"),
-             link.arr.type = "big.arrow",
-             link.target.prop = FALSE,
-             diffHeight = 0.08,
-             link.sort = TRUE)
-title("All marine and inland trade")
-circos.clear()
-
-## Marine capture trade network
-plot_df <- S_net %>%
-  filter(prod_method == "Capture", habitat == "Marine") %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(trade = sum(quantity)) 
-
-plot_df <- abbrev_region(plot_df)
-
-chordDiagram(plot_df,
-             grid.col = region.colors,
-             annotationTrack = c("name", "grid"),
-             directional = 1,
-             direction.type = c("diffHeight", "arrows"),
[…]

### 2022-08-12 sciname_analysis.R
sciname counts and percents traded vs produced

@@ -1,0 +1,79 @@
+
+library(tidyverse)
+library(DBI)
+library(rmarkdown)
+library(countrycode)
+
+# Establish connection to database
+con <- dbConnect(
+  RPostgres::Postgres(),
+  dbname = Sys.getenv("DB_NAME"),
+  host = Sys.getenv("localhost"),
+  port = Sys.getenv("DB_PORT"),
+  user = Sys.getenv("DB_USERNAME"),
+  password = Sys.getenv("DB_PASSWORD")
+)
+
+# Get initial data pull from database
+artis <- dbGetQuery(con,
+'SELECT hs_version, year, sciname, SUM(live_weight_t)
+FROM snet WHERE (snet_est = \'mid\')
+GROUP BY hs_version, year, sciname')
+
+# Production Data
+prod <- dbGetQuery(con,
+                   'SELECT * FROM production')
+
+# Species / Species Groups metadata
+sciname <- dbGetQuery(con,
+                      'SELECT * FROM sciname')
+
+# Disconnect from database
+dbDisconnect(con)
+
+# Threshold for amount traded or produced
+threshold <- 1000
+
+trade_sciname <- artis %>%
+  rename(live_weight_t = sum) %>%
+  filter(live_weight_t >= threshold)
+
+prod_sciname <- prod %>%
+  group_by(year, sciname) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  filter(live_weight_t >= threshold) %>%
+  rename(production = live_weight_t) %>%
+  pivot_longer(cols = c(production), names_to = "hs_version", values_to = "live_weight_t")
+
+sciname_summary <- trade_sciname %>%
+  full_join(prod_sciname) %>%
+  rename("Source" = hs_version)
+
+sciname_counts <- sciname_summary %>%
+  group_by(year, Source) %>%
+  tally() %>%
+  pivot_wider(names_from = Source, values_from = n)
+
+write.csv(sciname_summary, "sciname_counts_summary.csv", row.names = FALSE)
+
+sciname_counts %>%
+  pivot_longer(-year, names_to = "Source", values_to = "n") %>%
+  ggplot(aes(x = year, y = n, color = Source)) +
+  geom_line() +
+  labs(x = "Year", y = "")
+
+
+sciname_percents <- sciname_counts %>%
+  mutate_at(.vars = vars(HS96, HS02, HS07, HS12, HS17),
+            .funs = list(~((. / production) * 100)))
+
[…]

### 2022-08-19 Methods.Rmd
add ms figures file

@@
 
+###______________________________________________________________________________________###
+Old text
 To convert reported bilateral trade of seafood products to species trade flows, we first identify the mix of species going into each trade code.
@@
 Since the exports are of an unknown origin, we assign the most resolved common taxonomic name for the species falling within the code to the exports derived from the error term.
-[ADD EXAMPLE]
+###______________________________________________________________________________________###
+
+To convert bilateral trade data expressed as commodities into species flows, we must trace flows back to their source country to link the commodities with the appropriate species mix.
+To do this, we draw from multi-regional input-output methods and first construct a flow matrix $F$, with sectors representing all country-commodity combinations and the origin ("from") sector on the row and destination ("to") sector on the column.
+To construct $F$, we multiply the estimated conversion of products for each country $i$ ($W_i$) element wise by the vector of imports by country-commodity combination, expressed as live weight equivalents.
+This then represents the bilateral flow of product entering each country and remaining in its current form and being processed into a new form according to the appropriate estimated national $W$ matrix.
+
+Our final demand vector is comprised of domestic production of each commodity ($X_i p_i$) plus imports of each commodity expressed as the final product form ($W_i g$) plus the estimated error export quantity ($\epsilon$).
 

### 2022-08-19 ms_figures.Rmd
add ms figures file

---
title: "ms_figures"
author: "Jessica Gephart"
date: "2022-08-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)
library(vegan)

```

```{r load_database_data}
# Load ARTIS data from SQL local database
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("POSTGRES_DB"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("POSTGRES_USER"),
                 password=Sys.getenv("POSTGRES_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# Pull all ARTIS and production data
artis <- dbGetQuery(con, "SELECT * FROM snet")

artis <- artis %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id) %>%
  # Remove 2020 for now
  filter(year < 2020)

# Close database connection
dbDisconnect(con)

rm(list = c("con"))
```

```{r set_directories}
# Set time series dates
date_nodash <- "20220614"
date_dash <- "2022-06-14"

# Set directories 
# datadir <-  paste("/Volumes/jgephart/ARTIS/Outputs/S_net/", date_nodash, "_ARTIS-timeseries", sep="")
model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_archive_20220718"
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"

# Set color palettes
# All colors from palette: #FFE6BE (light yellow), #F7AF75 (light orange), #F37542 (medium orange), 
# #E24027 (dark orange), #0F2D59 (navy), #1B708F (blue), #2E8D9A (teal), #86ADA7 (light blue)
method_colors <- c("#F7AF75", "#2E8D9A", "#86ADA7") # light orange, teal, light blue
source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue

```

```{r load_drive_data}
code_max_resolved_taxa <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/code_max_resolved_taxa.csv") %>%
              mutate(hs6 = as.character(hs6))

sciname_metadata <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/sciname_metadata.csv")

```

```{r join_data}
# Load cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method" = "prod_method", "environment", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, environment) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

# Join production data to ARTIS
artis <- artis %>%
  filter(live_weight_t >= 0.1) %>%
# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                           "sciname", "method", "environment")) %>%
  ungroup()

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  # FIX IT: Change metadata to HSXX format and then remove this line
  mutate(hs_version = as.integer(parse_number(hs_version))) %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = c("sciname" = "SciName")) %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  ))

# Remove data frames after merged with ARTIS
rm(list = c("code_max_resolved_taxa", "sciname_metadata"))

```

# Figure 1: Increases in global export of blue foods

## Figure 1 a: Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019

## Figure 1 b: Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports

## Figure 1 c: Degree or measure of degree distribution by production method and environment

# Figure 2: Trade patterns

```{r SI_top_traders}
total_country_exports <- artis %>% 
  group_by(exporter_iso3c, method, environment) %>%
  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
  rename("iso3c" = "exporter_iso3c")

total_country_imports <- artis %>% 
  group_by(importer_iso3c, method, environment) %>%
  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
  rename("iso3c" = "importer_iso3c")

total_country_trade <- total_country_exports %>%
  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(total = sum(trade_live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 9*20, order_by = total)

total_net_import <- total_country_trade %>%
  group_by(iso3c) %>%
  summarise(export_live_weight_t = sum(export_live_weight_t),
            import_live_weight_t = sum(import_live_weight_t)) %>%
  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
  select(iso3c, net_import)

total_country_trade %>% 
  left_join(total_net_import, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
  ggplot() +
  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  scale_fill_manual(values = method_colors) +
  geom_line(aes(y = net_import, x = iso3c), group = 1) +
  geom_hline(yintercept = 0) +
  labs(x = "", y = "Trade (million t, live weight)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

```{r top_imp_exp}
# Top exporters, averaged over beginning and end time periods
top_exp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_exp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(exporter_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood exports (million t, live weight)", y = "",
       title = "Top 20 seafood \nexporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_exp_1, top_exp_2, common.legend = TRUE)

# Top importers, averaged over beginning and end time periods
top_imp_1 <- artis %>% 
  filter(year %in% 1996:2000) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (1996-2000)", fill = "Production method") +
  theme_minimal()

top_imp_2 <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
  ungroup() %>%
  group_by(importer_iso3c) %>%
  mutate(total = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
  slice_max(n = 60, order_by = total) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
  geom_bar(stat="identity", position = "stack") + 
  scale_fill_manual(values = method_colors) +
  lims(x = c(0, 10)) +
  labs(x = "Seafood imports (million t, live weight)", y = "",
       title = "Top 20 seafood \nimporters (2015-2019)", fill = "Production method") +
  theme_minimal()

ggarrange(top_imp_1, top_imp_2, common.legend = TRUE)

ggarrange(top_exp_2, top_imp_2, common.legend = TRUE)

```

```{r regional_bilateral_trends}
artis_region <- artis %>%
  mutate(importer_region = countrycode(importer_iso3c, origin = "iso3c", destination = "continent"),
         exporter_region = countrycode(exporter_iso3c, origin = "iso3c", destination = "continent")) %>%
  filter(!is.na(importer_region), !is.na(exporter_region)) %>%
  group_by(year, exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

region_slopes <- artis_region %>%
  split(.$exporter_importer) %>% 
  map(~lm(live_weight_t~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'exporter_importer') %>%
  filter(term == 'year') %>%
  select(exporter_importer, "slope" = "estimate")

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer")

```

```{r region_trend_matrix, fig.height=6, fig.width=6}
artis_region %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, color = slope/1000)) +
  geom_line(size = 2) +
  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
  labs(x = "", y = "Exports (million t)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "bottom") +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5))

```

# Figure 3:

```{r calculate_supply}
exports <- artis %>%
  group_by(year, exporter_iso3c, dom_source, sciname, CommonName, method, environment) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
  pivot_wider(names_from = "dom_source", values_from = "live_weight_t")

imports <- artis %>%
  group_by(year, importer_iso3c, sciname, CommonName, method, environment) %>%
  summarise(imports_live_weight_t = sum(live_weight_t))

supply <- exports %>%
  full_join(imports, by = c("year", "exporter_iso3c" = "importer_iso3c", 
                            "sciname", "CommonName", "method", "environment")) %>%
  full_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                          "sciname", "method", "environment"))

supply$foreign_export[is.na(supply$foreign_export)] <- 0
supply$domestic_export[is.na(supply$domestic_export)] <- 0
supply$error_export[is.na(supply$error_export)] <- 0
supply$imports_live_weight_t[is.na(supply$imports_live_weight_t)] <- 0
supply$production_t[is.na(supply$production_t)] <- 0

supply <- supply %>%
  mutate(supply = production_t + imports_live_weight_t - domestic_export - foreign_export - error_export,
         supply_no_error = production_t + imports_live_weight_t - domestic_export - foreign_export, 
         supply_domestic = production_t - domestic_export,
         supply_foreign = imports_live_weight_t - foreign_export) %>%
  rename("iso3c" = "exporter_iso3c") %>%
  ungroup() %>%
  mutate(supply_no_error = replace_na(supply_no_error, 0),
         supply_domestic = replace_na(supply_domestic, 0),
         supply_foreign = replace_na(supply_foreign, 0))

```

Foreign dependence trends
Where does trade have a positive versus negative effect on per capita supply for capture versus aquaculture? 
Does trade increase the diversity of supply and if so, where?

## Effect of trade on supply diversity

```{r supply_diversity}
# FIX IT: The new prod data is in here, so it doesn't merge correctly with the trade data (potentially leading to many negative supply values)

# FIX IT: Make sure supply calc includes supply of non-traded species

diversity <- supply %>%
  # FIX IT: temporarily remove negative supply values
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(iso3c, year, sciname) %>%
  summarise(supply_no_error = sum(supply_no_error), 
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t)) %>%
  ungroup(sciname) %>%
  mutate(shannon_supply_no_error = diversity(supply_no_error),
         shannon_supply_domestic = diversity(supply_domestic),
         shannon_supply_foreign = diversity(supply_foreign),
         shannon_prod = diversity(production_t),
         shannon_imports = diversity(imports_live_weight_t)) %>%
  select(iso3c, year, shannon_supply_no_error,
         shannon_supply_domestic, shannon_supply_foreign,
         shannon_prod, shannon_imports) %>%
  distinct() %>%
  mutate(region = countrycode(iso3c, origin = "iso3c", destination = "region"))

total_trade <- supply %>%
  # FIX IT: temporarily remove negative supply values
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(iso3c, year) %>%
  summarise(supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t),
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export))

diversity <- diversity %>%
  left_join(total_trade, by = c("iso3c", "year"))

# Compare diversity of foreign and domestic supply
diversity %>%
  filter(year == 2018, !is.na(region)) %>%
  ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
  labs(x = "Domestic supply diversity (Shannon index)", 
       y = "Foreign supply diversity (Shannon index)") +
    theme_minimal()

# Slope less than 1 means domestic diversity is higher than foreign-sourced diversity
summary(lm(diversity$shannon_supply_foreign~diversity$shannon_supply_domestic))

# Compare production and import diversity trends
diversity %>%
  filter(!is.na(region)) %>%
  group_by(year, region) %>%
  summarise(shannon_prod = mean(shannon_prod), 
            shannon_imports = mean(shannon_imports)) %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = shannon_prod), color = "hotpink") +
  geom_line(aes(y = shannon_imports), color = "darkgreen")  +
  labs(x = "", y = "Shannon index (pink = prod; green = import)") +
  theme_minimal() +
  facet_wrap(~region)

# Compare total imports and supply diversity
diversity %>%
  filter(year == 2018, !is.na(region)) %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
  geom_point() +
  scale_x_log10() +
  theme_minimal() +
  labs(x = "Total imports (million t, live weight)",
       y = "Supply diversity (Shannon index)")

diversity %>%
  filter(!is.na(region)) %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error)) +
  geom_point() +
  scale_x_log10() +
  theme_minimal() +
  labs(x = "Total imports (million t, live weight)",
       y = "Supply diversity (Shannon index)") +
  facet_wrap(~region)

diversity %>%
  mutate(dom_vs_foreign = case_when(
    shannon_supply_domestic >= shannon_supply_foreign ~ "dom",
    shannon_supply_domestic < shannon_supply_foreign ~ "foreign"
  )) %>% 
  group_by(dom_vs_foreign) %>%
  tally()

```

### 2022-08-19 prelim_results.Rmd
add ms figures file

@@
 author: "Jessica Gephart"
-output: html_document
+output: 
+  pdf_document
 ---
@@
 library(viridis)
+library(broom)
+library(exploreARTIS)
 
@@
 # Pull all ARTIS and production data
-artis <- dbGetQuery(con, 
-"SELECT * FROM snet 
-WHERE (snet_est = \'mid\') AND
-((\"year\" <= 2003 AND hs_version = \'HS96\') OR
-((\"year\" > 2003 AND \"year\" <= 2009) AND hs_version = \'HS02\') OR
-((\"year\" > 2009 AND \"year\" <= 2012) AND hs_version = \'HS07\') OR
-((\"year\" > 2012 AND \"year\" <= 2020) AND hs_version = \'HS12\'))")
+artis <- dbGetQuery(con, "SELECT * FROM snet")
 
 artis <- artis %>%
-  select(-record_id, -snet_est)
+  select(-record_id) %>%
+  # Remove 2020 for now
+  filter(year < 2020)
 
@@
 prod <- prod %>%
-  select(-record_id)
+  select(-record_id) %>%
+  # Remove 2020 for now
+  filter(year < 2020)
 
@@
 
-# Trend in global seafood trade
+# Trends in global seafood trade
 
@@
 
+artis_fmfo_ts <- artis %>% 
+  mutate(food_or_fmfo = case_when(
+    hs6 == "230120" ~ "fishmeal",
+    hs6 != "230120" ~ "nonfishmeal"
+  )) %>%
+  group_by(year, food_or_fmfo) %>%
+  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE))
+
+artis_fmfo_ts %>%
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = food_or_fmfo)) +
+  geom_area() +
+  scale_fill_manual(values = method_colors) +
+  labs(x = "", y = "Seafood exports (million t, live weight)", fill = "") +
+  theme_minimal()
+
+artis_fmfo_ts_summary <- artis_fmfo_ts %>%
+  pivot_wider(names_from = food_or_fmfo, values_from = live_weight_t) %>%
+  mutate(percent_fishmeal = 100*fishmeal/(fishmeal+nonfishmeal))
+
 total_1996 <- round(sum(artis %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
 total_2019 <- round(sum(artis %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)
+
+total_capture_1996 <- round(sum(artis %>% filter(year == 1996, method == "capture") %>% pull(live_weight_t))/1000000, 1)
+total_capture_2019 <- round(sum(artis %>% filter(year == 2019, method == "capture") %>% pull(live_weight_t))/1000000, 1)
+
+total_aquaculture_1996 <- round(sum(artis %>% filter(year == 1996, method == "aquaculture") %>% pull(live_weight_t))/1000000, 1)
+total_aquaculture_2019 <- round(sum(artis %>% filter(year == 2019, method == "aquaculture") %>% pull(live_weight_t))/1000000, 1)
+
 ```
@@
 Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.
+
+Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period.
+In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.
+
+Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period.
+In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.
 
@@
 kable(artis %>% 
+        filter(live_weight_t>10) %>%
         select(year, exporter_iso3c, method) %>%
@@
 
+```{r capture_aqua_sankey, eval = FALSE}
+artis %>% 
+  filter(method == "capture", year %in% 2015:2019) %>% 
+  plot_sankey(prop_flow_cutoff = 0.03)
+
+artis %>% 
+  filter(method == "aquaculture", year %in% 2015:2019) %>% 
+  plot_sankey(prop_flow_cutoff = 0.03)
+```
+
 ```{r regional_bilateral_trends}
@@
          exporter_region = countrycode(exporter_iso3c, origin = "iso3c", destination = "region")) %>%
+  filter(!is.na(importer_region), !is.na(exporter_region)) %>%
   group_by(year, exporter_region, importer_region) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
+
+region_slopes <- artis_region %>%
+  split(.$exporter_importer) %>% 
+  map(~lm(live_weight_t~year, data = .x)) %>% 
+  map_df(broom::tidy, .id = 'exporter_importer') %>%
+  filter(term == 'year') %>%
+  select(exporter_importer, "slope" = "estimate")
+
+artis_region <- artis_region %>%
+  left_join(region_slopes, by = "exporter_importer")
+```
+
+```{r region_trend_lollipop, fig.height=7, fig.width=5}
+region_slopes %>%
+  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
+  ggplot(aes(x = slope/1000, y = exporter_importer, color = slope, fill = slope)) +
+  geom_segment(aes(xend = 0, yend = exporter_importer)) +
+  geom_point() +
+  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
+  scale_fill_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
+  labs(y = "", x = "Ave. annual change in export (1000 t live weight)") +
+  theme_bw() +
+  theme(legend.position = "blank")
+
+```
+
+```{r region_trend_matrix, fig.height=7, fig.width=7}
+artis_region %>%
+  ggplot(aes(x = year, y = live_weight_t/1000000, color = slope/1000)) +
+  geom_line() +
+  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
+  labs(x = "", y = "Exports (million t)", color = "Ave. annual change in export \n(1000 t live weight)") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
+        legend.position = "bottom") +
+  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), scales = "free",
+             labeller = label_wrap_gen(width = 5))
+
+```
+
+```{r region_trend_heatmap}
+region_slopes %>%
+  separate(exporter_importer, into = c("exporter", "importer"), sep = " to ") %>% 
+  ggplot(aes(x = exporter, y = importer, fill = slope/1000)) +
+  geom_tile() +
+  scale_fill_gradient2(low = "#E24027", mid = "grey90", high = "#1B708F") +
+  labs(x = "Exporting region", y = "Importing region", fill = "Ave. annual change in export \n(1000 t live weight)") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
+        legend.position = "bottom")
+```
+
+```{r summarize_region_slopes}
+kable(region_slopes %>%
+  separate(exporter_importer, into = c("exporter", "importer"), sep = " to ") %>% 
+  mutate(slope_dir = case_when(
+    slope > 0 ~ 1, 
+    slope < 0 ~ 0
+  )) %>% 
+  group_by(exporter) %>%
+  summarise(total_increases = sum(slope_dir)) %>%
+  arrange(desc(total_increases)), col.names = c("Exporting region", "Number of increases"))
+
+```
+
+```{r econ_group_trends}
+# LCD countries according to UN (https://www.un.org/development/desa/dpad/least-developed-country-category/ldcs-at-a-glance.html)
+LCD_countries <- c("Afghanistan", "Angola", "Bangladesh", "Benin", "Bhutan", "Burkina Faso", "Burundi", "Cambodia", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Djibouti", "Eritrea", "Ethiopia", "Gambia", "Guinea", "Guinea-Bissau", "Haiti", "Kiribati", "Lao People’s Dem.
+Republic", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Myanmar", "Nepal", "Niger", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Solomon Islands", "Somalia", "South Sudan", "Sudan", "Timor-Leste", "Togo", "Tuvalu", "Uganda", "United Republic of Tanzania", "Yemen", "Zambia")
+
+LCD_countries <- countrycode(LCD_countries, origin = "country.name", destination = "iso3c")
+
+artis_lcd <- artis %>%
+  mutate(
+    importer_region = case_when(
+    (importer_iso3c %in% LCD_countries) ~ "LCD",
+    !(importer_iso3c %in% LCD_countries) ~ "non-LCD"),
+     exporter_region = case_when(
+    (exporter_iso3c %in% LCD_countries) ~ "LCD",
+    !(exporter_iso3c %in% LCD_countries) ~ "non-LCD")) %>%
+  group_by(year, exporter_region, importer_region, method) %>%
   summarise(live_weight_t = sum(live_weight_t))
 
-artis_region %>%
-  filter(!is.na(exporter_region), !is.na(importer_region)) %>%
-  ggplot(aes(x = year, y = live_weight_t/1000000)) +
-  geom_line() +
+artis_lcd %>%
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = method)) +
+  geom_area() +
+  scale_fill_manual(values = method_colors) +
   labs(x = "", y = "Exports (million t)") +
@@
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_grid(exporter_region~importer_region, scales = "free",
-             labeller = label_wrap_gen(width = 5))
-
-```
-
-```{r top_bilateral_flows}
+  facet_grid(rows = vars(exporter_region), cols = vars(importer_region))
+
+artis_lcd %>% 
+  group_by(importer_region, year, method) %>%
+  summarise(total_imports = sum(live_weight_t)) %>%
+  rename("region" = "importer_region") %>%
+  left_join(artis_lcd %>% 
+              group_by(exporter_region, year, method) %>%
+              summarise(total_exports = sum(live_weight_t)) %>%
+              rename("region" = "exporter_region"),
+            by = c("year", "region", "method")) %>%
+  filter(region == "LCD") %>%
+  mutate(net_export = total_exports - total_imports) %>%
+  ggplot(aes(x = year, y = net_export, color = method)) +
+  scale_color_manual(values = method_colors) +
+  geom_line() +
+  labs(x = "", y = "Net export (t, live weight)") +
+  ylim(c(0, 3100000)) +
+  theme_bw()
+```
+
+```{r top_bilateral_flows, fig.height=4, fig.width=7}
 top_flows_1 <- artis %>%
@@
 
+```{r country_bilateral_trends, eval=FALSE}
+
+country_slopes <- artis %>%
+  mutate(exporter_importer = paste(exporter_iso3c, " to ", importer_iso3c, sep = "")) %>%
+  split(.$exporter_importer) %>% 
+  map(~lm(live_weight_t~year, data = .x)) %>% 
+  map_df(broom::tidy, .id = 'exporter_importer') %>%
+  filter(term == 'year') %>%
+  select(exporter_importer, "slope" = "estimate")
+
+```
+
 # Exports from domestic, imported and unknown sources
@@
       caption = "Percent of exports by export source")
-```
+
+total_domestic_1996 <- round(sum(artis %>% filter(year == 1996, dom_source == "domestic export") %>% pull(live_weight_t))/1000000, 1)
+total_domestic_2019 <- round(sum(artis %>% filter(year == 2019, dom_source == "domestic export") %>% pull(live_weight_t))/1000000, 1)
+
+total_foreign_1996 <- round(sum(artis %>% filter(year == 1996, dom_source == "foreign export") %>% pull(live_weight_t))/1000000, 1)
+total_foreign_2019 <- round(sum(artis %>% filter(year == 2019, dom_source == "foreign export") %>% pull(live_weight_t))/1000000, 1)
+
+```
+
+Global domestic exports increased from `r total_domestic_1996` million t (live weight) in 1996 to `r total_domestic_2019` million t (live weight) in 2019, representing a `r round(100*((total_domestic_2019-total_domestic_1996)/total_domestic_1996),0)`% increase over that period.
+In 1996, domestic exports represented `r round(100*total_domestic_1996/total_1996)`% of all  exports and by 2019 domestic exports represented `r round(100*total_domestic_2019/total_2019)`% of all seafood exports.
+
+Global foreign (i.e., imported and re-exported) exports increased from `r total_foreign_1996` million t (live weight) in 1996 to `r total_foreign_2019` million t (live weight) in 2019, representing a `r round(100*((total_foreign_2019-total_foreign_1996)/total_foreign_1996),0)`% increase over that period.
+In 1996, foreign exports represented `r round(100*total_foreign_1996/total_1996)`% of all  exports and by 2019 foreign exports represented `r round(100*total_foreign_2019/total_2019)`% of all seafood exports.
 
@@
 
-```{r exports_percent_by_species}
+```{r exports_percent_by_species, fig.width=5, fig.height=7}
 prod_1 <- artis %>%
@@
   geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
-  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
+  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
   facet_wrap(~iso3c) +
   theme_bw()
+```
+
+```{r supply_source_by_region}
 
@@
   geom_line(aes(x = year, y = supply_foreign), color = "#E24027") +
-  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
+  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
   facet_wrap(~region) +
@@
   geom_line(aes(x = year, y = percent_foreign), color = "#E24027") +
-  labs(x = "", y = "Percent of supply (blue = domestic; red = domestic)") +
+  labs(x = "", y = "Percent of supply (blue = domestic; red = foreign)") +
   facet_wrap(~CommonName) +
@@
 
-```{r explore_country_error}
+```{r explore_country_error, eval=FALSE}
 # Plot time series of species with largest volume error exports
@@
 
-```{r explore_species_error}
+```{r explore_species_error, eval=FALSE}
 # Plot time series of species with largest volume error exports

### 2022-11-14 ms_figures.Rmd
update ms figures and results summary

@@
 library(vegan)
-
 ```
@@
 
+country_metadata <- dbGetQuery(con, "SELECT * FROM countries")
+country_metadata <- country_metadata %>%
+  select(-record_id)
+
 # Close database connection
@@
 # Set time series dates
-date_nodash <- "20220614"
-date_dash <- "2022-06-14"
+date_nodash <- "20220926"
+date_dash <- "2022-09-26"
 
 # Set directories 
-# datadir <-  paste("/Volumes/jgephart/ARTIS/Outputs/S_net/", date_nodash, "_ARTIS-timeseries", sep="")
-model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_archive_20220718"
+model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_20220926"
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
@@
 source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue
+environment_colors <- c( "#2E8D9A", "#0F2D59","#86ADA7") # teal, navy, light blue
+
+habitat_source_colors <- c("#1B708F", # blue = marine capture
+                          "#86ADA7", # light blue = freshwater capture
+                          "#F37542", # medium orange = marine aquaculture
+                          "#F7AF75", # light orange = inland aquaculture
+                          "grey50" # unknown
+                          )
+
+region_colors <- c(
+  "#741A32", # maroon
+  "#B34232", # burnt orange
+  "#D38F35", #  orange
+  "#D4B95f", # khaki
+  "#4FA2A2", # teal
+  "#114F59" # dark teal
+)
 
@@
 
+pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
+  left_join(country_metadata, by = c("iso3c")) %>%
+  select(iso3c, year, "region" = "owid_region", pop)
+
 ```
@@
   select("year", "iso3c", "sciname", 
-         "method" = "prod_method", "environment", 
+         "method", "habitat", 
          "production_t" = "live_weight_t") %>%
-  group_by(year, iso3c, sciname, method, environment) %>%
+  group_by(year, iso3c, sciname, method, habitat) %>%
   summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
@@
   filter(live_weight_t >= 0.1) %>%
+  # Add importer and exporter region
+  left_join(country_metadata %>% 
+              select("exporter_iso3c" = "iso3c", "exporter_region" = "owid_region"), by = c("exporter_iso3c")) %>%
+  left_join(country_metadata %>% 
+              select("importer_iso3c" = "iso3c", "importer_region" = "owid_region"), by = c("importer_iso3c")) %>%
 # FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
   left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
-                           "sciname", "method", "environment")) %>%
+                           "sciname", "method", "habitat")) %>%
   ungroup()
@@
   left_join(sciname_metadata, 
-            by = c("sciname" = "SciName")) %>%
+            by = "sciname") %>%
   # Leave any missing sciname_hs_modified as original sciname
@@
     TRUE ~ sciname_hs_modified
-  ))
+  )) %>%
+  rename(environment = habitat) %>%
+  # Add habitat-method column
+  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
+  mutate(habitat_method = case_when(
+    str_detect(habitat_method, "unknown") ~ "unknown", 
+    TRUE ~ habitat_method
+  )) %>% 
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" )))
 
@@
 
+```{r fig1a}
+global_export_by_source <- artis %>% 
+  mutate(food_or_fmfo = case_when(
+    hs6 == "230120" ~ "fishmeal",
+    hs6 != "230120" ~ "nonfishmeal"
+  )) %>%
+  group_by(year, habitat_method, food_or_fmfo) %>%
+  summarise(live_weight_t = sum(live_weight_t))
+
+write.csv(global_export_by_source, "outputs/global_export_by_source.csv", row.names = FALSE)
+
+global_export_by_source <- global_export_by_source %>%
+  group_by(year, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
+  geom_area() +
+  scale_fill_manual(values = habitat_source_colors) +
+  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+```
+
 ## Figure 1 b: Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports
+
+```{r fig1b}
+prod_by_source <- prod %>%
+  group_by(year, method, habitat) %>% 
+  summarise(production_t = sum(production_t)) %>%
+  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
+    # Set factor levels
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture",
+                                            "marine aquaculture", "inland aquaculture"))) %>%
+  ungroup() %>%
+  select(-habitat, -method)
+
+dom_export_by_source <- artis %>% 
+  filter(dom_source == "domestic export") %>%
+  group_by(year, habitat_method, environment) %>%
+  summarise(dom_live_weight_t = sum(live_weight_t))
+
+# One attempt to include unknown - would probably need to revisit this (perhaps by country, though that would ignore the possibility of it being sourced from elsewhere)
+
+# unknown_habitat_export<- artis %>% 
+#   filter(dom_source == "error export", environment == "unknown") %>%
+#   group_by(year) %>%
+#   summarise(unknown_habitat_live_weight_t = sum(live_weight_t)) 
+# 
+# unknown_method_export<- artis %>% 
+#   filter(dom_source == "error export", environment != "unknown") %>%
+#   group_by(year, environment) %>%
+#   summarise(unknown_method_live_weight_t = sum(live_weight_t))
+
+# dom_export_by_source %>%
+#   left_join(unknown_habitat_export, by = c("year")) %>%
+#   left_join(unknown_method_export, by = c("year", "environment")) %>%
+#   left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
+#   mutate(unknown_method_live_weight_t = ifelse(is.na(unknown_method_live_weight_t), 0, unknown_method_live_weight_t)) %>%
+#   mutate(percent_export_lower = 100*dom_live_weight_t/production_t,
+#          percent_export_upper = 100*(dom_live_weight_t + unknown_habitat_live_weight_t + unknown_method_live_weight_t)/production_t) %>%
+#   ggplot(aes(x = year, color = habitat_method)) +
+#   geom_line(aes(y = percent_export_lower)) +
+#   geom_line(aes(y = percent_export_upper)) +
+#   scale_color_manual(values = habitat_source_colors) +
+#   labs(y = "Percent of global production exported", x = "") +
+#   facet_wrap(~habitat_method) +
+#   theme_bw() +
+#   theme(legend.position = "none")
+
+percent_export_by_source <- dom_export_by_source %>%
+  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
+  mutate(percent_export = 100*dom_live_weight_t/production_t) %>%
+  ggplot(aes(x = year, y = percent_export, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(y = "% production exported", x = "", color = "Source") +
+  theme_bw()
+
+```
 
@@
 
-# Figure 2: Trade patterns
+```{r fig1c}
+all_degree_summary <-  artis %>% 
+  select(year, exporter_iso3c, importer_iso3c) %>%
+  distinct() %>%
+  group_by(year, exporter_iso3c) %>%
+  tally() %>% 
+  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
+  full_join(artis %>% 
+              select(year, exporter_iso3c, importer_iso3c) %>%
+              distinct() %>%
+              group_by(year, importer_iso3c) %>%
+              tally() %>%
+              rename(all_in_degree = n, iso3c = importer_iso3c), 
+            by = c("year", "iso3c"))
 
-```{r SI_top_traders}
-total_country_exports <- artis %>% 
-  group_by(exporter_iso3c, method, environment) %>%
-  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
-  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
-  rename("iso3c" = "exporter_iso3c")
+write.csv(all_degree_summary, "outputs/all_degree_summary.csv", row.names = FALSE)
 
-total_country_imports <- artis %>% 
-  group_by(importer_iso3c, method, environment) %>%
-  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
-  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
-  rename("iso3c" = "importer_iso3c")
+habitat_method_degree_summary <-  artis %>% 
+  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
+  distinct() %>%
+  group_by(year, exporter_iso3c, habitat_method) %>%
+  tally() %>% 
+  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
+  full_join(artis %>% 
+              select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
+              distinct() %>%
+              group_by(year, importer_iso3c, habitat_method) %>%
+              tally() %>%
+              rename(all_in_degree = n, iso3c = importer_iso3c), 
+            by = c("year", "iso3c", "habitat_method"))
 
-total_country_trade <- total_country_exports %>%
-  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
-  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
-  ungroup() %>%
-  group_by(iso3c) %>%
-  mutate(total = sum(trade_live_weight_t)) %>%
-  ungroup() %>%
-  slice_max(n = 9*20, order_by = total)
+write.csv(habitat_method_degree_summary, "outputs/habitat_method_degree_summary.csv", row.names = FALSE)
 
-total_net_import <- total_country_trade %>%
-  group_by(iso3c) %>%
-  summarise(export_live_weight_t = sum(export_live_weight_t),
-            import_live_weight_t = sum(import_live_weight_t)) %>%
-  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
-  select(iso3c, net_import)
+out_degree <- artis %>% 
+  filter(habitat_method != "unknown") %>%
+  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
+  distinct() %>%
+  group_by(year, exporter_iso3c, habitat_method) %>%
+  tally() %>%
+  group_by(year, habitat_method) %>%
+  ggplot(aes(x = year, y = n, color = habitat_method)) +
+  geom_smooth() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(y = "Out degree", fill = "Source") + 
+  theme_bw()
 
-total_country_trade %>% 
-  left_join(total_net_import, by = "iso3c") %>%
-  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
-  ggplot() +
-  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
-  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
-  scale_fill_manual(values = method_colors) +
-  geom_line(aes(y = net_import, x = iso3c), group = 1) +
-  geom_hline(yintercept = 0) +
-  labs(x = "", y = "Trade (million t, live weight)") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+in_degree <- artis %>% 
+  filter(habitat_method != "unknown") %>%
+  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
+  distinct() %>%
+  group_by(year, importer_iso3c, habitat_method) %>%
+  tally() %>%
+  group_by(year, habitat_method) %>%
+  ggplot(aes(x = year, y = n, color = habitat_method)) +
+  geom_smooth() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(y = "In degree", fill = "Source") + 
+  theme_bw()
 
@@
 
-```{r top_imp_exp}
+```{r fig1, fig.height = 3, fig.width = 4.5}
+ggarrange(global_export_by_source, 
+          ggarrange(percent_export_by_source, 
+                    out_degree, 
+                    ncol = 2, 
+                    legend = "none",
+                    labels = c("b", "c")), 
+          common.legend = TRUE, legend = "bottom",
+          ncol = 1, labels = c("a", ""))
+
+```
+
+# Figure 2: Trade patterns
+
+```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
+
+# Summarize total imports and exports by habitat and environment
+bilateral_habitat_method_summary <- artis %>%
+  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t))
+
+write.csv(bilateral_habitat_method_summary, "outputs/bilateral_habitat_method_summary.csv", row.names = FALSE)
+
 # Top exporters, averaged over beginning and end time periods
@@
 
+# Top capture and aquaculture exporters, averaged over beginning and end time periods
+top_capture_exp_1 <- artis %>% 
+  filter(year %in% 1996:2000, method == "capture") %>%
+  group_by(exporter_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
+  ungroup() %>%
+  group_by(exporter_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
+  geom_bar(stat="identity", position = "stack") + 
+  scale_fill_manual(values = environment_colors) +
+  lims(x = c(0, 4.75)) +
+  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment", 
+       title = "Top 15 capture \nexporters (1996-2000)") +
+  theme_minimal()
+
+top_capture_exp_2 <- artis %>% 
+  filter(year %in% 2015:2019, method == "capture") %>%
+  group_by(exporter_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
+  ungroup() %>%
+  group_by(exporter_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
+  geom_bar(stat="identity", position = "stack") + 
+  scale_fill_manual(values = environment_colors) +
+  lims(x = c(0, 4.75)) +
+  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 capture \nexporters (2015-2019)") +
+  theme_minimal()
+
+top_aqua_exp_1 <- artis %>% 
+  filter(year %in% 1996:2000, method == "aquaculture") %>%
+  group_by(exporter_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
+  ungroup() %>%
+  group_by(exporter_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
+  scale_fill_manual(values = environment_colors) +
+  geom_bar(stat="identity", position = "stack") + 
+  lims(x = c(0, 4.75)) +
+  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 aquaculture \nexporters (1996-2000)") +
+  theme_minimal()
+
+top_aqua_exp_2 <- artis %>% 
+  filter(year %in% 2015:2019, method == "aquaculture") %>%
+  group_by(exporter_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
+  ungroup() %>%
+  group_by(exporter_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
+  scale_fill_manual(values = environment_colors) +
+  geom_bar(stat="identity", position = "stack") + 
+  lims(x = c(0, 4.75)) +
+  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 aquaculture \nexporters (2015-2019)") +
+  theme_minimal()
+
+ggarrange(top_capture_exp_1, top_capture_exp_2, top_aqua_exp_1, top_aqua_exp_2, common.legend = TRUE, legend = "bottom")
+
+# Top capture and aquaculture importers, averaged over beginning and end time periods
+top_capture_imp_1 <- artis %>% 
+  filter(year %in% 1996:2000, method == "capture") %>%
+  group_by(importer_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
+  ungroup() %>%
+  group_by(importer_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
+  geom_bar(stat="identity", position = "stack") + 
+  scale_fill_manual(values = environment_colors) +
+  lims(x = c(0, 8)) +
+  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment", 
+       title = "Top 15 capture \nimporters (1996-2000)") +
+  theme_minimal()
+
+top_capture_imp_2 <- artis %>% 
+  filter(year %in% 2015:2019, method == "capture") %>%
+  group_by(importer_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
+  ungroup() %>%
+  group_by(importer_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
+  geom_bar(stat="identity", position = "stack") + 
+  scale_fill_manual(values = environment_colors) +
+  lims(x = c(0, 8)) +
+  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 capture \nimporters (2015-2019)") +
+  theme_minimal()
+
+top_aqua_imp_1 <- artis %>% 
+  filter(year %in% 1996:2000, method == "aquaculture") %>%
+  group_by(importer_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
+  ungroup() %>%
+  group_by(importer_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
+  scale_fill_manual(values = environment_colors) +
+  geom_bar(stat="identity", position = "stack") + 
+  lims(x = c(0, 8)) +
+  labs(x = "Imports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 aquaculture \nimporters (1996-2000)") +
+  theme_minimal()
+
+top_aqua_imp_2 <- artis %>% 
+  filter(year %in% 2015:2019, method == "aquaculture") %>%
+  group_by(importer_iso3c, environment) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
+  ungroup() %>%
+  group_by(importer_iso3c) %>%
+  mutate(total = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
+  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
+  slice_max(n = 2*15, order_by = total) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
+  scale_fill_manual(values = environment_colors) +
+  geom_bar(stat="identity", position = "stack") + 
+  lims(x = c(0, 8)) +
+  labs(x = "Imports (million t, live weight)", y = "", fill = "Production environment",
+       title = "Top 15 aquaculture \nimporters (2015-2019)") +
+  theme_minimal()
+
+ggarrange(top_capture_imp_1, top_capture_imp_2, top_aqua_imp_1, top_aqua_imp_2, common.legend = TRUE, legend = "bottom")
+```
+
+```{r SI_top_traders}
+total_country_exports <- artis %>% 
+  filter(year %in% 2015:2019) %>%
+  group_by(exporter_iso3c, method, environment) %>%
+  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
+  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
+  rename("iso3c" = "exporter_iso3c")
+
+total_country_imports <- artis %>% 
+  filter(year %in% 2015:2019) %>%
+  group_by(importer_iso3c, method, environment) %>%
+  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
+  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
+  rename("iso3c" = "importer_iso3c")
+
+total_country_trade <- total_country_exports %>%
+  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
+  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
+  ungroup() %>%
+  group_by(iso3c) %>%
+  mutate(total = sum(trade_live_weight_t)) %>%
+  ungroup() %>%
+  slice_max(n = 9*20, order_by = total)
+
+total_net_import <- total_country_trade %>%
+  group_by(iso3c) %>%
+  summarise(export_live_weight_t = sum(export_live_weight_t),
+            import_live_weight_t = sum(import_live_weight_t)) %>%
+  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
+  select(iso3c, net_import)
+
+total_country_trade %>% 
+  left_join(total_net_import, by = "iso3c") %>%
+  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
+  ggplot() +
+  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
+  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
+  scale_fill_manual(values = method_colors) +
+  geom_line(aes(y = net_import, x = iso3c), group = 1) +
+  geom_hline(yintercept = 0) +
+  labs(x = "", y = "Trade (million t, live weight)") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+```
+
+## Regional bilateral flows
+```{r region_trend_matrix_by_method, fig.height=5, fig.width=5}
+artis_region_method <- artis %>%
+  filter(!is.na(importer_region), !is.na(exporter_region), 
+         importer_region != "Other nei", exporter_region != "Other nei") %>%
+  group_by(year, exporter_region, importer_region, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
+
+region_trend_matrix_method <- artis_region_method %>% 
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
+  geom_area() +
+  scale_fill_manual(values = habitat_source_colors) +
+  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
+        legend.position = "none",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        strip.text.y = element_blank()) +
+  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
+             labeller = label_wrap_gen(width = 5), scales = "free")
+
+# Create basic plot for the legend
+artis_region_method %>%
+  group_by(year, exporter_region) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
+  geom_area() +
+  scale_fill_manual(values = region_colors) +
+  labs(fill = "Region") +
+  theme(legend.position = "bottom")
 ```
@@
 ```{r regional_bilateral_trends}
-artis_region <- artis %>%
-  mutate(importer_region = countrycode(importer_iso3c, origin = "iso3c", destination = "continent"),
-         exporter_region = countrycode(exporter_iso3c, origin = "iso3c", destination = "continent")) %>%
-  filter(!is.na(importer_region), !is.na(exporter_region)) %>%
+artis_region <- artis_region_method %>%
   group_by(year, exporter_region, importer_region) %>%
@@
 
-```{r region_trend_matrix, fig.height=6, fig.width=6}
-artis_region %>%
-  ggplot(aes(x = year, y = live_weight_t/1000000, color = slope/1000)) +
-  geom_line(size = 2) +
+```{r region_trend_lollipop, fig.height=4, fig.width=8}
+region_slopes_ranked <- region_slopes %>%
+  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
+  ggplot(aes(x = slope/1000, y = exporter_importer, color = slope, fill = slope)) +
+  geom_segment(aes(xend = 0, yend = exporter_importer)) +
+  geom_point() +
   scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
-  labs(x = "", y = "Exports (million t)", color = "Ave. annual change in export \n(1000 t live weight)") +
+  scale_fill_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
+  labs(y = "", x = "Ave. annual change in export (1000 t live weight)") +
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
-        legend.position = "bottom") +
-  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
-             labeller = label_wrap_gen(width = 5))
+  theme(legend.position = "blank")
+
+ggarrange(region_trend_matrix_method, region_slopes_ranked, widths = c(1.5,1))
+```
+
+```{r, fig.height=3, fig.width=4}
+exports_by_region <- artis_region %>% 
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
+  geom_area() +
+  labs(x = "", y = "", fill = "Importer") +
+  scale_fill_manual(values = region_colors) +
+  facet_wrap(~exporter_region, ncol = 1) +
+  theme_bw() +
+    theme(legend.position = "none",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+imports_by_region <- artis_region %>% 
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
+  geom_area() +
+  labs(x = "", y = "", fill = "Exporter") +
+  scale_fill_manual(values = region_colors) +
+  facet_wrap(~importer_region, nrow = 1) +
+  theme_bw() +
+  theme(legend.position = "none",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        axis.text.x = element_blank())
+
+ggarrange(ggarrange(imports_by_region, NULL,
+                    ncol = 2, nrow = 1,
+                    widths = c(4, 1)),
+          ggarrange(region_trend_matrix_method, exports_by_region, 
+                    ncol = 2, nrow = 1,
+                    widths = c(4, 1)),
+          ncol = 1, nrow = 2, 
+          heights = c(1, 4))
 
@@
 
-# Figure 3:
+```{r africa_trade_by_method}
+africa_exports <- artis_region_method %>%
+  filter(exporter_region == "Africa", 
+         importer_region != "Africa") %>%
+  group_by(year, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)/1000000) %>%
+  ggplot(aes(x = year, y = live_weight_t, color = habitat_method)) +
+  geom_line() +
+  lims(y = c(0, 3.25)) +
+  scale_color_manual(values = habitat_source_colors) +
+  theme_minimal() + 
+  labs(x = "", y = "Exports (million t, live weight)", color = "Source")
+
+africa_imports <- artis_region_method %>%
+  filter(importer_region == "Africa", 
+         exporter_region != "Africa") %>%
+  group_by(year, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)/1000000) %>%
+  ggplot(aes(x = year, y = live_weight_t, color = habitat_method)) +
+  geom_line() +
+  lims(y = c(0, 3.25)) +
+  scale_color_manual(values = habitat_source_colors) +
+  theme_minimal() + 
+  labs(x = "", y = "Exports (million t, live weight)", color = "Source")
+
+ggarrange(africa_exports, africa_imports, common.legend = TRUE, labels = c("Africa exports", "Africa imports"))
+```
+
+# Figure 3: Supply patterns
 
 ```{r calculate_supply}
-exports <- artis %>%
-  group_by(year, exporter_iso3c, dom_source, sciname, CommonName, method, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
-  pivot_wider(names_from = "dom_source", values_from = "live_weight_t")
-
-imports <- artis %>%
-  group_by(year, importer_iso3c, sciname, CommonName, method, environment) %>%
-  summarise(imports_live_weight_t = sum(live_weight_t))
-
-supply <- exports %>%
-  full_join(imports, by = c("year", "exporter_iso3c" = "importer_iso3c", 
-                            "sciname", "CommonName", "method", "environment")) %>%
-  full_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
-                          "sciname", "method", "environment"))
-
-supply$foreign_export[is.na(supply$foreign_export)] <- 0
-supply$domestic_export[is.na(supply$domestic_export)] <- 0
-supply$error_export[is.na(supply$error_export)] <- 0
-supply$imports_live_weight_t[is.na(supply$imports_live_weight_t)] <- 0
-supply$production_t[is.na(supply$production_t)] <- 0
-
-supply <- supply %>%
-  mutate(supply = production_t + imports_live_weight_t - domestic_export - foreign_export - error_export,
-         supply_no_error = production_t + imports_live_weight_t - domestic_export - foreign_export, 
-         supply_domestic = production_t - domestic_export,
-         supply_foreign = imports_live_weight_t - foreign_export) %>%
-  rename("iso3c" = "exporter_iso3c") %>%
-  ungroup() %>%
-  mutate(supply_no_error = replace_na(supply_no_error, 0),
-         supply_domestic = replace_na(supply_domestic, 0),
-         supply_foreign = replace_na(supply_foreign, 0))
+supply <- calculate_supply(artis %>% 
+                             # Remove FM trade from supply calculations
+                             #filter(hs6 != "230120") %>%
+                             rename(habitat = environment), 
+                           prod %>% 
+                             rename(live_weight_t = production_t)) %>%
+  # Add habitat-method column
+  mutate(habitat_method = paste(habitat, method, sep =" ")) %>% 
+  mutate(habitat_method = case_when(
+    str_detect(habitat_method, "unknown") ~ "unknown", 
+    TRUE ~ habitat_method
+  )) %>% 
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>% 
+  left_join(country_metadata %>% 
+              select(iso3c, "region" = "owid_region"), by = "iso3c") %>% 
+      mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
+                                     TRUE ~ supply_no_error))
 
@@
 
-Foreign dependence trends
+# Fig 3a: Foreign dependence trends
+
+```{r supply_source_by_region}
+region_supply_trends <- supply %>% 
+  filter(region != "Other nei") %>%
+  group_by(region, year) %>%
+  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
+            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
+  pivot_longer(cols = supply_domestic:supply_foreign, 
+               names_to = "supply_source", values_to = "supply_percent") %>%
+  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
+  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
+  geom_line() +
+  scale_color_manual(values = c("#F37542", "#0F2D59")) +
+  labs(x = "", y = "Percent of supply", color = "Origin") +
+  facet_wrap(~region) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+global_supply_trends <- supply %>% 
+  filter(region != "Other nei") %>%
+  group_by( year) %>%
+  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
+            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
+  pivot_longer(cols = supply_domestic:supply_foreign, 
+               names_to = "supply_source", values_to = "supply_percent") %>%
+  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
+  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
+  geom_line() +
+  scale_color_manual(values = c("#F37542", "#0F2D59")) +
+  labs(x = "", y = "Percent of supply", color = "Origin") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+supply %>% 
+  filter(region != "Other nei") %>%
+  filter(year == 2019) %>%
+  group_by(iso3c) %>%
+  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
+            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
+  arrange(desc(supply_foreign))
+
+```
+
+```{r}
+# Calculate per capita supply by region and source
+supply_total <- supply %>% 
+  group_by(iso3c, region, year, habitat_method) %>%
+  summarise(supply = sum(supply),
+            supply_no_error = sum(supply_no_error),
+            supply_domestic = sum(supply_domestic),
+            supply_foreign = sum(supply_foreign),
+            import = sum(imports_live_weight_t), 
+            domestic_export = sum(domestic_export),
+            foreign_export = sum(foreign_export),
+            error_export = sum(error_export), 
+            production_t = sum(production_t)
+            ) %>%
+  left_join(pop %>% select(-region), by = c("iso3c", "year"))
+
+# Global per capita supply
+global_percap_supply_area <- supply_total %>%
+  filter(region != "Other nei") %>%
+  group_by(year, habitat_method) %>%
+  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
+  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
+  geom_area() + 
+  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
+  scale_fill_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+global_percap_supply_line <- supply_total %>%
+  filter(region != "Other nei") %>%
+  group_by(year, habitat_method) %>%
+  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
+  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
+  geom_line() + 
+  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
+  scale_color_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+# Regional per capita supply
+regional_percap_supply_area <- supply_total %>%
+  filter(region != "Other nei") %>%
+  group_by(year, region, habitat_method) %>%
+  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
+  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
+  geom_area() + 
+  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
+  scale_fill_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~region)
+
+regional_percap_supply_line <- supply_total %>%
+  filter(region != "Other nei") %>%
+  group_by(year, region, habitat_method) %>%
+  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
+  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
+  geom_line() + 
+  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
+  scale_color_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~region)
+
+```
+
 Where does trade have a positive versus negative effect on per capita supply for capture versus aquaculture?
-Does trade increase the diversity of supply and if so, where?
 
-## Effect of trade on supply diversity
+```{r}
+# Plot relationship between per capita supply and imports by prod source and region
+country_supply <- supply %>% 
+  filter(year == 2018, region != "Other nei") %>% 
+  group_by(iso3c, region, year) %>% 
+  summarise(supply = sum(supply),
+            supply_no_error = sum(supply_no_error),
+            supply_domestic = sum(supply_domestic),
+            supply_foreign = sum(supply_foreign),
+            import = sum(imports_live_weight_t), 
+            domestic_export = sum(domestic_export),
+            foreign_export = sum(foreign_export),
+            error_export = sum(error_export), 
+            production_t = sum(production_t)
+  ) %>% 
+  left_join(pop %>% select(-region), by = c("iso3c", "year")) %>%
+  mutate(supply_per_cap = supply_no_error/pop) 
+```
+
+```{r, fig.height = 5, fig.width = 3}
+country_supply %>% 
+  ungroup() %>%
+  mutate(iso3c = fct_reorder(iso3c, supply_per_cap)) %>%
+  filter(supply_per_cap > 50) %>%
+  ggplot(aes(x = supply_per_cap, y = iso3c, fill = region)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = region_colors) +
+  theme_bw() +
+  theme(legend.position = "none")
+
+```
+
+```{r}
+import_supply <- country_supply %>%
+  # FIX IT: Temporarily remove very large (unrealistic) supply values
+  filter(supply_per_cap < 100) %>%
+  ggplot(aes(x = import/1000, y = supply_per_cap, color = region)) +
+  geom_point(alpha = 0.7) +
+  scale_x_continuous(trans='log10') +
+  labs(x = "Import (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
+  scale_color_manual(values = region_colors) +
+  theme_bw()
+
+export_supply <- country_supply %>%
+  # FIX IT: Temporarily remove very large (unrealistic) supply values
+  filter(supply_per_cap < 100) %>%
+  ggplot(aes(x = (domestic_export + foreign_export + error_export)/1000, y = supply_per_cap, color = region)) +
+  geom_point(alpha = 0.7) +
+  scale_x_continuous(trans='log10') +
+  labs(x = "Export (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
+  scale_color_manual(values = region_colors) +
+  theme_bw()
+```
+
+```{r, fig.height = 4, fig.width = 4}
+# Join figures
+ggarrange(
+  ggarrange(global_percap_supply_line, regional_percap_supply_area, 
+            common.legend = TRUE, nrow = 1, widths = c(1, 2),
+            labels = c("a", "b")),
+  ggarrange(global_supply_trends, region_supply_trends,
+            common.legend = TRUE, nrow = 1, widths = c(1, 2),
+            labels = c("c", "d")),
+  ggarrange(import_supply, export_supply, 
+            common.legend = TRUE, nrow = 1, widths = c(1, 1),
+            labels = c("e", "f")),
+  nrow = 3
+)
+```
+
+## Fig 4: Effect of trade on supply diversity
 
 ```{r supply_diversity}
-# FIX IT: The new prod data is in here, so it doesn't merge correctly with the trade data (potentially leading to many negative supply values)
-
-# FIX IT: Make sure supply calc includes supply of non-traded species
-
+# Calculate diversity indices
 diversity <- supply %>%
-  # FIX IT: temporarily remove negative supply values
   filter(supply_domestic > 0, supply_foreign > 0) %>%
-  group_by(iso3c, year, sciname) %>%
+  group_by(iso3c, region, year, sciname) %>%
   summarise(supply_no_error = sum(supply_no_error), 
@@
          shannon_prod, shannon_imports) %>%
-  distinct() %>%
-  mutate(region = countrycode(iso3c, origin = "iso3c", destination = "region"))
+  distinct()
 
 total_trade <- supply %>%
-  # FIX IT: temporarily remove negative supply values
   filter(supply_domestic > 0, supply_foreign > 0) %>%
-  group_by(iso3c, year) %>%
+  group_by(region, iso3c, year) %>%
   summarise(supply_domestic = sum(supply_domestic),
@@
 diversity <- diversity %>%
-  left_join(total_trade, by = c("iso3c", "year"))
+  left_join(total_trade, by = c("region", "iso3c", "year"))
 
+```
+
+# Fig 4a: diversity of foreign and domestic supply
+
+```{r}
 # Compare diversity of foreign and domestic supply
-diversity %>%
-  filter(year == 2018, !is.na(region)) %>%
+foreign_vs_domestic_diversity <- diversity %>%
+  filter(year == 2018, region != "Other nei") %>%
   ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
-    geom_point() +
+    geom_point(alpha = 0.7) +
     geom_abline(slope = 1, intercept = 0) +
-  labs(x = "Domestic supply diversity (Shannon index)", 
-       y = "Foreign supply diversity (Shannon index)") +
+  scale_color_manual(values = region_colors) +
+  labs(x = "Domestic supply diversity", 
+       y = "Foreign supply diversity",
+       color = "") +
     theme_minimal()
@@
 summary(lm(diversity$shannon_supply_foreign~diversity$shannon_supply_domestic))
+```
 
+# Fig 4b: Production and import diversity
+```{r}
 # Compare production and import diversity trends
-diversity %>%
-  filter(!is.na(region)) %>%
+prod_import_diversity_trends <- diversity %>%
+  filter(region != "Other nei") %>%
   group_by(year, region) %>%
@@
             shannon_imports = mean(shannon_imports)) %>%
-  ggplot(aes(x = year)) +
-  geom_line(aes(y = shannon_prod), color = "hotpink") +
-  geom_line(aes(y = shannon_imports), color = "darkgreen")  +
-  labs(x = "", y = "Shannon index (pink = prod; green = import)") +
+  pivot_longer(cols = shannon_prod:shannon_imports,
+               names_to = "source", values_to = "shannon") %>%
+  mutate(source = case_when(
+    source == "shannon_prod" ~ "Production diversity",
+    source == "shannon_imports" ~ "Import diversity"
+  )) %>%
+  ggplot(aes(x = year, y = shannon, color = source)) +
+  geom_line() +
+  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
+  labs(x = "", y = "Shannon index", color = "") +
   theme_minimal() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
   facet_wrap(~region)
@@
 # Compare total imports and supply diversity
-diversity %>%
-  filter(year == 2018, !is.na(region)) %>%
+imports_vs_supply_diversity <- diversity %>%
+  filter(year == 2018, region != "Other nei") %>%
   ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
-  geom_point() +
+  geom_point(alpha = 0.7) +
   scale_x_log10() +
+  scale_color_manual(values = region_colors) +
   theme_minimal() +
   labs(x = "Total imports (million t, live weight)",
-       y = "Supply diversity (Shannon index)")
+       y = "Supply diversity",
+       color = "")
 
 diversity %>%
-  filter(!is.na(region)) %>%
-  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error)) +
-  geom_point() +
+  filter(region != "Other nei") %>%
+  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
+  geom_point(alpha = 0.5) +
   scale_x_log10() +
+  scale_color_manual(values = region_colors) +
+    labs(x = "Total imports (million t, live weight)",
+       y = "Supply Shannon diversity",
+       color = "") +
   theme_minimal() +
-  labs(x = "Total imports (million t, live weight)",
-       y = "Supply diversity (Shannon index)") +
+  theme(legend.position = "bottom") +
   facet_wrap(~region)
@@
 ```
+
+```{r, fig.height = 3, fig.width = 3}
+ggarrange(prod_import_diversity_trends, 
+          ggarrange(foreign_vs_domestic_diversity,
+                    imports_vs_supply_diversity,
+                    nrow = 1,
+                    common.legend = TRUE, 
+                    legend = "bottom",
+                    labels = c("b", "c")),
+          nrow = 2,
+          labels = c("a", "", "")
+          )
+```
+
+```{r}
+
+```

### 2022-11-14 ms_results.Rmd
update ms figures and results summary

---
title: "ms_results"
author: "Jessica Gephart"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Load packages
library(tidyverse)
library(countrycode)
library(kableExtra)
library(ggpubr)

# Set color palettes 
habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )
```

# Trends blue food globalization

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Load data summary files generated by ms_results
global_export_by_source <- read.csv("outputs/global_export_by_source.csv")
all_degree_summary <- read.csv("outputs/all_degree_summary.csv")
habitat_method_degree_summary <- read.csv("outputs/habitat_method_degree_summary.csv")

# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)

total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)

total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)

all_degree_1996 <- mean(all_degree_summary %>% filter(year == 1996) %>% pull(all_out_degree), na.rm = TRUE)
all_degree_2019 <- mean(all_degree_summary %>% filter(year == 2019) %>% pull(all_out_degree), na.rm = TRUE)

```

## Text for MS

Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships. Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`% (from `r total_1996` to `r total_2019` mil t) and the average number of trade partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`). Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports. Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively. From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`%.

## Other summary stats
Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.

Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period. In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.

Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period. In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.

# Shifts in global flows of blue foods
```{r, echo=FALSE, warning=FALSE, message=FALSE}
bilateral_habitat_method_summary <- read.csv("outputs/bilateral_habitat_method_summary.csv") %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
cumulative_percent_threshold <- 75
# Concentration in exports
## All
export_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_n_countries <- export_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
export_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_habitat_method_n_countries <- export_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

export_concentration_plot <- ggplot() +
  geom_line(data = export_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = export_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

# Concentration in imports
## All
import_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_n_countries <- import_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
import_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_habitat_method_n_countries <- import_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

import_concentration_plot <- ggplot() +
  geom_line(data = import_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = import_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

ggarrange(export_concentration_plot, import_concentration_plot, common.legend = TRUE, legend = "bottom", labels = c("a", "b"))
```

Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)`countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)`countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Aquaculture exports and imports are generally concentrated among fewer countries than capture exports and imports, though inland aquaculture imports became less concentrated over the period.

```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Top trade tables
top_n_countries <- 5
top_n_countries_flows <- 10
top_n_regions <- 5
top_n_regions_flows <- 10

# Top exporters
## Countries
top_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture exporters
## Countries
top_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture exporters
## Countries
top_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top importers
## Countries
top_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture importers
## Countries
top_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture importers
## Countries
top_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top bilateral flows
## Countries
top_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top capture bilateral flows
## Countries
top_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top aquaculture bilateral flows
## Countries
top_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Create tables
## Top exporter countries
top_exporter_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = exporter_iso3c), by = "rank")

kable(top_exporter_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries, "exporting countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top exporter regions
top_exporter_region_table <- data.frame(rank = 1:top_n_regions) %>%
  left_join(top_region_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_2019 = exporter_region), by = "rank")

kable(top_exporter_region_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions, "exporting regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top importer countries
top_importer_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = importer_iso3c), by = "rank")

kable(top_importer_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries, "importing countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top importer regions
top_importer_region_table <- data.frame(rank = 1:top_n_regions) %>%
  left_join(top_region_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_1996 = importer_region), by = "rank") %>%
  left_join(top_region_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_2019 = importer_region), by = "rank") %>%
  left_join(top_region_capture_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_1996 = importer_region), by = "rank") %>%
  left_join(top_region_capture_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_2019 = importer_region), by = "rank") %>%
  left_join(top_region_aquaculture_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_1996 = importer_region), by = "rank") %>%
  left_join(top_region_aquaculture_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_2019 = importer_region), by = "rank")

kable(top_importer_region_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions, "importing regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top country flows
top_country_flow_table <- data.frame(rank = 1:top_n_countries_flows) %>%
  left_join(top_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_total_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_total_1996), by = "rank") %>%
  left_join(top_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_total_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_total_2019), by = "rank") %>%
  left_join(top_capture_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_capture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_capture_1996), by = "rank") %>%
  left_join(top_capture_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_capture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_capture_2019), by = "rank") %>%
  left_join(top_aquaculture_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_aquaculture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_aquaculture_1996), by = "rank") %>%
  left_join(top_aquaculture_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_aquaculture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_aquaculture_2019), by = "rank")

kable(top_country_flow_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top region flows
top_region_flow_table <- data.frame(rank = 1:top_n_regions_flows) %>%
  left_join(top_region_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_total_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_total_1996), by = "rank") %>%
  left_join(top_region_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_total_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_total_2019), by = "rank") %>%
  left_join(top_region_capture_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_capture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_capture_1996), by = "rank") %>%
  left_join(top_region_capture_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_capture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_capture_2019), by = "rank") %>%
  left_join(top_region_aquaculture_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_aquaculture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_aquaculture_1996), by = "rank") %>%
  left_join(top_region_aquaculture_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_aquaculture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_aquaculture_2019), by = "rank")

kable(top_region_flow_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
  filter(year %in% 2015:2019) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  group_by(exporter_region) %>%
  slice_max(n=2, order_by = live_weight_t)

bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
  filter(year %in% 2015:2019) %>%
  group_by(exporter_region, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  group_by(exporter_region, habitat_method) %>%
  slice_max(n=1, order_by = live_weight_t)
```

For all regions other than Oceania and South America, interregional trade is either the highest or second high

## Text for MS

## Other summary stats

# Role of trade in national blue food supplies

## Text for MS

## Other summary stats

# Trade and blue food supply diversity

## Text for MS

## Other summary stats

### 2023-01-06 ms_figures.Rmd
text and figure edits for full draft

@@
 library(vegan)
+library(tidytext)
 ```
@@
 artis <- artis %>%
-  select(-record_id) %>%
-  # Remove 2020 for now
-  filter(year < 2020)
+  select(-record_id)
 
@@
 prod <- prod %>%
-  select(-record_id) %>%
-  # Remove 2020 for now
-  filter(year < 2020)
+  select(-record_id)
 
@@
 country_metadata <- country_metadata %>%
+  select(-record_id)
+
+code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
+              mutate(hs6 = as.character(hs6)) %>%
+  select(-record_id)
+
+sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
   select(-record_id)
@@
 # Set directories 
-model_inputs_dir <- "/Volumes/jgephart/ARTIS/Outputs/model_inputs_20220926"
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
@@
 ```{r load_drive_data}
-code_max_resolved_taxa <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/code_max_resolved_taxa.csv") %>%
-              mutate(hs6 = as.character(hs6))
-
-sciname_metadata <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/sciname_metadata.csv")
 
@@
 ```{r join_data}
-# Load cleaned FAO production data
+# Summarize cleaned FAO production data
 prod <- prod %>% 
@@
 artis <- artis %>%
-  # FIX IT: Change metadata to HSXX format and then remove this line
-  mutate(hs_version = as.integer(parse_number(hs_version))) %>%
   left_join(code_max_resolved_taxa,
@@
 
-global_export_by_source <- global_export_by_source %>%
+fig1a <- global_export_by_source %>%
   group_by(year, habitat_method) %>%
@@
 
-# One attempt to include unknown - would probably need to revisit this (perhaps by country, though that would ignore the possibility of it being sourced from elsewhere)
-
-# unknown_habitat_export<- artis %>% 
-#   filter(dom_source == "error export", environment == "unknown") %>%
-#   group_by(year) %>%
-#   summarise(unknown_habitat_live_weight_t = sum(live_weight_t)) 
-# 
-# unknown_method_export<- artis %>% 
-#   filter(dom_source == "error export", environment != "unknown") %>%
-#   group_by(year, environment) %>%
-#   summarise(unknown_method_live_weight_t = sum(live_weight_t))
-
-# dom_export_by_source %>%
-#   left_join(unknown_habitat_export, by = c("year")) %>%
-#   left_join(unknown_method_export, by = c("year", "environment")) %>%
-#   left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
-#   mutate(unknown_method_live_weight_t = ifelse(is.na(unknown_method_live_weight_t), 0, unknown_method_live_weight_t)) %>%
-#   mutate(percent_export_lower = 100*dom_live_weight_t/production_t,
-#          percent_export_upper = 100*(dom_live_weight_t + unknown_habitat_live_weight_t + unknown_method_live_weight_t)/production_t) %>%
-#   ggplot(aes(x = year, color = habitat_method)) +
-#   geom_line(aes(y = percent_export_lower)) +
-#   geom_line(aes(y = percent_export_upper)) +
-#   scale_color_manual(values = habitat_source_colors) +
-#   labs(y = "Percent of global production exported", x = "") +
-#   facet_wrap(~habitat_method) +
-#   theme_bw() +
-#   theme(legend.position = "none")
-
 percent_export_by_source <- dom_export_by_source %>%
   left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
+  mutate(percent_export = 100*dom_live_weight_t/production_t)
+
+write.csv(percent_export_by_source, "outputs/percent_export_by_source_summary.csv", row.names = FALSE)
+
+percent_export_total <- dom_export_by_source %>%
+  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
+  group_by(year) %>% 
+  summarise(dom_live_weight_t = sum(dom_live_weight_t), 
+            production_t = sum(production_t)) %>%
   mutate(percent_export = 100*dom_live_weight_t/production_t) %>% 
-  ggplot(aes(x = year, y = percent_export, color = habitat_method)) +
-  geom_line() +
+  mutate(calc_type = "domestic exports") %>% 
+  select(year, percent_export, calc_type)
+
+percent_export_total_all_source <- artis %>% 
+  group_by(year) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  left_join(prod_by_source %>% 
+              group_by(year) %>% 
+              summarise(production_t = sum(production_t)), 
+            by = c("year")) %>% 
+  mutate(percent_export = 100*live_weight_t/production_t) %>% 
+  mutate(calc_type = "all export sources") %>% 
+  select(year, percent_export, calc_type)
+
+percent_export_total <- percent_export_total %>%
+  bind_rows(percent_export_total_all_source)
+
+write.csv(percent_export_total, "outputs/percent_export_total_summary.csv", row.names = FALSE)
+
+fig1b <- ggplot() +
+  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
   scale_color_manual(values = habitat_source_colors) +
@@
 
-out_degree <- artis %>% 
+fig1c_out_degree <- artis %>% 
   filter(habitat_method != "unknown") %>%
@@
 
-in_degree <- artis %>% 
+fig1c_in_degree <- artis %>% 
   filter(habitat_method != "unknown") %>%
@@
 ```{r fig1, fig.height = 3, fig.width = 4.5}
-ggarrange(global_export_by_source, 
-          ggarrange(percent_export_by_source, 
-                    out_degree, 
-                    ncol = 2, 
+png("outputs/fig1.png")
+ggarrange(fig1a,
+          ggarrange(fig1b,
+                    fig1c_out_degree,
+                    ncol = 2, nrow = 1,
                     legend = "none",
@@
           common.legend = TRUE, legend = "bottom",
-          ncol = 1, labels = c("a", ""))
-
+          ncol = 1, nrow = 2, labels = c("a", ""))
+
+dev.off()
 ```
@@
 ```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
-
 # Summarize total imports and exports by habitat and environment
 bilateral_habitat_method_summary <- artis %>%
+  # Remove if we decide to leave FMFO in
+  filter(hs6 != "230120") %>%
   group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
@@
 
-# Top exporters, averaged over beginning and end time periods
-top_exp_1 <- artis %>% 
-  filter(year %in% 1996:2000) %>%
-  group_by(exporter_iso3c, method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
-  slice_max(n = 60, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = method_colors) +
-  lims(x = c(0, 10)) +
-  labs(x = "Seafood exports (million t, live weight)", y = "",
-       title = "Top 20 seafood \nexporters (1996-2000)", fill = "Production method") +
-  theme_minimal()
-
-top_exp_2 <- artis %>% 
-  filter(year %in% 2015:2019) %>%
-  group_by(exporter_iso3c, method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = fct_reorder(as.factor(exporter_iso3c), total)) %>%
-  slice_max(n = 60, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = method)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = method_colors) +
-  lims(x = c(0, 10)) +
-  labs(x = "Seafood exports (million t, live weight)", y = "",
-       title = "Top 20 seafood \nexporters (2015-2019)", fill = "Production method") +
-  theme_minimal()
-
-ggarrange(top_exp_1, top_exp_2, common.legend = TRUE)
-
-# Top importers, averaged over beginning and end time periods
-top_imp_1 <- artis %>% 
-  filter(year %in% 1996:2000) %>%
-  group_by(importer_iso3c, method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
-  slice_max(n = 60, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = method_colors) +
-  lims(x = c(0, 10)) +
-  labs(x = "Seafood imports (million t, live weight)", y = "",
-       title = "Top 20 seafood \nimporters (1996-2000)", fill = "Production method") +
-  theme_minimal()
-
-top_imp_2 <- artis %>% 
-  filter(year %in% 2015:2019) %>%
-  group_by(importer_iso3c, method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = fct_reorder(as.factor(importer_iso3c), total)) %>%
-  slice_max(n = 60, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = method)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = method_colors) +
-  lims(x = c(0, 10)) +
-  labs(x = "Seafood imports (million t, live weight)", y = "",
-       title = "Top 20 seafood \nimporters (2015-2019)", fill = "Production method") +
-  theme_minimal()
-
-ggarrange(top_imp_1, top_imp_2, common.legend = TRUE)
-
-ggarrange(top_exp_2, top_imp_2, common.legend = TRUE)
-
-# Top capture and aquaculture exporters, averaged over beginning and end time periods
-top_capture_exp_1 <- artis %>% 
-  filter(year %in% 1996:2000, method == "capture") %>%
-  group_by(exporter_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = environment_colors) +
-  lims(x = c(0, 4.75)) +
-  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment", 
-       title = "Top 15 capture \nexporters (1996-2000)") +
-  theme_minimal()
-
-top_capture_exp_2 <- artis %>% 
-  filter(year %in% 2015:2019, method == "capture") %>%
-  group_by(exporter_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = environment_colors) +
-  lims(x = c(0, 4.75)) +
-  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 capture \nexporters (2015-2019)") +
-  theme_minimal()
-
-top_aqua_exp_1 <- artis %>% 
-  filter(year %in% 1996:2000, method == "aquaculture") %>%
-  group_by(exporter_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
-  scale_fill_manual(values = environment_colors) +
-  geom_bar(stat="identity", position = "stack") + 
-  lims(x = c(0, 4.75)) +
-  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 aquaculture \nexporters (1996-2000)") +
-  theme_minimal()
-
-top_aqua_exp_2 <- artis %>% 
-  filter(year %in% 2015:2019, method == "aquaculture") %>%
-  group_by(exporter_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(exporter_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter = countrycode(exporter_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(exporter = fct_reorder(as.factor(exporter), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter, fill = environment)) +
-  scale_fill_manual(values = environment_colors) +
-  geom_bar(stat="identity", position = "stack") + 
-  lims(x = c(0, 4.75)) +
-  labs(x = "Ave. annual exports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 aquaculture \nexporters (2015-2019)") +
-  theme_minimal()
-
-ggarrange(top_capture_exp_1, top_capture_exp_2, top_aqua_exp_1, top_aqua_exp_2, common.legend = TRUE, legend = "bottom")
-
-# Top capture and aquaculture importers, averaged over beginning and end time periods
-top_capture_imp_1 <- artis %>% 
-  filter(year %in% 1996:2000, method == "capture") %>%
-  group_by(importer_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = environment_colors) +
-  lims(x = c(0, 8)) +
-  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment", 
-       title = "Top 15 capture \nimporters (1996-2000)") +
-  theme_minimal()
-
-top_capture_imp_2 <- artis %>% 
-  filter(year %in% 2015:2019, method == "capture") %>%
-  group_by(importer_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
-  geom_bar(stat="identity", position = "stack") + 
-  scale_fill_manual(values = environment_colors) +
-  lims(x = c(0, 8)) +
-  labs(x = "Ave. annual imports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 capture \nimporters (2015-2019)") +
-  theme_minimal()
-
-top_aqua_imp_1 <- artis %>% 
-  filter(year %in% 1996:2000, method == "aquaculture") %>%
-  group_by(importer_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
-  scale_fill_manual(values = environment_colors) +
-  geom_bar(stat="identity", position = "stack") + 
-  lims(x = c(0, 8)) +
-  labs(x = "Imports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 aquaculture \nimporters (1996-2000)") +
-  theme_minimal()
-
-top_aqua_imp_2 <- artis %>% 
-  filter(year %in% 2015:2019, method == "aquaculture") %>%
-  group_by(importer_iso3c, environment) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2015:2019)) %>% 
-  ungroup() %>%
-  group_by(importer_iso3c) %>%
-  mutate(total = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer = countrycode(importer_iso3c, origin = "iso3c", destination = "country.name")) %>%
-  mutate(importer = fct_reorder(as.factor(importer), total)) %>%
-  slice_max(n = 2*15, order_by = total) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer, fill = environment)) +
-  scale_fill_manual(values = environment_colors) +
-  geom_bar(stat="identity", position = "stack") + 
-  lims(x = c(0, 8)) +
-  labs(x = "Imports (million t, live weight)", y = "", fill = "Production environment",
-       title = "Top 15 aquaculture \nimporters (2015-2019)") +
-  theme_minimal()
-
-ggarrange(top_capture_imp_1, top_capture_imp_2, top_aqua_imp_1, top_aqua_imp_2, common.legend = TRUE, legend = "bottom")
+top_traders_recent <- bilateral_habitat_method_summary %>% 
+  filter(year %in% 2016:2020, habitat_method != "unknown") %>%
+  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(2016:2020))
+
+top_importers_recent <- top_traders_recent %>%
+  group_by(importer_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+top_exporters_recent <- top_traders_recent %>%
+  group_by(exporter_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(top_exporters_recent, top_importers_recent, ncol = 2, labels = c("Export", "Import"))
+
+top_traders_past <- bilateral_habitat_method_summary %>% 
+  filter(year %in% 1996:2000, habitat_method != "unknown") %>%
+  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000))
+
+top_importers_past <- top_traders_past %>%
+  group_by(importer_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+top_exporters_past <- top_traders_past %>%
+  group_by(exporter_iso3c, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(top_exporters_past, top_importers_past, ncol = 2, labels = c("Export", "Import"))
+
+# Largest export increases
+top_exporter_change <- bilateral_habitat_method_summary %>% 
+  filter(year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_iso3c, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = end-beginning) %>% 
+  arrange(desc(change))
+
+# Largest increases
+top_export_increases <- top_exporter_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Largest decreases
+top_export_decreases <- top_exporter_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_min(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(top_export_decreases, top_export_increases, nrow = 1)
+
+# Largest import increases
+top_importer_change <- bilateral_habitat_method_summary %>% 
+  filter(year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, importer_iso3c, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = end-beginning) %>% 
+  arrange(desc(change))
+
+# Largest increases
+top_import_increases <-top_importer_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Largest decreases
+top_import_decreases <-top_importer_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_min(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(top_import_decreases, top_import_increases, nrow = 1)
+
+# Repeat for regional trends
+# Largest export increases
+top_exporter_change <- bilateral_habitat_method_summary %>% 
+  filter(year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_region, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = end-beginning) %>% 
+  arrange(desc(change))
+
+# Largest increases
+top_region_export <- top_exporter_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(exporter_region = reorder_within(exporter_region, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Largest import increases
+top_importer_change <- bilateral_habitat_method_summary %>% 
+  filter(year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, importer_region, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = end-beginning) %>% 
+  arrange(desc(change))
+
+# Largest increases
+top_region_import <-top_importer_change %>% 
+  filter(habitat_method != "unknown") %>%
+  ungroup() %>%
+  group_by(habitat_method) %>%
+  slice_max(n = 10, order_by = change) %>%
+  ungroup() %>%
+  mutate(importer_region = reorder_within(importer_region, change, habitat_method)) %>%
+  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(top_region_export, top_region_import, nrow = 1, labels = c("Export", "Import"))
+
 ```
@@
 
+png("outputs/fig2a.png")
 ggarrange(ggarrange(imports_by_region, NULL,
@@
           heights = c(1, 4))
-
-```
-
-```{r africa_trade_by_method}
-africa_exports <- artis_region_method %>%
-  filter(exporter_region == "Africa", 
-         importer_region != "Africa") %>%
-  group_by(year, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/1000000) %>%
-  ggplot(aes(x = year, y = live_weight_t, color = habitat_method)) +
-  geom_line() +
-  lims(y = c(0, 3.25)) +
-  scale_color_manual(values = habitat_source_colors) +
-  theme_minimal() + 
-  labs(x = "", y = "Exports (million t, live weight)", color = "Source")
-
-africa_imports <- artis_region_method %>%
-  filter(importer_region == "Africa", 
-         exporter_region != "Africa") %>%
-  group_by(year, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/1000000) %>%
-  ggplot(aes(x = year, y = live_weight_t, color = habitat_method)) +
-  geom_line() +
-  lims(y = c(0, 3.25)) +
-  scale_color_manual(values = habitat_source_colors) +
-  theme_minimal() + 
-  labs(x = "", y = "Exports (million t, live weight)", color = "Source")
-
-ggarrange(africa_exports, africa_imports, common.legend = TRUE, labels = c("Africa exports", "Africa imports"))
+dev.off()
 ```
@@
 
-```{r, fig.height = 4, fig.width = 4}
+```{r, fig.height = 4, fig.width = 5}
 # Join figures
+png("outputs/fig3.png")
 ggarrange(
@@
 )
+dev.off()
 ```
@@
 ```{r, fig.height = 3, fig.width = 3}
+png("outputs/fig4.png")
 ggarrange(prod_import_diversity_trends, 
@@
           )
-```
-
-```{r}
-
-```
+dev.off()
+```

### 2023-01-06 ms_results.Rmd
text and figure edits for full draft

@@
 ```{r, echo=FALSE, warning=FALSE, message=FALSE}
+knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
+
 # Load packages
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r trends_summary}
+# Generate summary statistics for total trade trends
+
 # Load data summary files generated by ms_results
@@
 
-all_degree_1996 <- mean(all_degree_summary %>% filter(year == 1996) %>% pull(all_out_degree), na.rm = TRUE)
-all_degree_2019 <- mean(all_degree_summary %>% filter(year == 2019) %>% pull(all_out_degree), na.rm = TRUE)
+```
+
+```{r percent_export_by_source_trends}
+percent_export_by_source <- read.csv("outputs/percent_export_by_source_summary.csv")
+percent_export_total <- read.csv("outputs/percent_export_total_summary.csv")
+
+percent_export_total_dom_1996 <- percent_export_total %>% 
+  filter(year == 1996, calc_type == "domestic exports") %>% 
+  pull(percent_export)
+percent_export_total_dom_2019 <- percent_export_total %>% 
+  filter(year == 2019, calc_type == "domestic exports") %>% 
+  pull(percent_export)
+
+percent_export_total_1996 <- percent_export_total %>% 
+  filter(year == 1996, calc_type == "all export sources") %>% 
+  pull(percent_export)
+percent_export_total_2019 <- percent_export_total %>% 
+  filter(year == 2019, calc_type == "all export sources") %>% 
+  pull(percent_export)
+
+percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
+  filter(year == 1996, habitat_method == "marine capture") %>%
+  pull(percent_export)
+percent_marine_capture_exported_2019 <- percent_export_by_source %>% 
+  filter(year == 2019, habitat_method == "marine capture") %>%
+  pull(percent_export)
+
+percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
+  filter(year == 1996, habitat_method == "marine aquaculture") %>%
+  pull(percent_export)
+percent_marine_aqua_exported_2019 <- percent_export_by_source %>% 
+  filter(year == 2019, habitat_method == "marine aquaculture") %>%
+  pull(percent_export)
+
+percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
+  filter(year == 1996, habitat_method == "inland capture") %>%
+  pull(percent_export)
+percent_inland_capture_exported_2019 <- percent_export_by_source %>% 
+  filter(year == 2019, habitat_method == "inland capture") %>%
+  pull(percent_export)
+
+percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
+  filter(year == 1996, habitat_method == "inland aquaculture") %>%
+  pull(percent_export)
+percent_inland_aqua_exported_2019 <- percent_export_by_source %>% 
+  filter(year == 2019, habitat_method == "inland aquaculture") %>%
+  pull(percent_export)
 
@@
 
-## Text for MS
+```{r trends_concentration}
+# Calculate import/export concentration stats
 
-Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships.
-Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`% (from `r total_1996` to `r total_2019` mil t) and the average number of trade partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`).
-Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
-However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
-Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively.
-From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`%.
-
-## Other summary stats
-Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.
-
-Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period.
-In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.
-
-Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period.
-In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.
-
-# Shifts in global flows of blue foods
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Load data as bilateral flows by habitat and method
 bilateral_habitat_method_summary <- read.csv("outputs/bilateral_habitat_method_summary.csv") %>% 
@@
                                                             "unknown" )))
-```
-
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+# Set percent threshold
 cumulative_percent_threshold <- 75
@@
 
-Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)`countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)`countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
-Aquaculture exports and imports are generally concentrated among fewer countries than capture exports and imports, though inland aquaculture imports became less concentrated over the period.
+```{r degree_summary_stats}
+all_degree_1996 <- all_degree_summary %>%
+  filter(year == 1996) %>% 
+  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
+  pull(all_out_degree)
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+all_degree_2019 <- all_degree_summary %>%
+  filter(year == 2019) %>% 
+  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
+  pull(all_out_degree)
+
+```
+
+## Text for MS
+
+Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships.
+Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`) since 1996.
+Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
+However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
+Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively.
+From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period.
+Consequently, the share of global blue food exports represented by fishmeal declined from `r abs(round(100*(total_fmfo_1996)/(total_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019), 1))`% in 2019.
+
+Another measure of the degree of globalization is the share of production exported.
+Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports (i.e., domestic plus exports of foreign and unknown origin) reached `r round(percent_export_total_2019, 1)`% of all production.
+For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022).
+Marine capture products have the greatest share of production destined for export (domestic exports represented `r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, while inland aquaculture had the lowest share of production destined for export (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).
+
+One concern for increasing globalization is that it may create systemic risks from shocks, particularly when trade is highly concentrated.
+Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
+Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022).
+Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production.
+However, inland aquaculture imports have become less concentrated over time.
+The more dispersed trade patterns can be attributed to expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing blue food demand in many countries (Naylor et al.
+2021; SOFIA 2022).
+[Roz - maybe we can come up with an illustrative example from the BFA demand paper for here?]
+
+## Other summary stats
+Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.
+
+Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period.
+In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.
+
+Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period.
+In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.
+
+```{r covid_effect}
+artis %>% 
+  filter(year %in% 2019:2020) %>% 
+  group_by(year, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>% 
+  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
+  mutate(change = (year2019-year2020)/year2020) %>% 
+  arrange(desc(change))
+```
+
+# Shifts in global flows of blue foods
+
+```{r top_traders}
 # Top trade tables
@@
 
-```{r, echo=FALSE, warning=FALSE, message=FALSE}
+```{r top_bilateral_flows}
+
 bilateral_habitat_method_summary %>% 
@@
   slice_max(n=1, order_by = live_weight_t)
+
+# Regionalization trends
+# Changes in trade flow volumes
+bilateral_habitat_method_summary %>% 
+  filter(exporter_region != "Other nei", importer_region != "Other nei",
+         year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_region, importer_region) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = end-beginning,
+         from_to = paste(exporter_region, " to ", importer_region), 
+         intraregion = case_when(
+           exporter_region == importer_region ~ "intraregional trade",
+           exporter_region != importer_region ~ "interegional trade"
+         )) %>% 
+  arrange(desc(change)) %>% 
+  ungroup() %>%
+  ggplot(aes(x = change/1000000, y = reorder(from_to, change), color = intraregion)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
+  geom_point(size=3) +
+  scale_y_reordered() +
+  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
+  labs(y = "", x = "Live weight (million t)", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+# Changes in trade flow percents
+bilateral_habitat_method_summary %>% 
+  filter(exporter_region != "Other nei", importer_region != "Other nei",
+         year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_region, importer_region) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  mutate(change = 100*(end-beginning)/beginning,
+         from_to = paste(exporter_region, " to ", importer_region), 
+         intraregion = case_when(
+           exporter_region == importer_region ~ "intraregional trade",
+           exporter_region != importer_region ~ "interegional trade"
+         )) %>% 
+  arrange(desc(change)) %>% 
+  ungroup() %>%
+  ggplot(aes(x = change, y = reorder(from_to, change), color = intraregion)) +
+  geom_segment(aes(x=0, xend=change, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
+  geom_point(size=3) +
+  scale_y_reordered() +
+  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
+  labs(y = "", x = "Percent change", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
 ```
-
-For all regions other than Oceania and South America, interregional trade is either the highest or second high
 
@@
 
+The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
+In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
+Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
+At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
+Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
+Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
+
+Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
+This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
+[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
+
+Since 1996, nearly all regional trade pairs increased, other than South America.
+The largest increases occurred for X, Y and Z.
+At the country level, X% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
+The declining exports from South America mirror declining marine catch over this period.
+[Suggestions for discussion points and context related to this?].
+
+Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
+Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
+High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
+2016).
+
 ## Other summary stats
 
-# Role of trade in national blue food supplies
+# National blue food supplies
 
@@
 
-## Other summary stats
+Since direct measurements of consumption are not collected globally, our understanding of global consumption patterns are based on apparent consumption, which generally represents the sum of production and imports minus exports and waste.
+Consequently, trade is central to estimating consumption and the resolution of consumption patterns is limited by the resolution of the production and trade data.
+By estimating species level trade, we are able to estimate apparent consumption of blue foods by species and production method.
 
-# Trade and blue food supply diversity
+Globally, blue food consumption increased from X in 1996 to Y in 2020.
+Blue food consumption also increased across all regions outside of South America, where per capita blue food consumption decreased X%, from X% in 1996 to Y% in 2020.
+[Double check that FMFO was excluded and discuss this pattern].
+Most of the global increase was from inland and marine aquaculture, which increased by X and Y%, respectively, over that period while marine capture consumption declined by X%.
+Marine capture still makes up X% of global blue food consumption and its contribution to regional blue food supplies ranges from X% in Y to X% in Asia, where farmed consumption overtook wild consumption in [year].
 
-## Text for MS
+High reliance on foreign-sourced foods can pose a food security risk, motivating policies to protect domestic supplies such as the development of grain stocks and support for domestic food production (REFs).
+For example, high seafood import dependence in the US was used to propose a suite of policy changes to boost domestic production, including expanding aquaculture production and opening marine protected areas to fishing (Gephart et al.
+2019).
+Since we estimate exports from domestic and foreign sources, we are able to track changes in reliance on foreign sourced products.
+Globally, the share of consumption that is foreign sourced has increased from X% to Y% (Fig 3c), but patterns vary across regions with Asia and S America dominated by domestic supply, but Europe dominated by foreign supply (Fig 3d).
+[Add points about patterns across production methods].
+
+The diversity of blue food production globally enables high diversity in blue food consumption.
+However, one concern about food trade is that it leads to homogenization of diets (REFS).
+To explore this issue, we compared the diversity (represented by the Shannon index) of production versus imports and domestic supply versus foreign supply, as well as the diversity of total supply versus total import volume (Fig 4).
+One notable limitation of this measure of supply diversity is that instances where production or trade is only estimated at the genus or higher level will be interpreted in the calculation as a single species.
+Consequently, countries with low species resolution will generally appear to have less diverse production, limiting the ability to compare supply diversity across countries.
+We find that across all regions and over time, production diversity is higher than import diversity (Fig 4a).
+Additionally, the diversity of domestic supply is generally higher than the diversity of foreign supply, with the domestic diversity is higher in 68% of cases across all years (Fig 4b).
+Nevertheless, total imports are still positively associated with supply diversity across all regions (Fig 4c).
+This suggests that although domestic production diversity is generally higher than the diversity of imported products, trade is still bringing in new species and increasing the diversity of products available on the market.
+Although imports appear to currently add to supply diversity, it is important for it not to displace domestic production diversity, particularly given that domestic production still contributes more to the diversity of the overall blue food supply.
 

### 2023-01-06 ms_figures.Rmd
ms manuscript changes

@@
 ```{r set_directories}
-# Set time series dates
-date_nodash <- "20220926"
-date_dash <- "2022-09-26"
-
 # Set directories 
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
+outdir <- "outputs"
 
@@
   labs(fill = "Region") +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom") +
+  theme_bw()
 ```
@@
 
+write.csv(supply, file.path(outdir, "supply.csv"), row.names = FALSE)
 ```
@@
   group_by(year, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
+  summarise(supply_per_cap = 1000 * (sum(supply_no_error))/sum(pop)) %>%
   ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
@@
   left_join(pop %>% select(-region), by = c("iso3c", "year")) %>%
-  mutate(supply_per_cap = supply_no_error/pop) 
+  mutate(supply_per_cap = supply_no_error/pop) %>%
+  # Converting from tonnes to kg
+  mutate(supply_per_cap = 1000 * supply_per_cap)
 ```
@@
   mutate(iso3c = fct_reorder(iso3c, supply_per_cap)) %>%
-  filter(supply_per_cap > 50) %>%
+  #filter(supply_per_cap > 50) %>%
   ggplot(aes(x = supply_per_cap, y = iso3c, fill = region)) +
@@
             labels = c("c", "d")),
-  ggarrange(import_supply, export_supply, 
-            common.legend = TRUE, nrow = 1, widths = c(1, 1),
-            labels = c("e", "f")),
+  # ggarrange(import_supply, export_supply, 
+  #           common.legend = TRUE, nrow = 1, widths = c(1, 1),
+  #           labels = c("e", "f")),
   nrow = 3

### 2023-01-06 ms_results.Rmd
ms manuscript changes

@@
 library(ggpubr)
+library(tidytext)
 
@@
 ```{r covid_effect}
-artis %>% 
-  filter(year %in% 2019:2020) %>% 
-  group_by(year, habitat_method) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>% 
-  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
-  mutate(change = (year2019-year2020)/year2020) %>% 
-  arrange(desc(change))
+# artis %>% 
+#   filter(year %in% 2019:2020) %>% 
+#   group_by(year, habitat_method) %>% 
+#   summarise(live_weight_t = sum(live_weight_t)) %>% 
+#   pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
+#   mutate(change = (year2019-year2020)/year2020) %>% 
+#   arrange(desc(change))
 ```
@@
 
+```{r consumption}
+
+supply <- read.csv("outputs/supply.csv")
+
+global_supply <- supply %>%
+  group_by(year) %>%
+  summarize(supply_no_error = sum(supply_no_error) / 1000000) %>%
+  ungroup()
+
+global_supply_1996 <- global_supply %>%
+  filter(year == 1996) %>%
+  pull(supply_no_error)
+
+global_supply_2020 <- global_supply %>%
+  filter(year == 2020) %>%
+  pull(supply_no_error)
+
+regional_supply <- supply %>%
+  group_by(year, region) %>%
+  summarize(supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  group_by(year) %>%
+  mutate(supply_no_error_total = sum(supply_no_error)) %>%
+  ungroup() %>%
+  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)
+
+habitat_method_supply <- supply %>%
+  group_by(year, habitat, method) %>%
+  summarize(supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  group_by(year) %>%
+  mutate(supply_no_error_total = sum(supply_no_error)) %>%
+  ungroup() %>%
+  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)
+
+tmp <- habitat_method_supply %>%
+  select(habitat, method, year, supply_no_error) %>%
+  filter(year == 1996 | year == 2020) %>%
+  mutate(year = paste("year", year, sep = "_")) %>%
+  pivot_wider(names_from = year, values_from = supply_no_error) %>%
+  mutate(percent_diff = 100 * (year_2020 - year_1996) / year_1996)
+
+inland_aqua_supply_change <- tmp %>%
+  filter(habitat == "inland" & method == "aquaculture") %>%
+  pull(percent_diff)
+
+marine_aqua_supply_change <- tmp %>%
+  filter(habitat == "marine" & method == "aquaculture") %>%
+  pull(percent_diff)
+
+marine_cap_supply_change <- tmp %>%
+  filter(habitat == "marine" & method == "capture") %>%
+  mutate(percent_diff = abs(percent_diff)) %>%
+  pull(percent_diff)
+
+marine_cap_supply_percent <- habitat_method_supply %>%
+  filter(year == 2020) %>%
+  filter(percent_supply == max(percent_supply)) %>%
+  pull(percent_supply)
+
+method_supplies <- supply %>%
+  filter(year == 2020) %>%
+  group_by(method) %>%
+  summarize(supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  mutate(total_supply = sum(supply_no_error)) %>%
+  mutate(percent_supply = 100 * supply_no_error / total_supply)
+
+regional_method_supplies <- supply %>%
+  filter(year == 2020) %>%
+  group_by(year, region, method) %>%
+  summarize(supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  group_by(year, region) %>%
+  mutate(region_supply = sum(supply_no_error)) %>%
+  ungroup() %>%
+  mutate(percent_supply = 100 * supply_no_error / region_supply) %>%
+  select(region, region_supply, method, percent_supply) %>%
+  pivot_wider(names_from = "method", values_from = "percent_supply")
+
+oceania_cap_supply_percent <- regional_method_supplies %>%
+  filter(region == "Oceania") %>%
+  pull(capture)
+
+asia_cap_supply_percent <- regional_method_supplies %>%
+  filter(region == "Asia") %>%
+  pull(capture)
+
+cap_supply_percent <- method_supplies %>%
+  filter(method == "capture") %>%
+  pull(percent_supply)
+
+asia_farmed_year <- supply %>%
+  filter(region == "Asia") %>%
+  group_by(year, method) %>%
+  summarize(supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  filter(method == "capture" | method == "aquaculture") %>%
+  pivot_wider(names_from = method, values_from = supply_no_error) %>%
+  filter(aquaculture > capture) %>%
+  filter(year == min(year)) %>%
+  pull(year)
+
+consumption_foreign <- supply %>%
+  filter(year == 1996 | year == 2020) %>%
+  group_by(year) %>%
+  summarize(supply_foreign = sum(supply_foreign),
+            supply_no_error = sum(supply_no_error)) %>%
+  ungroup() %>%
+  mutate(percent_supply = 100 * supply_foreign / supply_no_error)
+
+foreign_consumption_start <- consumption_foreign %>%
+  filter(year == 1996) %>%
+  pull(percent_supply)
+
+foreign_consumption_end <- consumption_foreign %>%
+  filter(year == 2020) %>%
+  pull(percent_supply)
+
+```
+
 ## Text for MS
@@
 
-Globally, blue food consumption increased from X in 1996 to Y in 2020.
+Globally, blue food consumption increased from `r signif(global_supply_1996, digit = 3)` (million tonnes) in 1996 to `r signif(global_supply_2020, digit = 3)` (million tonnes) in 2020.
 Blue food consumption also increased across all regions outside of South America, where per capita blue food consumption decreased X%, from X% in 1996 to Y% in 2020.
-[Double check that FMFO was excluded and discuss this pattern].
-Most of the global increase was from inland and marine aquaculture, which increased by X and Y%, respectively, over that period while marine capture consumption declined by X%.
-Marine capture still makes up X% of global blue food consumption and its contribution to regional blue food supplies ranges from X% in Y to X% in Asia, where farmed consumption overtook wild consumption in [year].
+[say this is supply included non food uses] [Double check that FMFO was excluded and discuss this pattern].
+Most of the global increase was from inland and marine aquaculture, which increased by `r signif(inland_aqua_supply_change, digits = 3)` and `r signif(inland_aqua_supply_change, digits = 3)`%, respectively, over that period while marine capture consumption declined by `r signif(marine_cap_supply_change, digits = 1)`%.
+Capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global blue food consumption and its contribution to regional blue food supplies ranges from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.
 
@@
 Since we estimate exports from domestic and foreign sources, we are able to track changes in reliance on foreign sourced products.
-Globally, the share of consumption that is foreign sourced has increased from X% to Y% (Fig 3c), but patterns vary across regions with Asia and S America dominated by domestic supply, but Europe dominated by foreign supply (Fig 3d).
+Globally, the share of consumption that is foreign sourced has increased from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c), but patterns vary across regions with Asia and S America dominated by domestic supply, but Europe dominated by foreign supply (Fig 3d).
 [Add points about patterns across production methods].

### 2023-01-06 ms_figures.Rmd
update trade flow trend section and figure

--- 2023-01-06 ms_figures.Rmd
@@
 
+# Create basic plot for the legend
+# artis_region_method %>%
+#   group_by(year, exporter_region) %>%
+#   summarise(live_weight_t = sum(live_weight_t)) %>%
+#   ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
+#   geom_area() +
+#   scale_fill_manual(values = region_colors) +
+#   labs(fill = "Region") +
+#   theme(legend.position = "bottom")
+```
+
+```{r regional_bilateral_trends}
+artis_region <- artis_region_method %>%
+  group_by(year, exporter_region, importer_region) %>%
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
+
+region_slopes <- artis_region %>%
+  split(.$exporter_importer) %>% 
+  map(~lm(live_weight_t~year, data = .x)) %>% 
+  map_df(broom::tidy, .id = 'exporter_importer') %>%
+  filter(term == 'year') %>%
+  select(exporter_importer, "slope" = "estimate")
+
+artis_region <- artis_region %>%
+  left_join(region_slopes, by = "exporter_importer") %>% 
+  mutate(intraregion = case_when(
+           exporter_region == importer_region ~ "intraregional trade",
+           exporter_region != importer_region ~ "interegional trade"))
+
+```
+
+# SI Fig X: Bilateral region trends lollypop plot
+
+```{r region_trend_lollipop, fig.height=4, fig.width=3}
+region_slopes_ranked <- artis_region %>%
+  select(-year, -live_weight_t) %>%
+  distinct() %>%
+  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
+  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
+  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
+  geom_point(size=3) +
+  scale_y_reordered() +
+  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
+  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+region_slopes_ranked
+
+```
+
+```{r, fig.height=3, fig.width=5}
 region_trend_matrix_method <- artis_region_method %>% 
@@
         strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        strip.text.y = element_blank()) +
+        plot.margin=unit(c(0,0,0,0), "cm")) +
   facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
@@
 
-# Create basic plot for the legend
-artis_region_method %>%
-  group_by(year, exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
-  geom_area() +
-  scale_fill_manual(values = region_colors) +
-  labs(fill = "Region") +
-  theme(legend.position = "bottom")
-```
-
-```{r regional_bilateral_trends}
-artis_region <- artis_region_method %>%
-  group_by(year, exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
-
-region_slopes <- artis_region %>%
-  split(.$exporter_importer) %>% 
-  map(~lm(live_weight_t~year, data = .x)) %>% 
-  map_df(broom::tidy, .id = 'exporter_importer') %>%
-  filter(term == 'year') %>%
-  select(exporter_importer, "slope" = "estimate")
-
-artis_region <- artis_region %>%
-  left_join(region_slopes, by = "exporter_importer")
-
-```
-
-```{r region_trend_lollipop, fig.height=4, fig.width=8}
-region_slopes_ranked <- region_slopes %>%
-  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
-  ggplot(aes(x = slope/1000, y = exporter_importer, color = slope, fill = slope)) +
-  geom_segment(aes(xend = 0, yend = exporter_importer)) +
-  geom_point() +
-  scale_color_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
-  scale_fill_gradient2(low = "#E24027", mid = "grey70", high = "#1B708F") +
-  labs(y = "", x = "Ave. annual change in export (1000 t live weight)") +
-  theme_bw() +
-  theme(legend.position = "blank")
-
-ggarrange(region_trend_matrix_method, region_slopes_ranked, widths = c(1.5,1))
-```
-
-```{r, fig.height=3, fig.width=4}
 exports_by_region <- artis_region %>% 
@@
         strip.text.x = element_blank(),
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.margin=unit(c(0,0.5,0,0), "cm"))
 
@@
         strip.text.x = element_blank(),
-        axis.text.x = element_blank())
-
-png("outputs/fig2a.png")
+        axis.text.x = element_blank(),
+        plot.margin=unit(c(0.5,0,0,0), "cm"))
+
+png("outputs/fig2a.png", width = 6.5, height = 5, res = 500, units = "in")
 ggarrange(ggarrange(imports_by_region, NULL,
                     ncol = 2, nrow = 1,
-                    widths = c(4, 1)),
-          ggarrange(region_trend_matrix_method, exports_by_region, 
+                    widths = c(4.5, 1.25), labels = c("a", "")),
+          ggarrange(region_trend_matrix_method, 
+                    ggarrange(NULL, exports_by_region,
+                              ncol = 1, heights = c(0.3, 4)), 
                     ncol = 2, nrow = 1,
-                    widths = c(4, 1)),
+                    widths = c(4.5, 1), labels = c("b", "c")),
           ncol = 1, nrow = 2, 
-          heights = c(1, 4))
+          heights = c(1, 4.5))
 dev.off()

### 2023-01-06 ms_results.Rmd
update trade flow trend section and figure

--- 2023-01-06 ms_results.Rmd
@@
   summarise(live_weight_t = sum(live_weight_t)) %>% 
-  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
+  pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year", values_fill = 0) %>% 
   mutate(change = (year2019-year2020)/year2020) %>% 
@@
   summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
   mutate(change = end-beginning,
@@
 
-# Changes in trade flow percents
-bilateral_habitat_method_summary %>% 
+# Calculate percent of pairs with increased trade flows
+bilateral_flow_change <- bilateral_habitat_method_summary %>% 
   filter(exporter_region != "Other nei", importer_region != "Other nei",
@@
     year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_region, importer_region) %>% 
+  group_by(year_group, exporter_iso3c, importer_iso3c) %>% 
   summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = 100*(end-beginning)/beginning,
-         from_to = paste(exporter_region, " to ", importer_region), 
-         intraregion = case_when(
-           exporter_region == importer_region ~ "intraregional trade",
-           exporter_region != importer_region ~ "interegional trade"
-         )) %>% 
-  arrange(desc(change)) %>% 
-  ungroup() %>%
-  ggplot(aes(x = change, y = reorder(from_to, change), color = intraregion)) +
-  geom_segment(aes(x=0, xend=change, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
-  geom_point(size=3) +
-  scale_y_reordered() +
-  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
-  labs(y = "", x = "Percent change", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
+  mutate(change = end-beginning) %>%
+  arrange(desc(change))
+
+percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)
+
+# Calculate percent of pairs with increased trade flows by production method and habitat
+bilateral_flow_change_method <- bilateral_habitat_method_summary %>% 
+  filter(exporter_region != "Other nei", importer_region != "Other nei",
+         year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_iso3c, importer_iso3c, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
+  mutate(change = end-beginning) %>%
+  arrange(desc(change))
+
+percent_flow_increases_marine_capture <- 100*(nrow(bilateral_flow_change_method %>% 
+                                                     filter(habitat_method == "marine capture",
+                                                            change > 0)))/nrow(bilateral_flow_change)
 
@@
 
-Since 1996, nearly all regional trade pairs increased, other than South America.
-The largest increases occurred for X, Y and Z.
-At the country level, X% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
+Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America.
+The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia.
+At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
 The declining exports from South America mirror declining marine catch over this period.

### 2023-01-06 ms_figures.Rmd

@@
 # Create basic plot for the legend
+
 # artis_region_method %>%

### 2023-01-08 ms_results.Rmd
updates ahead of circulating among coauthors

--- 2023-01-06 ms_results.Rmd
@@
     year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_region, importer_region) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = 100*(end-beginning)/beginning,
-         from_to = paste(exporter_region, " to ", importer_region), 
-         intraregion = case_when(
-           exporter_region == importer_region ~ "intraregional trade",
-           exporter_region != importer_region ~ "interegional trade"
-         )) %>% 
-  arrange(desc(change)) %>% 
-  ungroup() %>%
-  ggplot(aes(x = change, y = reorder(from_to, change), color = intraregion)) +
-  geom_segment(aes(x=0, xend=change, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
-  geom_point(size=3) +
-  scale_y_reordered() +
-  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
-  labs(y = "", x = "Percent change", color = "") +
+  group_by(year_group, exporter_iso3c, importer_iso3c) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
+  mutate(change = end-beginning) %>%
+  arrange(desc(change))
+
+percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)
+
+# Calculate percent of pairs with increased trade flows by production method and habitat
+bilateral_flow_change_method <- bilateral_habitat_method_summary %>% 
+  filter(exporter_region != "Other nei", importer_region != "Other nei",
+         year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_iso3c, importer_iso3c, habitat_method) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
+  mutate(change = end-beginning) %>%
+  arrange(desc(change))
+
+percent_flow_increases_marine_capture <- 100*(nrow(bilateral_flow_change_method %>% 
+                                                     filter(habitat_method == "marine capture",
+                                                            change > 0)))/nrow(bilateral_flow_change)
+
+```
+
+```{r LDC_patterns}
+# LDC countries according to UN (https://www.un.org/development/desa/dpad/least-developed-country-category/ldcs-at-a-glance.html)
+LDC_countries <- c("Afghanistan", "Angola", "Bangladesh", "Benin", "Bhutan", "Burkina Faso", "Burundi", "Cambodia", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Djibouti", "Eritrea", "Ethiopia", "Gambia", "Guinea", "Guinea-Bissau", "Haiti", "Kiribati", "Lao People’s Dem.
+Republic", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Myanmar", "Nepal", "Niger", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Solomon Islands", "Somalia", "South Sudan", "Sudan", "Timor-Leste", "Togo", "Tuvalu", "Uganda", "United Republic of Tanzania", "Yemen", "Zambia")
+
+LDC_countries <- countrycode(LDC_countries, origin = "country.name", destination = "iso3c")
+
+ldc_habitat_method_summary <- bilateral_habitat_method_summary %>%
+  mutate(
+    importer_ldc = case_when(
+    (importer_iso3c %in% LDC_countries) ~ "LDC",
+    !(importer_iso3c %in% LDC_countries) ~ "non-LDC"),
+     exporter_ldc = case_when(
+    (exporter_iso3c %in% LDC_countries) ~ "LDC",
+    !(exporter_iso3c %in% LDC_countries) ~ "non-LDC")) %>%
+  group_by(year, exporter_ldc, importer_ldc, habitat_method) %>%
+  summarise(live_weight_t = sum(live_weight_t))
+
+ldc_net_habitat_method_summary <- ldc_habitat_method_summary %>%
+  ungroup() %>% 
+  filter(importer_ldc == "LDC", exporter_ldc == "non-LDC") %>% 
+  select(year, habitat_method, "ldc_import" = "live_weight_t") %>%
+  left_join(ldc_habitat_method_summary %>%
+              ungroup() %>%
+              filter(importer_ldc == "non-LDC", exporter_ldc == "LDC") %>% 
+              select(year, habitat_method, "ldc_export" = "live_weight_t"),
+            by = c("year", "habitat_method")) %>%
+  mutate(net_import = ldc_import-ldc_export,
+         net_export = ldc_export-ldc_import)
+
+ldc_net_habitat_method_summary %>%
+  filter(habitat_method != "unknown") %>%
+  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(x = "", y = "Net export (million t)", color = "") +
   theme_minimal() +
@@
 
+ldc_net_habitat_method_summary %>%
+  group_by(year) %>% 
+  summarise(net_export = sum(net_export))
+
 ```
+
+## Text for MS
+
+The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
+In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
+Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
+At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
+Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
+Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
+
+Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
+This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
+[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
+
+Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America.
+The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia.
+At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
+The declining exports from South America mirror declining marine catch over this period.
+[Suggestions for discussion points and context related to this?].
+
+Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
+Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
+High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
+2016).
+
+## Other summary stats
+
+# National blue food supplies
 
@@
 ```
-
-## Text for MS
-
-The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
-In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
-Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
-At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
-Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
-Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
-
-Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
-This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
-[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
-
-Since 1996, nearly all regional trade pairs increased, other than South America.
-The largest increases occurred for X, Y and Z.
-At the country level, X% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
-The declining exports from South America mirror declining marine catch over this period.
-[Suggestions for discussion points and context related to this?].
-
-Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
-Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
-High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
-2016).
-
-## Other summary stats
-
-# National blue food supplies
 

### 2023-01-17 ms_figures.Rmd
new ms figures pipeline

--- 2023-01-06 ms_figures.Rmd
@@
 
-```{r load_database_data}
-# Load ARTIS data from SQL local database
-con <- dbConnect(RPostgres::Postgres(),
-                 dbname=Sys.getenv("POSTGRES_DB"),
-                 host="localhost",
-                 port="5432",
-                 user=Sys.getenv("POSTGRES_USER"),
-                 password=Sys.getenv("POSTGRES_PASSWORD"))
+```{r load_data}
+# Set directories
+datadir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/outputs_20220111"
 
-# Check that connection is established by checking which tables are present
-dbListTables(con)
-
-# Pull all ARTIS and production data
-artis <- dbGetQuery(con, "SELECT * FROM snet")
-
-artis <- artis %>%
-  select(-record_id)
-
-prod <- dbGetQuery(con, "SELECT * FROM production")
-prod <- prod %>%
-  select(-record_id)
-
-country_metadata <- dbGetQuery(con, "SELECT * FROM countries")
-country_metadata <- country_metadata %>%
-  select(-record_id)
-
-code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
-              mutate(hs6 = as.character(hs6)) %>%
-  select(-record_id)
-
-sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
-  select(-record_id)
-
-# Close database connection
-dbDisconnect(con)
-
-rm(list = c("con"))
 ```
 
-```{r set_directories}
-# Set time series dates
-date_nodash <- "20220926"
-date_dash <- "2022-09-26"
-
-# Set directories 
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
-
+```{r color_palettes}
 # Set color palettes
@@
 
-```{r join_data}
-# Summarize cleaned FAO production data
-prod <- prod %>% 
-  select("year", "iso3c", "sciname", 
-         "method", "habitat", 
-         "production_t" = "live_weight_t") %>%
-  group_by(year, iso3c, sciname, method, habitat) %>%
-  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
-  ungroup()
-
-# Join production data to ARTIS
-artis <- artis %>%
-  filter(live_weight_t >= 0.1) %>%
-  # Add importer and exporter region
-  left_join(country_metadata %>% 
-              select("exporter_iso3c" = "iso3c", "exporter_region" = "owid_region"), by = c("exporter_iso3c")) %>%
-  left_join(country_metadata %>% 
-              select("importer_iso3c" = "iso3c", "importer_region" = "owid_region"), by = c("importer_iso3c")) %>%
-# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
-  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
-                           "sciname", "method", "habitat")) %>%
-  ungroup()
-
-# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
-artis <- artis %>%
-  left_join(code_max_resolved_taxa,
-            by = c("hs_version", "hs6", "sciname")) %>%
-  left_join(sciname_metadata, 
-            by = "sciname") %>%
-  # Leave any missing sciname_hs_modified as original sciname
-  mutate(sciname_hs_modified = case_when(
-    is.na(sciname_hs_modified) ~ sciname,
-    TRUE ~ sciname_hs_modified
-  )) %>%
-  rename(environment = habitat) %>%
-  # Add habitat-method column
-  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
-  mutate(habitat_method = case_when(
-    str_detect(habitat_method, "unknown") ~ "unknown", 
-    TRUE ~ habitat_method
-  )) %>% 
-  # Set factor levels
-  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
-                                                            "marine aquaculture", "inland aquaculture",
-                                                            "unknown" )))
-
-# Remove data frames after merged with ARTIS
-rm(list = c("code_max_resolved_taxa", "sciname_metadata"))
-
-```
-
 # Figure 1: Increases in global export of blue foods
@@
 ```{r fig1a}
-global_export_by_source <- artis %>% 
-  mutate(food_or_fmfo = case_when(
-    hs6 == "230120" ~ "fishmeal",
-    hs6 != "230120" ~ "nonfishmeal"
-  )) %>%
-  group_by(year, habitat_method, food_or_fmfo) %>%
-  summarise(live_weight_t = sum(live_weight_t))
+fig1a_data <- read.csv(file.path(datadir, "fig_1a_data.csv"))
 
-write.csv(global_export_by_source, "outputs/global_export_by_source.csv", row.names = FALSE)
-
-fig1a <- global_export_by_source %>%
-  group_by(year, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
+fig1a <- fig1a_data %>%
   ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
@@
 
+fig1a
 ```
@@
 ```{r fig1b}
-prod_by_source <- prod %>%
-  group_by(year, method, habitat) %>% 
-  summarise(production_t = sum(production_t)) %>%
-  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
-    # Set factor levels
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture",
-                                            "marine aquaculture", "inland aquaculture"))) %>%
-  ungroup() %>%
-  select(-habitat, -method)
-
-dom_export_by_source <- artis %>% 
-  filter(dom_source == "domestic export") %>%
-  group_by(year, habitat_method, environment) %>%
-  summarise(dom_live_weight_t = sum(live_weight_t))
-
-percent_export_by_source <- dom_export_by_source %>%
-  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
-  mutate(percent_export = 100*dom_live_weight_t/production_t)
-
-write.csv(percent_export_by_source, "outputs/percent_export_by_source_summary.csv", row.names = FALSE)
-
-percent_export_total <- dom_export_by_source %>%
-  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
-  group_by(year) %>% 
-  summarise(dom_live_weight_t = sum(dom_live_weight_t), 
-            production_t = sum(production_t)) %>%
-  mutate(percent_export = 100*dom_live_weight_t/production_t) %>% 
-  mutate(calc_type = "domestic exports") %>% 
-  select(year, percent_export, calc_type)
-
-percent_export_total_all_source <- artis %>% 
-  group_by(year) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  left_join(prod_by_source %>% 
-              group_by(year) %>% 
-              summarise(production_t = sum(production_t)), 
-            by = c("year")) %>% 
-  mutate(percent_export = 100*live_weight_t/production_t) %>% 
-  mutate(calc_type = "all export sources") %>% 
-  select(year, percent_export, calc_type)
-
-percent_export_total <- percent_export_total %>%
-  bind_rows(percent_export_total_all_source)
-
-write.csv(percent_export_total, "outputs/percent_export_total_summary.csv", row.names = FALSE)
+fig1b_data <- read.csv(file.path(datadir, "fig_1b_data.csv"))
 
@@
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "% production exported", x = "", color = "Source") +
+  labs(y = "% Production Exported", x = "", color = "Source") +
   theme_bw()
 
+fig1b
 ```
@@
 ```{r fig1c}
-all_degree_summary <-  artis %>% 
-  select(year, exporter_iso3c, importer_iso3c) %>%
-  distinct() %>%
-  group_by(year, exporter_iso3c) %>%
-  tally() %>% 
-  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
-  full_join(artis %>% 
-              select(year, exporter_iso3c, importer_iso3c) %>%
-              distinct() %>%
-              group_by(year, importer_iso3c) %>%
-              tally() %>%
-              rename(all_in_degree = n, iso3c = importer_iso3c), 
-            by = c("year", "iso3c"))
+fig1c_data <- read.csv(file.path(datadir, "fig_1c_data.csv"))
 
-write.csv(all_degree_summary, "outputs/all_degree_summary.csv", row.names = FALSE)
-
-habitat_method_degree_summary <-  artis %>% 
-  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
-  distinct() %>%
-  group_by(year, exporter_iso3c, habitat_method) %>%
-  tally() %>% 
-  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
-  full_join(artis %>% 
-              select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
-              distinct() %>%
-              group_by(year, importer_iso3c, habitat_method) %>%
-              tally() %>%
-              rename(all_in_degree = n, iso3c = importer_iso3c), 
-            by = c("year", "iso3c", "habitat_method"))
-
-write.csv(habitat_method_degree_summary, "outputs/habitat_method_degree_summary.csv", row.names = FALSE)
-
-fig1c_out_degree <- artis %>% 
-  filter(habitat_method != "unknown") %>%
-  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
-  distinct() %>%
-  group_by(year, exporter_iso3c, habitat_method) %>%
-  tally() %>%
-  group_by(year, habitat_method) %>%
-  ggplot(aes(x = year, y = n, color = habitat_method)) +
+fig1c_out_degree <- fig1c_data %>%
+  ggplot(aes(x = year, y = all_in_degree, color = habitat_method)) +
   geom_smooth() +
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "Out degree", fill = "Source") + 
+  labs(y = "Out degree", color = "Source") + 
   theme_bw()
 
-fig1c_in_degree <- artis %>% 
-  filter(habitat_method != "unknown") %>%
-  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
-  distinct() %>%
-  group_by(year, importer_iso3c, habitat_method) %>%
-  tally() %>%
-  group_by(year, habitat_method) %>%
-  ggplot(aes(x = year, y = n, color = habitat_method)) +
-  geom_smooth() +
-  scale_color_manual(values = habitat_source_colors) +
-  labs(y = "In degree", fill = "Source") + 
-  theme_bw()
-
+fig1c_out_degree
 ```
@@
 ```{r fig1, fig.height = 3, fig.width = 4.5}
-png("outputs/fig1.png")
 ggarrange(fig1a,
@@
 
-dev.off()
+ggsave(file.path(outdir, "fig1.png"))
 ```
@@
 
-```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
-# Summarize total imports and exports by habitat and environment
-bilateral_habitat_method_summary <- artis %>%
-  # Remove if we decide to leave FMFO in
-  filter(hs6 != "230120") %>%
-  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t))
-
-write.csv(bilateral_habitat_method_summary, "outputs/bilateral_habitat_method_summary.csv", row.names = FALSE)
-
-top_traders_recent <- bilateral_habitat_method_summary %>% 
-  filter(year %in% 2016:2020, habitat_method != "unknown") %>%
-  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(2016:2020))
-
-top_importers_recent <- top_traders_recent %>%
-  group_by(importer_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = live_weight_t) %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  mutate(ranking = rank(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-top_exporters_recent <- top_traders_recent %>%
-  group_by(exporter_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = live_weight_t) %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  mutate(ranking = rank(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(top_exporters_recent, top_importers_recent, ncol = 2, labels = c("Export", "Import"))
-
-top_traders_past <- bilateral_habitat_method_summary %>% 
-  filter(year %in% 1996:2000, habitat_method != "unknown") %>%
-  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000))
-
-top_importers_past <- top_traders_past %>%
-  group_by(importer_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = live_weight_t) %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  mutate(ranking = rank(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-top_exporters_past <- top_traders_past %>%
-  group_by(exporter_iso3c, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = live_weight_t) %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  mutate(ranking = rank(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(top_exporters_past, top_importers_past, ncol = 2, labels = c("Export", "Import"))
-
-# Largest export increases
-top_exporter_change <- bilateral_habitat_method_summary %>% 
-  filter(year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_iso3c, habitat_method) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = end-beginning) %>% 
-  arrange(desc(change))
-
-# Largest increases
-top_export_increases <- top_exporter_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-# Largest decreases
-top_export_decreases <- top_exporter_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_min(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(top_export_decreases, top_export_increases, nrow = 1)
-
-# Largest import increases
-top_importer_change <- bilateral_habitat_method_summary %>% 
-  filter(year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, importer_iso3c, habitat_method) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = end-beginning) %>% 
-  arrange(desc(change))
-
-# Largest increases
-top_import_increases <-top_importer_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-# Largest decreases
-top_import_decreases <-top_importer_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_min(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(top_import_decreases, top_import_increases, nrow = 1)
-
-# Repeat for regional trends
-# Largest export increases
-top_exporter_change <- bilateral_habitat_method_summary %>% 
-  filter(year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_region, habitat_method) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = end-beginning) %>% 
-  arrange(desc(change))
-
-# Largest increases
-top_region_export <- top_exporter_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(exporter_region = reorder_within(exporter_region, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-# Largest import increases
-top_importer_change <- bilateral_habitat_method_summary %>% 
-  filter(year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, importer_region, habitat_method) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = end-beginning) %>% 
-  arrange(desc(change))
-
-# Largest increases
-top_region_import <-top_importer_change %>% 
-  filter(habitat_method != "unknown") %>%
-  ungroup() %>%
-  group_by(habitat_method) %>%
-  slice_max(n = 10, order_by = change) %>%
-  ungroup() %>%
-  mutate(importer_region = reorder_within(importer_region, change, habitat_method)) %>%
-  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(top_region_export, top_region_import, nrow = 1, labels = c("Export", "Import"))
-
+```{r fig2_data}
+artis_region_method <- read.csv(file.path(datadir, "artis_region_method.csv"))
+artis_region <- read.csv(file.path(datadir, "artis_region.csv"))
 ```
 
-```{r SI_top_traders}
-total_country_exports <- artis %>% 
-  filter(year %in% 2015:2019) %>%
-  group_by(exporter_iso3c, method, environment) %>%
-  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
-  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
-  rename("iso3c" = "exporter_iso3c")
+```{r fig2a}
+# Imports by Region
+fig2a <- artis_region %>% 
+  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
+  geom_area() +
+  labs(x = "", y = "", fill = "Exporter") +
+  scale_fill_manual(values = region_colors) +
+  facet_wrap(~importer_region, nrow = 1) +
+  theme_bw() +
+  theme(legend.position = "none",
+        strip.background = element_blank(),
+        strip.text.x = element_blank(),
+        axis.text.x = element_blank(),
+        plot.margin=unit(c(0.5,0,0,0), "cm"))
 
-total_country_imports <- artis %>% 
-  filter(year %in% 2015:2019) %>%
-  group_by(importer_iso3c, method, environment) %>%
-  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
-  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
-  rename("iso3c" = "importer_iso3c")
-
-total_country_trade <- total_country_exports %>%
-  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
-  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
-  ungroup() %>%
-  group_by(iso3c) %>%
-  mutate(total = sum(trade_live_weight_t)) %>%
-  ungroup() %>%
-  slice_max(n = 9*20, order_by = total)
-
-total_net_import <- total_country_trade %>%
-  group_by(iso3c) %>%
-  summarise(export_live_weight_t = sum(export_live_weight_t),
-            import_live_weight_t = sum(import_live_weight_t)) %>%
-  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
-  select(iso3c, net_import)
-
-total_country_trade %>% 
-  left_join(total_net_import, by = "iso3c") %>%
-  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
-  ggplot() +
-  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
-  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
-  scale_fill_manual(values = method_colors) +
-  geom_line(aes(y = net_import, x = iso3c), group = 1) +
-  geom_hline(yintercept = 0) +
-  labs(x = "", y = "Trade (million t, live weight)") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
+fig2a
 ```
 
-## Regional bilateral flows
-```{r region_trend_matrix_by_method, fig.height=5, fig.width=5}
-artis_region_method <- artis %>%
-  filter(!is.na(importer_region), !is.na(exporter_region), 
-         importer_region != "Other nei", exporter_region != "Other nei") %>%
-  group_by(year, exporter_region, importer_region, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
-
-# Create basic plot for the legend
-# artis_region_method %>%
-#   group_by(year, exporter_region) %>%
-#   summarise(live_weight_t = sum(live_weight_t)) %>%
-#   ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
-#   geom_area() +
-#   scale_fill_manual(values = region_colors) +
-#   labs(fill = "Region") +
-#   theme(legend.position = "bottom")
-```
-
-```{r regional_bilateral_trends}
-artis_region <- artis_region_method %>%
-  group_by(year, exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))
-
-region_slopes <- artis_region %>%
-  split(.$exporter_importer) %>% 
-  map(~lm(live_weight_t~year, data = .x)) %>% 
-  map_df(broom::tidy, .id = 'exporter_importer') %>%
-  filter(term == 'year') %>%
-  select(exporter_importer, "slope" = "estimate")
-
-artis_region <- artis_region %>%
-  left_join(region_slopes, by = "exporter_importer") %>% 
-  mutate(intraregion = case_when(
-           exporter_region == importer_region ~ "intraregional trade",
-           exporter_region != importer_region ~ "interegional trade"))
-
-```
-
-# SI Fig X: Bilateral region trends lollypop plot
-
-```{r region_trend_lollipop, fig.height=4, fig.width=3}
-region_slopes_ranked <- artis_region %>%
-  select(-year, -live_weight_t) %>%
-  distinct() %>%
-  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
-  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
-  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
-  geom_point(size=3) +
-  scale_y_reordered() +
-  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
-  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-region_slopes_ranked
-
-```
-
-```{r, fig.height=3, fig.width=5}
-region_trend_matrix_method <- artis_region_method %>% 
+```{r fig2b}
+# Matrix of plots showcasing inter-regional trade trends
+fig2b <- artis_region_method %>% 
   ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
@@
 
-exports_by_region <- artis_region %>% 
+fig2b
+```
+
+```{r fig 2c}
+# Figure 2c
+# Exports by Region
+fig2c <- artis_region %>% 
   ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
@@
 
-imports_by_region <- artis_region %>% 
-  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
-  geom_area() +
-  labs(x = "", y = "", fill = "Exporter") +
-  scale_fill_manual(values = region_colors) +
-  facet_wrap(~importer_region, nrow = 1) +
-  theme_bw() +
-  theme(legend.position = "none",
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        axis.text.x = element_blank(),
-        plot.margin=unit(c(0.5,0,0,0), "cm"))
+fig2c
+```
 
-png("outputs/fig2a.png", width = 6.5, height = 5, res = 500, units = "in")
-ggarrange(ggarrange(imports_by_region, NULL,
+```{r fig2}
+# Overall Figure 2
+ggarrange(ggarrange(fig2a, NULL,
                     ncol = 2, nrow = 1,
                     widths = c(4.5, 1.25), labels = c("a", "")),
-          ggarrange(region_trend_matrix_method, 
-                    ggarrange(NULL, exports_by_region,
+          ggarrange(fig2b, 
+                    ggarrange(NULL, fig2c,
                               ncol = 1, heights = c(0.3, 4)), 
@@
           heights = c(1, 4.5))
-dev.off()
+
+ggsave(file.path(outdir, "fig2.png"))
 ```
 
-# Figure 3: Supply patterns
+# Figure 3
+```{r fig3a}
+# Global per capita supply
+fig3a_data <- read.csv(file.path(datadir, "fig3a_data.csv"))
 
-```{r calculate_supply}
-supply <- calculate_supply(artis %>% 
-                             # Remove FM trade from supply calculations
-                             #filter(hs6 != "230120") %>%
-                             rename(habitat = environment), 
-                           prod %>% 
-                             rename(live_weight_t = production_t)) %>%
-  # Add habitat-method column
-  mutate(habitat_method = paste(habitat, method, sep =" ")) %>% 
-  mutate(habitat_method = case_when(
-    str_detect(habitat_method, "unknown") ~ "unknown", 
-    TRUE ~ habitat_method
-  )) %>% 
-  # Set factor levels
-  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
-                                                            "marine aquaculture", "inland aquaculture",
-                                                            "unknown" ))) %>% 
-  left_join(country_metadata %>% 
-              select(iso3c, "region" = "owid_region"), by = "iso3c") %>% 
-      mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
-                                     TRUE ~ supply_no_error))
+fig3a <- fig3a_data %>%
+  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
+  geom_line() + 
+  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
+  scale_color_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
 
+fig3a
 ```
 
-# Fig 3a: Foreign dependence trends
+```{r fig3b}
+# Per capita supply by region
+fig3b_data <- read.csv(file.path(datadir, "fig3b_data.csv"))
 
-```{r supply_source_by_region}
-region_supply_trends <- supply %>% 
-  filter(region != "Other nei") %>%
-  group_by(region, year) %>%
-  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
-            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
-  pivot_longer(cols = supply_domestic:supply_foreign, 
-               names_to = "supply_source", values_to = "supply_percent") %>%
-  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
+fig3b <- fig3b_data %>%
+  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
+  geom_area() + 
+  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
+  scale_fill_manual(values = habitat_source_colors) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  facet_wrap(~region)
+
+fig3b
+```
+
+```{r fig3c}
+# Global percent of supply by domestic / foreign exports
+fig3c_data <- read.csv(file.path(datadir, "fig3c_data.csv"))
+
+fig3c <- fig3c_data %>%
+  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
+  geom_line() +
+  scale_color_manual(values = c("#F37542", "#0F2D59")) +
+  labs(x = "", y = "Percent of supply", color = "Origin") +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+
+fig3c
+```
+
+```{r fig3d}
+fig3d_data <- read.csv(file.path(datadir, "fig3d_data.csv"))
+
+fig3d <- fig3d_data %>%
   ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
@@
 
-global_supply_trends <- supply %>% 
-  filter(region != "Other nei") %>%
-  group_by( year) %>%
-  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
-            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
-  pivot_longer(cols = supply_domestic:supply_foreign, 
-               names_to = "supply_source", values_to = "supply_percent") %>%
-  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
-  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
-  geom_line() +
-  scale_color_manual(values = c("#F37542", "#0F2D59")) +
-  labs(x = "", y = "Percent of supply", color = "Origin") +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-supply %>% 
-  filter(region != "Other nei") %>%
-  filter(year == 2019) %>%
-  group_by(iso3c) %>%
-  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
-            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
-  arrange(desc(supply_foreign))
-
+fig3d
 ```
 
-```{r}
-# Calculate per capita supply by region and source
-supply_total <- supply %>% 
-  group_by(iso3c, region, year, habitat_method) %>%
-  summarise(supply = sum(supply),
-            supply_no_error = sum(supply_no_error),
-            supply_domestic = sum(supply_domestic),
-            supply_foreign = sum(supply_foreign),
-            import = sum(imports_live_weight_t), 
-            domestic_export = sum(domestic_export),
-            foreign_export = sum(foreign_export),
-            error_export = sum(error_export), 
-            production_t = sum(production_t)
-            ) %>%
-  left_join(pop %>% select(-region), by = c("iso3c", "year"))
+```{r fig3}
+fig3 <- ggarrange(
+  ggarrange(fig3a, fig3b, 
+            common.legend = TRUE, nrow = 1, widths = c(1, 2),
+            labels = c("a", "b")),
+  ggarrange(fig3c, fig3d,
+            common.legend = TRUE, nrow = 1, widths = c(1, 2),
+            labels = c("c", "d")),
+  nrow = 2
+)
 
-# Global per capita supply
-global_percap_supply_area <- supply_total %>%
-  filter(region != "Other nei") %>%
-  group_by(year, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
-  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
-  geom_area() + 
-  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
-  scale_fill_manual(values = habitat_source_colors) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+fig3
 
-global_percap_supply_line <- supply_total %>%
-  filter(region != "Other nei") %>%
-  group_by(year, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
-  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
+ggsave(file.path(outdir, "fig3.png"))
+```
+
+```{r fig4a}
+fig4a_data <- read.csv(file.path(datadir, "figa4a_data.csv"))
+
+fig4a <- fig4a_data %>%
+  ggplot(aes(x = year, y = shannon, color = source)) +
   geom_line() +
-  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
-  scale_color_manual(values = habitat_source_colors) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
-
-# Regional per capita supply
-regional_percap_supply_area <- supply_total %>%
-  filter(region != "Other nei") %>%
-  group_by(year, region, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
-  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
-  geom_area() + 
-  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
-  scale_fill_manual(values = habitat_source_colors) +
-  theme_bw() +
+  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
+  labs(x = "", y = "Shannon index", color = "") +
+  theme_minimal() +
   theme(legend.position = "bottom",
@@
 
-regional_percap_supply_line <- supply_total %>%
-  filter(region != "Other nei") %>%
-  group_by(year, region, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
-  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
-  geom_line() + 
-  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
-  scale_color_manual(values = habitat_source_colors) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~region)
-
+fig4a
 ```
 
-Where does trade have a positive versus negative effect on per capita supply for capture versus aquaculture?
+```{r fig4b}
+fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))
 
-```{r}
-# Plot relationship between per capita supply and imports by prod source and region
-country_supply <- supply %>% 
-  filter(year == 2018, region != "Other nei") %>% 
-  group_by(iso3c, region, year) %>% 
-  summarise(supply = sum(supply),
-            supply_no_error = sum(supply_no_error),
-            supply_domestic = sum(supply_domestic),
-            supply_foreign = sum(supply_foreign),
-            import = sum(imports_live_weight_t), 
-            domestic_export = sum(domestic_export),
-            foreign_export = sum(foreign_export),
-            error_export = sum(error_export), 
-            production_t = sum(production_t)
-  ) %>% 
-  left_join(pop %>% select(-region), by = c("iso3c", "year")) %>%
-  mutate(supply_per_cap = supply_no_error/pop) 
-```
-
-```{r, fig.height = 5, fig.width = 3}
-country_supply %>% 
-  ungroup() %>%
-  mutate(iso3c = fct_reorder(iso3c, supply_per_cap)) %>%
-  filter(supply_per_cap > 50) %>%
-  ggplot(aes(x = supply_per_cap, y = iso3c, fill = region)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = region_colors) +
-  theme_bw() +
-  theme(legend.position = "none")
-
-```
-
-```{r}
-import_supply <- country_supply %>%
-  # FIX IT: Temporarily remove very large (unrealistic) supply values
-  filter(supply_per_cap < 100) %>%
-  ggplot(aes(x = import/1000, y = supply_per_cap, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_continuous(trans='log10') +
-  labs(x = "Import (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
-  scale_color_manual(values = region_colors) +
-  theme_bw()
-
-export_supply <- country_supply %>%
-  # FIX IT: Temporarily remove very large (unrealistic) supply values
-  filter(supply_per_cap < 100) %>%
-  ggplot(aes(x = (domestic_export + foreign_export + error_export)/1000, y = supply_per_cap, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_continuous(trans='log10') +
-  labs(x = "Export (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
-  scale_color_manual(values = region_colors) +
-  theme_bw()
-```
-
-```{r, fig.height = 4, fig.width = 5}
-# Join figures
-png("outputs/fig3.png")
-ggarrange(
-  ggarrange(global_percap_supply_line, regional_percap_supply_area, 
-            common.legend = TRUE, nrow = 1, widths = c(1, 2),
-            labels = c("a", "b")),
-  ggarrange(global_supply_trends, region_supply_trends,
-            common.legend = TRUE, nrow = 1, widths = c(1, 2),
-            labels = c("c", "d")),
-  ggarrange(import_supply, export_supply, 
-            common.legend = TRUE, nrow = 1, widths = c(1, 1),
-            labels = c("e", "f")),
-  nrow = 3
-)
-dev.off()
-```
-
-## Fig 4: Effect of trade on supply diversity
-
-```{r supply_diversity}
-# Calculate diversity indices
-diversity <- supply %>%
-  filter(supply_domestic > 0, supply_foreign > 0) %>%
-  group_by(iso3c, region, year, sciname) %>%
-  summarise(supply_no_error = sum(supply_no_error), 
-            supply_domestic = sum(supply_domestic),
-            supply_foreign = sum(supply_foreign),
-            production_t = sum(production_t), 
-            imports_live_weight_t = sum(imports_live_weight_t)) %>%
-  ungroup(sciname) %>%
-  mutate(shannon_supply_no_error = diversity(supply_no_error),
-         shannon_supply_domestic = diversity(supply_domestic),
-         shannon_supply_foreign = diversity(supply_foreign),
-         shannon_prod = diversity(production_t),
-         shannon_imports = diversity(imports_live_weight_t)) %>%
-  select(iso3c, year, shannon_supply_no_error,
-         shannon_supply_domestic, shannon_supply_foreign,
-         shannon_prod, shannon_imports) %>%
-  distinct()
-
-total_trade <- supply %>%
-  filter(supply_domestic > 0, supply_foreign > 0) %>%
-  group_by(region, iso3c, year) %>%
-  summarise(supply_domestic = sum(supply_domestic),
-            supply_foreign = sum(supply_foreign),
-            production_t = sum(production_t), 
-            imports_live_weight_t = sum(imports_live_weight_t),
-            domestic_export = sum(domestic_export),
-            foreign_export = sum(foreign_export),
-            error_export = sum(error_export))
-
-diversity <- diversity %>%
-  left_join(total_trade, by = c("region", "iso3c", "year"))
-
-```
-
-# Fig 4a: diversity of foreign and domestic supply
-
-```{r}
-# Compare diversity of foreign and domestic supply
-foreign_vs_domestic_diversity <- diversity %>%
-  filter(year == 2018, region != "Other nei") %>%
+fig4b <- fig4b_data %>%
   ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
@@
 
-# Slope less than 1 means domestic diversity is higher than foreign-sourced diversity
-summary(lm(diversity$shannon_supply_foreign~diversity$shannon_supply_domestic))
+fig4b
 ```
 
-# Fig 4b: Production and import diversity
-```{r}
-# Compare production and import diversity trends
-prod_import_diversity_trends <- diversity %>%
-  filter(region != "Other nei") %>%
-  group_by(year, region) %>%
-  summarise(shannon_prod = mean(shannon_prod), 
-            shannon_imports = mean(shannon_imports)) %>%
-  pivot_longer(cols = shannon_prod:shannon_imports,
-               names_to = "source", values_to = "shannon") %>%
-  mutate(source = case_when(
-    source == "shannon_prod" ~ "Production diversity",
-    source == "shannon_imports" ~ "Import diversity"
-  )) %>%
-  ggplot(aes(x = year, y = shannon, color = source)) +
-  geom_line() +
-  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
-  labs(x = "", y = "Shannon index", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~region)
+```{r fig4c}
+fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
 
-# Compare total imports and supply diversity
-imports_vs_supply_diversity <- diversity %>%
-  filter(year == 2018, region != "Other nei") %>%
+fig4c <- fig4c_data %>%
   ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
@@
 
-diversity %>%
-  filter(region != "Other nei") %>%
-  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
-  geom_point(alpha = 0.5) +
-  scale_x_log10() +
-  scale_color_manual(values = region_colors) +
-    labs(x = "Total imports (million t, live weight)",
-       y = "Supply Shannon diversity",
-       color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom") +
-  facet_wrap(~region)
-
-diversity %>%
-  mutate(dom_vs_foreign = case_when(
-    shannon_supply_domestic >= shannon_supply_foreign ~ "dom",
-    shannon_supply_domestic < shannon_supply_foreign ~ "foreign"
-  )) %>% 
-  group_by(dom_vs_foreign) %>%
-  tally()
-
+fig4c
 ```
-
-```{r, fig.height = 3, fig.width = 3}
-png("outputs/fig4.png")
-ggarrange(prod_import_diversity_trends, 
-          ggarrange(foreign_vs_domestic_diversity,
-                    imports_vs_supply_diversity,
+```{r fig4, fig.height = 3, fig.width = 3}
+fig4 <- ggarrange(fig4a, 
+          ggarrange(fig4b,
+                    fig4c,
                     nrow = 1,
@@
           )
-dev.off()
+
+fig4
 ```
+
+# Supplementary Figures
+
+```{r si_top_exporters, fig.height = 4, fig.width = 4}
+
+top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"), row.names = FALSE)
+top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"), row.names = FALSE)
+
+exporters_old_plot <- top_exporters_old %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+exporters_recent_plot <- top_exporters_recent %>%
+  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(
+  exporters_old_plot,
+  exporters_recent_plot,
+  ncol = 2,
+  labels = c("Old", "Recent")
+)
+
+ggsave(file.path(outdir, "si_top_exporters.png"))
+```
+
+```{r si_top_importers, fig.height = 4, fig.width = 4}
+
+top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"), row.names = FALSE)
+top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"), row.names = FALSE)
+
+importers_old_plot <- top_importers_old %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+importers_recent_plot <- top_importers_recent %>%
+  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(
+  importers_old_plot,
+  importers_recent_plot,
+  ncol = 2,
+  labels = c("Old", "Recent")
+)
+
+ggsave(file.path(outdir, "si_top_importers.png"))
+```
+
+```{r si_regional_changes, fig.height = 4, fig.width = 4}
+
+# Largest regional export changes
+top_regional_export_change <- read.csv(file.path(datadir, "top_regional_export_change.csv"))
+
+regional_export_change_plot <- top_regional_export_change %>%
+  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+# Largest regional import changes
+top_regional_import_change <- read.csv(file.path(datadir, "top_regional_import_change.csv"))
+
+regional_import_change_plot <- top_regional_import_change %>%
+  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(
+  regional_export_change_plot,
+  regional_import_change_plot,
+  ncol = 2,
+  labels = c("Export", "Import")
+)
+
+ggsave(file.path(outdir, "regional_changes.png"))
+```
+
+```{r si_export_decreases_increases, fig.height = 4, fig.width = 4}
+top_exporter_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
+top_exporter_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))
+
+exporter_increases_plot <- top_export_increases %>%
+  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+exporter_decreases_plot <- top_export_decreases %>%
+  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(
+  exporter_decreases_plot, exporter_increases_plot, nrow = 1
+)
+
+ggsave(file.path(outdir, "si_export_changes.png"))
+```
+
+```{r si_import_decreases_increases, fig.height = 4, fig.width = 4}
+top_importer_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
+top_importer_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))
+
+importer_increases_plot <- top_import_increases %>%
+  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+importer_decreases_plot <- top_import_decreases %>%
+  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  geom_point(size=4) +
+  scale_color_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  labs(y = "", x = "Live weight (million t)") +
+  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
+ggarrange(
+  importer_decreases_plot, importer_increases_plot, nrow = 1
+)
+
+ggsave(file.path(outdir, "si_import_changes.png"))
+```
+
+```{r regional_avg_annual_change, fig.height = 4, fig.width = 4}
+regional_avg_annual_change <- read.csv(file.path(datadir, "regional_avg_annual_change.csv"))
+
+regional_slopes_ranked <- regional_avg_annual_change %>%
+  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
+  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
+  geom_point(size=3) +
+  scale_y_reordered() +
+  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
+  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+regional_slopes_ranked
+```
+
+```{r custom_artis_ts}
+
+artis_ts <- read.csv(file.path(datadir, "artis_ts.csv"))
+custom_ts <- read.csv(file.path(datadir, "custom_ts.csv"))
+
+ggplot() +
+  geom_line(data = artis_ts, aes(x = year, y = live_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
+  geom_line(data = custom_ts, aes(x = year, y = live_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
+  scale_color_manual(
+    values = c("Custom ARTIS Time series" = "black", "HS96" = "#741A32", "HS02" = "#B34232", "HS07" = "#D38F35", "HS12" = "red", "HS17" = "#114F59")
+  ) +
+  labs(x = "Year", y = "Live Weight (tonnes)", color = "HS Version") +
+  theme_bw() +
+  theme(legend.position = "bottom")
+
+ggsave(file.path(outdir, "custom_artis_ts.png"))
+```

### 2023-01-17 prep_data.R

@@ -1,0 +1,677 @@
+
+library(tidyverse)
+library(countrycode)
+library(circlize)
+library(here)
+library(ggpubr)
+library(kableExtra)
+library(DBI)
+library(janitor)
+library(viridis)
+library(broom)
+library(exploreARTIS)
+library(vegan)
+library(tidytext)
+
+#-------------------------------------------------------------------------------
+# Directory listing
+
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
+
+#-------------------------------------------------------------------------------
+# Initial database pulls
+
+# Database connection
+con <- dbConnect(RPostgres::Postgres(),
+                 dbname=Sys.getenv("POSTGRES_DB"),
+                 host="localhost",
+                 port="5432",
+                 user=Sys.getenv("POSTGRES_USER"),
+                 password=Sys.getenv("POSTGRES_PASSWORD"))
+
+# Check that connection is established by checking which tables are present
+dbListTables(con)
+
+# ARTIS dataframe
+artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
+  select(-record_id)
+
+# Production dataframe
+prod <- dbGetQuery(con, "SELECT * FROM production") %>%
+  select(-record_id)
+
+# Country metadata
+country_metadata <- dbGetQuery(con, "SELECT * FROM countries") %>%
+  select(-record_id)
+
+# Code max resolved taxa
+code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
+  select(-record_id) %>%
+  mutate(hs6 = as.character(hs6)) %>%
+  mutate(hs6 = case_when(
+    str_length(hs6) == 5 ~ paste("0", hs6, sep = ""),
+    TRUE ~ hs6
+  ))
+
+# Sciname metadata
+sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
+  select(-record_id)
+
+# Close database connection
+dbDisconnect(con)
+rm(list = c("con"))
+
+# Population data
+pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
+  left_join(country_metadata, by = c("iso3c")) %>%
+  select(iso3c, year, "region" = "owid_region", pop)
+
+# write.csv(artis, file.path(outdir, "artis.csv"), row.names = FALSE)
[…]

### 2023-01-20 ms_figures.Rmd
new pipeline for preparing manuscript ms results and ms figures

@@
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/outputs_20220111"
-
 ```
@@
 
+```{r ldc_net_exports_habitat_method}
+ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv"))
+
+ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
+  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(x = "", y = "Net export (million t)", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+ldc_net_exports_by_source_plot
+```
+
+```{r trade_concentration_n_countries}
+# Concentration of trade by production source, measured as the number of countries comprising 75%
+
+export_n_countries_concentration <- read.csv(file.path(datadir, "export_concentration_n_countries_by_source.csv"))
+import_n_countries_concentration <- read.csv(file.path(datadir, "import_concentration_n_countries_by_source.csv"))
+
+export_n_countries_by_source_plot <- export_n_countries_concentration %>%
+  ggplot(aes(x = year, y = n, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  geom_line(data = export_concentration_n_countries, aes(x = year, y = n), color = "black") +
+  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+export_n_countries_by_source_plot
+
+import_n_countries_by_source_plot <- import_n_countries_concentration %>%
+  ggplot(aes(x = year, y = n, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  geom_line(data = import_concentration_n_countries, aes(x = year, y = n), color = "black") +
+  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+import_n_countries_by_source_plot
+```
+
 ```{r custom_artis_ts}

### 2023-01-20 ms_results.Rmd
new pipeline for preparing manuscript ms results and ms figures

--- 2023-01-06 ms_results.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-
 # Load packages
@@
 library(tidytext)
+```
 
-# Set color palettes 
-habitat_source_colors <- c("#1B708F", # blue = marine capture
-                          "#86ADA7", # light blue = freshwater capture
-                          "#F37542", # medium orange = marine aquaculture
-                          "#F7AF75", # light orange = inland aquaculture
-                          "grey50" # unknown
-                          )
+```{r load_data}
+# Directory
+datadir <- "outputs"
+
+global_export_by_source <- read.csv(file.path(datadir, "global_export_by_source.csv"))
+all_degree_summary <- read.csv(file.path(datadir, "all_degree_summary.csv"))
+habitat_method_degree_summary <- read.csv(file.path(datadir, "habitat_method_degree_summary.csv"))
 ```
@@
 ```{r trends_summary}
-# Generate summary statistics for total trade trends
-
-# Load data summary files generated by ms_results
-global_export_by_source <- read.csv("outputs/global_export_by_source.csv")
-all_degree_summary <- read.csv("outputs/all_degree_summary.csv")
-habitat_method_degree_summary <- read.csv("outputs/habitat_method_degree_summary.csv")
-
 # Generate stats for text
@@
 
+# Total trade by environment and production method
+# Total trade for marine capture
 total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
 total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
-
 total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
 total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
+# Total trade for marine aquaculture
+total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
+total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
+# Total trade for inland aquaculture
+total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
+total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
 
+# Total trade for all aquaculture
 total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
@@
 
-total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
-total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
-
-total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
-total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
-
+# Total trade for FMFO
 total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
 total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
-
 ```
@@
 ```{r percent_export_by_source_trends}
-percent_export_by_source <- read.csv("outputs/percent_export_by_source_summary.csv")
-percent_export_total <- read.csv("outputs/percent_export_total_summary.csv")
-
+percent_export_by_source <- read.csv(file.path(datadir, "domestic_export_percent_production_by_habitat_method.csv"))
+percent_export_total <- read.csv(file.path(datadir, "total_export_percent_prod.csv"))
 percent_export_total_dom_1996 <- percent_export_total %>% 
@@
   pull(percent_export)
-
 percent_export_total_1996 <- percent_export_total %>% 
@@
   pull(percent_export)
-
 percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
@@
   pull(percent_export)
-
 percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
@@
   pull(percent_export)
-
 percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
@@
   pull(percent_export)
-
 percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
@@
   pull(percent_export)
-
 ```
@@
 ```{r trends_concentration}
-# Calculate import/export concentration stats
+# Export / Imports concentration
+cumulative_percent_threshold <- 75
+export_concentration_n_countries <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))
 
-# Load data as bilateral flows by habitat and method
-bilateral_habitat_method_summary <- read.csv("outputs/bilateral_habitat_method_summary.csv") %>% 
-  # Set factor levels
-  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
-                                                            "marine aquaculture", "inland aquaculture",
-                                                            "unknown" )))
-# Set percent threshold
-cumulative_percent_threshold <- 75
-# Concentration in exports
-## All
-export_concentration <- bilateral_habitat_method_summary %>%
-  group_by(year, exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(year, desc(live_weight_t)) %>%
-  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))
-
-export_concentration_n_countries <- export_concentration %>%
-  filter(cumulative_percent < cumulative_percent_threshold) %>%
-  group_by(year) %>% 
-  tally()
-
-## By habitat and method
-export_concentration_habitat_method <- bilateral_habitat_method_summary %>%
-  group_by(year, habitat_method, exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(year, habitat_method, desc(live_weight_t)) %>%
-  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))
-
-export_concentration_habitat_method_n_countries <- export_concentration_habitat_method %>%
-  filter(cumulative_percent < cumulative_percent_threshold) %>%
-  group_by(year, habitat_method) %>% 
-  tally()
-
-export_concentration_plot <- ggplot() +
-  geom_line(data = export_concentration_habitat_method_n_countries %>% 
-              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
-  scale_color_manual(values = habitat_source_colors) +
-  geom_line(data = export_concentration_n_countries, aes(x = year, y = n), color = "black") +
-  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
-  lims(y = c(0, 30)) +
-  theme_bw()
-
-# Concentration in imports
-## All
-import_concentration <- bilateral_habitat_method_summary %>%
-  group_by(year, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(year, desc(live_weight_t)) %>%
-  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))
-
-import_concentration_n_countries <- import_concentration %>%
-  filter(cumulative_percent < cumulative_percent_threshold) %>%
-  group_by(year) %>% 
-  tally()
-
-## By habitat and method
-import_concentration_habitat_method <- bilateral_habitat_method_summary %>%
-  group_by(year, habitat_method, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(year, habitat_method, desc(live_weight_t)) %>%
-  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))
-
-import_concentration_habitat_method_n_countries <- import_concentration_habitat_method %>%
-  filter(cumulative_percent < cumulative_percent_threshold) %>%
-  group_by(year, habitat_method) %>% 
-  tally()
-
-import_concentration_plot <- ggplot() +
-  geom_line(data = import_concentration_habitat_method_n_countries %>% 
-              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
-  scale_color_manual(values = habitat_source_colors) +
-  geom_line(data = import_concentration_n_countries, aes(x = year, y = n), color = "black") +
-  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
-  lims(y = c(0, 30)) +
-  theme_bw()
-
-ggarrange(export_concentration_plot, import_concentration_plot, common.legend = TRUE, legend = "bottom", labels = c("a", "b"))
+import_concentration_n_countries <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
 ```
@@
   pull(all_out_degree)
-
 all_degree_2019 <- all_degree_summary %>%
@@
   pull(all_out_degree)
-
 ```
@@
 
+![Figure 1: Increases in global export of blue foods.
+a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019.
+b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports.
+c) Out degree by production method and environment.](outputs/fig1.png)
+
 ## Other summary stats
@@
 
-```{r covid_effect}
-# artis %>% 
-#   filter(year %in% 2019:2020) %>% 
-#   group_by(year, habitat_method) %>% 
-#   summarise(live_weight_t = sum(live_weight_t)) %>% 
-#   pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
-#   mutate(change = (year2019-year2020)/year2020) %>% 
-#   arrange(desc(change))
+# Shifts in global flows of blue foods
+
+```{r}
+bilateral_flow_change <- read.csv(file.path(datadir, "bilateral_flow_change.csv"))
+
+percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)
 ```
 
-# Shifts in global flows of blue foods
+## Text for MS
 
-```{r top_traders}
-# Top trade tables
-top_n_countries <- 5
-top_n_countries_flows <- 10
-top_n_regions <- 5
-top_n_regions_flows <- 10
+The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
+In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
+Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
+At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
+Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
+Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
 
-# Top exporters
-## Countries
-top_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
+Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
+This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
+[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
 
-top_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
+Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America.
+The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia.
+At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
+The declining exports from South America mirror declining marine catch over this period.
+[Suggestions for discussion points and context related to this?].
 
-## Regions
-top_region_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
+Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
+Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
+High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
+2016).
 
-top_region_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
+![Figure 2: Regional trade flows by trade partner and production source.
+a) Total imports to each region colored by trade partner.
+b) Bilateral flows colored by production method and habitat source.
+c) Total exports from each region colored by trade partner.
+Values represent million tonnes in live weight equivalents.](outputs/fig2.png)
 
-# Top capture exporters
-## Countries
-top_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
+## Other summary stats
 
-top_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
+# National blue food supplies
+```{r}
+supply <- read.csv(file.path(datadir, "supply.csv"))
 
-## Regions
-top_region_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-top_region_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-# Top aquaculture exporters
-## Countries
-top_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-top_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-## Regions
-top_region_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-top_region_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-# Top importers
-## Countries
-top_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-top_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-## Regions
-top_region_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-top_region_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-# Top capture importers
-## Countries
-top_capture_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-top_capture_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-## Regions
-top_region_capture_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-top_region_capture_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-# Top aquaculture importers
-## Countries
-top_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-top_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries, order_by = live_weight_t)
-
-## Regions
-top_region_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-top_region_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions, order_by = live_weight_t)
-
-# Top bilateral flows
-## Countries
-top_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-top_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-## Regions
-top_region_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-top_region_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-# Top capture bilateral flows
-## Countries
-top_capture_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-top_capture_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-## Regions
-top_region_capture_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-top_region_capture_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-# Top aquaculture bilateral flows
-## Countries
-top_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-top_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_iso3c, importer_iso3c) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_countries_flows, order_by = live_weight_t)
-
-## Regions
-top_region_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
-  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-top_region_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
-  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  arrange(desc(live_weight_t)) %>%
-  slice_max(n = top_n_regions_flows, order_by = live_weight_t)
-
-# Create tables
-## Top exporter countries
-top_exporter_table <- data.frame(rank = 1:top_n_countries) %>%
-  left_join(top_exporters_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_total_1996 = exporter_iso3c), by = "rank") %>%
-  left_join(top_exporters_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_total_2019 = exporter_iso3c), by = "rank") %>%
-  left_join(top_capture_exporters_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_capture_1996 = exporter_iso3c), by = "rank") %>%
-  left_join(top_capture_exporters_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_capture_2019 = exporter_iso3c), by = "rank") %>%
-  left_join(top_aquaculture_exporters_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_aquaculture_1996 = exporter_iso3c), by = "rank") %>%
-  left_join(top_aquaculture_exporters_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_aquaculture_2019 = exporter_iso3c), by = "rank")
-
-kable(top_exporter_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_countries, "exporting countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-## Top exporter regions
-top_exporter_region_table <- data.frame(rank = 1:top_n_regions) %>%
-  left_join(top_region_exporters_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_total_1996 = exporter_region), by = "rank") %>%
-  left_join(top_region_exporters_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_total_2019 = exporter_region), by = "rank") %>%
-  left_join(top_region_capture_exporters_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_capture_1996 = exporter_region), by = "rank") %>%
-  left_join(top_region_capture_exporters_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_capture_2019 = exporter_region), by = "rank") %>%
-  left_join(top_region_aquaculture_exporters_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_aquaculture_1996 = exporter_region), by = "rank") %>%
-  left_join(top_region_aquaculture_exporters_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_aquaculture_2019 = exporter_region), by = "rank")
-
-kable(top_exporter_region_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_regions, "exporting regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-## Top importer countries
-top_importer_table <- data.frame(rank = 1:top_n_countries) %>%
-  left_join(top_importers_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_total_1996 = importer_iso3c), by = "rank") %>%
-  left_join(top_importers_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_total_2019 = importer_iso3c), by = "rank") %>%
-  left_join(top_capture_importers_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_capture_1996 = importer_iso3c), by = "rank") %>%
-  left_join(top_capture_importers_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_capture_2019 = importer_iso3c), by = "rank") %>%
-  left_join(top_aquaculture_importers_1996 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_aquaculture_1996 = importer_iso3c), by = "rank") %>%
-  left_join(top_aquaculture_importers_2019 %>% 
-              mutate(rank = 1:top_n_countries) %>%
-              select(rank, top_aquaculture_2019 = importer_iso3c), by = "rank")
-
-kable(top_importer_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_countries, "importing countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-## Top importer regions
-top_importer_region_table <- data.frame(rank = 1:top_n_regions) %>%
-  left_join(top_region_importers_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_total_1996 = importer_region), by = "rank") %>%
-  left_join(top_region_importers_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_total_2019 = importer_region), by = "rank") %>%
-  left_join(top_region_capture_importers_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_capture_1996 = importer_region), by = "rank") %>%
-  left_join(top_region_capture_importers_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_capture_2019 = importer_region), by = "rank") %>%
-  left_join(top_region_aquaculture_importers_1996 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_aquaculture_1996 = importer_region), by = "rank") %>%
-  left_join(top_region_aquaculture_importers_2019 %>% 
-              mutate(rank = 1:top_n_regions) %>%
-              select(rank, top_aquaculture_2019 = importer_region), by = "rank")
-
-kable(top_importer_region_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_regions, "importing regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-## Top country flows
-top_country_flow_table <- data.frame(rank = 1:top_n_countries_flows) %>%
-  left_join(top_flows_1996 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_total_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_total_1996), by = "rank") %>%
-  left_join(top_flows_2019 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_total_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_total_2019), by = "rank") %>%
-  left_join(top_capture_flows_1996 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_capture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_capture_1996), by = "rank") %>%
-  left_join(top_capture_flows_2019 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_capture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_capture_2019), by = "rank") %>%
-  left_join(top_aquaculture_flows_1996 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_aquaculture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_aquaculture_1996), by = "rank") %>%
-  left_join(top_aquaculture_flows_2019 %>% 
-              mutate(rank = 1:top_n_countries_flows,
-                     top_aquaculture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
-              select(rank, top_aquaculture_2019), by = "rank")
-
-kable(top_country_flow_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_countries_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-## Top region flows
-top_region_flow_table <- data.frame(rank = 1:top_n_regions_flows) %>%
-  left_join(top_region_flows_1996 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_total_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_total_1996), by = "rank") %>%
-  left_join(top_region_flows_2019 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_total_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_total_2019), by = "rank") %>%
-  left_join(top_region_capture_flows_1996 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_capture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_capture_1996), by = "rank") %>%
-  left_join(top_region_capture_flows_2019 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_capture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_capture_2019), by = "rank") %>%
-  left_join(top_region_aquaculture_flows_1996 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_aquaculture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_aquaculture_1996), by = "rank") %>%
-  left_join(top_region_aquaculture_flows_2019 %>% 
-              mutate(rank = 1:top_n_regions_flows,
-                     top_aquaculture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
-              select(rank, top_aquaculture_2019), by = "rank")
-
-kable(top_region_flow_table, 
-      col.names = c("Rank", "All 1996", "All 2019", 
-                    "Capture 1996", "Capture 2019", 
-                    "Aqua 1996", "Aqua 2019"),
-      caption = paste("Top", top_n_regions_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
-  kable_styling()
-
-```
-
-```{r top_bilateral_flows}
-
-bilateral_habitat_method_summary %>% 
-  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
-  filter(year %in% 2015:2019) %>%
-  group_by(exporter_region, importer_region) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  group_by(exporter_region) %>%
-  slice_max(n=2, order_by = live_weight_t)
-
-bilateral_habitat_method_summary %>% 
-  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
-  filter(year %in% 2015:2019) %>%
-  group_by(exporter_region, importer_region, habitat_method) %>%
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  ungroup() %>%
-  group_by(exporter_region, habitat_method) %>%
-  slice_max(n=1, order_by = live_weight_t)
-
-# Regionalization trends
-# Changes in trade flow volumes
-bilateral_habitat_method_summary %>% 
-  filter(exporter_region != "Other nei", importer_region != "Other nei",
-         year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_region, importer_region) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = end-beginning,
-         from_to = paste(exporter_region, " to ", importer_region), 
-         intraregion = case_when(
-           exporter_region == importer_region ~ "intraregional trade",
-           exporter_region != importer_region ~ "interegional trade"
-         )) %>% 
-  arrange(desc(change)) %>% 
-  ungroup() %>%
-  ggplot(aes(x = change/1000000, y = reorder(from_to, change), color = intraregion)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
-  geom_point(size=3) +
-  scale_y_reordered() +
-  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
-  labs(y = "", x = "Live weight (million t)", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-# Changes in trade flow percents
-bilateral_habitat_method_summary %>% 
-  filter(exporter_region != "Other nei", importer_region != "Other nei",
-         year %in% c(1996:2000, 2016:2020)) %>%
-  mutate(year_group = case_when(
-    year %in% 1996:2000 ~ "beginning", 
-    year %in% 2016:2020 ~ "end")) %>%
-  group_by(year_group, exporter_region, importer_region) %>% 
-  summarise(live_weight_t = sum(live_weight_t)) %>%
-  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
-  mutate(change = 100*(end-beginning)/beginning,
-         from_to = paste(exporter_region, " to ", importer_region), 
-         intraregion = case_when(
-           exporter_region == importer_region ~ "intraregional trade",
-           exporter_region != importer_region ~ "interegional trade"
-         )) %>% 
-  arrange(desc(change)) %>% 
-  ungroup() %>%
-  ggplot(aes(x = change, y = reorder(from_to, change), color = intraregion)) +
-  geom_segment(aes(x=0, xend=change, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
-  geom_point(size=3) +
-  scale_y_reordered() +
-  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
-  labs(y = "", x = "Percent change", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom")
-
-```
-
-```{r consumption}
-
-supply <- read.csv("outputs/supply.csv")
-
-global_supply <- supply %>%
-  group_by(year) %>%
-  summarize(supply_no_error = sum(supply_no_error) / 1000000) %>%
-  ungroup()
+global_supply <- read.csv(file.path(datadir, "global_supply.csv"))
 
@@
 
-regional_supply <- supply %>%
-  group_by(year, region) %>%
-  summarize(supply_no_error = sum(supply_no_error)) %>%
-  ungroup() %>%
-  group_by(year) %>%
-  mutate(supply_no_error_total = sum(supply_no_error)) %>%
-  ungroup() %>%
-  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)
-
-habitat_method_supply <- supply %>%
-  group_by(year, habitat, method) %>%
-  summarize(supply_no_error = sum(supply_no_error)) %>%
-  ungroup() %>%
-  group_by(year) %>%
-  mutate(supply_no_error_total = sum(supply_no_error)) %>%
-  ungroup() %>%
-  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)
+habitat_method_supply <- read.csv(file.path(datadir, "habitat_method_supply.csv"))
 
@@
 
-method_supplies <- supply %>%
-  filter(year == 2020) %>%
-  group_by(method) %>%
-  summarize(supply_no_error = sum(supply_no_error)) %>%
-  ungroup() %>%
-  mutate(total_supply = sum(supply_no_error)) %>%
-  mutate(percent_supply = 100 * supply_no_error / total_supply)
-
-regional_method_supplies <- supply %>%
-  filter(year == 2020) %>%
-  group_by(year, region, method) %>%
-  summarize(supply_no_error = sum(supply_no_error)) %>%
-  ungroup() %>%
-  group_by(year, region) %>%
-  mutate(region_supply = sum(supply_no_error)) %>%
-  ungroup() %>%
-  mutate(percent_supply = 100 * supply_no_error / region_supply) %>%
-  select(region, region_supply, method, percent_supply) %>%
-  pivot_wider(names_from = "method", values_from = "percent_supply")
+regional_method_supplies <- read.csv(file.path(datadir, "regional_method_supplies.csv"))
 
@@
 
+method_supplies <- read.csv(file.path(datadir, "method_supplies.csv"))
+
 cap_supply_percent <- method_supplies %>%
   filter(method == "capture") %>%
+  pull(percent_supply)
+
+consumption_foreign <- read.csv(file.path(datadir, "consumption_foreign.csv"))
+
+foreign_consumption_start <- consumption_foreign %>%
+  filter(year == 1996) %>%
+  pull(percent_supply)
+
+foreign_consumption_end <- consumption_foreign %>%
+  filter(year == 2020) %>%
   pull(percent_supply)
@@
 
-consumption_foreign <- supply %>%
-  filter(year == 1996 | year == 2020) %>%
-  group_by(year) %>%
-  summarize(supply_foreign = sum(supply_foreign),
-            supply_no_error = sum(supply_no_error)) %>%
-  ungroup() %>%
-  mutate(percent_supply = 100 * supply_foreign / supply_no_error)
-
-foreign_consumption_start <- consumption_foreign %>%
-  filter(year == 1996) %>%
-  pull(percent_supply)
-
-foreign_consumption_end <- consumption_foreign %>%
-  filter(year == 2020) %>%
-  pull(percent_supply)
-
 ```
-
-## Text for MS
-
-The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
-In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
-Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
-At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
-Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
-Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
-
-Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
-This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
-[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
-
-Since 1996, nearly all regional trade pairs increased, other than South America.
-The largest increases occurred for X, Y and Z.
-At the country level, X% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
-The declining exports from South America mirror declining marine catch over this period.
-[Suggestions for discussion points and context related to this?].
-
-Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
-Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
-High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
-2016).
-
-## Other summary stats
-
-# National blue food supplies
 
@@
 ## Other summary stats
+
+![Figure 3: Aquatic resource supply trends and regional patterns.
+a) Global blue food supply by production source over time.
+b) Regional blue food supply by production source over time.
+c) Global blue food domestic versus foreign sourcing over time.
+d) Regional blue food domestic versus foreign sourcing over time.
+Note these supply figures include non-food uses.](outputs/fig3.png)
+
+![Figure 4: Relationship between trade and supply diversity.
+a) Time series of the Shannon diversity of imported blue foods and produced blue foods by region.
+b) Foreign supply Shannon diversity versus domestic supply Shannon diversity for 2018.
+The black line represents the 1-to-1 line, such that points falling below the line indicate countries where the domestic supply diversity is greater than the foreign supply diversity.
+c) Relationship between total imports and overall supply diversity for 2018.](outputs/fig4.png)

### 2023-01-20 prep_data.R

@@ -17,6 +17,7 @@ library(tidytext)
 # Directory listing
 
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
+outdir <- "outputs"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -433,6 +434,21 @@ bilateral_habitat_method_summary <- artis %>%
 
 write.csv(bilateral_habitat_method_summary, file.path(outdir, "bilateral_habitat_method_summary.csv"), row.names = FALSE)
 
+# Calculate percent of pairs with increased trade flows
+bilateral_flow_change <- bilateral_habitat_method_summary %>% 
+  filter(exporter_region != "Other nei", importer_region != "Other nei",
+         year %in% c(1996:2000, 2016:2020)) %>%
+  mutate(year_group = case_when(
+    year %in% 1996:2000 ~ "beginning", 
+    year %in% 2016:2020 ~ "end")) %>%
+  group_by(year_group, exporter_iso3c, importer_iso3c) %>% 
+  summarise(live_weight_t = sum(live_weight_t)) %>%
+  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
+  mutate(change = end-beginning) %>%
+  arrange(desc(change))
+
+write.csv(bilateral_flow_change, file.path(outdir, "bilateral_flow_change.csv"), row.names = FALSE)
+
 # Top Exporters by habitat and production method
 # Year groups 1996 - 2000 and 2016 - 2020
 top_exporters_old <- bilateral_habitat_method_summary %>%
@@ -511,7 +527,7 @@ top_regional_export_change <- bilateral_habitat_method_summary %>%
   ungroup() %>%
   mutate(exporter_region = reorder_within(exporter_region, change, habitat_method))
 
-write.csv(top_regional_export_change, file.path(datadir, "top_regional_export_change.csv"), row.names = FALSE)
+write.csv(top_regional_export_change, file.path(outdir, "top_regional_export_change.csv"), row.names = FALSE)
 
 # Top Regional Import changes
 top_regional_import_change <- bilateral_habitat_method_summary %>% 
@@ -531,7 +547,7 @@ top_regional_import_change <- bilateral_habitat_method_summary %>%
   ungroup() %>%
   mutate(importer_region = reorder_within(importer_region, change, habitat_method))
 
-write.csv(top_regional_import_change, file.path(datadir, "top_regional_import_change.csv"), row.names = FALSE)
[…]

### 2023-01-20 Archive/ms_figures.Rmd
archived results and figures

---
title: "ms_figures"
author: "Jessica Gephart"
date: "2022-08-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(countrycode)
library(circlize)
library(here)
library(ggpubr)
library(kableExtra)
library(DBI)
library(janitor)
library(viridis)
library(broom)
library(exploreARTIS)
library(vegan)
library(tidytext)
```

```{r load_database_data}
# Load ARTIS data from SQL local database
con <- dbConnect(RPostgres::Postgres(),
                 dbname=Sys.getenv("POSTGRES_DB"),
                 host="localhost",
                 port="5432",
                 user=Sys.getenv("POSTGRES_USER"),
                 password=Sys.getenv("POSTGRES_PASSWORD"))

# Check that connection is established by checking which tables are present
dbListTables(con)

# Pull all ARTIS and production data
artis <- dbGetQuery(con, "SELECT * FROM snet")

artis <- artis %>%
  select(-record_id)

prod <- dbGetQuery(con, "SELECT * FROM production")
prod <- prod %>%
  select(-record_id)

country_metadata <- dbGetQuery(con, "SELECT * FROM countries")
country_metadata <- country_metadata %>%
  select(-record_id)

code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
              mutate(hs6 = as.character(hs6)) %>%
  select(-record_id)

sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
  select(-record_id)

# Close database connection
dbDisconnect(con)

rm(list = c("con"))
```

```{r set_directories}
# Set directories 
outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization"
outdir <- "outputs"

# Set color palettes
# All colors from palette: #FFE6BE (light yellow), #F7AF75 (light orange), #F37542 (medium orange), 
# #E24027 (dark orange), #0F2D59 (navy), #1B708F (blue), #2E8D9A (teal), #86ADA7 (light blue)
method_colors <- c("#F7AF75", "#2E8D9A", "#86ADA7") # light orange, teal, light blue
source_colors <- c("#F37542", "#0F2D59", "#86ADA7") # medium orange, navy, light blue
environment_colors <- c( "#2E8D9A", "#0F2D59","#86ADA7") # teal, navy, light blue

habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )

region_colors <- c(
  "#741A32", # maroon
  "#B34232", # burnt orange
  "#D38F35", #  orange
  "#D4B95f", # khaki
  "#4FA2A2", # teal
  "#114F59" # dark teal
)

```

```{r load_drive_data}

pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
  left_join(country_metadata, by = c("iso3c")) %>%
  select(iso3c, year, "region" = "owid_region", pop)

```

```{r join_data}
# Summarize cleaned FAO production data
prod <- prod %>% 
  select("year", "iso3c", "sciname", 
         "method", "habitat", 
         "production_t" = "live_weight_t") %>%
  group_by(year, iso3c, sciname, method, habitat) %>%
  summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
  ungroup()

# Join production data to ARTIS
artis <- artis %>%
  filter(live_weight_t >= 0.1) %>%
  # Add importer and exporter region
  left_join(country_metadata %>% 
              select("exporter_iso3c" = "iso3c", "exporter_region" = "owid_region"), by = c("exporter_iso3c")) %>%
  left_join(country_metadata %>% 
              select("importer_iso3c" = "iso3c", "importer_region" = "owid_region"), by = c("importer_iso3c")) %>%
# FIX IT: Determine if this should be deleted. Production of only species exports may not be relevant
  left_join(prod, by = c("year", "exporter_iso3c" = "iso3c", 
                           "sciname", "method", "habitat")) %>%
  ungroup()

# Update artis data to lowest resolution by comparing code taxa_level and prod taxa_level
artis <- artis %>%
  left_join(code_max_resolved_taxa,
            by = c("hs_version", "hs6", "sciname")) %>%
  left_join(sciname_metadata, 
            by = "sciname") %>%
  # Leave any missing sciname_hs_modified as original sciname
  mutate(sciname_hs_modified = case_when(
    is.na(sciname_hs_modified) ~ sciname,
    TRUE ~ sciname_hs_modified
  )) %>%
  rename(environment = habitat) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(environment, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))

# Remove data frames after merged with ARTIS
rm(list = c("code_max_resolved_taxa", "sciname_metadata"))

```

# Figure 1: Increases in global export of blue foods

## Figure 1 a: Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019

```{r fig1a}
global_export_by_source <- artis %>% 
  mutate(food_or_fmfo = case_when(
    hs6 == "230120" ~ "fishmeal",
    hs6 != "230120" ~ "nonfishmeal"
  )) %>%
  group_by(year, habitat_method, food_or_fmfo) %>%
  summarise(live_weight_t = sum(live_weight_t))

write.csv(global_export_by_source, "outputs/global_export_by_source.csv", row.names = FALSE)

fig1a <- global_export_by_source %>%
  group_by(year, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
  theme_bw() +
  theme(legend.position = "bottom")

```

## Figure 1 b: Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports

```{r fig1b}
prod_by_source <- prod %>%
  group_by(year, method, habitat) %>% 
  summarise(production_t = sum(production_t)) %>%
  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
    # Set factor levels
  mutate(habitat_method = factor(habitat_method, 
                                 levels = c("marine capture", "inland capture",
                                            "marine aquaculture", "inland aquaculture"))) %>%
  ungroup() %>%
  select(-habitat, -method)

dom_export_by_source <- artis %>% 
  filter(dom_source == "domestic export") %>%
  group_by(year, habitat_method, environment) %>%
  summarise(dom_live_weight_t = sum(live_weight_t))

percent_export_by_source <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  mutate(percent_export = 100*dom_live_weight_t/production_t)

write.csv(percent_export_by_source, "outputs/percent_export_by_source_summary.csv", row.names = FALSE)

percent_export_total <- dom_export_by_source %>%
  left_join(prod_by_source, by = c("year", "habitat_method")) %>% 
  group_by(year) %>% 
  summarise(dom_live_weight_t = sum(dom_live_weight_t), 
            production_t = sum(production_t)) %>%
  mutate(percent_export = 100*dom_live_weight_t/production_t) %>% 
  mutate(calc_type = "domestic exports") %>% 
  select(year, percent_export, calc_type)

percent_export_total_all_source <- artis %>% 
  group_by(year) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  left_join(prod_by_source %>% 
              group_by(year) %>% 
              summarise(production_t = sum(production_t)), 
            by = c("year")) %>% 
  mutate(percent_export = 100*live_weight_t/production_t) %>% 
  mutate(calc_type = "all export sources") %>% 
  select(year, percent_export, calc_type)

percent_export_total <- percent_export_total %>%
  bind_rows(percent_export_total_all_source)

write.csv(percent_export_total, "outputs/percent_export_total_summary.csv", row.names = FALSE)

fig1b <- ggplot() +
  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "% production exported", x = "", color = "Source") +
  theme_bw()

```

## Figure 1 c: Degree or measure of degree distribution by production method and environment

```{r fig1c}
all_degree_summary <-  artis %>% 
  select(year, exporter_iso3c, importer_iso3c) %>%
  distinct() %>%
  group_by(year, exporter_iso3c) %>%
  tally() %>% 
  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
  full_join(artis %>% 
              select(year, exporter_iso3c, importer_iso3c) %>%
              distinct() %>%
              group_by(year, importer_iso3c) %>%
              tally() %>%
              rename(all_in_degree = n, iso3c = importer_iso3c), 
            by = c("year", "iso3c"))

write.csv(all_degree_summary, "outputs/all_degree_summary.csv", row.names = FALSE)

habitat_method_degree_summary <-  artis %>% 
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>% 
  rename(all_out_degree = n, iso3c = exporter_iso3c) %>%
  full_join(artis %>% 
              select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
              distinct() %>%
              group_by(year, importer_iso3c, habitat_method) %>%
              tally() %>%
              rename(all_in_degree = n, iso3c = importer_iso3c), 
            by = c("year", "iso3c", "habitat_method"))

write.csv(habitat_method_degree_summary, "outputs/habitat_method_degree_summary.csv", row.names = FALSE)

fig1c_out_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, exporter_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "Out degree", fill = "Source") + 
  theme_bw()

fig1c_in_degree <- artis %>% 
  filter(habitat_method != "unknown") %>%
  select(year, exporter_iso3c, importer_iso3c, habitat_method) %>%
  distinct() %>%
  group_by(year, importer_iso3c, habitat_method) %>%
  tally() %>%
  group_by(year, habitat_method) %>%
  ggplot(aes(x = year, y = n, color = habitat_method)) +
  geom_smooth() +
  scale_color_manual(values = habitat_source_colors) +
  labs(y = "In degree", fill = "Source") + 
  theme_bw()

```

```{r fig1, fig.height = 3, fig.width = 4.5}
png("outputs/fig1.png")
ggarrange(fig1a,
          ggarrange(fig1b,
                    fig1c_out_degree,
                    ncol = 2, nrow = 1,
                    legend = "none",
                    labels = c("b", "c")),
          common.legend = TRUE, legend = "bottom",
          ncol = 1, nrow = 2, labels = c("a", ""))

dev.off()
```

# Figure 2: Trade patterns

```{r SI_top_capture_aqua_imp_exp, fig.height=4, fig.width=4}
# Summarize total imports and exports by habitat and environment
bilateral_habitat_method_summary <- artis %>%
  # Remove if we decide to leave FMFO in
  filter(hs6 != "230120") %>%
  group_by(year, exporter_iso3c, exporter_region, importer_iso3c, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t))

write.csv(bilateral_habitat_method_summary, "outputs/bilateral_habitat_method_summary.csv", row.names = FALSE)

top_traders_recent <- bilateral_habitat_method_summary %>% 
  filter(year %in% 2016:2020, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(2016:2020))

top_importers_recent <- top_traders_recent %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_recent <- top_traders_recent %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_recent, top_importers_recent, ncol = 2, labels = c("Export", "Import"))

top_traders_past <- bilateral_habitat_method_summary %>% 
  filter(year %in% 1996:2000, habitat_method != "unknown") %>%
  group_by(importer_iso3c, exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)/length(1996:2000))

top_importers_past <- top_traders_past %>%
  group_by(importer_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

top_exporters_past <- top_traders_past %>%
  group_by(exporter_iso3c, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = live_weight_t) %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  mutate(ranking = rank(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method)) %>%
  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_exporters_past, top_importers_past, ncol = 2, labels = c("Export", "Import"))

# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_export_increases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_export_decreases <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_iso3c = reorder_within(exporter_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_export_decreases, top_export_increases, nrow = 1)

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_import_increases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest decreases
top_import_decreases <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_min(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_iso3c = reorder_within(importer_iso3c, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_import_decreases, top_import_increases, nrow = 1)

# Repeat for regional trends
# Largest export increases
top_exporter_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_export <- top_exporter_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(exporter_region = reorder_within(exporter_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

# Largest import increases
top_importer_change <- bilateral_habitat_method_summary %>% 
  filter(year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, importer_region, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t) %>% 
  mutate(change = end-beginning) %>% 
  arrange(desc(change))

# Largest increases
top_region_import <-top_importer_change %>% 
  filter(habitat_method != "unknown") %>%
  ungroup() %>%
  group_by(habitat_method) %>%
  slice_max(n = 10, order_by = change) %>%
  ungroup() %>%
  mutate(importer_region = reorder_within(importer_region, change, habitat_method)) %>%
  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
  geom_point(size=4) +
  scale_color_manual(values = habitat_source_colors) +
  scale_y_reordered() +
  labs(y = "", x = "Live weight (million t)") +
  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none")

ggarrange(top_region_export, top_region_import, nrow = 1, labels = c("Export", "Import"))

```

```{r SI_top_traders}
total_country_exports <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(exporter_iso3c, method, environment) %>%
  summarise(export_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(export_live_weight_t = export_live_weight_t/1000000) %>%
  rename("iso3c" = "exporter_iso3c")

total_country_imports <- artis %>% 
  filter(year %in% 2015:2019) %>%
  group_by(importer_iso3c, method, environment) %>%
  summarise(import_live_weight_t = sum(live_weight_t)/length(2015:2019)) %>%
  mutate(import_live_weight_t = import_live_weight_t/1000000) %>%
  rename("iso3c" = "importer_iso3c")

total_country_trade <- total_country_exports %>%
  full_join(total_country_imports, by = c("iso3c", "method", "environment")) %>%
  mutate(trade_live_weight_t = export_live_weight_t + import_live_weight_t) %>%
  ungroup() %>%
  group_by(iso3c) %>%
  mutate(total = sum(trade_live_weight_t)) %>%
  ungroup() %>%
  slice_max(n = 9*20, order_by = total)

total_net_import <- total_country_trade %>%
  group_by(iso3c) %>%
  summarise(export_live_weight_t = sum(export_live_weight_t),
            import_live_weight_t = sum(import_live_weight_t)) %>%
  mutate(net_import = import_live_weight_t - export_live_weight_t,) %>%
  select(iso3c, net_import)

total_country_trade %>% 
  left_join(total_net_import, by = "iso3c") %>%
  mutate(iso3c = fct_reorder(as.factor(iso3c), -net_import)) %>%
  ggplot() +
  geom_bar(aes(y = -1*export_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  geom_bar(aes(y = import_live_weight_t, x = iso3c, fill = method), stat = "identity", position = "stack") +
  scale_fill_manual(values = method_colors) +
  geom_line(aes(y = net_import, x = iso3c), group = 1) +
  geom_hline(yintercept = 0) +
  labs(x = "", y = "Trade (million t, live weight)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

## Regional bilateral flows
```{r region_trend_matrix_by_method, fig.height=5, fig.width=5}
artis_region_method <- artis %>%
  filter(!is.na(importer_region), !is.na(exporter_region), 
         importer_region != "Other nei", exporter_region != "Other nei") %>%
  group_by(year, exporter_region, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

# Create basic plot for the legend

# artis_region_method %>%
#   group_by(year, exporter_region) %>%
#   summarise(live_weight_t = sum(live_weight_t)) %>%
#   ggplot(aes(x = year, y = live_weight_t, fill = exporter_region)) +
#   geom_area() +
#   scale_fill_manual(values = region_colors) +
#   labs(fill = "Region") +
#   theme(legend.position = "bottom")
```

```{r regional_bilateral_trends}
artis_region <- artis_region_method %>%
  group_by(year, exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  mutate(exporter_importer = paste(exporter_region, " to ", importer_region, sep = ""))

region_slopes <- artis_region %>%
  split(.$exporter_importer) %>% 
  map(~lm(live_weight_t~year, data = .x)) %>% 
  map_df(broom::tidy, .id = 'exporter_importer') %>%
  filter(term == 'year') %>%
  select(exporter_importer, "slope" = "estimate")

artis_region <- artis_region %>%
  left_join(region_slopes, by = "exporter_importer") %>% 
  mutate(intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"))

```

# SI Fig X: Bilateral region trends lollypop plot

```{r region_trend_lollipop, fig.height=4, fig.width=3}
region_slopes_ranked <- artis_region %>%
  select(-year, -live_weight_t) %>%
  distinct() %>%
  mutate(exporter_importer = fct_reorder(exporter_importer, slope)) %>%
  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Ave. annual change in export (1000 t live weight)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

region_slopes_ranked

```

```{r, fig.height=3, fig.width=5}
region_trend_matrix_method <- artis_region_method %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
  geom_area() +
  scale_fill_manual(values = habitat_source_colors) +
  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), 
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(0,0,0,0), "cm")) +
  facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
             labeller = label_wrap_gen(width = 5), scales = "free")

exports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Importer") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~exporter_region, ncol = 1) +
  theme_bw() +
    theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.margin=unit(c(0,0.5,0,0), "cm"))

imports_by_region <- artis_region %>% 
  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
  geom_area() +
  labs(x = "", y = "", fill = "Exporter") +
  scale_fill_manual(values = region_colors) +
  facet_wrap(~importer_region, nrow = 1) +
  theme_bw() +
  theme(legend.position = "none",
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.text.x = element_blank(),
        plot.margin=unit(c(0.5,0,0,0), "cm"))

png("outputs/fig2a.png", width = 6.5, height = 5, res = 500, units = "in")
ggarrange(ggarrange(imports_by_region, NULL,
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1.25), labels = c("a", "")),
          ggarrange(region_trend_matrix_method, 
                    ggarrange(NULL, exports_by_region,
                              ncol = 1, heights = c(0.3, 4)), 
                    ncol = 2, nrow = 1,
                    widths = c(4.5, 1), labels = c("b", "c")),
          ncol = 1, nrow = 2, 
          heights = c(1, 4.5))
dev.off()
```

# Figure 3: Supply patterns

```{r calculate_supply}
supply <- calculate_supply(artis %>% 
                             # Remove FM trade from supply calculations
                             #filter(hs6 != "230120") %>%
                             rename(habitat = environment), 
                           prod %>% 
                             rename(live_weight_t = production_t)) %>%
  # Add habitat-method column
  mutate(habitat_method = paste(habitat, method, sep =" ")) %>% 
  mutate(habitat_method = case_when(
    str_detect(habitat_method, "unknown") ~ "unknown", 
    TRUE ~ habitat_method
  )) %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" ))) %>% 
  left_join(country_metadata %>% 
              select(iso3c, "region" = "owid_region"), by = "iso3c") %>% 
      mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
                                     TRUE ~ supply_no_error))

write.csv(supply, file.path(outdir, "supply.csv"), row.names = FALSE)
```

# Fig 3a: Foreign dependence trends

```{r supply_source_by_region}
region_supply_trends <- supply %>% 
  filter(region != "Other nei") %>%
  group_by(region, year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  pivot_longer(cols = supply_domestic:supply_foreign, 
               names_to = "supply_source", values_to = "supply_percent") %>%
  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  facet_wrap(~region) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

global_supply_trends <- supply %>% 
  filter(region != "Other nei") %>%
  group_by( year) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  pivot_longer(cols = supply_domestic:supply_foreign, 
               names_to = "supply_source", values_to = "supply_percent") %>%
  mutate(supply_source = gsub("supply_", "", supply_source)) %>%
  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
  geom_line() +
  scale_color_manual(values = c("#F37542", "#0F2D59")) +
  labs(x = "", y = "Percent of supply", color = "Origin") +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

supply %>% 
  filter(region != "Other nei") %>%
  filter(year == 2019) %>%
  group_by(iso3c) %>%
  summarise(supply_domestic = 100*sum(supply_domestic)/sum(supply_no_error),
            supply_foreign = 100*sum(supply_foreign)/sum(supply_no_error)) %>%
  arrange(desc(supply_foreign))

```

```{r}
# Calculate per capita supply by region and source
supply_total <- supply %>% 
  group_by(iso3c, region, year, habitat_method) %>%
  summarise(supply = sum(supply),
            supply_no_error = sum(supply_no_error),
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            import = sum(imports_live_weight_t), 
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export), 
            production_t = sum(production_t)
            ) %>%
  left_join(pop %>% select(-region), by = c("iso3c", "year"))

# Global per capita supply
global_percap_supply_area <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

global_percap_supply_line <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, habitat_method) %>%
  summarise(supply_per_cap = 1000 * (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
  geom_line() + 
  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
  scale_color_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Regional per capita supply
regional_percap_supply_area <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, region, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
  geom_area() + 
  labs(y = "Per capita supply (kg)", x = "", fill = "Source") +
  scale_fill_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

regional_percap_supply_line <- supply_total %>%
  filter(region != "Other nei") %>%
  group_by(year, region, habitat_method) %>%
  summarise(supply_per_cap = (sum(supply_no_error))/sum(pop)) %>%
  ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
  geom_line() + 
  labs(y = "Per capita supply (kg)", x = "", color = "Source") +
  scale_color_manual(values = habitat_source_colors) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

```

Where does trade have a positive versus negative effect on per capita supply for capture versus aquaculture?

```{r}
# Plot relationship between per capita supply and imports by prod source and region
country_supply <- supply %>% 
  filter(year == 2018, region != "Other nei") %>% 
  group_by(iso3c, region, year) %>% 
  summarise(supply = sum(supply),
            supply_no_error = sum(supply_no_error),
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            import = sum(imports_live_weight_t), 
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export), 
            production_t = sum(production_t)
  ) %>% 
  left_join(pop %>% select(-region), by = c("iso3c", "year")) %>%
  mutate(supply_per_cap = supply_no_error/pop) %>%
  # Converting from tonnes to kg
  mutate(supply_per_cap = 1000 * supply_per_cap)
```

```{r, fig.height = 5, fig.width = 3}
country_supply %>% 
  ungroup() %>%
  mutate(iso3c = fct_reorder(iso3c, supply_per_cap)) %>%
  #filter(supply_per_cap > 50) %>%
  ggplot(aes(x = supply_per_cap, y = iso3c, fill = region)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = region_colors) +
  theme_bw() +
  theme(legend.position = "none")

```

```{r}
import_supply <- country_supply %>%
  # FIX IT: Temporarily remove very large (unrealistic) supply values
  filter(supply_per_cap < 100) %>%
  ggplot(aes(x = import/1000, y = supply_per_cap, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(trans='log10') +
  labs(x = "Import (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
  scale_color_manual(values = region_colors) +
  theme_bw()

export_supply <- country_supply %>%
  # FIX IT: Temporarily remove very large (unrealistic) supply values
  filter(supply_per_cap < 100) %>%
  ggplot(aes(x = (domestic_export + foreign_export + error_export)/1000, y = supply_per_cap, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(trans='log10') +
  labs(x = "Export (1000 t, live weight)", y = "Supply (per capita)", color = "Region") +
  scale_color_manual(values = region_colors) +
  theme_bw()
```

```{r, fig.height = 5, fig.width = 7}
# Join figures
png("outputs/fig3.png", width = 7, height = 5, units = "in", res = 500)
ggarrange(
  ggarrange(global_percap_supply_line, regional_percap_supply_area, 
            common.legend = TRUE, nrow = 1, widths = c(1, 2),
            labels = c("a", "b")),
  ggarrange(global_supply_trends, region_supply_trends,
            common.legend = TRUE, nrow = 1, widths = c(1, 2),
            labels = c("c", "d")),
  nrow = 2
)
dev.off()
```

## Fig 4: Effect of trade on supply diversity

```{r supply_diversity}
# Calculate diversity indices
diversity <- supply %>%
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(iso3c, region, year, sciname) %>%
  summarise(supply_no_error = sum(supply_no_error), 
            supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t)) %>%
  ungroup(sciname) %>%
  mutate(shannon_supply_no_error = diversity(supply_no_error),
         shannon_supply_domestic = diversity(supply_domestic),
         shannon_supply_foreign = diversity(supply_foreign),
         shannon_prod = diversity(production_t),
         shannon_imports = diversity(imports_live_weight_t)) %>%
  select(iso3c, year, shannon_supply_no_error,
         shannon_supply_domestic, shannon_supply_foreign,
         shannon_prod, shannon_imports) %>%
  distinct()

total_trade <- supply %>%
  filter(supply_domestic > 0, supply_foreign > 0) %>%
  group_by(region, iso3c, year) %>%
  summarise(supply_domestic = sum(supply_domestic),
            supply_foreign = sum(supply_foreign),
            production_t = sum(production_t), 
            imports_live_weight_t = sum(imports_live_weight_t),
            domestic_export = sum(domestic_export),
            foreign_export = sum(foreign_export),
            error_export = sum(error_export))

diversity <- diversity %>%
  left_join(total_trade, by = c("region", "iso3c", "year"))

```

# Fig 4a: diversity of foreign and domestic supply

```{r}
# Compare diversity of foreign and domestic supply
foreign_vs_domestic_diversity <- diversity %>%
  filter(year == 2018, region != "Other nei") %>%
  ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
    geom_point(alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0) +
  scale_color_manual(values = region_colors) +
  labs(x = "Domestic supply diversity", 
       y = "Foreign supply diversity",
       color = "") +
    theme_minimal()

# Slope less than 1 means domestic diversity is higher than foreign-sourced diversity
summary(lm(diversity$shannon_supply_foreign~diversity$shannon_supply_domestic))
```

# Fig 4b: Production and import diversity
```{r}
# Compare production and import diversity trends
prod_import_diversity_trends <- diversity %>%
  filter(region != "Other nei") %>%
  group_by(year, region) %>%
  summarise(shannon_prod = mean(shannon_prod), 
            shannon_imports = mean(shannon_imports)) %>%
  pivot_longer(cols = shannon_prod:shannon_imports,
               names_to = "source", values_to = "shannon") %>%
  mutate(source = case_when(
    source == "shannon_prod" ~ "Production diversity",
    source == "shannon_imports" ~ "Import diversity"
  )) %>%
  ggplot(aes(x = year, y = shannon, color = source)) +
  geom_line() +
  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
  labs(x = "", y = "Shannon index", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

# Compare total imports and supply diversity
imports_vs_supply_diversity <- diversity %>%
  filter(year == 2018, region != "Other nei") %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_log10() +
  scale_color_manual(values = region_colors) +
  theme_minimal() +
  labs(x = "Total imports (million t, live weight)",
       y = "Supply diversity",
       color = "")

diversity %>%
  filter(region != "Other nei") %>%
  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
  geom_point(alpha = 0.5) +
  scale_x_log10() +
  scale_color_manual(values = region_colors) +
    labs(x = "Total imports (million t, live weight)",
       y = "Supply Shannon diversity",
       color = "") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  facet_wrap(~region)

diversity %>%
  mutate(dom_vs_foreign = case_when(
    shannon_supply_domestic >= shannon_supply_foreign ~ "dom",
    shannon_supply_domestic < shannon_supply_foreign ~ "foreign"
  )) %>% 
  group_by(dom_vs_foreign) %>%
  tally()

```

```{r, fig.height = 3, fig.width = 3}
png("outputs/fig4.png")
ggarrange(prod_import_diversity_trends, 
          ggarrange(foreign_vs_domestic_diversity,
                    imports_vs_supply_diversity,
                    nrow = 1,
                    common.legend = TRUE, 
                    legend = "bottom",
                    labels = c("b", "c")),
          nrow = 2,
          labels = c("a", "", "")
          )
dev.off()
```

### 2023-01-20 Archive/ms_results.Rmd
archived results and figures

---
title: "ms_results"
author: "Jessica Gephart"
output: html_document
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(tidyverse)
library(countrycode)
library(kableExtra)
library(ggpubr)
library(tidytext)

# Set color palettes 
habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )
```

# Trends blue food globalization

```{r trends_summary}
# Generate summary statistics for total trade trends

# Load data summary files generated by ms_results
global_export_by_source <- read.csv("outputs/global_export_by_source.csv")
all_degree_summary <- read.csv("outputs/all_degree_summary.csv")
habitat_method_degree_summary <- read.csv("outputs/habitat_method_degree_summary.csv")

# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)

total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)

total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)

```

```{r percent_export_by_source_trends}
percent_export_by_source <- read.csv("outputs/percent_export_by_source_summary.csv")
percent_export_total <- read.csv("outputs/percent_export_total_summary.csv")

percent_export_total_dom_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_dom_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "domestic exports") %>% 
  pull(percent_export)

percent_export_total_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_export_total_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "all export sources") %>% 
  pull(percent_export)

percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine capture") %>%
  pull(percent_export)

percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_marine_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine aquaculture") %>%
  pull(percent_export)

percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland capture") %>%
  pull(percent_export)

percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
percent_inland_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland aquaculture") %>%
  pull(percent_export)

```

```{r trends_concentration}
# Calculate import/export concentration stats

# Load data as bilateral flows by habitat and method
bilateral_habitat_method_summary <- read.csv("outputs/bilateral_habitat_method_summary.csv") %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))
# Set percent threshold
cumulative_percent_threshold <- 75
# Concentration in exports
## All
export_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_n_countries <- export_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
export_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_habitat_method_n_countries <- export_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

export_concentration_plot <- ggplot() +
  geom_line(data = export_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = export_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

# Concentration in imports
## All
import_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_n_countries <- import_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
import_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_habitat_method_n_countries <- import_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

import_concentration_plot <- ggplot() +
  geom_line(data = import_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = import_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

ggarrange(export_concentration_plot, import_concentration_plot, common.legend = TRUE, legend = "bottom", labels = c("a", "b"))
```

```{r degree_summary_stats}
all_degree_1996 <- all_degree_summary %>%
  filter(year == 1996) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)

all_degree_2019 <- all_degree_summary %>%
  filter(year == 2019) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)

```

## Text for MS

Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships. Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`) since 1996. Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%. However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports. Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively. From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period. Consequently, the share of global blue food exports represented by fishmeal declined from `r abs(round(100*(total_fmfo_1996)/(total_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019), 1))`% in 2019.

Another measure of the degree of globalization is the share of production exported. Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports (i.e., domestic plus exports of foreign and unknown origin) reached `r round(percent_export_total_2019, 1)`% of all production. For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022). Marine capture products have the greatest share of production destined for export (domestic exports represented `r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, while inland aquaculture had the lowest share of production destined for export (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).

One concern for increasing globalization is that it may create systemic risks from shocks, particularly when trade is highly concentrated. Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019. Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022). Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production. However, inland aquaculture imports have become less concentrated over time. The more dispersed trade patterns can be attributed to expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing blue food demand in many countries (Naylor et al. 2021; SOFIA 2022). [Roz - maybe we can come up with an illustrative example from the BFA demand paper for here?]

## Other summary stats
Global seafood trade increased from `r total_1996` million t (live weight) in 1996 to `r total_2019` million t (live weight) in 2019, representing a `r round(100*((total_2019-total_1996)/total_1996),0)`% increase over that period.

Global capture fishery trade increased from `r total_capture_1996` million t (live weight) in 1996 to `r total_capture_2019` million t (live weight) in 2019, representing a `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`% increase over that period. In 1996, capture fishery trade represented `r round(100*total_capture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_capture_2019/total_2019)`% of all seafood exports.

Global aquaculture fishery trade increased from `r total_aquaculture_1996` million t (live weight) in 1996 to `r total_aquaculture_2019` million t (live weight) in 2019, representing a `r round(100*((total_aquaculture_2019-total_aquaculture_1996)/total_aquaculture_1996),0)`% increase over that period. In 1996, aquaculture fishery trade represented `r round(100*total_aquaculture_1996/total_1996)`% of all seafood exports and by 2019 capture fishery exports represented `r round(100*total_aquaculture_2019/total_2019)`% of all seafood exports.

```{r covid_effect}
# artis %>% 
#   filter(year %in% 2019:2020) %>% 
#   group_by(year, habitat_method) %>% 
#   summarise(live_weight_t = sum(live_weight_t)) %>% 
#   pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
#   mutate(change = (year2019-year2020)/year2020) %>% 
#   arrange(desc(change))
```

# Shifts in global flows of blue foods

```{r top_traders}
# Top trade tables
top_n_countries <- 5
top_n_countries_flows <- 10
top_n_regions <- 5
top_n_regions_flows <- 10

# Top exporters
## Countries
top_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture exporters
## Countries
top_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture exporters
## Countries
top_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top importers
## Countries
top_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture importers
## Countries
top_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture importers
## Countries
top_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top bilateral flows
## Countries
top_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top capture bilateral flows
## Countries
top_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top aquaculture bilateral flows
## Countries
top_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Create tables
## Top exporter countries
top_exporter_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = exporter_iso3c), by = "rank")

kable(top_exporter_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries, "exporting countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top exporter regions
top_exporter_region_table <- data.frame(rank = 1:top_n_regions) %>%
  left_join(top_region_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_2019 = exporter_region), by = "rank")

kable(top_exporter_region_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions, "exporting regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top importer countries
top_importer_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = importer_iso3c), by = "rank")

kable(top_importer_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries, "importing countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top importer regions
top_importer_region_table <- data.frame(rank = 1:top_n_regions) %>%
  left_join(top_region_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_1996 = importer_region), by = "rank") %>%
  left_join(top_region_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_2019 = importer_region), by = "rank") %>%
  left_join(top_region_capture_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_1996 = importer_region), by = "rank") %>%
  left_join(top_region_capture_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_2019 = importer_region), by = "rank") %>%
  left_join(top_region_aquaculture_importers_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_1996 = importer_region), by = "rank") %>%
  left_join(top_region_aquaculture_importers_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_2019 = importer_region), by = "rank")

kable(top_importer_region_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions, "importing regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top country flows
top_country_flow_table <- data.frame(rank = 1:top_n_countries_flows) %>%
  left_join(top_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_total_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_total_1996), by = "rank") %>%
  left_join(top_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_total_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_total_2019), by = "rank") %>%
  left_join(top_capture_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_capture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_capture_1996), by = "rank") %>%
  left_join(top_capture_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_capture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_capture_2019), by = "rank") %>%
  left_join(top_aquaculture_flows_1996 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_aquaculture_1996 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_aquaculture_1996), by = "rank") %>%
  left_join(top_aquaculture_flows_2019 %>% 
              mutate(rank = 1:top_n_countries_flows,
                     top_aquaculture_2019 = paste(exporter_iso3c, "to", importer_iso3c, sep =" ")) %>%
              select(rank, top_aquaculture_2019), by = "rank")

kable(top_country_flow_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top region flows
top_region_flow_table <- data.frame(rank = 1:top_n_regions_flows) %>%
  left_join(top_region_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_total_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_total_1996), by = "rank") %>%
  left_join(top_region_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_total_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_total_2019), by = "rank") %>%
  left_join(top_region_capture_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_capture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_capture_1996), by = "rank") %>%
  left_join(top_region_capture_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_capture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_capture_2019), by = "rank") %>%
  left_join(top_region_aquaculture_flows_1996 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_aquaculture_1996 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_aquaculture_1996), by = "rank") %>%
  left_join(top_region_aquaculture_flows_2019 %>% 
              mutate(rank = 1:top_n_regions_flows,
                     top_aquaculture_2019 = paste(exporter_region, "to", importer_region, sep = " ")) %>%
              select(rank, top_aquaculture_2019), by = "rank")

kable(top_region_flow_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions_flows, "country trade flows for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

```

```{r top_bilateral_flows}

bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
  filter(year %in% 2015:2019) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  group_by(exporter_region) %>%
  slice_max(n=2, order_by = live_weight_t)

bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei") %>%
  filter(year %in% 2015:2019) %>%
  group_by(exporter_region, importer_region, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  group_by(exporter_region, habitat_method) %>%
  slice_max(n=1, order_by = live_weight_t)

# Regionalization trends
# Changes in trade flow volumes
bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_region, importer_region) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning,
         from_to = paste(exporter_region, " to ", importer_region), 
         intraregion = case_when(
           exporter_region == importer_region ~ "intraregional trade",
           exporter_region != importer_region ~ "interegional trade"
         )) %>% 
  arrange(desc(change)) %>% 
  ungroup() %>%
  ggplot(aes(x = change/1000000, y = reorder(from_to, change), color = intraregion)) +
  geom_segment(aes(x=0, xend=change/1000000, y=reorder(from_to, change), yend=reorder(from_to, change), color = intraregion)) +
  geom_point(size=3) +
  scale_y_reordered() +
  scale_color_manual(values = habitat_source_colors[c(1,4)]) +
  labs(y = "", x = "Live weight (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

# Calculate percent of pairs with increased trade flows
bilateral_flow_change <- bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, importer_iso3c) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning) %>%
  arrange(desc(change))

percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)

# Calculate percent of pairs with increased trade flows by production method and habitat
bilateral_flow_change_method <- bilateral_habitat_method_summary %>% 
  filter(exporter_region != "Other nei", importer_region != "Other nei",
         year %in% c(1996:2000, 2016:2020)) %>%
  mutate(year_group = case_when(
    year %in% 1996:2000 ~ "beginning", 
    year %in% 2016:2020 ~ "end")) %>%
  group_by(year_group, exporter_iso3c, importer_iso3c, habitat_method) %>% 
  summarise(live_weight_t = sum(live_weight_t)) %>%
  pivot_wider(names_from = year_group, values_from = live_weight_t, values_fill = 0) %>% 
  mutate(change = end-beginning) %>%
  arrange(desc(change))

percent_flow_increases_marine_capture <- 100*(nrow(bilateral_flow_change_method %>% 
                                                     filter(habitat_method == "marine capture",
                                                            change > 0)))/nrow(bilateral_flow_change)

```

```{r LDC_patterns}
# LDC countries according to UN (https://www.un.org/development/desa/dpad/least-developed-country-category/ldcs-at-a-glance.html)
LDC_countries <- c("Afghanistan", "Angola", "Bangladesh", "Benin", "Bhutan", "Burkina Faso", "Burundi", "Cambodia", "Central African Republic", "Chad", "Comoros", "Democratic Republic of the Congo", "Djibouti", "Eritrea", "Ethiopia", "Gambia", "Guinea", "Guinea-Bissau", "Haiti", "Kiribati", "Lao People’s Dem. Republic", "Lesotho", "Liberia", "Madagascar", "Malawi", "Mali", "Mauritania", "Mozambique", "Myanmar", "Nepal", "Niger", "Rwanda", "Sao Tome and Principe", "Senegal", "Sierra Leone", "Solomon Islands", "Somalia", "South Sudan", "Sudan", "Timor-Leste", "Togo", "Tuvalu", "Uganda", "United Republic of Tanzania", "Yemen", "Zambia")

LDC_countries <- countrycode(LDC_countries, origin = "country.name", destination = "iso3c")

ldc_habitat_method_summary <- bilateral_habitat_method_summary %>%
  mutate(
    importer_ldc = case_when(
    (importer_iso3c %in% LDC_countries) ~ "LDC",
    !(importer_iso3c %in% LDC_countries) ~ "non-LDC"),
     exporter_ldc = case_when(
    (exporter_iso3c %in% LDC_countries) ~ "LDC",
    !(exporter_iso3c %in% LDC_countries) ~ "non-LDC")) %>%
  group_by(year, exporter_ldc, importer_ldc, habitat_method) %>%
  summarise(live_weight_t = sum(live_weight_t))

ldc_net_habitat_method_summary <- ldc_habitat_method_summary %>%
  ungroup() %>% 
  filter(importer_ldc == "LDC", exporter_ldc == "non-LDC") %>% 
  select(year, habitat_method, "ldc_import" = "live_weight_t") %>%
  left_join(ldc_habitat_method_summary %>%
              ungroup() %>%
              filter(importer_ldc == "non-LDC", exporter_ldc == "LDC") %>% 
              select(year, habitat_method, "ldc_export" = "live_weight_t"),
            by = c("year", "habitat_method")) %>%
  mutate(net_import = ldc_import-ldc_export,
         net_export = ldc_export-ldc_import)

ldc_net_habitat_method_summary %>%
  filter(habitat_method != "unknown") %>%
  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
  geom_line() +
  scale_color_manual(values = habitat_source_colors) +
  labs(x = "", y = "Net export (million t)", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

ldc_net_habitat_method_summary %>%
  group_by(year) %>% 
  summarise(net_export = sum(net_export))

```

## Text for MS

The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX). In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX). Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX). At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method. Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX). Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).

Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF). This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America. [Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.

Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America. The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia. At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture. The declining exports from South America mirror declining marine catch over this period. [Suggestions for discussion points and context related to this?].

Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure). Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF). High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al. 2016).

## Other summary stats

# National blue food supplies

```{r consumption}

supply <- read.csv("outputs/supply.csv")

global_supply <- supply %>%
  group_by(year) %>%
  summarize(supply_no_error = sum(supply_no_error) / 1000000) %>%
  ungroup()

global_supply_1996 <- global_supply %>%
  filter(year == 1996) %>%
  pull(supply_no_error)

global_supply_2020 <- global_supply %>%
  filter(year == 2020) %>%
  pull(supply_no_error)

regional_supply <- supply %>%
  group_by(year, region) %>%
  summarize(supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(supply_no_error_total = sum(supply_no_error)) %>%
  ungroup() %>%
  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)

habitat_method_supply <- supply %>%
  group_by(year, habitat, method) %>%
  summarize(supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(supply_no_error_total = sum(supply_no_error)) %>%
  ungroup() %>%
  mutate(percent_supply = 100 * supply_no_error / supply_no_error_total)

tmp <- habitat_method_supply %>%
  select(habitat, method, year, supply_no_error) %>%
  filter(year == 1996 | year == 2020) %>%
  mutate(year = paste("year", year, sep = "_")) %>%
  pivot_wider(names_from = year, values_from = supply_no_error) %>%
  mutate(percent_diff = 100 * (year_2020 - year_1996) / year_1996)

inland_aqua_supply_change <- tmp %>%
  filter(habitat == "inland" & method == "aquaculture") %>%
  pull(percent_diff)

marine_aqua_supply_change <- tmp %>%
  filter(habitat == "marine" & method == "aquaculture") %>%
  pull(percent_diff)

marine_cap_supply_change <- tmp %>%
  filter(habitat == "marine" & method == "capture") %>%
  mutate(percent_diff = abs(percent_diff)) %>%
  pull(percent_diff)

marine_cap_supply_percent <- habitat_method_supply %>%
  filter(year == 2020) %>%
  filter(percent_supply == max(percent_supply)) %>%
  pull(percent_supply)

method_supplies <- supply %>%
  filter(year == 2020) %>%
  group_by(method) %>%
  summarize(supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  mutate(total_supply = sum(supply_no_error)) %>%
  mutate(percent_supply = 100 * supply_no_error / total_supply)

regional_method_supplies <- supply %>%
  filter(year == 2020) %>%
  group_by(year, region, method) %>%
  summarize(supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  group_by(year, region) %>%
  mutate(region_supply = sum(supply_no_error)) %>%
  ungroup() %>%
  mutate(percent_supply = 100 * supply_no_error / region_supply) %>%
  select(region, region_supply, method, percent_supply) %>%
  pivot_wider(names_from = "method", values_from = "percent_supply")

oceania_cap_supply_percent <- regional_method_supplies %>%
  filter(region == "Oceania") %>%
  pull(capture)

asia_cap_supply_percent <- regional_method_supplies %>%
  filter(region == "Asia") %>%
  pull(capture)

cap_supply_percent <- method_supplies %>%
  filter(method == "capture") %>%
  pull(percent_supply)

asia_farmed_year <- supply %>%
  filter(region == "Asia") %>%
  group_by(year, method) %>%
  summarize(supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  filter(method == "capture" | method == "aquaculture") %>%
  pivot_wider(names_from = method, values_from = supply_no_error) %>%
  filter(aquaculture > capture) %>%
  filter(year == min(year)) %>%
  pull(year)

consumption_foreign <- supply %>%
  filter(year == 1996 | year == 2020) %>%
  group_by(year) %>%
  summarize(supply_foreign = sum(supply_foreign),
            supply_no_error = sum(supply_no_error)) %>%
  ungroup() %>%
  mutate(percent_supply = 100 * supply_foreign / supply_no_error)

foreign_consumption_start <- consumption_foreign %>%
  filter(year == 1996) %>%
  pull(percent_supply)

foreign_consumption_end <- consumption_foreign %>%
  filter(year == 2020) %>%
  pull(percent_supply)

```

## Text for MS

Since direct measurements of consumption are not collected globally, our understanding of global consumption patterns are based on apparent consumption, which generally represents the sum of production and imports minus exports and waste. Consequently, trade is central to estimating consumption and the resolution of consumption patterns is limited by the resolution of the production and trade data. By estimating species level trade, we are able to estimate apparent consumption of blue foods by species and production method.

Globally, blue food consumption increased from `r signif(global_supply_1996, digit = 3)` (million tonnes) in 1996 to `r signif(global_supply_2020, digit = 3)` (million tonnes) in 2020. Blue food consumption also increased across all regions outside of South America, where per capita blue food consumption decreased X%, from X% in 1996 to Y% in 2020. [say this is supply included non food uses] [Double check that FMFO was excluded and discuss this pattern]. Most of the global increase was from inland and marine aquaculture, which increased by `r signif(inland_aqua_supply_change, digits = 3)` and `r signif(inland_aqua_supply_change, digits = 3)`%, respectively, over that period while marine capture consumption declined by `r signif(marine_cap_supply_change, digits = 1)`%. Capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global blue food consumption and its contribution to regional blue food supplies ranges from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.

High reliance on foreign-sourced foods can pose a food security risk, motivating policies to protect domestic supplies such as the development of grain stocks and support for domestic food production (REFs). For example, high seafood import dependence in the US was used to propose a suite of policy changes to boost domestic production, including expanding aquaculture production and opening marine protected areas to fishing (Gephart et al. 2019). Since we estimate exports from domestic and foreign sources, we are able to track changes in reliance on foreign sourced products. Globally, the share of consumption that is foreign sourced has increased from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c), but patterns vary across regions with Asia and S America dominated by domestic supply, but Europe dominated by foreign supply (Fig 3d). [Add points about patterns across production methods].

The diversity of blue food production globally enables high diversity in blue food consumption. However, one concern about food trade is that it leads to homogenization of diets (REFS). To explore this issue, we compared the diversity (represented by the Shannon index) of production versus imports and domestic supply versus foreign supply, as well as the diversity of total supply versus total import volume (Fig 4). One notable limitation of this measure of supply diversity is that instances where production or trade is only estimated at the genus or higher level will be interpreted in the calculation as a single species. Consequently, countries with low species resolution will generally appear to have less diverse production, limiting the ability to compare supply diversity across countries.  
We find that across all regions and over time, production diversity is higher than import diversity (Fig 4a). Additionally, the diversity of domestic supply is generally higher than the diversity of foreign supply, with the domestic diversity is higher in 68% of cases across all years (Fig 4b). Nevertheless, total imports are still positively associated with supply diversity across all regions (Fig 4c). This suggests that although domestic production diversity is generally higher than the diversity of imported products, trade is still bringing in new species and increasing the diversity of products available on the market. Although imports appear to currently add to supply diversity, it is important for it not to displace domestic production diversity, particularly given that domestic production still contributes more to the diversity of the overall blue food supply.

## Other summary stats

### 2023-02-04 ms_figures.Rmd
Update analyses to exclude non-human food

--- 2023-01-17 ms_figures.Rmd
@@
 # Set directories
-datadir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/outputs_20220111"
-
+datadir <- "outputs"
+outdir <- "outputs"
 ```
@@
 
-```{r load_drive_data}
-
-pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
-  left_join(country_metadata, by = c("iso3c")) %>%
-  select(iso3c, year, "region" = "owid_region", pop)
-
-```
-
 # Figure 1: Increases in global export of blue foods
@@
 fig1b <- ggplot() +
-  geom_line(data = percent_export_by_source, aes(x = year, y = percent_export, color = habitat_method)) +
+  geom_line(data = fig1b_data, aes(x = year, y = percent_export, color = habitat_method)) +
   scale_color_manual(values = habitat_source_colors) +
@@
 
-```{r fig2a}
+```{r fig2a_v2}
+
+region_net_import <- artis_region_method %>% 
+  group_by(year, importer_region) %>%
+  summarise(import_live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  rename(region = importer_region) %>%
+  left_join(artis_region_method %>% 
+              group_by(year, exporter_region) %>%
+              summarise(export_live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+              rename(region = exporter_region), by = c("year", "region")) %>%
+  mutate(net_import = import_live_weight_t-export_live_weight_t)
+
+fig2a <- ggplot() +
+  geom_area(data = artis_region_method %>% 
+              select(-exporter_importer) %>%
+              pivot_longer(exporter_region:importer_region, 
+                           names_to = c("flow", "drop"), 
+                           values_to = "region", names_sep = "_") %>% 
+              filter(flow == "exporter") %>%
+              group_by(year, region, habitat_method) %>%
+              summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)), 
+            aes(x = year, y = -1*live_weight_t/1000000, fill = habitat_method)) +
+  geom_area(data = artis_region_method %>% 
+              select(-exporter_importer) %>%
+              pivot_longer(exporter_region:importer_region, 
+                           names_to = c("flow", "drop"), 
+                           values_to = "region", names_sep = "_") %>% 
+              filter(flow == "importer") %>%
+              group_by(year, region, habitat_method) %>%
+              summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE)), 
+            aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
+  scale_fill_manual(values = habitat_source_colors) +
+  geom_hline(yintercept = 0, color = "black", linewidth = 0.5, linetype = "dashed") +
+  geom_line(data = region_net_import, aes(x = year, y = net_import/1000000, color = "Net import"), linewidth = 1) +
+  scale_color_manual(values = "black") +
+  labs(x = "", y = "Trade (million t, live weight)", fill = "Source", color = "") +
+  guides(fill = guide_legend(nrow = 2)) +
+  theme_bw() +
+  theme(legend.position = "top",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank()) +
+  facet_wrap(~region, nrow = 1)
+
+fig2a
+```
+
+```{r fig2a_old, eval=FALSE}
 # Imports by Region
@@
   scale_fill_manual(values = habitat_source_colors) +
-  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)") +
+  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)",
+       fill = "Source") +
   theme_bw() +
@@
         legend.position = "none",
-        strip.background = element_blank(),
-        plot.margin=unit(c(0,0,0,0), "cm")) +
+        strip.background = element_blank()) +
   facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
@@
 
-```{r fig 2c}
+```{r fig2c, eval=FALSE}
 # Figure 2c
@@
 
-```{r fig2}
+```{r fig2_old, eval=FALSE}
 # Overall Figure 2
@@
           heights = c(1, 4.5))
+
+ggsave(file.path(outdir, "fig2.png"))
+```
+
+```{r fig2, fig.height=5, fig.width=4}
+# Overall Figure 2
+ggarrange(ggarrange(fig2a, NULL,
+                    ncol = 2, nrow = 1,
+                    widths = c(4.5, 0.2), labels = c("a", "")), 
+          fig2b, labels = c("", "b"),
+          ncol = 1, heights = c(1.75, 4))
 
@@
   theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank()) +
   facet_wrap(~region)
@@
   theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank())
 
@@
 ```{r fig4a}
-fig4a_data <- read.csv(file.path(datadir, "figa4a_data.csv"))
+fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))
 
@@
 
-top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"), row.names = FALSE)
-top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"), row.names = FALSE)
+top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"))
+top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"))
 
@@
 
-top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"), row.names = FALSE)
-top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"), row.names = FALSE)
+top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"))
+top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"))
 
@@
 ```{r si_export_decreases_increases, fig.height = 4, fig.width = 4}
-top_exporter_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
-top_exporter_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))
+top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
+top_export_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))
 
@@
 ```{r si_import_decreases_increases, fig.height = 4, fig.width = 4}
-top_importer_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
-top_importer_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))
+top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
+top_import_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))
 
@@
 
+```{r ldc_net_exports_habitat_method}
+ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv"))
+
+ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
+  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
+  geom_line() +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(x = "", y = "Net export (million t)", color = "") +
+  theme_minimal() +
+  theme(legend.position = "bottom")
+
+ldc_net_exports_by_source_plot
+```
+
+```{r trade_concentration_n_countries}
+# Concentration of trade by production source, measured as the number of countries comprising 75%
+cumulative_percent_threshold <- 75
+
+export_n_countries_concentration <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))
+import_n_countries_concentration <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
+
+export_n_countries_concentration_by_source <- read.csv(file.path(datadir, "export_concentration_n_countries_by_source.csv"))
+import_n_countries_concentration_by_source <- read.csv(file.path(datadir, "import_concentration_n_countries_by_source.csv"))
+
+export_n_countries_by_source_plot <- ggplot() +
+  geom_line(data = export_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
+  scale_color_manual(values = habitat_source_colors) +
+  geom_line(data = export_n_countries_concentration, aes(x = year, y = n), color = "black") +
+  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+export_n_countries_by_source_plot
+
+import_n_countries_by_source_plot <- ggplot() +
+  geom_line(data = import_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
+  scale_color_manual(values = habitat_source_colors) +
+  geom_line(data = import_n_countries_concentration, aes(x = year, y = n), color = "black") +
+  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+import_n_countries_by_source_plot
+```
+
 ```{r custom_artis_ts}

### 2023-02-04 ms_results.Rmd
Update analyses to exclude non-human food

@@
 total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)
+total_2020 <- round(sum(global_export_by_source %>% filter(year == 2020) %>% pull(live_weight_t))/1000000, 1)
 
@@
 
+# Total trade for all capture
+total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
+total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
+total_capture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
+
 # Total trade for all aquaculture
@@
 total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
+total_aquaculture_2020 <- round(sum(global_export_by_source %>% filter(year == 2020, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
 
 # Total trade for FMFO
-total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
-total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
+artis_fm <- read.csv(file.path(datadir, "artis_fm.csv"))
+
+total_fmfo_1996 <- artis_fm %>%
+  filter(year == 1996) %>%
+  mutate(live_weight_t = live_weight_t/1000000) %>%
+  pull(live_weight_t)
+
+total_fmfo_2019 <- artis_fm %>%
+  filter(year == 2019) %>%
+  mutate(live_weight_t = live_weight_t/1000000) %>%
+  pull(live_weight_t)
+
 ```
@@
 Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships.
-Blue foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`) since 1996.
+Aquatic foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c) since 1996.
 Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
-However, despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
+Corresponding with the first year of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019),1)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),1)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a).
+Despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
 Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively.
-From 1996 to 2019, fishmeal exports remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period.
-Consequently, the share of global blue food exports represented by fishmeal declined from `r abs(round(100*(total_fmfo_1996)/(total_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019), 1))`% in 2019.
+Aquatic food trade outpaced fishmeal (for non-human consumption) trade, which remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period.
+Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.
 
 Another measure of the degree of globalization is the share of production exported.
-Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports (i.e., domestic plus exports of foreign and unknown origin) reached `r round(percent_export_total_2019, 1)`% of all production.
+Domestic exports, which represent exports sourced from that nation’s production irrespective of where their production occurred, increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports, which represents domestic plus exports of foreign and unknown origin, reached `r round(percent_export_total_2019, 1)`% of all production.
 For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022).
@@
 
-One concern for increasing globalization is that it may create systemic risks from shocks, particularly when trade is highly concentrated.
-Since 1996, blue food exports and imports have become somewhat less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 and `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
+Globalization may create systemic risks from shocks, particularly when trade is highly concentrated.
+Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
 Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022).
+The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig SX).
 Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production.
-However, inland aquaculture imports have become less concentrated over time.
-The more dispersed trade patterns can be attributed to expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing blue food demand in many countries (Naylor et al.
+Import concentration declined more dramatically, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig SX).
+The more dispersed import patterns are likely associated with expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al.
 2021; SOFIA 2022).
-[Roz - maybe we can come up with an illustrative example from the BFA demand paper for here?]
 
-![Figure 1: Increases in global export of blue foods.
+![Figure 1: Increases in global export of aquatic foods.
 a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019.
@@
 
+Aquatic foods have become increasingly globalized, with the percent of production exported increasing by `r 100*round((percent_export_total_dom_2019-percent_export_total_dom_1996)/percent_export_total_dom_1996, 1)` and the volume and number of trade partnerships approximately doubling.
+However, trends differ across aquatic food groups, with marine capture remaining the most highly globalized group, but with marine aquaculture showing faster trade growth.
+
 # Shifts in global flows of blue foods
@@
 
-The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating blue food trade (Fig SX).
+The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating aquatic food trade (Fig SX).
 In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
@@
 Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
-Inland blue food trade is dominated by aquaculture, with the largest exports from Vietnam and China and the largest imports by the United States, Japan, and South Korea (Fig SX).
+Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig SX).
 
-Intraregional trade is generally stronger due to shorter transport distances, historical ties, and established regional trade agreements (REF).
-This pattern generally holds for blue food trade as intraregional trade is either the highest or second highest for all regions other than Oceania and South America.
-[Suggestions for discussion points around Oceania and South America exports?] Regional trade strengthened within Asia, Europe and Africa, where intraregional trade increased more than interegional trade.
+Intraregional trade tends to be higher than interregional trade due to shorter transport distances, historical ties, and established regional trade agreements (REF).
+This pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America.
+Our disaggregated trade network is based on a reconciled version of reported trade (Gaulier and Zignago 2010) and therefore does not capture informal and unreported trade networks that often exist among neighboring countries, notably in Africa (Benassi et al.
+2019).
+Since 1996, trade increased between nearly all regional trade pairs, other than South America to Europe and Asia and within North America.
+At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs.
+Despite trade increasing with partners across the globe, regional trade within Asia, Europe and Africa outpaced interregional trade with those regions.
+The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia.
 
-Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America.
-The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia.
-At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture.
-The declining exports from South America mirror declining marine catch over this period.
-[Suggestions for discussion points and context related to this?].
-
-Least developed countries are net importers of blue foods [true of all production methods?] and this pattern has increased X% since 1996 (Fig SX - need to create this figure).
-Net imports of blue foods has previously been identified as a threat to nutrition security in these nations as [check ref for reasoning here] (REF).
-High net imports is also associated with risk of receiving shocks propagated through trade (Gephart et al.
+Least developed countries, as defined by the UN, are net exporters of aquatic foods across all production methods, with net exports more than tripling between 1996 and their peak in 2018 (Fig SX).
+The pattern is largely driven by marine capture products, which nearly quadrupled, while other aquatic food sector net exports remained relatively stable.
+Net exports of aquatic foods may be beneficial to least developed countries where high value species are exported and revenue used to purchase other foods (Asche et al, 2015).
+However, recent work exploring movement of nutrients derived from fisheries suggests international trade is driving redistribution of essential micronutrients prevalent and bioavailable in aquatic foods from areas of high deficiency in middle and low income countries to developed nations with greater nutrient security (Nash et al, 2022).
+Further, Nash et al (2022) found evidence of high vulnerability to future shifts in trade practices among middle and low income nations currently benefiting from trade in aquatic foods; high net imports are associated with the risk of receiving shocks propagated through trade (Gephart et al.
 2016).

### 2023-02-04 prep_data.R

@@ -35,7 +35,21 @@ dbListTables(con)
 
 # ARTIS dataframe
 artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
-  select(-record_id)
+  select(-record_id) 
+
+artis_fm <- artis %>% 
+  filter(hs6 == "230120") %>% 
+  group_by(year) %>%
+  summarise(live_weight_t = sum(live_weight_t, na.rm = TRUE),
+            product_weight_t = sum(product_weight_t, na.rm = TRUE))
+
+artis_nonfood <- artis %>% 
+  # Keep only FMFO and ornamental species
+  filter(hs6 %in% c("230120", "051191", "030110", "030111", "030119"))
+
+artis <- artis %>% 
+  # Remove FMFO and ornamental species
+  filter(!(hs6 %in% c("230120", "051191", "030110", "030111", "030119")))
 
 # Production dataframe
 prod <- dbGetQuery(con, "SELECT * FROM production") %>%
@@ -69,6 +83,12 @@ pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.c
 
 # write.csv(artis, file.path(outdir, "artis.csv"), row.names = FALSE)
 write.csv(prod, file.path(outdir, "prod.csv"), row.names = FALSE)
+
+# Write out the fishmeal trade data, aggregated by year, for summary stats
+write.csv(artis_fm, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
+
+# Write out the bilater non-food trade data to use in supply calculations
+write.csv(artis_nonfood, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
 
 #-------------------------------------------------------------------------------
 # Summarize cleaned FAO production data
@@ -267,10 +287,51 @@ write.csv(artis_region, file.path(outdir, "artis_region.csv"), row.names = FALSE
 #-------------------------------------------------------------------------------
 # Figure 3
 
+# Create calculate_supply function (small difference from version in explore_artis)
+calculate_supply <- function(artis_data, production_data){
+  exports <- artis_data %>%
+    group_by(year, exporter_iso3c, dom_source, sciname, method, habitat) %>%
+    summarise(live_weight_t = sum(live_weight_t)) %>%
+    mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
+    pivot_wider(names_from = "dom_source", values_from = "live_weight_t")
+  
[…]

### 2023-02-06 ms_figures.Rmd
revised figure 2 to include net imports and corrected source factors to have consistent colors for fig 1 and 2 (still need to update other figures)

@@
 fig1a <- fig1a_data %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown"))) %>%
   ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
@@
 ```{r fig1b}
-fig1b_data <- read.csv(file.path(datadir, "fig_1b_data.csv"))
+fig1b_data <- read.csv(file.path(datadir, "fig_1b_data.csv")) %>%
+    mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
@@
 fig1c_out_degree <- fig1c_data %>%
+    mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown"))) %>%
   ggplot(aes(x = year, y = all_in_degree, color = habitat_method)) +
@@
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "Out degree", color = "Source") + 
+  labs(y = "Average no. export partners", color = "Source") + 
   theme_bw()
@@
 
-```{r fig1, fig.height = 3, fig.width = 4.5}
+```{r fig1de}
+habitat_source_colors_v2 <- c("#F7AF75", # light orange = inland aquaculture
+                              "#86ADA7", # light blue = freshwater capture
+                              "#F37542", # medium orange = marine aquaculture
+                              "#1B708F", # blue = marine capture
+                              "black" # unknown
+                          )
+
+# Concentration of trade by production source, measured as the number of countries comprising 75%
+cumulative_percent_threshold <- 75
+
+export_n_countries_concentration <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))
+import_n_countries_concentration <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
+
+export_n_countries_concentration_by_source <- read.csv(file.path(datadir, "export_concentration_n_countries_by_source.csv"))  
+import_n_countries_concentration_by_source <- read.csv(file.path(datadir, "import_concentration_n_countries_by_source.csv"))  %>% 
+      mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture")))
+
+fig1d <- ggplot() +
+  geom_line(data = export_n_countries_concentration_by_source %>% 
+      mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture"))), 
+      aes(x = year, y = n, color = habitat_method)) +
+  geom_line(data = export_n_countries_concentration, aes(x = year, y = n, color = "total"), linewidth = 1) +
+  scale_color_manual(values = habitat_source_colors_v2) +
+  labs(x = "", y = paste("No. countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+fig1d
+
+fig1e <- ggplot() +
+  geom_line(data = import_n_countries_concentration_by_source %>% 
+              mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown"))), 
+      aes(x = year, y = n, color = habitat_method)) +
+  geom_line(data = import_n_countries_concentration, aes(x = year, y = n, color = "total"), linewidth = 1) +
+    scale_color_manual(values = habitat_source_colors_v2) +
+  labs(x = "", y = paste("No. countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
+  lims(y = c(0, 30)) +
+  theme_bw()
+
+fig1e
+```
+
+```{r fig1, fig.height = 5.75, fig.width = 4.5}
 ggarrange(fig1a,
@@
                     labels = c("b", "c")),
+          ggarrange(fig1d,
+                    fig1e,
+                    ncol = 2, nrow = 1,
+                    legend = "none",
+                    labels = c("d", "e")),
           common.legend = TRUE, legend = "bottom",
-          ncol = 1, nrow = 2, labels = c("a", ""))
+          ncol = 1, nrow = 3, labels = c("a", ""))
 
@@
 
-```{r fig2a_v2}
+```{r fig2a}
 
@@
   geom_area(data = artis_region_method %>% 
+                mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown"))) %>%
               select(-exporter_importer) %>%
@@
 fig2b <- artis_region_method %>% 
+    mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown"))) %>%
   ggplot(aes(x = year, y = live_weight_t/1000000, fill = habitat_method)) +
@@
 
-```{r fig2c, eval=FALSE}
+```{r fig2c_old, eval=FALSE}
 # Figure 2c
@@
 
-```{r trade_concentration_n_countries}
-# Concentration of trade by production source, measured as the number of countries comprising 75%
-cumulative_percent_threshold <- 75
-
-export_n_countries_concentration <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))
-import_n_countries_concentration <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
-
-export_n_countries_concentration_by_source <- read.csv(file.path(datadir, "export_concentration_n_countries_by_source.csv"))
-import_n_countries_concentration_by_source <- read.csv(file.path(datadir, "import_concentration_n_countries_by_source.csv"))
-
-export_n_countries_by_source_plot <- ggplot() +
-  geom_line(data = export_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
-  scale_color_manual(values = habitat_source_colors) +
-  geom_line(data = export_n_countries_concentration, aes(x = year, y = n), color = "black") +
-  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
-  lims(y = c(0, 30)) +
-  theme_bw()
-
-export_n_countries_by_source_plot
-
-import_n_countries_by_source_plot <- ggplot() +
-  geom_line(data = import_n_countries_concentration_by_source, aes(x = year, y = n, color = habitat_method)) +
-  scale_color_manual(values = habitat_source_colors) +
-  geom_line(data = import_n_countries_concentration, aes(x = year, y = n), color = "black") +
-  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
-  lims(y = c(0, 30)) +
-  theme_bw()
-
-import_n_countries_by_source_plot
-```
-
 ```{r custom_artis_ts}

### 2023-02-06 prep_data.R
new supply and diversity data in prep data workflow

@@ -331,47 +331,43 @@ calculate_supply <- function(artis_data, production_data){
 }
 
 # Supply / Consumption data
-supply <- calculate_supply(artis %>% 
-                             rename(habitat = environment), 
-                           prod %>% 
-                             rename(live_weight_t = production_t)) %>%
-  # Add habitat-method column
-  mutate(habitat_method = paste(habitat, method, sep =" ")) %>% 
+consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_max_per_capita_100kg"
+consumption_dir <- "../ARTIS/consumption/consumption_max_per_capita_100kg"
+supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
+
+supply <- supply %>%
+  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
   mutate(habitat_method = case_when(
     str_detect(habitat_method, "unknown") ~ "unknown", 
     TRUE ~ habitat_method
-  )) %>% 
+  )) %>%
   # Set factor levels
   mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                             "marine aquaculture", "inland aquaculture",
-                                                            "unknown" ))) %>% 
-  left_join(country_metadata %>% 
-              select(iso3c, "region" = "owid_region"), by = "iso3c") %>% 
-  mutate(supply_no_error = case_when(supply_no_error < 0 ~ 0,
-                                     TRUE ~ supply_no_error))
+                                                            "unknown" ))) %>%
+  rename(supply_domestic = domestic_consumption_t,
+         supply_foreign = foreign_consumption_t)
 
-write.csv(supply, file.path(outdir, "supply.csv"), row.names = FALSE)
+supply <- supply %>%
+  left_join(
+    pop,
+    by = c("iso3c", "year")
+  )
 
 # Figure 3a
 # Calculate per capita supply by region and source
 supply_total <- supply %>% 
   group_by(iso3c, region, year, habitat_method) %>%
   summarise(supply = sum(supply),
-            supply_no_error = sum(supply_no_error),
             supply_domestic = sum(supply_domestic),
             supply_foreign = sum(supply_foreign),
-            import = sum(imports_live_weight_t), 
[…]

### 2023-02-06 prep_data.R
editing prep data

@@ -17,7 +17,7 @@ library(tidytext)
 # Directory listing
 
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
-outdir <- "outputs"
+outdir <- "outputs_20230206"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -336,6 +336,7 @@ consumption_dir <- "../ARTIS/consumption/consumption_max_per_capita_100kg"
 supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
 
 supply <- supply %>%
+  filter(!is.na(habitat) & !is.na(method)) %>%
   mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
   mutate(habitat_method = case_when(
     str_detect(habitat_method, "unknown") ~ "unknown", 
@@ -354,20 +355,36 @@ supply <- supply %>%
     by = c("iso3c", "year")
   )
 
+global_pop <- pop %>%
+  filter(region != "Other nei") %>%
+  group_by(year) %>%
+  summarize(pop = sum(pop, na.rm = TRUE)) %>%
+  ungroup()
+
 # Figure 3a
 # Calculate per capita supply by region and source
-supply_total <- supply %>% 
+supply_total <- supply %>%
+  filter(!is.na(habitat_method)) %>%
   group_by(iso3c, region, year, habitat_method) %>%
-  summarise(supply = sum(supply),
+  summarise(supply = sum(supply_test),
             supply_domestic = sum(supply_domestic),
             supply_foreign = sum(supply_foreign),
             pop = sum(pop)
-  )
+  ) %>%
+  ungroup()
 
 fig3a_data <- supply_total %>%
+  filter(!is.na(habitat_method)) %>%
   filter(region != "Other nei") %>%
   group_by(year, habitat_method) %>%
-  summarise(supply_per_cap = 1000 * (sum(supply, na.rm = TRUE)) / sum(pop, na.rm = TRUE))
+  summarize(supply = sum(supply)) %>%
+  ungroup() %>%
+  left_join(
+    global_pop %>%
+      rename(year_global_pop = pop),
+    by = c("year")
+  ) %>%
+  mutate(supply_per_cap = 1000 * supply / year_global_pop)
 
 write.csv(fig3a_data, file.path(outdir, "fig3a_data.csv"), row.names = FALSE)
 
[…]

### 2023-02-06 ms_figures.Rmd

@@
 library(tidytext)
+library(cowplot)
 ```
@@
         legend.position = "none",
-        strip.background = element_blank()) +
+        strip.background = element_blank(), 
+        plot.margin = margin(t = 8, r = 8, l = 2, unit = "mm")) +
   facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
              labeller = label_wrap_gen(width = 5), scales = "free")
+
+fig2b <- ggdraw(fig2b) +
+  draw_label("Exporting region", x = 0.975, y = 0.5, angle = 270) +
+  draw_label("Importing region", x = 0.5, y = 0.975)
 
@@
 
-```{r fig4b}
+```{r fig4b_old, eval=FALSE}
 fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))
@@
 fig4b
+```
+
+```{r fig4b}
+# NOTE: Rahul - we need to have the correct file read in here.
+For reference, the file currently generating this and the new fig 4d is "/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv"
+# I am not sure we should go with these colors, but didn't have time to keep messing with this
+# These comments also apply to the new fig4c
+
+fig4b <- diversity %>% 
+  filter(year == 2019) %>%
+  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
+  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
+  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
+  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
+  geom_violin() + 
+  geom_point(alpha = 0.5) +
+  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
+  labs(x = "", y = "Shannon index", fill = "") +
+  facet_wrap(~region, nrow=1) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank())
+
+fig4b
+```
+
+```{r fig4c}
+fig4c <- diversity %>% 
+  filter(year == 2019) %>%
+  select(iso3c, region, "Domestic" = "shannon_domestic", "Foreign" = "shannon_foreign") %>% 
+  pivot_longer(Domestic:Foreign, names_to = "shannon_var", values_to = "shannon_index") %>%
+  mutate(shannon_var = factor(shannon_var, levels = c("Domestic", "Foreign"))) %>%
+  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
+  geom_violin() + 
+  geom_point(alpha = 0.5) +
+  scale_fill_manual(values = c(habitat_source_colors[2], habitat_source_colors[4])) +
+  labs(x = "", y = "Shannon index", fill = "") +
+  facet_wrap(~region, nrow=1) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank())
+
+fig4c
 ```
@@
 ```
+
 ```{r fig4, fig.height = 3, fig.width = 3}

### 2023-02-06 ms_results.Rmd
ms results formatted correctly

@@
 # Directory
-datadir <- "outputs"
+datadir <- "outputs_20230206"
 
@@
   mutate(live_weight_t = live_weight_t/1000000) %>%
-  pull(live_weight_t)
+  pull(live_weight_t) %>%
+  sum(na.rm = TRUE)
 
@@
   filter(year == 1996) %>%
-  pull(supply_no_error)
+  pull(supply)
 
@@
   filter(year == 2020) %>%
-  pull(supply_no_error)
+  pull(supply)
 
@@
 tmp <- habitat_method_supply %>%
-  select(habitat, method, year, supply_no_error) %>%
+  select(habitat, method, year, supply) %>%
   filter(year == 1996 | year == 2020) %>%
   mutate(year = paste("year", year, sep = "_")) %>%
-  pivot_wider(names_from = year, values_from = supply_no_error) %>%
+  pivot_wider(names_from = year, values_from = supply) %>%
   mutate(percent_diff = 100 * (year_2020 - year_1996) / year_1996)
@@
   group_by(year, method) %>%
-  summarize(supply_no_error = sum(supply_no_error)) %>%
+  summarize(supply = sum(supply)) %>%
   ungroup() %>%
   filter(method == "capture" | method == "aquaculture") %>%
-  pivot_wider(names_from = method, values_from = supply_no_error) %>%
+  pivot_wider(names_from = method, values_from = supply) %>%
   filter(aquaculture > capture) %>%

### 2023-02-07 ms_figures.Rmd
all figures except for custom artis timeline running

@@
 # Set directories
-datadir <- "outputs"
-outdir <- "outputs"
+datadir <- "outputs_20230206"
+outdir <- "outputs_20230206"
 ```
@@
 
-```{r fig1, fig.height = 5.75, fig.width = 4.5}
+```{r fig1, fig.height = 12, fig.width = 7.5}
 ggarrange(fig1a,
@@
 
-```{r fig2, fig.height=5, fig.width=4}
+```{r fig2, fig.height=8, fig.width=7}
 # Overall Figure 2
@@
 fig3a <- fig3a_data %>%
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>%
   ggplot(aes(x = year, y = supply_per_cap, color = habitat_method)) +
@@
 fig3b <- fig3b_data %>%
+   # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>%
   ggplot(aes(x = year, y = supply_per_cap, fill = habitat_method)) +
@@
 
-```{r fig3}
+```{r fig3, fig.height = 6, weight = 8}
 fig3 <- ggarrange(
@@
 fig4b <- fig4b_data %>%
-  ggplot(aes(x = shannon_supply_domestic, y = shannon_supply_foreign, color = region)) +
+  ggplot(aes(x = shannon_domestic, y = shannon_foreign, color = region)) +
     geom_point(alpha = 0.7) +
@@
 
-```{r fig4c}
+```{r fig4c_old}
 fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
@@
 fig4c <- fig4c_data %>%
-  ggplot(aes(x = imports_live_weight_t/1000000, y = shannon_supply_no_error, color = region)) +
+  ggplot(aes(x = imports/1000000, y = shannon_supply, color = region)) +
   geom_point(alpha = 0.7) +
@@
 
-```{r fig4, fig.height = 3, fig.width = 3}
+```{r fig4, fig.height = 8, fig.width = 12}
 fig4 <- ggarrange(fig4a, 
@@
 fig4
+
+ggsave(file.path(outdir, "fig4.png"))
 ```
@@
 
-```{r si_top_exporters, fig.height = 4, fig.width = 4}
+```{r si_top_exporters, fig.height = 12, fig.width = 5}
 
-top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv"))
-top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv"))
+top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
+top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 exporters_old_plot <- top_exporters_old %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  ggplot(aes(x = live_weight_t/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
   geom_bar(stat = "identity") +
@@
 exporters_recent_plot <- top_exporters_recent %>%
-  ggplot(aes(x = live_weight_t/1000000, y = exporter_iso3c, fill = habitat_method)) +
+  ggplot(aes(x = live_weight_t/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
   geom_bar(stat = "identity") +
@@
 
-```{r si_top_importers, fig.height = 4, fig.width = 4}
+```{r si_top_importers, fig.height = 12, fig.width = 5}
 
-top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv"))
-top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv"))
+top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
+top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 importers_old_plot <- top_importers_old %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  ggplot(aes(x = live_weight_t/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method)) +
   geom_bar(stat = "identity") +
@@
 importers_recent_plot <- top_importers_recent %>%
-  ggplot(aes(x = live_weight_t/1000000, y = importer_iso3c, fill = habitat_method)) +
+  ggplot(aes(x = live_weight_t/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method)) +
   geom_bar(stat = "identity") +
@@
 
-```{r si_regional_changes, fig.height = 4, fig.width = 4}
+```{r si_regional_changes, fig.height = 12, fig.width = 5}
 
 # Largest regional export changes
-top_regional_export_change <- read.csv(file.path(datadir, "top_regional_export_change.csv"))
+top_regional_export_change <- read.csv(file.path(datadir, "top_regional_export_change.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 regional_export_change_plot <- top_regional_export_change %>%
-  ggplot(aes(x = change/1000000, y = exporter_region, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_region, yend=exporter_region)) +
+  ggplot(aes(x = change/1000000, y = reorder(exporter_region, ranking), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=reorder(exporter_region, ranking), yend=exporter_region)) +
   geom_point(size=4) +
@@
 # Largest regional import changes
-top_regional_import_change <- read.csv(file.path(datadir, "top_regional_import_change.csv"))
+top_regional_import_change <- read.csv(file.path(datadir, "top_regional_import_change.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 regional_import_change_plot <- top_regional_import_change %>%
-  ggplot(aes(x = change/1000000, y = importer_region, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_region, yend=importer_region)) +
+  ggplot(aes(x = change/1000000, y = reorder(importer_region, change), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y=reorder(importer_region, change), yend=importer_region)) +
   geom_point(size=4) +
@@
 
-```{r si_export_decreases_increases, fig.height = 4, fig.width = 4}
-top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv"))
-top_export_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv"))
+```{r si_export_decreases_increases, fig.height = 12, fig.width = 5}
+top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
+top_export_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 exporter_increases_plot <- top_export_increases %>%
-  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  ggplot(aes(x = change/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y= reorder(exporter_iso3c, ranking), yend=exporter_iso3c)) +
   geom_point(size=4) +
@@
 exporter_decreases_plot <- top_export_decreases %>%
-  ggplot(aes(x = change/1000000, y = exporter_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=exporter_iso3c, yend=exporter_iso3c)) +
+  ggplot(aes(x = change/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y= reorder(exporter_iso3c, ranking), yend=exporter_iso3c)) +
   geom_point(size=4) +
@@
 
-```{r si_import_decreases_increases, fig.height = 4, fig.width = 4}
-top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv"))
-top_import_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv"))
+```{r si_import_decreases_increases, fig.height = 12, fig.width = 5}
+top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
+top_import_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv")) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "unknown")))
 
 importer_increases_plot <- top_import_increases %>%
-  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  ggplot(aes(x = change/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y= reorder(importer_iso3c, ranking), yend=importer_iso3c)) +
   geom_point(size=4) +
@@
 importer_decreases_plot <- top_import_decreases %>%
-  ggplot(aes(x = change/1000000, y = importer_iso3c, fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=importer_iso3c, yend=importer_iso3c)) +
+  ggplot(aes(x = change/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
+  geom_segment(aes(x=0, xend=change/1000000, y= reorder(importer_iso3c, ranking), yend=importer_iso3c)) +
   geom_point(size=4) +
@@
 
-```{r regional_avg_annual_change, fig.height = 4, fig.width = 4}
+```{r regional_avg_annual_change, fig.height = 10, fig.width = 6}
 regional_avg_annual_change <- read.csv(file.path(datadir, "regional_avg_annual_change.csv"))
@@
 regional_slopes_ranked <- regional_avg_annual_change %>%
-  ggplot(aes(x = slope/1000, y = exporter_importer, color = intraregion)) +
-  geom_segment(aes(x=0, xend=slope/1000, y=exporter_importer, yend=exporter_importer, color = intraregion)) +
+  ggplot(aes(x = slope/1000, y = reorder(exporter_importer, slope), color = intraregion)) +
+  geom_segment(aes(x=0, xend=slope/1000, y=reorder(exporter_importer, slope), yend=exporter_importer, color = intraregion)) +
   geom_point(size=3) +

### 2023-02-07 prep_data.R
prep data updates to reorder graphs

@@ -361,12 +361,22 @@ global_pop <- pop %>%
   summarize(pop = sum(pop, na.rm = TRUE)) %>%
   ungroup()
 
+global_supply_per_cap <- supply %>%
+  group_by(year) %>%
+  summarize(supply = sum(supply)) %>%
+  ungroup() %>%
+  left_join(
+    global_pop,
+    by = c("year")
+  ) %>%
+  mutate(per_cap = 1000 * supply / pop)
+
 # Figure 3a
 # Calculate per capita supply by region and source
 supply_total <- supply %>%
   filter(!is.na(habitat_method)) %>%
   group_by(iso3c, region, year, habitat_method) %>%
-  summarise(supply = sum(supply_test),
+  summarise(supply = sum(supply),
             supply_domestic = sum(supply_domestic),
             supply_foreign = sum(supply_foreign),
             pop = sum(pop)
@@ -389,10 +399,26 @@ fig3a_data <- supply_total %>%
 write.csv(fig3a_data, file.path(outdir, "fig3a_data.csv"), row.names = FALSE)
 
 # Figure 3b
-fig3b_data <- supply_total %>%
+# fig3b_data <- supply_total %>%
+#   filter(region != "Other nei") %>%
+#   group_by(year, region, habitat_method) %>%
+#   summarise(supply_per_cap = 1000 * (sum(supply, na.rm = TRUE)) / sum(pop, na.rm = TRUE))
+
+regional_pop <- pop %>%
+  group_by(region, year) %>%
+  summarize(pop = sum(pop)) %>%
+  ungroup()
+
+fig3b_data <- supply %>%
   filter(region != "Other nei") %>%
   group_by(year, region, habitat_method) %>%
-  summarise(supply_per_cap = (sum(supply, na.rm = TRUE)) / sum(pop, na.rm = TRUE))
+  summarize(supply = sum(supply)) %>%
+  ungroup() %>%
+  left_join(
+    regional_pop,
+    by = c("region", "year")
+  ) %>%
+  mutate(supply_per_cap = 1000 * supply / pop)
 
 write.csv(fig3b_data, file.path(outdir, "fig3b_data.csv"), row.names = FALSE)
 
@@ -567,7 +593,10 @@ top_regional_export_change <- bilateral_habitat_method_summary %>%
   group_by(habitat_method) %>%
   slice_max(n = 10, order_by = change) %>%
   ungroup() %>%
[…]

### 2023-02-07 ms_figures.Rmd
updates to diversity and custom artis ts

@@
 # These comments also apply to the new fig4c
-
+diversity <- read.csv("/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv")
 fig4b <- diversity %>% 

### 2023-02-07 prep_data.R

@@ -850,30 +850,26 @@ write.csv(import_concentration_habitat_method_n_countries, file.path(outdir, "im
 # Custom ARTIS timeseries (based on different HS versions used)
 
 hs_version_datadir <- "/Volumes/jgephart/ARTIS/Outputs/S_net/snet_20221129/snet"
-hs_version_datadir <- "outputs_20230206"
 
 # Based on snet midpoint estimation
 hs_versions <- c("96", "02", "07", "12", "17")
 
 artis_ts <- data.frame()
 
-artis_ts <- artis %>%
-  group_by(year, hs_version) %>%
-  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE))
+for (i in 1:length(hs_versions)) {
+  curr_hs <- hs_versions[i]
+  print(curr_hs)
+  curr_fp <- file.path(hs_version_datadir, paste("HS", curr_hs, "/midpoint_artis_habitat_prod_ts_HS", curr_hs, ".csv", sep = ""))
+  print(curr_fp)
+  curr_artis <- read.csv(curr_fp) %>%
+    mutate(hs_version = paste("HS", curr_hs, sep = "")) %>%
+    group_by(year, hs_version) %>%
+    summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+    ungroup()
 
-# for (i in 1:length(hs_versions)) {
-#   curr_hs <- hs_versions[i]
-#   print(curr_hs)
-#   curr_fp <- file.path(hs_version_datadir, paste("HS", curr_hs, "/midpoint_artis_habitat_prod_ts_HS", curr_hs, ".csv", sep = ""))
-#   curr_artis <- read.csv(curr_fp) %>%
-#     mutate(hs_version = paste("HS", curr_hs, sep = "")) %>%
-#     group_by(year, hs_version) %>%
-#     summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
-#     ungroup()
-#   
-#   artis_ts <- artis_ts %>%
-#     bind_rows(curr_artis)
-# }
+  artis_ts <- artis_ts %>%
+    bind_rows(curr_artis)
+}
 
 custom_ts <- artis_ts %>%
   filter(

### 2023-02-07 prep_data.R
update supply file path to local copy to read in faster

@@ -331,9 +331,9 @@ calculate_supply <- function(artis_data, production_data){
 }
 
 # Supply / Consumption data
-consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_max_per_capita_100kg"
-consumption_dir <- "../ARTIS/consumption/consumption_max_per_capita_100kg"
-supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
+#consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_max_per_capita_100kg"
+#consumption_dir <- "../ARTIS/consumption/consumption_max_per_capita_100kg"
+supply <- read.csv("summary_consumption.csv")
 
 supply <- supply %>%
   filter(!is.na(habitat) & !is.na(method)) %>%

### 2023-02-07 ms_figures.Rmd
updated ms figures

--- 2023-02-07 ms_figures.Rmd
@@
 library(cowplot)
+library(scales)
 ```
@@
 
-```{r fig1, fig.height = 12, fig.width = 7.5}
+```{r fig1, fig.height = 7, fig.width = 5}
 ggarrange(fig1a,
@@
 
-```{r fig2, fig.height=8, fig.width=7}
+```{r fig2, fig.height=4, fig.width=3}
 # Overall Figure 2
@@
 
-```{r fig3, fig.height = 6, weight = 8}
+```{r fig3, fig.height = 4, fig.width = 5}
 fig3 <- ggarrange(
@@
 
-```{r fig4a}
+```{r fig4a_old, eval = FALSE}
 fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))
@@
   facet_wrap(~region)
+
+fig4a
+```
+
+```{r fig4a}
+diversity <- read.csv("/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv")
+fig4a <- diversity %>% 
+  filter(year == 2019) %>%
+  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
+  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
+  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
+  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
+  geom_violin() + 
+  geom_point(alpha = 0.5) +
+  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
+  labs(x = "", y = "Shannon index", fill = "") +
+  facet_wrap(~region, nrow=1) +
+  theme_bw() +
+  theme(legend.position = "bottom",
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        strip.background = element_blank())
 
@@
 ```{r fig4b}
-# NOTE: Rahul - we need to have the correct file read in here.
-For reference, the file currently generating this and the new fig 4d is "/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv"
-# I am not sure we should go with these colors, but didn't have time to keep messing with this
-# These comments also apply to the new fig4c
-
 fig4b <- diversity %>% 
-  filter(year == 2019) %>%
-  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
-  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
-  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
-  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
-  geom_violin() + 
-  geom_point(alpha = 0.5) +
-  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
-  labs(x = "", y = "Shannon index", fill = "") +
-  facet_wrap(~region, nrow=1) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank())
-
-fig4b
-```
-
-```{r fig4c}
-fig4c <- diversity %>% 
   filter(year == 2019) %>%
@@
 
-```{r fig4c_old}
-fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
+```{r fig4d}
+fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
 
-fig4c <- fig4c_data %>%
+fig4d <- fig4d_data %>%
+  # total supply greater than 1 tonne (FIXIT update manuscript caption)
+  filter(supply > 1) %>%
   ggplot(aes(x = imports/1000000, y = shannon_supply, color = region)) +
   geom_point(alpha = 0.7) +
-  scale_x_log10() +
+  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
+              labels = trans_format("log10", math_format(10^.x))) +
+  annotation_logticks(sides = "b") +
   scale_color_manual(values = region_colors) +
@@
 
-fig4c
+#
+
+fig4d
+```
+```{r fig4e}
+fig4e_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
+
+fig4e <- fig4e_data %>%
+  # total supply greater than 1 tonne (FIXIT update manuscript caption)
+  filter(supply > 1) %>%
+  ggplot(aes(x = exports/1000000, y = shannon_supply, color = region)) +
+  geom_point(alpha = 0.7) +
+  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
+              labels = trans_format("log10", math_format(10^.x))) +
+  annotation_logticks(sides = "b") +
+  scale_color_manual(values = region_colors) +
+  theme_minimal() +
+  labs(x = "Total exports (million t, live weight)",
+       y = "Supply diversity",
+       color = "")
+
+#
+
+fig4e
 ```
 
-```{r fig4, fig.height = 8, fig.width = 12}
+```{r fig4, fig.height = 7, fig.width = 7}
 fig4 <- ggarrange(fig4a,
-          ggarrange(fig4b,
-                    fig4c,
+                  fig4b,
+          ggarrange(fig4d,
+                    fig4e,
                     nrow = 1,
@@
                     legend = "bottom",
-                    labels = c("b", "c")),
-          nrow = 2,
-          labels = c("a", "", "")
+                    labels = c("c", "d")),
+          nrow = 3,
+          labels = c("a", "b", "")
           )
@@
   labs(y = "", x = "Live weight (million t)") +
+  xlim(c(0,6)) +
   facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
@@
   ncol = 2,
-  labels = c("Old", "Recent")
+  labels = c("'96-'00", "'16-'20")
 )

### 2023-02-07 prep_data.R
regional artis by source data

@@ -492,6 +492,27 @@ write.csv(fig4c_data, file.path(outdir, "fig4c_data.csv"), row.names = FALSE)
 # SUPPLEMENTARY FIGURES
 #-------------------------------------------------------------------------------
 
+regional_artis <- artis %>%
+  rename(habitat = environment) %>%
+  group_by(exporter_iso3c, importer_iso3c, habitat, method, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  left_join(
+    country_metadata %>%
+      select(exporter_iso3c = iso3c, exporter_region = owid_region),
+    by = c("exporter_iso3c")
+  ) %>%
+  left_join(
+    country_metadata %>%
+      select(importer_iso3c = iso3c, importer_region = owid_region),
+    by = c("importer_iso3c")
+  ) %>%
+  group_by(exporter_region, importer_region, habitat, method, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup()
+
+write.csv(regional_artis, file.path(outdir, "regional_artis_by_source.csv"), row.names = FALSE)
+
 # Top Trading Partners old (1996 - 2000) vs recent (2016 - 2020)
 # Summarize total imports and exports by habitat and environment
 bilateral_habitat_method_summary <- artis %>%

### 2023-02-07 ms_results.Rmd
Update results with new text

@@
 total_2020 <- round(sum(global_export_by_source %>% filter(year == 2020) %>% pull(live_weight_t))/1000000, 1)
-
-total_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
-total_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "capture")) %>% pull(live_weight_t))/1000000, 1)
 
@@
   filter(year == 2019, calc_type == "domestic exports") %>% 
+  pull(percent_export)
+percent_export_total_dom_1996 <- percent_export_total %>% 
+  filter(year == 1996, calc_type == "domestic exports") %>% 
   pull(percent_export)
@@
 
-Globalization describes the degree of international connectedness, which can be characterized by the increasing flow of goods and number of trade relationships.
-Aquatic foods became increasingly globalized from 1996-2019, with exports increasing `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a) and the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c) since 1996.
-Over that period, both farmed and wild exports increased, though aquaculture exports grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
-Corresponding with the first year of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019),1)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),1)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a).
-Despite aquaculture now comprising half of blue food production, capture fishery products still dominate blue food exports, with capture products responsible for `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
-Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively.
-Aquatic food trade outpaced fishmeal (for non-human consumption) trade, which remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period.
-Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.
+Globalization describes the degree of international connectedness, which can be characterized by increasing flows of input, intermediate and final products among countries.
+Aquatic food exports increased `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019.
+Over that period, both farmed and wild exports increased, though aquaculture grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
+Corresponding with the start of the COVID-19 pandemic, global aquatic food exports declined `r -1*round(100*((total_2020-total_2019)/total_2019), 0)`% overall in 2020 relative to 2019, with a `r -1*round(100*((total_capture_2020-total_capture_2019)/total_capture_2019),0)`% decline in capture exports, but a `r round(100*((total_aquaculture_2020-total_aquaculture_2019)/total_aquaculture_2019),1)`% increase in aquaculture exports (Fig 1a).
+Despite aquaculture comprising half of aquatic food production, capture fishery products constitute `r 100*round(total_capture_2019/total_2019, 1)`% of exports.
 
 Another measure of the degree of globalization is the share of production exported.
-Domestic exports, which represent exports sourced from that nation’s production irrespective of where their production occurred, increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019 while total exports, which represents domestic plus exports of foreign and unknown origin, reached `r round(percent_export_total_2019, 1)`% of all production.
+Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_2019, 1)`% in 1996.
+while total exports reached 33% of all production in 2019.
 For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022).
-Marine capture products have the greatest share of production destined for export (domestic exports represented `r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, while inland aquaculture had the lowest share of production destined for export (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).
+Increasing marine capture export volumes despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, and largest increases in the share exported (Fig 2b).
+Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, leading to an increasing share exported (Fig 2b).
+Despite increases, inland aquaculture still had the lowest share of production destined for export in 2019 (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).
+This finding clarifies prior debates about the orientation of aquaculture, while export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al.
+2017; Belton et al 2018; Gephart et al.
+2019).
 
-Globalization may create systemic risks from shocks, particularly when trade is highly concentrated.
+Globalization exposes countries to external shocks, while also serving as a buffer against local shocks.
+Recent work exploring trade characteristics associated with systemic risk to shocks suggests higher exposure when networks are highly connected and concentrated, and when countries are more dependent on imports (Gephart et al.
+2016;  Marchand et al.
+2016; d’Amour et al.
+2016).
+From 1996-2019, aquatic food trade became more connected with the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c).
+Marine capture networks are most highly connected, followed by marine aquaculture, with inland capture and aquaculture trade being the least connected.
+
 Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
 Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022).
-The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig SX).
-Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production.
-Import concentration declined more dramatically, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig SX).
-The more dispersed import patterns are likely associated with expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al.
+The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig 2d).
+Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production to a few key regions.
+Similarly, the concentration of trade for individual species tends to be much higher.
+The divergent trends in trade features and differences among aquatic food groups suggests differences in the degree and types of trade shock risks across the sector.
+Such differences were observed in responses to COVID-19 (Love et al.
+2021).
+However, understanding risk to shocks across foods is a priority area for future work (Davis et al.
+2021).
+
+Though aquatic food production, distribution, and consumption remain highly uneven (Hicks et al.
+2022), we found declinding import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e).
+The more dispersed import patterns are likely associated with growing populations and expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al.
 2021; SOFIA 2022).
+However, the relationship between aquatic food demand and income varies across aquatic foods, with demand generally increasing with income for higher quality fish, but falling for lower quality fish (Naylor et al.
+2021).
 
@@
 b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports.
-c) Out degree by production method and environment.](outputs/fig1.png)
+c) Average number of export partners (out degree) by production method and environment.
+d) Number of countries comprising 75% of global exports by production method and environment with the global total in the black line.
+e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](outputs/fig1.png)
 
@@
 
+Marine-sourced products dominate both capture and aquaculture exports, comprising `r round(100*total_marine_capture_2019/(total_marine_capture_2019+total_inland_capture_2019), 1)`% and `r round(100*total_marine_aquaculture_2019/(total_marine_aquaculture_2019+total_inland_aquaculture_2019), 1)`%, respectively.
+Aquatic food trade outpaced fishmeal (for non-human consumption) trade, which remained relatively stable, declining `r abs(round(100*(total_fmfo_2019-total_fmfo_1996)/(total_fmfo_1996), 1))`% over the time period.
+Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.
+
 # Shifts in global flows of blue foods
@@
 
-The top importers, exporters, and bilateral flows vary by production method, underscoring the importance of disaggregating aquatic food trade (Fig SX).
-In recent years, the marine capture and aquaculture trade networks have been dominated by Asia and Europe, and to a lesser extent, North America (Fig SX).
-Since 1996, the marine trade networks shifted more toward Asia and Europe, with large increases in marine capture imports and exports by Africa and large increases in marine aquaculture exports from South America and imports to North America (Fig SX).
-At the country level, although some countries rank among the top traders across production methods, such as China for exports and the USA for imports, many countries are only top traders for one production method.
-Meanwhile, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig SX).
-Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig SX).
+Given the geographic patchiness of wild fish and aquaculture production, trade helps meet aquatic food demand in countries.
+Aquatic food imports are especially important where per capita demand is rising, aquaculture is limited, and wild fishery catch is stagnant.
+Corresponding to the geographical variability, we find the top importers, exporters, and bilateral flows to vary by habitat and farmed versus wild source, underscoring the importance of disaggregating trade (Fig S1-3).
+For example, Asia and Europe, and to a lesser extent, North America dominated marine capture and aquaculture trade networks in recent years (Fig S1).
+Since 1996, marine trade shifted toward Asia and Europe, with large increases in marine capture imports to and exports from Africa (Fig 2b).
 
-Intraregional trade tends to be higher than interregional trade due to shorter transport distances, historical ties, and established regional trade agreements (REF).
-This pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America.
-Our disaggregated trade network is based on a reconciled version of reported trade (Gaulier and Zignago 2010) and therefore does not capture informal and unreported trade networks that often exist among neighboring countries, notably in Africa (Benassi et al.
+At the country level, although some countries rank among the top traders across production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one production method.
+Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig S2-3).
+In general, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).
+
+Intraregional trade tends to be higher than interregional trade for all products due to shorter transport distances, historical ties, patterns of aquatic food preferences and established regional trade agreements (Natale et al.
+2015).
+We find that this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America.
+Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America.
+At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs.
+Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa outpaced interregional trade with those regions.
+The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia (Fig 2a; Fig S6).
+Our trade estimates are ultimately from reported trade and therefore do not capture informal and unreported trade networks.
+Though estimated unreported trade is not globally available, it can be significant, especially for neighboring countries.
+For example, informal exports from Benin to Nigeria are estimated to be more than five times the formal exports (Benassi et al.
 2019).
-Since 1996, trade increased between nearly all regional trade pairs, other than South America to Europe and Asia and within North America.
-At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs.
-Despite trade increasing with partners across the globe, regional trade within Asia, Europe and Africa outpaced interregional trade with those regions.
-The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia.
+Including informal trade would therefore likely strengthen intraregional trade patterns.
 
-Least developed countries, as defined by the UN, are net exporters of aquatic foods across all production methods, with net exports more than tripling between 1996 and their peak in 2018 (Fig SX).
-The pattern is largely driven by marine capture products, which nearly quadrupled, while other aquatic food sector net exports remained relatively stable.
-Net exports of aquatic foods may be beneficial to least developed countries where high value species are exported and revenue used to purchase other foods (Asche et al, 2015).
-However, recent work exploring movement of nutrients derived from fisheries suggests international trade is driving redistribution of essential micronutrients prevalent and bioavailable in aquatic foods from areas of high deficiency in middle and low income countries to developed nations with greater nutrient security (Nash et al, 2022).
-Further, Nash et al (2022) found evidence of high vulnerability to future shifts in trade practices among middle and low income nations currently benefiting from trade in aquatic foods; high net imports are associated with the risk of receiving shocks propagated through trade (Gephart et al.
-2016).
+Across regions, Europe and North America have the highest net imports while South America has the highest net exports (Fig 2a).
+Least developed countries collectively are net exporters of aquatic foods across all production methods, with net exports almost tripling between 1996 and their 2018 peak (Fig S7).
+Least developed country net exports are dominated by marine capture products, which nearly quadrupled from 1996-2019.
+Net exports of aquatic foods may be economically beneficial to least developed countries where high value species are exported and revenue used to purchase other foods (Asche et al, 2015).
+However, economic and political barriers inhibit these wealth-based benefits from being realized (Hicks et al.
+2022).
+Further, recent work exploring movement of nutrients derived from fisheries suggests international trade is driving redistribution of essential micronutrients prevalent and bioavailable in aquatic foods from areas of high deficiency in middle and low income countries to developed nations with greater nutrient security (Nash et al, 2022).
+Further, Nash et al (2022) found evidence of high vulnerability to future shifts in trade practices among middle and low income nations currently benefiting from trade in aquatic foods.
 
-![Figure 2: Regional trade flows by trade partner and production source.
-a) Total imports to each region colored by trade partner.
-b) Bilateral flows colored by production method and habitat source.
-c) Total exports from each region colored by trade partner.
-Values represent million tonnes in live weight equivalents.](outputs/fig2.png)
+![Figure 2: Regional trade flows by production source (habitat and method).
+a) Total imports (positive) and exports (negative) colored by source, with net import trend in black.
+b) Bilateral flows colored by production source with exporting region along the rows and importing region along the columns.
+Values represent million tonnes in live weight equivalents.
+Note the scale difference across rows.](outputs/fig2.png)
 
@@
 
-global_supply_1996 <- global_supply %>%
+global_supply_per_cap <- read.csv(file.path(datadir, "global_supply_per_cap.csv"))
+
+regional_supply_per_cap <- read.csv(file.path(datadir, "fig3b_data.csv"))
+
+# Pull global supply numbers from 1996 and 2019
+global_supply_1996 <- global_supply_per_cap %>%
   filter(year == 1996) %>%
-  pull(supply)
+  pull(per_cap)
 
-global_supply_2020 <- global_supply %>%
-  filter(year == 2020) %>%
-  pull(supply)
+global_supply_2019 <- global_supply_per_cap %>%
+  filter(year == 2019) %>%
+  pull(per_cap)
+
+# Check trends across regions
+regional_supply_trends_summary <- regional_supply_per_cap %>%
+  select(year, region, habitat_method, supply_per_cap) %>%
+  group_by(year, region) %>% 
+  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
+  filter(year %in% c(1996, 2019)) %>%
+  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
+  mutate(diff = year2019-year1996, 
+         percent_diff = 100*diff/year1996)
+
+regional_supply_trends_habitatmethod_summary <- regional_supply_per_cap %>%
+  select(year, region, habitat_method, supply_per_cap) %>%
+  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
+  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
+  mutate(diff = year2019-year1996, 
+         percent_diff = 100*diff/year1996)
+
+habitatmethod_supply_trends_summary <- regional_supply_per_cap %>%
+  select(year, habitat_method, supply_per_cap) %>%
+  group_by(year, habitat_method) %>% 
+  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
+  filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
+  pivot_wider(names_from = year, values_from = supply_per_cap, names_prefix = "year") %>%
+  mutate(diff = year2019-year1996, 
+         percent_diff = 100*diff/year1996)
 
@@
 
+foreign_consumption_region <- read.csv(file.path(datadir, "fig3d_data.csv"))
+
+```
+
+```{r}
+national_supply_diversity <- read.csv(file.path(datadir, "fig4b_data.csv"))
+
+n_import_exceeds_prod <- national_supply_diversity %>%
+  filter(shannon_imports > shannon_prod) %>%
+  nrow()
+
 ```
@@
 
-Since direct measurements of consumption are not collected globally, our understanding of global consumption patterns are based on apparent consumption, which generally represents the sum of production and imports minus exports and waste.
-Consequently, trade is central to estimating consumption and the resolution of consumption patterns is limited by the resolution of the production and trade data.
-By estimating species level trade, we are able to estimate apparent consumption of blue foods by species and production method.
+Since direct measurements of human consumption (e.g.
+dietary intake) are not collected globally, consumption is generally represented by apparent consumption.
+Apparent consumption is generally calculated as production plus imports minus exports and waste.
+Trade is therefore central to estimating consumption and historically limited understanding of aquatic food consumption patterns.
+By estimating species level trade, we are able to estimate apparent consumption of aquatic foods by species, production method, and geographical origin.
 
-Globally, blue food consumption increased from `r signif(global_supply_1996, digit = 3)` (million tonnes) in 1996 to `r signif(global_supply_2020, digit = 3)` (million tonnes) in 2020.
-Blue food consumption also increased across all regions outside of South America, where per capita blue food consumption decreased X%, from X% in 1996 to Y% in 2020.
-[say this is supply included non food uses] [Double check that FMFO was excluded and discuss this pattern].
-Most of the global increase was from inland and marine aquaculture, which increased by `r signif(inland_aqua_supply_change, digits = 3)` and `r signif(inland_aqua_supply_change, digits = 3)`%, respectively, over that period while marine capture consumption declined by `r signif(marine_cap_supply_change, digits = 1)`%.
-Capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global blue food consumption and its contribution to regional blue food supplies ranges from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.
+Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019.
+Our estimates are slightly lower than FAOSTAT, which reports global aquatic food consumption at 15.6 kg/capita/year in 1996 and 20.7 kg/capita/year in 2019.
+We found that aquatic food consumption increased across all regions outside of North America, which was relatively stable and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b).
+Global increase were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively.
+Meanwhile, inland capture consumption grew modestly (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, a `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% increase), while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019).
+Nevertheless, capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global aquatic food consumption, with its contribution to regional aquatic food consumption ranging from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.
 
-High reliance on foreign-sourced foods can pose a food security risk, motivating policies to protect domestic supplies such as the development of grain stocks and support for domestic food production (REFs).
-For example, high seafood import dependence in the US was used to propose a suite of policy changes to boost domestic production, including expanding aquaculture production and opening marine protected areas to fishing (Gephart et al.
+Since we estimate exports from domestic and foreign sources, we can track changes in reliance on foreign sourced products.
+Globally the share of consumption that is foreign-sourced increased modestly, from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c).
+Patterns vary greatly across regions with Asia and S America dominated by domestic supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Asia", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% and `r signif(foreign_consumption_region %>% filter(year == 2019, region == "South America", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019, respectively), but Europe dominated by foreign supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Europe", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019) (Fig 3d).
+High reliance on foreign-sourced foods can pose a food security risk (Baer-Nawrocka and Sadowski 2019; Kummu et al.
+2020), though it is not clear the extent to which these risks exist across aquatic foods.
+Nevertheless, countries have enacted policies to protect domestic supplies, including developing food stocks and subsidizing domestic food production (Wood et al.
+2023).
+The United States used foreign dependence on aquatic foods as a motivation for a suite of policy changes to boost domestic production, including expanding aquaculture production and opening marine protected areas to fishing (Gephart et al.
 2019).
-Since we estimate exports from domestic and foreign sources, we are able to track changes in reliance on foreign sourced products.
-Globally, the share of consumption that is foreign sourced has increased from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c), but patterns vary across regions with Asia and S America dominated by domestic supply, but Europe dominated by foreign supply (Fig 3d).
-[Add points about patterns across production methods].
 
-The diversity of blue food production globally enables high diversity in blue food consumption.
-However, one concern about food trade is that it leads to homogenization of diets (REFS).
-To explore this issue, we compared the diversity (represented by the Shannon index) of production versus imports and domestic supply versus foreign supply, as well as the diversity of total supply versus total import volume (Fig 4).
-One notable limitation of this measure of supply diversity is that instances where production or trade is only estimated at the genus or higher level will be interpreted in the calculation as a single species.
-Consequently, countries with low species resolution will generally appear to have less diverse production, limiting the ability to compare supply diversity across countries.
-We find that across all regions and over time, production diversity is higher than import diversity (Fig 4a).
-Additionally, the diversity of domestic supply is generally higher than the diversity of foreign supply, with the domestic diversity is higher in 68% of cases across all years (Fig 4b).
-Nevertheless, total imports are still positively associated with supply diversity across all regions (Fig 4c).
-This suggests that although domestic production diversity is generally higher than the diversity of imported products, trade is still bringing in new species and increasing the diversity of products available on the market.
-Although imports appear to currently add to supply diversity, it is important for it not to displace domestic production diversity, particularly given that domestic production still contributes more to the diversity of the overall blue food supply.
+High global diversity of aquatic food production enables high diversity in consumption, which is important for supplying a range of micronutrients, supporting a range of livelihoods, and is thought to support food system resilience (Seekell et al.
+2017).
+However, one concern about food trade is that it homogenizes diets through two related pathways: by introducing foods that make diets more similar across countries and/or by out-competing diverse local foods.
+To explore this issue for aquatic foods, we compared the Shannon diversity index, of production versus imports and and exports (Fig 4a) and domestic supply versus foreign supply (Fig 4b).
+Across regions, import diversity is generally higher than production diversity and export diversity is similar to production diversity across regions (Fig 4a).
+Import diversity exceeds production diversity in `r n_import_exceeds_prod` out of `r nrow(national_supply_diversity)` countries.
+Although the diversity of imported goods tends to be higher than production, diversity of domestic-sourced consumption tends to be higher than  diversity of foreign sourced consumption (Fig 4b).
+Total imports and exports are positively associated with supply diversity across all regions (Fig 4c).
+This suggests trade is contributing new species to many national markets.
+Collectively, our diversity findings point to a need to protect domestic production diversity while still benefiting from the diversity of imported products.
 
@@
 
-![Figure 3: Aquatic resource supply trends and regional patterns.
-a) Global blue food supply by production source over time.
-b) Regional blue food supply by production source over time.
-c) Global blue food domestic versus foreign sourcing over time.
-d) Regional blue food domestic versus foreign sourcing over time.
-Note these supply figures include non-food uses.](outputs/fig3.png)
+![Figure 3: Aquatic food apparent consumption (supply) trends and regional patterns.
+a) Global aquatic food apparent consumption by production source over time.
+b) Regional aquatic food apparent consumption by production source over time.
+c) Global aquatic food domestic versus foreign sourcing over time.
+d) Regional aquatic food domestic versus foreign sourcing over time.
+](outputs/fig3.png)
 
 ![Figure 4: Relationship between trade and supply diversity.
-a) Time series of the Shannon diversity of imported blue foods and produced blue foods by region.
-b) Foreign supply Shannon diversity versus domestic supply Shannon diversity for 2018.
-The black line represents the 1-to-1 line, such that points falling below the line indicate countries where the domestic supply diversity is greater than the foreign supply diversity.
-c) Relationship between total imports and overall supply diversity for 2018.](outputs/fig4.png)
+a) Violin plots of the Shannon diversity of production, imports and exports for 2019.
+b) Violin plots of the Shannon diversity of domestic and foreign consumption for 2019.
+c) Relationship between total imports and overall supply diversity for 2019.
+d) Relationship between total exports and overall supply diversity for 2019.
+Note for figures 4c and 4d, only countries with a supply of at least 1 tonne were included.](outputs/fig4.png)

### 2023-02-07 prep_data.R

@@ -371,6 +371,8 @@ global_supply_per_cap <- supply %>%
   ) %>%
   mutate(per_cap = 1000 * supply / pop)
 
+write.csv(global_supply_per_cap, file.path(outdir, "global_supply_per_cap.csv"), row.names = FALSE)
+
 # Figure 3a
 # Calculate per capita supply by region and source
 supply_total <- supply %>%

### 2023-02-07 ms_figures.Rmd

@@
         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank())
+        strip.background = element_blank(),
+        text = element_text(size = 18)
+        )
 
@@
         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank())
+        strip.background = element_blank(),
+        text = element_text(size = 18))
 
-fig4c
+fig4b
 ```
 
-```{r fig4d}
-fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
+```{r fig4c}
+fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
 
-fig4d <- fig4d_data %>%
+fig4c <- fig4c_data %>%
   # total supply greater than 1 tonne (FIXIT update manuscript caption)
@@
        y = "Supply diversity",
-       color = "")
+       color = "") +
+  theme(
+    text = element_text(size = 18)
+  )
 
@@
 
-fig4d
+fig4c
 ```
-```{r fig4e}
-fig4e_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
+```{r fig4d}
+fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
 
-fig4e <- fig4e_data %>%
+fig4d <- fig4d_data %>%
   # total supply greater than 1 tonne (FIXIT update manuscript caption)
@@
 
-fig4e
+fig4d
 ```
 
-```{r fig4, fig.height = 7, fig.width = 7}
+```{r fig4, fig.height = 5, fig.width = 7}
+fig4 <- ggarrange(
+  ggarrange(
+    fig4a, fig4b,
+    ncol = 1,
+    nrow = 2,
+    labels = c("a", "b")
+  ),
+  ggarrange(
+    fig4c, fig4d,
+    ncol = 1,
+    nrow = 2,
+    common.legend = TRUE,
+    legend = "bottom",
+    labels = c("c", "d")
+  ),
+  nrow = 1, ncol = 2,
+  widths = c(1, 0.5)
+)
+
+fig4
+
+ggsave(file.path(outdir, "fig4.png"))
+```
+
+```{r fig4_old, fig.height = 7, fig.width = 7}
 fig4 <- ggarrange(fig4a,
@@
 
-```{r si_top_exporters, fig.height = 12, fig.width = 5}
+```{r si_top_exporters, fig.height = 5, fig.width = 4}
 
@@
 
-```{r si_top_importers, fig.height = 12, fig.width = 5}
+```{r si_top_importers, fig.height = 5, fig.width = 4}
 
@@
   scale_y_reordered() +
+  xlim(c(0, 5.5)) +
   labs(y = "", x = "Live weight (million t)") +
@@
 
-```{r si_regional_changes, fig.height = 12, fig.width = 5}
+```{r si_regional_changes, fig.height = 4, fig.width = 4}
 
@@
 
-```{r si_export_decreases_increases, fig.height = 12, fig.width = 5}
+```{r si_export_decreases_increases, fig.height = 4, fig.width = 4.5}
 top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv")) %>%
@@
 
-```{r si_import_decreases_increases, fig.height = 12, fig.width = 5}
+```{r si_import_decreases_increases, fig.height = 4, fig.width = 4.5}
 top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv")) %>%
@@
 
-```{r regional_avg_annual_change, fig.height = 10, fig.width = 6}
+```{r regional_avg_annual_change, fig.height = 4, fig.width = 5}
 regional_avg_annual_change <- read.csv(file.path(datadir, "regional_avg_annual_change.csv"))
@@
 regional_slopes_ranked
+
+ggsave(file.path(outdir, "regional_avg_annual_change.png"))
 ```
@@
 ldc_net_exports_by_source_plot
+
+ggsave(file.path(outdir, "ldc_net_exports_by_source.png"))
 ```

### 2023-02-07 ms_figures.Rmd
regional sankey by source figure

@@
   scale_fill_manual(values = habitat_source_colors) +
-  labs(y = "Exports (mil. t, live weight)", fill = "Source", x = "") +
+  labs(y = "Exports\n(mil. t, live weight)", fill = "Source", x = "") +
   theme_bw() +
-  theme(legend.position = "bottom")
+  theme(legend.position = "bottom",
+        axis.title.y = element_text(size = 16),
+        axis.text.x = element_text(size = 12),
+        axis.text.y = element_text(size = 12),
+        legend.text = element_text(size = 9),
+        )
 
@@
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "% Production Exported", x = "", color = "Source") +
-  theme_bw()
+  labs(y = "% Production\nExported", x = "", color = "Source") +
+  theme_bw() +
+  theme(
+    text = element_text(size = 16)
+  )
 
@@
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "Average no. export partners", color = "Source") + 
-  theme_bw()
+  labs(y = "Average\nno. export partners", color = "Source") + 
+  theme_bw() +
+  theme(
+    text = element_text(size = 16),
+    axis.title.x = element_blank()
+  )
 
@@
   scale_color_manual(values = habitat_source_colors_v2) +
-  labs(x = "", y = paste("No. countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
+  labs(x = "", y = paste("No. countries\ncomprising ", cumulative_percent_threshold, "% exports", sep = ""), color = "Source") +
   lims(y = c(0, 30)) +
-  theme_bw()
+  theme_bw() +
+  theme(
+    text = element_text(size = 16)
+  )
 
@@
     scale_color_manual(values = habitat_source_colors_v2) +
-  labs(x = "", y = paste("No. countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
+  labs(x = "", y = paste("No. countries\ncomprising ", cumulative_percent_threshold, "% imports", sep = ""), color = "Source") +
   lims(y = c(0, 30)) +
-  theme_bw()
+  theme_bw() +
+  theme(
+    text = element_text(size = 16)
+  )
 
@@
 
-```{r fig1, fig.height = 7, fig.width = 5}
+```{r fig1, fig.height = 4, fig.width = 4}
 ggarrange(fig1a,
@@
 
+```{r regional_sankey, fig.height = 2, fig.width = 2}
+# regional_artis <- read.csv(file.path(datadir, "regional_artis_by_source.csv"))
+# plot_regional_sankey_method_habitat(regional_artis, 2019, 2019)
+
+marine_cap_regions <- regional_artis %>%
+  filter(habitat == "marine" & method == "capture") %>%
+  group_by(exporter_region, importer_region) %>%
+    summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+    ungroup() %>%
+    make_long(exporter_region, importer_region, value = live_weight_t)
+
+marine_cap_regions$node <- factor(marine_cap_regions$node,
+                              levels = c("Europe", "Oceania", "Asia",
+                                         "South America", "North America",
+                                         "Africa"))
+
+marine_cap_p <- marine_cap_regions %>%
+  ggplot(aes(x = x, 
+               next_x = next_x, 
+               node = node, 
+               next_node = next_node,
+               fill = factor(node),
+               value = value,
+               label = node)) +
+    geom_sankey(flow.alpha = 0.6) +
+    scale_fill_manual(values = region6_palette) +
+    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    theme_sankey(base_size = 18) +
+  labs(x = NULL) +
+    #labs(title = "Marine", x = NULL, y = "CAPTURE") +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_blank(),
+    plot.title = element_text(hjust = 0.5),
+    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    )
+
+marine_cap_p
+```
+
+```{r marine_aqua, fig.height = 2, fig.width = 2}
+marine_aqua_regions <- regional_artis %>%
+  filter(habitat == "marine" & method == "aquaculture") %>%
+  group_by(exporter_region, importer_region) %>%
+    summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+    ungroup() %>%
+    make_long(exporter_region, importer_region, value = live_weight_t)
+
+marine_aqua_regions$node <- factor(marine_aqua_regions$node,
+                              levels = c("Europe", "Oceania", "Asia",
+                                         "South America", "North America",
+                                         "Africa"))
+
+marine_aqua_p <- marine_aqua_regions %>%
+  ggplot(aes(x = x, 
+               next_x = next_x, 
+               node = node, 
+               next_node = next_node,
+               fill = factor(node),
+               value = value,
+               label = node)) +
+    geom_sankey(flow.alpha = 0.6) +
+    scale_fill_manual(values = region6_palette) +
+    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    theme_sankey(base_size = 18) +
+  labs(x = NULL) +
+    #labs(title = "Marine", x = NULL, y = "CAPTURE") +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_blank(),
+    plot.title = element_text(hjust = 0.5),
+    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    )
+
+marine_aqua_p
+```
+
+```{r inland_cap, fig.height = 2, fig.width = 2}
+inland_cap_regions <- regional_artis %>%
+  filter(habitat == "inland" & method == "capture") %>%
+  group_by(exporter_region, importer_region) %>%
+    summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+    ungroup() %>%
+    make_long(exporter_region, importer_region, value = live_weight_t)
+
+inland_cap_regions$node <- factor(inland_cap_regions$node,
+                              levels = c("Europe", "Oceania", "Asia",
+                                         "South America", "North America",
+                                         "Africa"))
+
+inland_cap_p <- inland_cap_regions %>%
+  ggplot(aes(x = x, 
+               next_x = next_x, 
+               node = node, 
+               next_node = next_node,
+               fill = factor(node),
+               value = value,
+               label = node)) +
+    geom_sankey(flow.alpha = 0.6) +
+    scale_fill_manual(values = region6_palette) +
+    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    theme_sankey(base_size = 18) +
+  labs(x = NULL) +
+    #labs(title = "Marine", x = NULL, y = "CAPTURE") +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_blank(),
+    plot.title = element_text(hjust = 0.5),
+    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    )
+
+inland_cap_p
+```
+
+```{r inland_aqua, fig.height = 2, fig.width = 2}
+inland_aqua_regions <- regional_artis %>%
+  filter(habitat == "inland" & method == "aquaculture") %>%
+  group_by(exporter_region, importer_region) %>%
+    summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+    ungroup() %>%
+    make_long(exporter_region, importer_region, value = live_weight_t)
+
+inland_aqua_regions$node <- factor(inland_aqua_regions$node,
+                              levels = c("Europe", "Oceania", "Asia",
+                                         "South America", "North America",
+                                         "Africa"))
+
+inland_aqua_p <- inland_aqua_regions %>%
+  ggplot(aes(x = x, 
+               next_x = next_x, 
+               node = node, 
+               next_node = next_node,
+               fill = factor(node),
+               value = value,
+               label = node)) +
+    geom_sankey(flow.alpha = 0.6) +
+    scale_fill_manual(values = region6_palette) +
+    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    theme_sankey(base_size = 18) +
+  labs(x = NULL) +
+    #labs(title = "Marine", x = NULL, y = "CAPTURE") +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_blank(),
+    plot.title = element_text(hjust = 0.5),
+    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    )
+
+inland_aqua_p
+```
+```{r regional_sankey_by_source, fig.height = 2.5, fig.width = 3}
+ggarrange(marine_cap_p, inland_cap_p,
+            marine_aqua_p, inland_aqua_p,
+            ncol = 2, nrow = 2,
+          labels = c("a", "b", "c", "d"))
+          # labels = c("Marine Capture", "Inland Capture",
+          #            "Marine Aquaculture", "Inland Aquaculture"))
+
+ggsave(file.path(outdir, "regional_sankey_by_source.png"))
+```
+
 ```{r si_top_exporters, fig.height = 5, fig.width = 4}

### 2023-02-07 ms_results.Rmd

@@
 Another measure of the degree of globalization is the share of production exported.
-Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_2019, 1)`% in 1996.
+Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_1996, 1)`% in 1996.
 while total exports reached 33% of all production in 2019.

### 2023-02-08 ms_figures.Rmd
updating with foreign export stats

@@
 library(scales)
+library(ggsankey)
+library(ggpubr)
 ```
@@
 
-```{r fig4_old, fig.height = 7, fig.width = 7}
+```{r fig4_old, fig.height = 7, fig.width = 7, eval = FALSE}
 fig4 <- ggarrange(fig4a,
@@
 ```{r regional_sankey, fig.height = 2, fig.width = 2}
-# regional_artis <- read.csv(file.path(datadir, "regional_artis_by_source.csv"))
+regional_artis <- read.csv(file.path(datadir, "regional_artis_by_source.csv"))
 # plot_regional_sankey_method_habitat(regional_artis, 2019, 2019)
@@
     geom_sankey(flow.alpha = 0.6) +
-    scale_fill_manual(values = region6_palette) +
+    scale_fill_manual(values = region_colors) +
     geom_sankey_label(size = 3, color = "white", fill = "gray40") +
@@
 ```
+
+```{r dom_source_ts, fig.height = 4, fig.width = 6}
+dom_source_ts <- read.csv(file.path(datadir, "dom_source_ts.csv"))
+
+dom_source_ts_p <- dom_source_ts %>%
+  filter(dom_source != "error export") %>%
+  ggplot(aes(x = year, y = live_weight_t / 1000000, color = dom_source)) +
+  geom_line() +
+  scale_color_manual(values = c("#D38F35", "#114F59")) +
+  theme_bw() +
+  labs(x = "", y = "Live weight\nmillion tonnes", color = "Source") +
+  theme(
+    legend.position = "bottom",
+    legend.box.margin=margin(t = -25)
+  )
+
+dom_source_ts_p
+```
+
+```{r dom_source_by_habitat_ts, fig.height = 4, fig.width = 7}
+dom_source_by_habitat_method <- read.csv(file.path(datadir, "dom_source_by_habitat_method.csv"))
+
+dom_source_by_habitat_method_p <- dom_source_by_habitat_method %>%
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>%
+  ggplot(aes(x = year, y = live_weight_t / 1000000, fill = habitat_method)) +
+  geom_area() +
+  scale_fill_manual(values = habitat_source_colors) +
+  theme_bw() +
+  labs(x = "", y = "Live weight\nmillion tonnes", fill = "Source") +
+  facet_wrap(~dom_source) +
+  theme(
+    legend.position = "bottom",
+    panel.spacing = unit(5, "mm"),
+    legend.box.margin=margin(t = -15)
+  )
+
+dom_source_by_habitat_method_p
+```
+
+```{r dom_source_overview, fig.height = 5, fig.width = 7}
+ggarrange(
+  dom_source_ts_p,
+  dom_source_by_habitat_method_p,
+  nrow = 2,
+  labels = c("a", "b")
+)
+
+ggsave("dom_source_2019_overview.png")
+```
+
+```{r top_processing_countries_by_source, fig.height = 4, fig.width = 6}
+
+top_processing_by_source <- read.csv(file.path(datadir, "top_processing_by_source.csv"))
+
+top_processing_by_source %>%
+  filter(year == 2019) %>%
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>%
+  ggplot(aes(x = live_weight_t / 1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  theme_bw() +
+  labs(x = "Live weight million tonnes", y = "", fill = "Source") +
+  facet_wrap(~habitat_method, scales = "free_y") +
+  theme(legend.position = "none")
+
+ggsave("foreign_exports_2019_by_source.png")
+```

### 2023-02-08 ms_results.Rmd
updating with foreign export stats

@@
 
+```{r dom_source_stats}
+dom_source_ts <- read.csv(file.path(datadir, "dom_source_ts.csv"))
+foreign_exports_1996 <- dom_source_ts %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 1996) %>%
+  pull(live_weight_t)
+
+foreign_exports_2019 <- dom_source_ts %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 2019) %>%
+  pull(live_weight_t)
+
+foreign_exports_percent_change <- 100 * (foreign_exports_2019 - foreign_exports_1996) / foreign_exports_2019
+
+# foreign exports as share of total production
+exports_percents <- dom_source_ts %>%
+  group_by(year) %>%
+  mutate(total_export = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(export_percent = 100 * live_weight_t / total_export)
+
+exports_percents_no_error <- dom_source_ts %>%
+  filter(dom_source != "error export") %>%
+  group_by(year) %>%
+  mutate(total_export = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  mutate(export_percent = 100 * live_weight_t / total_export)
+
+foreign_export_percent_1996 <- exports_percents %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 1996) %>%
+  pull(export_percent)
+
+foreign_export_percent_2019 <- exports_percents %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 2019) %>%
+  pull(export_percent)
+
+foreign_export_percent_1996_no_error <- exports_percents_no_error %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 1996) %>%
+  pull(export_percent)
+
+foreign_export_percent_2019_no_error <- exports_percents_no_error %>%
+  filter(dom_source == "foreign export") %>%
+  filter(year == 2019) %>%
+  pull(export_percent)
+```
+
 ## Text for MS
@@
 Consequently, fishmeal exports as a share of the sum of aquatic food and fishmeal exports declined from `r abs(round(100*(total_fmfo_1996)/(total_1996+total_fmfo_1996), 1))`% in 1996 to only `r abs(round(100*(total_fmfo_2019)/(total_2019+total_fmfo_2019), 1))`% in 2019.
+
+Foreign exports changed by `r round(foreign_exports_percent_change, digits = 1)`% (from `r round(foreign_exports_1996/1000000, digits = 2)` million tonnes in 1996 to `r round(foreign_exports_2019/1000000, digits = 2)` million tonnes in 2019).
+
+As a share of exports (error exports included), foreign exports increased by `r round(100*(foreign_export_percent_2019-foreign_export_percent_1996)/foreign_export_percent_1996, digits = 1)`% from `r round(foreign_export_percent_1996, digits = 1)`% in 1996 to `r round(foreign_export_percent_2019, digits = 1)`% in 2019.
+
+As a share of exports (error exports NOT included), foreign exports increased by `r round(100*(foreign_export_percent_2019_no_error-foreign_export_percent_1996_no_error)/foreign_export_percent_1996_no_error, digits = 1)`% from `r round(foreign_export_percent_1996_no_error, digits = 1)`% in 1996 to `r round(foreign_export_percent_2019_no_error, digits = 1)`% in 2019.
 

### 2023-02-08 prep_data.R

@@ -1077,3 +1077,64 @@ consumption_foreign <- supply %>%
 
 write.csv(consumption_foreign, file.path(outdir, "consumption_foreign.csv"), row.names = FALSE)
 
+
+# Processing
+
+dom_source_artis <- artis %>%
+  filter(dom_source != "error export")
+
+dom_source_ts <- artis %>%
+  group_by(dom_source, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup()
+
+write.csv(dom_source_ts, file.path(outdir, "dom_source_ts.csv"), row.names = FALSE)
+
+dom_source_by_habitat_method <- dom_source_artis %>%
+  group_by(dom_source, habitat_method, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup()
+
+write.csv(dom_source_by_habitat_method, file.path(outdir, "dom_source_by_habitat_method.csv"), row.names = FALSE)
+
+foreign_exports <- artis %>%
+  filter(dom_source == "foreign export")
+
+top_processing_countries <- foreign_exports %>%
+  group_by(exporter_iso3c, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  group_by(year) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup()
+
+top_processing_countries_by_source <- foreign_exports %>%
+  group_by(exporter_iso3c, habitat_method, year) %>%
+  summarize(live_weight_t = sum(live_weight_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  group_by(habitat_method, year) %>%
+  slice_max(n = 10, order_by = live_weight_t) %>%
+  mutate(ranking = rank(live_weight_t)) %>%
+  ungroup() %>%
+  mutate(exporter_iso3c = reorder_within(exporter_iso3c, ranking, habitat_method))
+
+write.csv(top_processing_countries_by_source, file.path(outdir, "top_processing_by_source.csv"), row.names = FALSE)
+
+top_processing_countries %>%
+  filter(year == 2019) %>%
+  ggplot(aes(x = live_weight_t, y = reorder(exporter_iso3c, ranking))) +
+  geom_bar(stat = "identity") +
+  scale_y_reordered() +
[…]

### 2023-03-31 ms_figures.Rmd
fixed output file path
Move supply diversity out of main MS to go into separate analysis

--- 2023-02-07 ms_figures.Rmd
@@
 # Set directories
-datadir <- "outputs_20230206"
-outdir <- "outputs_20230206"
+datadir <- "outputs_20230330"
+outdir <- "outputs_20230330"
 ```
@@
 
-```{r fig4a_old, eval = FALSE}
-fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))
-
-fig4a <- fig4a_data %>%
-  ggplot(aes(x = year, y = shannon, color = source)) +
-  geom_line() +
-  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
-  labs(x = "", y = "Shannon index", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~region)
-
-fig4a
-```
-
-```{r fig4a}
-diversity <- read.csv("/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv")
-fig4a <- diversity %>% 
-  filter(year == 2019) %>%
-  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
-  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
-  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
-  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
-  geom_violin() + 
-  geom_point(alpha = 0.5) +
-  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
-  labs(x = "", y = "Shannon index", fill = "") +
-  facet_wrap(~region, nrow=1) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank(),
-        text = element_text(size = 18)
-        )
-
-fig4a
-```
-
-```{r fig4b_old, eval=FALSE}
-fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))
-
-fig4b <- fig4b_data %>%
-  ggplot(aes(x = shannon_domestic, y = shannon_foreign, color = region)) +
-    geom_point(alpha = 0.7) +
-    geom_abline(slope = 1, intercept = 0) +
-  scale_color_manual(values = region_colors) +
-  labs(x = "Domestic supply diversity", 
-       y = "Foreign supply diversity",
-       color = "") +
-    theme_minimal()
-
-fig4b
-```
-
-```{r fig4b}
-fig4b <- diversity %>% 
-  filter(year == 2019) %>%
-  select(iso3c, region, "Domestic" = "shannon_domestic", "Foreign" = "shannon_foreign") %>% 
-  pivot_longer(Domestic:Foreign, names_to = "shannon_var", values_to = "shannon_index") %>%
-  mutate(shannon_var = factor(shannon_var, levels = c("Domestic", "Foreign"))) %>%
-  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
-  geom_violin() + 
-  geom_point(alpha = 0.5) +
-  scale_fill_manual(values = c(habitat_source_colors[2], habitat_source_colors[4])) +
-  labs(x = "", y = "Shannon index", fill = "") +
-  facet_wrap(~region, nrow=1) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank(),
-        text = element_text(size = 18))
-
-fig4b
-```
-
-```{r fig4c}
-fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
-
-fig4c <- fig4c_data %>%
-  # total supply greater than 1 tonne (FIXIT update manuscript caption)
-  filter(supply > 1) %>%
-  ggplot(aes(x = imports/1000000, y = shannon_supply, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
-              labels = trans_format("log10", math_format(10^.x))) +
-  annotation_logticks(sides = "b") +
-  scale_color_manual(values = region_colors) +
-  theme_minimal() +
-  labs(x = "Total imports (million t, live weight)",
-       y = "Supply diversity",
-       color = "") +
-  theme(
-    text = element_text(size = 18)
-  )
-
-#
-
-fig4c
-```
-```{r fig4d}
-fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
-
-fig4d <- fig4d_data %>%
-  # total supply greater than 1 tonne (FIXIT update manuscript caption)
-  filter(supply > 1) %>%
-  ggplot(aes(x = exports/1000000, y = shannon_supply, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
-              labels = trans_format("log10", math_format(10^.x))) +
-  annotation_logticks(sides = "b") +
-  scale_color_manual(values = region_colors) +
-  theme_minimal() +
-  labs(x = "Total exports (million t, live weight)",
-       y = "Supply diversity",
-       color = "") +
-  theme(
-    text = element_text(size = 18)
-  )
-
-#
-
-fig4d
-```
-
-```{r fig4, fig.height = 5, fig.width = 7}
-fig4 <- ggarrange(
-  ggarrange(
-    fig4a, fig4b,
-    ncol = 1,
-    nrow = 2,
-    labels = c("a", "b")
-  ),
-  ggarrange(
-    fig4c, fig4d,
-    ncol = 1,
-    nrow = 2,
-    common.legend = TRUE,
-    legend = "bottom",
-    labels = c("c", "d")
-  ),
-  nrow = 1, ncol = 2,
-  widths = c(1, 0.5)
-)
-
-fig4
-
-ggsave(file.path(outdir, "fig4.png"))
-```
-
-```{r fig4_old, fig.height = 7, fig.width = 7}
-fig4 <- ggarrange(fig4a,
-                  fig4b,
-          ggarrange(fig4d,
-                    fig4e,
-                    nrow = 1,
-                    common.legend = TRUE, 
-                    legend = "bottom",
-                    labels = c("c", "d")),
-          nrow = 3,
-          labels = c("a", "b", "")
-          )
-
-fig4
-
-ggsave(file.path(outdir, "fig4.png"))
-```
-
 # Supplementary Figures

### 2023-03-31 ms_results.Rmd
Move supply diversity out of main MS to go into separate analysis

--- 2023-02-07 ms_results.Rmd
@@
 # Directory
-datadir <- "outputs_20230206"
+datadir <- "outputs_20230330"
 
@@
 
-High global diversity of aquatic food production enables high diversity in consumption, which is important for supplying a range of micronutrients, supporting a range of livelihoods, and is thought to support food system resilience (Seekell et al.
-2017).
-However, one concern about food trade is that it homogenizes diets through two related pathways: by introducing foods that make diets more similar across countries and/or by out-competing diverse local foods.
-To explore this issue for aquatic foods, we compared the Shannon diversity index, of production versus imports and and exports (Fig 4a) and domestic supply versus foreign supply (Fig 4b).
-Across regions, import diversity is generally higher than production diversity and export diversity is similar to production diversity across regions (Fig 4a).
-Import diversity exceeds production diversity in `r n_import_exceeds_prod` out of `r nrow(national_supply_diversity)` countries.
-Although the diversity of imported goods tends to be higher than production, diversity of domestic-sourced consumption tends to be higher than  diversity of foreign sourced consumption (Fig 4b).
-Total imports and exports are positively associated with supply diversity across all regions (Fig 4c).
-This suggests trade is contributing new species to many national markets.
-Collectively, our diversity findings point to a need to protect domestic production diversity while still benefiting from the diversity of imported products.
-
 ## Other summary stats
@@
 ](outputs/fig3.png)
-
-![Figure 4: Relationship between trade and supply diversity.
-a) Violin plots of the Shannon diversity of production, imports and exports for 2019.
-b) Violin plots of the Shannon diversity of domestic and foreign consumption for 2019.
-c) Relationship between total imports and overall supply diversity for 2019.
-d) Relationship between total exports and overall supply diversity for 2019.
-Note for figures 4c and 4d, only countries with a supply of at least 1 tonne were included.](outputs/fig4.png)

### 2023-03-31 supply_diversity.Rmd
Move supply diversity out of main MS to go into separate analysis

---
title: "Supply diversity"
author: "Jessica Gephart"
date: "2023-03-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prep_data}
# Figure 4
# Relationship between trade and supply diversity
diversity_dir <- "/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg"
diversity <- read.csv(file.path(diversity_dir, "diversity.csv"))

# Figure 4a
# Time series of the Shannon diversity of imported blue foods and produced blue foods by region.
fig4a_data <- diversity %>%
  filter(region != "Other nei") %>%
  group_by(year, region) %>%
  summarise(shannon_prod = mean(shannon_prod),
            shannon_imports = mean(shannon_imports)) %>%
  pivot_longer(cols = shannon_prod:shannon_imports,
               names_to = "source", values_to = "shannon") %>%
  mutate(source = case_when(
    source == "shannon_prod" ~ "Production diversity",
    source == "shannon_imports" ~ "Import diversity"
  ))

write.csv(fig4a_data, file.path(outdir, "fig4a_data.csv"), row.names = FALSE)

# Figure 4b
# Foreign supply Shannon diversity versus domestic supply Shannon diversity for 2018.
# The black line represents the 1-to-1 line, such that points falling below the line
# indicate countries where the domestic supply diversity is greater than the foreign
# supply diversity.
fig4b_data <- diversity %>%
  filter(year == 2018, region != "Other nei")

write.csv(fig4b_data, file.path(outdir, "fig4b_data.csv"), row.names = FALSE)

# Figure 4c
# Relationship between total imports and overall supply diversity for 2018.
fig4c_data <- diversity %>%
  filter(year == 2018, region != "Other nei")

write.csv(fig4c_data, file.path(outdir, "fig4c_data.csv"), row.names = FALSE)
```

```{r fig4a_old}
fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))

fig4a <- fig4a_data %>%
  ggplot(aes(x = year, y = shannon, color = source)) +
  geom_line() +
  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
  labs(x = "", y = "Shannon index", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~region)

fig4a
```

```{r fig4a}
diversity <- read.csv("/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv")
fig4a <- diversity %>% 
  filter(year == 2019) %>%
  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
  geom_violin() + 
  geom_point(alpha = 0.5) +
  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
  labs(x = "", y = "Shannon index", fill = "") +
  facet_wrap(~region, nrow=1) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_blank(),
        text = element_text(size = 18)
        )

fig4a
```

```{r fig4b_old}
fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))

fig4b <- fig4b_data %>%
  ggplot(aes(x = shannon_domestic, y = shannon_foreign, color = region)) +
    geom_point(alpha = 0.7) +
    geom_abline(slope = 1, intercept = 0) +
  scale_color_manual(values = region_colors) +
  labs(x = "Domestic supply diversity", 
       y = "Foreign supply diversity",
       color = "") +
    theme_minimal()

fig4b
```

```{r fig4b}
fig4b <- diversity %>% 
  filter(year == 2019) %>%
  select(iso3c, region, "Domestic" = "shannon_domestic", "Foreign" = "shannon_foreign") %>% 
  pivot_longer(Domestic:Foreign, names_to = "shannon_var", values_to = "shannon_index") %>%
  mutate(shannon_var = factor(shannon_var, levels = c("Domestic", "Foreign"))) %>%
  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
  geom_violin() + 
  geom_point(alpha = 0.5) +
  scale_fill_manual(values = c(habitat_source_colors[2], habitat_source_colors[4])) +
  labs(x = "", y = "Shannon index", fill = "") +
  facet_wrap(~region, nrow=1) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_blank(),
        text = element_text(size = 18))

fig4b
```

```{r fig4c, eval=FALSE}
fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))

fig4c <- fig4c_data %>%
  # total supply greater than 1 tonne (FIXIT update manuscript caption)
  filter(supply > 1) %>%
  ggplot(aes(x = imports/1000000, y = shannon_supply, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides = "b") +
  scale_color_manual(values = region_colors) +
  theme_minimal() +
  labs(x = "Total imports (million t, live weight)",
       y = "Supply diversity",
       color = "") +
  theme(
    text = element_text(size = 18)
  )

#

fig4c
```
```{r fig4d, eval=FALSE}
fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))

fig4d <- fig4d_data %>%
  # total supply greater than 1 tonne (FIXIT update manuscript caption)
  filter(supply > 1) %>%
  ggplot(aes(x = exports/1000000, y = shannon_supply, color = region)) +
  geom_point(alpha = 0.7) +
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
              labels = trans_format("log10", math_format(10^.x))) +
  annotation_logticks(sides = "b") +
  scale_color_manual(values = region_colors) +
  theme_minimal() +
  labs(x = "Total exports (million t, live weight)",
       y = "Supply diversity",
       color = "") +
  theme(
    text = element_text(size = 18)
  )

#

fig4d
```

```{r fig4, fig.height = 5, fig.width = 7, eval=FALSE}
fig4 <- ggarrange(
  ggarrange(
    fig4a, fig4b,
    ncol = 1,
    nrow = 2,
    labels = c("a", "b")
  ),
  ggarrange(
    fig4c, fig4d,
    ncol = 1,
    nrow = 2,
    common.legend = TRUE,
    legend = "bottom",
    labels = c("c", "d")
  ),
  nrow = 1, ncol = 2,
  widths = c(1, 0.5)
)

fig4

ggsave(file.path(outdir, "fig4.png"))
```

```{r fig4_old, fig.height = 7, fig.width = 7, eval=FALSE}
fig4 <- ggarrange(fig4a,
                  fig4b,
          ggarrange(fig4d,
                    fig4e,
                    nrow = 1,
                    common.legend = TRUE, 
                    legend = "bottom",
                    labels = c("c", "d")),
          nrow = 3,
          labels = c("a", "b", "")
          )

fig4

ggsave(file.path(outdir, "fig4.png"))
```

High global diversity of aquatic food production enables high diversity in consumption, which is important for supplying a range of micronutrients, supporting a range of livelihoods, and is thought to support food system resilience (Seekell et al. 2017). However, one concern about food trade is that it homogenizes diets through two related pathways: by introducing foods that make diets more similar across countries and/or by out-competing diverse local foods. To explore this issue for aquatic foods, we compared the Shannon diversity index, of production versus imports and and exports (Fig 4a) and domestic supply versus foreign supply (Fig 4b). Across regions, import diversity is generally higher than production diversity and export diversity is similar to production diversity across regions (Fig 4a). Import diversity exceeds production diversity in `r n_import_exceeds_prod` out of `r nrow(national_supply_diversity)` countries. Although the diversity of imported goods tends to be higher than production, diversity of domestic-sourced consumption tends to be higher than  diversity of foreign sourced consumption (Fig 4b). Total imports and exports are positively associated with supply diversity across all regions (Fig 4c). This suggests trade is contributing new species to many national markets. Collectively, our diversity findings point to a need to protect domestic production diversity while still benefiting from the diversity of imported products.

![Figure 4: Relationship between trade and supply diversity. a) Violin plots of the Shannon diversity of production, imports and exports for 2019. b) Violin plots of the Shannon diversity of domestic and foreign consumption for 2019. c) Relationship between total imports and overall supply diversity for 2019. d) Relationship between total exports and overall supply diversity for 2019. Note for figures 4c and 4d, only countries with a supply of at least 1 tonne were included.](outputs/fig4.png)

```{r}
national_supply_diversity <- read.csv(file.path(datadir, "fig4b_data.csv"))

n_import_exceeds_prod <- national_supply_diversity %>%
  filter(shannon_imports > shannon_prod) %>%
  nrow()

```

### 2023-03-31 prep_data.R

@@ -17,7 +17,7 @@ library(tidytext)
 # Directory listing
 
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
-outdir <- "outputs_20230206"
+outdir <- "outputs_20230330"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -88,7 +88,7 @@ write.csv(prod, file.path(outdir, "prod.csv"), row.names = FALSE)
 write.csv(artis_fm, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
 
 # Write out the bilater non-food trade data to use in supply calculations
-write.csv(artis_nonfood, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
+write.csv(artis_nonfood, file.path(outdir, "artis_nonfood.csv"), row.names = FALSE)
 
 #-------------------------------------------------------------------------------
 # Summarize cleaned FAO production data
@@ -100,7 +100,7 @@ prod <- prod %>%
   summarise(production_t = sum(production_t, na.rm = TRUE)) %>%
   ungroup()
 
-# Join production data to ARTIS
+
 artis <- artis %>%
   filter(live_weight_t >= 0.1) %>%
   # Add importer and exporter region
@@ -449,46 +449,6 @@ fig3d_data <- supply %>%
   mutate(supply_source = gsub("supply_", "", supply_source))
 
 write.csv(fig3d_data, file.path(outdir, "fig3d_data.csv"), row.names = FALSE)
-
-#-------------------------------------------------------------------------------
-# Figure 4
-# Relationship between trade and supply diversity
-diversity_dir <- "/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg"
-diversity <- read.csv(file.path(diversity_dir, "diversity.csv"))
-
-# Figure 4a
-# Time series of the Shannon diversity of imported blue foods and produced blue foods by region.
-fig4a_data <- diversity %>%
-  filter(region != "Other nei") %>%
-  group_by(year, region) %>%
-  summarise(shannon_prod = mean(shannon_prod), 
-            shannon_imports = mean(shannon_imports)) %>%
-  pivot_longer(cols = shannon_prod:shannon_imports,
-               names_to = "source", values_to = "shannon") %>%
-  mutate(source = case_when(
-    source == "shannon_prod" ~ "Production diversity",
-    source == "shannon_imports" ~ "Import diversity"
-  ))
-
[…]

### 2023-03-31 ms_figures.Rmd

--- 2023-02-08 ms_figures.Rmd
@@
 ggsave(file.path(outdir, "fig3.png"))
-```
-
-```{r fig4a_old, eval = FALSE}
-fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))
-
-fig4a <- fig4a_data %>%
-  ggplot(aes(x = year, y = shannon, color = source)) +
-  geom_line() +
-  scale_color_manual(values = c(c("#F37542", "#0F2D59"))) +
-  labs(x = "", y = "Shannon index", color = "") +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  facet_wrap(~region)
-
-fig4a
-```
-
-```{r fig4a}
-diversity <- read.csv("/Volumes/jgephart/ARTIS/Outputs/diversity/max_per_capita_100kg/diversity.csv")
-fig4a <- diversity %>% 
-  filter(year == 2019) %>%
-  select(iso3c, region, "Production" = "shannon_prod", "Imports" = "shannon_imports", "Exports" = "shannon_exports") %>% 
-  pivot_longer(Production:Exports, names_to = "shannon_var", values_to = "shannon_index") %>%
-  mutate(shannon_var = factor(shannon_var, levels = c("Production", "Imports", "Exports"))) %>%
-  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
-  geom_violin() + 
-  geom_point(alpha = 0.5) +
-  scale_fill_manual(values = c(habitat_source_colors[1], habitat_source_colors[3], habitat_source_colors[4])) +
-  labs(x = "", y = "Shannon index", fill = "") +
-  facet_wrap(~region, nrow=1) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank(),
-        text = element_text(size = 18)
-        )
-
-fig4a
-```
-
-```{r fig4b_old, eval=FALSE}
-fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))
-
-fig4b <- fig4b_data %>%
-  ggplot(aes(x = shannon_domestic, y = shannon_foreign, color = region)) +
-    geom_point(alpha = 0.7) +
-    geom_abline(slope = 1, intercept = 0) +
-  scale_color_manual(values = region_colors) +
-  labs(x = "Domestic supply diversity", 
-       y = "Foreign supply diversity",
-       color = "") +
-    theme_minimal()
-
-fig4b
-```
-
-```{r fig4b}
-fig4b <- diversity %>% 
-  filter(year == 2019) %>%
-  select(iso3c, region, "Domestic" = "shannon_domestic", "Foreign" = "shannon_foreign") %>% 
-  pivot_longer(Domestic:Foreign, names_to = "shannon_var", values_to = "shannon_index") %>%
-  mutate(shannon_var = factor(shannon_var, levels = c("Domestic", "Foreign"))) %>%
-  ggplot(aes(x = shannon_var, y = shannon_index, fill = shannon_var)) +
-  geom_violin() + 
-  geom_point(alpha = 0.5) +
-  scale_fill_manual(values = c(habitat_source_colors[2], habitat_source_colors[4])) +
-  labs(x = "", y = "Shannon index", fill = "") +
-  facet_wrap(~region, nrow=1) +
-  theme_bw() +
-  theme(legend.position = "bottom",
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank(),
-        text = element_text(size = 18))
-
-fig4b
-```
-
-```{r fig4c}
-fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
-
-fig4c <- fig4c_data %>%
-  # total supply greater than 1 tonne (FIXIT update manuscript caption)
-  filter(supply > 1) %>%
-  ggplot(aes(x = imports/1000000, y = shannon_supply, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
-              labels = trans_format("log10", math_format(10^.x))) +
-  annotation_logticks(sides = "b") +
-  scale_color_manual(values = region_colors) +
-  theme_minimal() +
-  labs(x = "Total imports (million t, live weight)",
-       y = "Supply diversity",
-       color = "") +
-  theme(
-    text = element_text(size = 18)
-  )
-
-#
-
-fig4c
-```
-```{r fig4d}
-fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
-
-fig4d <- fig4d_data %>%
-  # total supply greater than 1 tonne (FIXIT update manuscript caption)
-  filter(supply > 1) %>%
-  ggplot(aes(x = exports/1000000, y = shannon_supply, color = region)) +
-  geom_point(alpha = 0.7) +
-  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
-              labels = trans_format("log10", math_format(10^.x))) +
-  annotation_logticks(sides = "b") +
-  scale_color_manual(values = region_colors) +
-  theme_minimal() +
-  labs(x = "Total exports (million t, live weight)",
-       y = "Supply diversity",
-       color = "") +
-  theme(
-    text = element_text(size = 18)
-  )
-
-#
-
-fig4d
-```
-
-```{r fig4, fig.height = 5, fig.width = 7}
-fig4 <- ggarrange(
-  ggarrange(
-    fig4a, fig4b,
-    ncol = 1,
-    nrow = 2,
-    labels = c("a", "b")
-  ),
-  ggarrange(
-    fig4c, fig4d,
-    ncol = 1,
-    nrow = 2,
-    common.legend = TRUE,
-    legend = "bottom",
-    labels = c("c", "d")
-  ),
-  nrow = 1, ncol = 2,
-  widths = c(1, 0.5)
-)
-
-fig4
-
-ggsave(file.path(outdir, "fig4.png"))
-```
-
-```{r fig4_old, fig.height = 7, fig.width = 7, eval = FALSE}
-fig4 <- ggarrange(fig4a,
-                  fig4b,
-          ggarrange(fig4d,
-                    fig4e,
-                    nrow = 1,
-                    common.legend = TRUE, 
-                    legend = "bottom",
-                    labels = c("c", "d")),
-          nrow = 3,
-          labels = c("a", "b", "")
-          )
-
-fig4
-
-ggsave(file.path(outdir, "fig4.png"))
 ```
@@
 
-ggsave("dom_source_2019_overview.png")
+ggsave(file.path(outdir, "dom_source_2019_overview.png"))
 ```
@@
 
-ggsave("foreign_exports_2019_by_source.png")
-```
+ggsave(file.path(outdir, "foreign_exports_2019_by_source.png"))
+```

### 2023-03-31 supply_diversity.Rmd

@@
 
-```{r fig4a_old}
+```{r fig4a_old, eval = FALSE}
 fig4a_data <- read.csv(file.path(datadir, "fig4a_data.csv"))
@@
 
-```{r fig4b_old}
+```{r fig4b_old, eval=FALSE}
 fig4b_data <- read.csv(file.path(datadir, "fig4b_data.csv"))
@@
 
-```{r fig4c, eval=FALSE}
+```{r fig4c}
 fig4c_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
@@
 ```
-```{r fig4d, eval=FALSE}
+```{r fig4d}
 fig4d_data <- read.csv(file.path(datadir, "fig4c_data.csv"))
@@
 
-```{r fig4, fig.height = 5, fig.width = 7, eval=FALSE}
+```{r fig4, fig.height = 5, fig.width = 7}
 fig4 <- ggarrange(
@@
 
-```{r fig4_old, fig.height = 7, fig.width = 7, eval=FALSE}
+```{r fig4_old, fig.height = 7, fig.width = 7, eval = FALSE}
 fig4 <- ggarrange(fig4a,

### 2023-03-31 prep_data.R
prep data working with most current snet and consumption files 20230328

@@ -16,19 +16,19 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20220111"
-outdir <- "outputs_20230330"
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/2023031"
+outdir <- "outputs_20230331"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
 
 # Database connection
 con <- dbConnect(RPostgres::Postgres(),
-                 dbname=Sys.getenv("POSTGRES_DB"),
+                 dbname=Sys.getenv("DB_NAME"),
                  host="localhost",
                  port="5432",
-                 user=Sys.getenv("POSTGRES_USER"),
-                 password=Sys.getenv("POSTGRES_PASSWORD"))
+                 user=Sys.getenv("DB_USERNAME"),
+                 password=Sys.getenv("DB_PASSWORD"))
 
 # Check that connection is established by checking which tables are present
 dbListTables(con)
@@ -179,7 +179,7 @@ prod_by_source <- prod %>%
 write.csv(prod_by_source, file.path(outdir, "prod_by_habitat_method.csv"), row.names = FALSE)
 
 dom_export_by_source <- artis %>% 
-  filter(dom_source == "domestic export") %>%
+  filter(dom_source == "domestic") %>%
   group_by(year, habitat_method, environment) %>%
   summarise(dom_live_weight_t = sum(live_weight_t))
 
@@ -332,8 +332,8 @@ calculate_supply <- function(artis_data, production_data){
 
 # Supply / Consumption data
 #consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_max_per_capita_100kg"
-#consumption_dir <- "../ARTIS/consumption/consumption_max_per_capita_100kg"
-supply <- read.csv("summary_consumption.csv")
+consumption_dir <- "../ARTIS/qa/consumption_20230328"
+supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
 
 supply <- supply %>%
   filter(!is.na(habitat) & !is.na(method)) %>%
@@ -604,6 +604,7 @@ top_regional_import_change <- bilateral_habitat_method_summary %>%
   ungroup() %>%
[…]

### 2023-03-31 ms_figures.Rmd
update file paths

@@
 # Set directories
-datadir <- "outputs_20230330"
-outdir <- "outputs_20230330"
+datadir <- "outputs"
+outdir <- "outputs"
 ```
@@
   geom_line() +
-  scale_color_manual(values = c("#D38F35", "#114F59")) +
+  scale_color_manual(values = c("#D38F35", "#114F59", "grey")) +
   theme_bw() +
@@
 ```
+
+```{r, eval=FALSE}
+# Plots moved from end of prep_data
+top_processing_countries %>%
+  filter(year == 2019) %>%
+  ggplot(aes(x = live_weight_t, y = reorder(exporter_iso3c, ranking))) +
+  geom_bar(stat = "identity") +
+  scale_y_reordered() +
+  theme_bw() +
+  labs(x = "Re-exports (Live weight tonnes)", y = "Processing Country")
+
+top_processing_countries_by_source %>%
+  filter(year == 2019) %>%
+  ggplot(aes(x = live_weight_t / 1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
+  geom_bar(stat = "identity") +
+  scale_fill_manual(values = habitat_source_colors) +
+  scale_y_reordered() +
+  theme_bw() +
+  labs(x = "Re-exports (Live weight million tonnes)", y = "Processing Country") +
+  facet_wrap(~habitat_method, scales = "free_y")
+
+```

### 2023-03-31 prep_data.R

@@ -17,7 +17,7 @@ library(tidytext)
 # Directory listing
 
 outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/2023031"
-outdir <- "outputs_20230331"
+outdir <- "outputs"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -287,52 +287,9 @@ write.csv(artis_region, file.path(outdir, "artis_region.csv"), row.names = FALSE
 #-------------------------------------------------------------------------------
 # Figure 3
 
-# Create calculate_supply function (small difference from version in explore_artis)
-calculate_supply <- function(artis_data, production_data){
-  exports <- artis_data %>%
-    group_by(year, exporter_iso3c, dom_source, sciname, method, habitat) %>%
-    summarise(live_weight_t = sum(live_weight_t)) %>%
-    mutate(dom_source = str_replace(dom_source, " ", "_")) %>%
-    pivot_wider(names_from = "dom_source", values_from = "live_weight_t")
-  
-  if (!("error_export" %in% colnames(exports))) {
-    exports$error_export <- NA
-  }
-  
-  imports <- artis_data %>%
-    group_by(year, importer_iso3c, sciname, method, habitat) %>%
-    summarise(imports_live_weight_t = sum(live_weight_t))
-  
-  supply <- exports %>%
-    full_join(imports, by = c("year", "exporter_iso3c" = "importer_iso3c", 
-                              "sciname", "method", "habitat")) %>%
-    full_join(production_data %>% 
-                rename(production_t = live_weight_t), by = c("year", "exporter_iso3c" = "iso3c", 
-                                                             "sciname", "method", "habitat")) 
-  
-  supply$foreign_export[is.na(supply$foreign_export)] <- 0
-  supply$domestic_export[is.na(supply$domestic_export)] <- 0
-  supply$error_export[is.na(supply$error_export)] <- 0
-  supply$imports_live_weight_t[is.na(supply$imports_live_weight_t)] <- 0
-  supply$production_t[is.na(supply$production_t)] <- 0
-  
-  supply <- supply %>%
[…]

### 2023-04-20 ms_figures.Rmd
Update results with new text

@@
 ```{r ldc_net_exports_habitat_method}
-ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv"))
+ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv")) %>%
+  select(year, habitat_method, net_export)
+
+ldc_net_total <- ldc_net_habitat_method_summary %>%
+  group_by(year) %>%
+  summarise(net_export = sum(net_export)) %>%
+  mutate(habitat_method = "total")
+
+ldc_net_habitat_method_summary <- ldc_net_habitat_method_summary %>%
+  bind_rows(ldc_net_total)
 
@@
   ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
-  geom_line() +
+  geom_line(size = 1) +
   scale_color_manual(values = habitat_source_colors) +
@@
 dom_source_ts_p <- dom_source_ts %>%
-  filter(dom_source != "error export") %>%
   ggplot(aes(x = year, y = live_weight_t / 1000000, color = dom_source)) +
   geom_line() +
-  scale_color_manual(values = c("#D38F35", "#114F59", "grey")) +
+  scale_color_manual(values = c("#D38F35", "grey40", "#114F59")) +
   theme_bw() +

### 2023-04-20 ms_results.Rmd
update file paths
Update results with new text

@@
 # Directory
-datadir <- "outputs_20230330"
+datadir <- "outputs"
 
@@
 Globalization describes the degree of international connectedness, which can be characterized by increasing flows of input, intermediate and final products among countries.
-Aquatic food exports increased `r round(100*((total_2019-total_1996)/total_1996),0)`%  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019.
+Aquatic food exports more than doubled from 1996-2019  (from `r total_1996` to `r total_2019` mil t; Fig 1a)  from 1996-2019.
 Over that period, both farmed and wild exports increased, though aquaculture grew faster, more than tripling, whereas capture exports grew by `r round(100*((total_capture_2019-total_capture_1996)/total_capture_1996),0)`%.
@@
 Another measure of the degree of globalization is the share of production exported.
-Domestic exports increased to `r round(percent_export_total_dom_2019, 1)`% of production in 2019, up from `r round(percent_export_total_dom_1996, 1)`% in 1996.
-while total exports reached 33% of all production in 2019.
+Domestic exports increased from `r round(percent_export_total_dom_1996, 1)`% to `r round(percent_export_total_dom_2019, 1)`% between 1996 and 2019, while total exports reached `r round(percent_export_total_2019, 1)`% of all production in 2019.
 For comparison, the share of cereal production exported grew from around 10% in the late 1990s to 17% in the 2020s (Clapp 2022).
-Increasing marine capture export volumes despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and experienced the largest increases in the share exported, and largest increases in the share exported (Fig 2b).
-Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, leading to an increasing share exported (Fig 2b).
+Increasing marine capture exports despite stagnating production resulted in marine capture products having the greatest share of production destined for export (`r round(percent_marine_capture_exported_2019, 1)`% of production) and the largest increases in the share exported (Fig 2b).
+Aquaculture production more than doubled from 1996 to 2019, but aquaculture exports grew even faster, increasing the share exported (Fig 2b).
 Despite increases, inland aquaculture still had the lowest share of production destined for export in 2019  (domestic exports represented only `r round(percent_inland_aqua_exported_2019, 1)`% of production) (Fig 1b).
-This finding clarifies prior debates about the orientation of aquaculture, while export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al.
+This finding clarifies standing debates about the orientation of aquaculture and export trends suggest a need to consider international markets when crafting nutrition-sensitive policies (Golden et al.
 2017; Belton et al 2018; Gephart et al.
@@
 2016).
-From 1996-2019, aquatic food trade became more connected with the average number of export partners increasing `r round(100*((all_degree_2019-all_degree_1996)/all_degree_1996),0)`%  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 4c).
+By disaggregating aquatic food trade, we can evaluate the structural features of aquatic food trade associated with resilience to shocks.
+First, we found that aquatic food trade became more connected with the average number of export partners nearly doubling from 1996-2019  (from `r round(all_degree_1996, 1)` to `r round(all_degree_2019, 1)`; Fig 1c).
 Marine capture networks are most highly connected, followed by marine aquaculture, with inland capture and aquaculture trade being the least connected.
 
-Since 1996, aquatic food exports and imports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
+Since 1996, aquatic food exports have become moderately less concentrated, with only `r export_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of exports in 1996 versus `r export_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019.
 Compare this with crops where just 7 countries and the EU account for 90% of wheat exports and just four countries accounting >80% of maize exports (Clapp 2022).
-The declining concentration is driven by capture fishery exports, whereas both inland and marine aquaculture exports became somewhat more concentrated (Fig 2d).
+Declining concentration is driven by capture fishery exports, whereas aquaculture exports became somewhat more concentrated (Fig 1d).
 Aquaculture exports are generally concentrated among fewer countries than capture exports, corresponding to the high concentration of aquaculture production to a few key regions.
@@
 Though aquatic food production, distribution, and consumption remain highly uneven (Hicks et al.
-2022), we found declinding import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e).
+2022), we found declining import concentration, with `r import_concentration_n_countries %>% filter(year == 1996) %>% pull(n)` countries comprising `r cumulative_percent_threshold`% of imports in 1996 versus `r import_concentration_n_countries %>% filter(year == 2019) %>% pull(n)` countries in 2019 (Fig 2e).
 The more dispersed import patterns are likely associated with growing populations and expanding middle classes and urbanization, particularly in low- and middle-income countries, which are responsible for increasing aquatic food demand in many countries (Naylor et al.
@@
 
-Given the geographic patchiness of wild fish and aquaculture production, trade helps meet aquatic food demand in countries.
+Given the geographic patchiness of capture and aquaculture production, trade helps meet aquatic food demand in countries.
 Aquatic food imports are especially important where per capita demand is rising, aquaculture is limited, and wild fishery catch is stagnant.
 Corresponding to the geographical variability, we find the top importers, exporters, and bilateral flows to vary by habitat and farmed versus wild source, underscoring the importance of disaggregating trade (Fig S1-3).
-For example, Asia and Europe, and to a lesser extent, North America dominated marine capture and aquaculture trade networks in recent years (Fig S1).
-Since 1996, marine trade shifted toward Asia and Europe, with large increases in marine capture imports to and exports from Africa (Fig 2b).
+For example, Asia and Europe, and to a lesser extent, North America recently dominated marine capture and aquaculture trade networks whereas Asia dominates inland aquatic food trade  (Fig S1).
 
-At the country level, although some countries rank among the top traders across production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one production method.
+At the country level, although some countries rank among the top traders across all production methods, such as China for exports and the United Stated for imports, many countries are only top traders for one.
+China and Russia are the top marine capture aquatic food exporters, with China and the United States as the top importers (Fig S2-3).
+Meanwhile, Norway and Chile rank highest in marine aquaculture exports, with the United States and Japan leading imports (Fig S2-3).
 Inland aquatic food trade is dominated by aquaculture, with the highest exports from Vietnam and China and highest imports by the United States, Japan, and South Korea (Fig S2-3).
-In general, inland farmed and wild production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).
+In general, inland production is oriented more toward domestic consumption and what is exported tends to stay within the region, particularly within Asia (Fig 2; Fig S1, Fig S6).
 
@@
 2015).
-We find that this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America.
-Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America.
-At the country level, trade increased between `r round(percent_flow_increases, 1)`% of trade pairs.
-Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa outpaced interregional trade with those regions.
-The largest average annual growth increases occurred for trade within Asia and within Europe, followed by trade between Europe and Asia (Fig 2a; Fig S6).
+We find this pattern generally holds for aquatic food trade as intraregional trade is the highest or second highest for all regions other than Oceania and South America (Fig 2b).
+Since 1996, trade increased or remained approximately stable between nearly all regional trade pairs, other than within North America (Fig S6).
+At the country level, trade increased between two thirds of trade pairs.
+Despite trade increasing with partners across the globe, trade within Asia, Europe and Africa grew faster.
+The largest average annual growth increases occurred for trade within Asia and Europe, followed by trade between Europe and Asia (Fig 2b; Fig S6).
 Our trade estimates are ultimately from reported trade and therefore do not capture informal and unreported trade networks.
@@
 habitatmethod_supply_trends_summary <- regional_supply_per_cap %>%
-  select(year, habitat_method, supply_per_cap) %>%
+  select(year, habitat_method, supply, pop) %>%
   group_by(year, habitat_method) %>% 
-  summarise(supply_per_cap = sum(supply_per_cap, na.rm = TRUE)) %>%
+  summarise(supply = sum(supply, na.rm = TRUE),
+            pop = sum(pop, na.rm = TRUE)) %>%
+  mutate(supply_per_cap = 1000 * supply/pop) %>%
+  select(-supply, -pop) %>%
   filter(year %in% c(1996, 2019), habitat_method != "unknown") %>%
@@
 
-```{r}
-national_supply_diversity <- read.csv(file.path(datadir, "fig4b_data.csv"))
-
-n_import_exceeds_prod <- national_supply_diversity %>%
-  filter(shannon_imports > shannon_prod) %>%
-  nrow()
-
-```
-
 ## Text for MS
@@
 
-Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019.
+Globally, annual aquatic food apparent consumption increased from `r signif(global_supply_1996, digit = 3)` kg per capita in 1996 to `r signif(global_supply_2019, digit = 3)` kg per capita in 2019 (Fig 3a).
 Our estimates are slightly lower than FAOSTAT, which reports global aquatic food consumption at 15.6 kg/capita/year in 1996 and 20.7 kg/capita/year in 2019.
-We found that aquatic food consumption increased across all regions outside of North America, which was relatively stable and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b).
-Global increase were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively.
-Meanwhile, inland capture consumption grew modestly (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, a `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% increase), while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019).
+We found aquatic food consumption increased across all regions outside of North America, which was relatively stable, and South America, where aquatic food consumption declined `r -1*signif(regional_supply_trends_summary %>% filter(region == "South America") %>% pull(percent_diff), digit = 3)`% (Fig 3b).
+Global increases were driven by inland and marine aquaculture, which increased by `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019) and `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine aquaculture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019), respectively.
+Meanwhile, inland capture consumption grew from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "inland capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019, while marine capture consumption declined `r -1*signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(percent_diff), digits = 3)`% (from `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year1996), digits = 3)` kg/capita/year in 1996 to `r signif(habitatmethod_supply_trends_summary %>% filter(habitat_method == "marine capture") %>% pull(year2019), digits = 3)` kg/capita/year in 2019).
 Nevertheless, capture still makes up `r signif(cap_supply_percent, digits = 2)`% of global aquatic food consumption, with its contribution to regional aquatic food consumption ranging from `r signif(oceania_cap_supply_percent, digits = 2)`% in Oceania to `r signif(asia_cap_supply_percent, digits = 2)`% in Asia, where farmed consumption overtook wild consumption in `r asia_farmed_year`.

### 2023-04-20 supply_diversity.Rmd
Update results with new text

@@
 ```
+
+# Methods
+
+## Estimating supply diversity
+	We separately calculated the Shannon diversity for the overall supply, the diversity of supply sourced from domestic products, and the supply sourced from foreign products.
+Additionally, we calculated the Shannon diversity of domestic production, imports, and exports.
+One notable limitation of applying this measure is that instances where production or trade is only estimated at the genus or higher level will be interpreted in the calculation as a single species.
+Consequently, countries with low species resolution will generally appear to have less diverse production, limiting the ability to compare supply diversity across countries.
+To address this issue, we first matched scientific groups to species based on data in FishBase50 and SealifeBase53.
+We then calculated the global production distribution within each phylogenetic group based on a lognormal distribution of produced species.
+Proportions for each potential species within a phylogenetic group were derived from the lognormal global distribution.
+These proportions were matched to species in two ways.
+For species that had been previously reported by a country they were matched in order by the volume reported for that year and reported species were followed by the unreported species.
+This meant that, for a given year, the highest proportion derived from the lognormal distribution was matched to the species that the country had reported the largest volume for

### 2023-04-20 prep_data.R

@@ -288,12 +288,13 @@ write.csv(artis_region, file.path(outdir, "artis_region.csv"), row.names = FALSE
 # Figure 3
 
 # Supply / Consumption data
-consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_20230328"
-#consumption_dir <- "../ARTIS/qa/consumption_20230328"
+consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_20230331"
+#consumption_dir <- "../ARTIS/qa/consumption_20230331"
 supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
 
 supply <- supply %>%
-  filter(!is.na(habitat) & !is.na(method)) %>%
+  # Remove observations where supply is 0
+  filter(supply > 0) %>%
   mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
   mutate(habitat_method = case_when(
     str_detect(habitat_method, "unknown") ~ "unknown", 
@@ -333,7 +334,6 @@ write.csv(global_supply_per_cap, file.path(outdir, "global_supply_per_cap.csv"),
 # Figure 3a
 # Calculate per capita supply by region and source
 supply_total <- supply %>%
-  filter(!is.na(habitat_method)) %>%
   group_by(iso3c, region, year, habitat_method) %>%
   summarise(supply = sum(supply),
             supply_domestic = sum(supply_domestic),
@@ -343,7 +343,6 @@ supply_total <- supply %>%
   ungroup()
 
 fig3a_data <- supply_total %>%
-  filter(!is.na(habitat_method)) %>%
   filter(region != "Other nei") %>%
   group_by(year, habitat_method) %>%
   summarize(supply = sum(supply)) %>%

### 2023-04-20 prep_data.R
fig3c min and max bands

@@ -16,7 +16,7 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/2023031"
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20230420"
 outdir <- "outputs"
 
 #-------------------------------------------------------------------------------
@@ -291,6 +291,8 @@ write.csv(artis_region, file.path(outdir, "artis_region.csv"), row.names = FALSE
 consumption_dir <- "/Volumes/jgephart/ARTIS/Outputs/consumption/consumption_20230331"
 #consumption_dir <- "../ARTIS/qa/consumption_20230331"
 supply <- read.csv(file.path(consumption_dir, "summary_consumption.csv"))
+supply_min <- read.csv(file.path(consumption_dir, "consumption_min", "summary_consumption.csv"))
+supply_max <- read.csv(file.path(consumption_dir, "consumption_max", "summary_consumption.csv"))
 
 supply <- supply %>%
   # Remove observations where supply is 0
@@ -307,7 +309,49 @@ supply <- supply %>%
   rename(supply_domestic = domestic_consumption_t,
          supply_foreign = foreign_consumption_t)
 
+supply_min <- supply_min %>%
+  # Remove observations where supply is 0
+  filter(supply > 0) %>%
+  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
+  mutate(habitat_method = case_when(
+    str_detect(habitat_method, "unknown") ~ "unknown", 
+    TRUE ~ habitat_method
+  )) %>%
+  # Set factor levels
+  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
+                                                            "marine aquaculture", "inland aquaculture",
+                                                            "unknown" ))) %>%
+  rename(supply_domestic = domestic_consumption_t,
+         supply_foreign = foreign_consumption_t)
+
+supply_max <- supply_max %>%
+  # Remove observations where supply is 0
+  filter(supply > 0) %>%
+  mutate(habitat_method = paste(habitat, method, sep = " ")) %>%
[…]

### 2023-04-20 ms_figures.Rmd
fig3c min max geom ribbon

@@
 
-fig3c <- fig3c_data %>%
-  ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
-  geom_line() +
+fig3c <- ggplot() +
+  geom_ribbon(data = fig3c_data %>%
+                filter(supply_source == "domestic"),
+              aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
+  geom_ribbon(data = fig3c_data %>%
+                filter(supply_source == "foreign"),
+              aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
+  geom_line(data = fig3c_data, aes(x = year, y = supply_percent, color = supply_source)) +
   scale_color_manual(values = c("#F37542", "#0F2D59")) +
@@
 ggplot() +
-  geom_line(data = artis_ts, aes(x = year, y = live_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
-  geom_line(data = custom_ts, aes(x = year, y = live_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
+  geom_line(data = artis_ts, aes(x = year, y = product_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
+  geom_line(data = custom_ts, aes(x = year, y = product_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
   scale_color_manual(

### 2023-04-28 ms_figures.Rmd
add lims to fig 3 and update text

--- 2023-04-20 ms_figures.Rmd
@@
   labs(x = "", y = "Percent of supply", color = "Origin") +
+  lims(y = c(0,100)) +
   theme_bw() +
@@
   facet_wrap(~region) +
+  lims(y = c(0,100)) +
   theme_bw() +

### 2023-04-28 prep_data.R
foreign exports check and min max ranges

@@ -90,6 +90,69 @@ write.csv(artis_fm, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
 # Write out the bilater non-food trade data to use in supply calculations
 write.csv(artis_nonfood, file.path(outdir, "artis_nonfood.csv"), row.names = FALSE)
 
+hs_versions <- c("96", "02", "07", "12", "17")
+datadir <- "../ARTIS/qa/model_inputs_20221129_NEW"
+
+hs_clade_match <- data.frame()
+for (i in 1:length(hs_versions)) {
+  curr_hs <- hs_versions[i]
+  print(curr_hs)
+  
+  # Determine most specific clade of each HS code (but if clade is not reported in production data (i.e., hs_taxa_match$SciName), return NA)
+  # To match to clade, even if not reported in production data, set match_to_prod to FALSE
+  curr_hs_clade_match <- match_hs_to_clade(hs_taxa_match = read.csv(file.path(datadir, paste("hs-taxa-match_HS", curr_hs, ".csv", sep = ""))) %>%
+                                             select(-c(sciname_habitat, code_habitat)),
+                                           prod_taxa_classification = read.csv(file.path(datadir, "clean_fao_taxa.csv")),
+                                           match_to_prod = FALSE) %>%
+    # pad HS codes with zeroes
+    mutate(Code = as.character(Code)) %>%
+    mutate(Code = if_else(str_detect(Code, "^30"), true = str_replace(Code, pattern = "^30", replacement = "030"),
+                          if_else(str_detect(Code, "^511"), true = str_replace(Code, pattern = "^511", replacement = "0511"),
+                                  false = Code)))# %>%
+  # # Filter down to just the codes in cc_m (recreated above from country_est)
+  # filter(Code %in% cc_m)
+  
+  curr_hs_clade_match <- curr_hs_clade_match %>%
+    mutate(hs_version = paste("HS", curr_hs, sep = ""))
+  
+  hs_clade_match <- hs_clade_match %>%
+    bind_rows(curr_hs_clade_match)
+}
+
+hs_clade_scinames <- hs_clade_match %>%
+  pull(hs_clade) %>%
+  unique()
+
+#-------------------------------------------------------------------------------
+
+# Finding number of foreign exports that exceeded 100% imports
[…]

### 2023-04-28 ms_results.Rmd

@@
 
-Since direct measurements of human consumption (e.g.
-dietary intake) are not collected globally, consumption is generally represented by apparent consumption.
-Apparent consumption is generally calculated as production plus imports minus exports and waste.
-Trade is therefore central to estimating consumption and historically limited understanding of aquatic food consumption patterns.
-By estimating species level trade, we are able to estimate apparent consumption of aquatic foods by species, production method, and geographical origin.
+Since direct measurements of human food consumption (e.g., dietary intake) are not collected globally, it is often represented by apparent consumption.
+Apparent consumption is calculated as production plus imports minus exports and waste.
+Trade is therefore central to estimating consumption and has historically limited understanding of aquatic food consumption patterns.
+By estimating species level trade, we estimate apparent consumption of aquatic foods by species/species group, production method, and geographical origin.
 
@@
 
-Since we estimate exports from domestic and foreign sources, we can track changes in reliance on foreign sourced products.
-Globally the share of consumption that is foreign-sourced increased modestly, from `r signif(foreign_consumption_start, digits = 3)`% to `r signif(foreign_consumption_end, digits = 3)`% (Fig 3c).
-Patterns vary greatly across regions with Asia and S America dominated by domestic supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Asia", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% and `r signif(foreign_consumption_region %>% filter(year == 2019, region == "South America", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019, respectively), but Europe dominated by foreign supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Europe", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019) (Fig 3d).
+Estimating the foreign versus domestic source of consumption requires identifying the share of production retained in the country and tracking products that undergo foreign processing but are imported again.
+By estimating the source of traded aquatic foods, we can therefore track changes in reliance on foreign-sourced products.
+Globally the share of foreign-sourced consumption increased modestly, from `r signif(foreign_consumption_start, digits = 3)`% in 1996 to `r signif(foreign_consumption_end, digits = 3)`% in 2019 (Fig 3c).
+However, patterns vary greatly across regions with Asia and South America dominated by domestic supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Asia", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% and `r signif(foreign_consumption_region %>% filter(year == 2019, region == "South America", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019, respectively), ), but countries in Europe dominated by foreign supply (`r signif(foreign_consumption_region %>% filter(year == 2019, region == "Europe", supply_source == "foreign") %>% pull(supply_percent), digits = 2)`% foreign in 2019) (Fig 3d).
 High reliance on foreign-sourced foods can pose a food security risk (Baer-Nawrocka and Sadowski 2019; Kummu et al.

### 2023-05-15 ms_figures.Rmd
clean up code for submission

@@
 fig3c <- ggplot() +
-  geom_ribbon(data = fig3c_data %>%
-                filter(supply_source == "domestic"),
-              aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
-  geom_ribbon(data = fig3c_data %>%
-                filter(supply_source == "foreign"),
-              aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
+  # geom_ribbon(data = fig3c_data %>%
+  #               filter(supply_source == "domestic"),
+  #             aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
+  # geom_ribbon(data = fig3c_data %>%
+  #               filter(supply_source == "foreign"),
+  #             aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
   geom_line(data = fig3c_data, aes(x = year, y = supply_percent, color = supply_source)) +
@@
 ldc_net_habitat_method_summary <- ldc_net_habitat_method_summary %>%
-  bind_rows(ldc_net_total)
+  bind_rows(ldc_net_total) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "total")))
 
@@
 ```
-
-```{r, eval=FALSE}
-# Plots moved from end of prep_data
-top_processing_countries %>%
-  filter(year == 2019) %>%
-  ggplot(aes(x = live_weight_t, y = reorder(exporter_iso3c, ranking))) +
-  geom_bar(stat = "identity") +
-  scale_y_reordered() +
-  theme_bw() +
-  labs(x = "Re-exports (Live weight tonnes)", y = "Processing Country")
-
-top_processing_countries_by_source %>%
-  filter(year == 2019) %>%
-  ggplot(aes(x = live_weight_t / 1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  theme_bw() +
-  labs(x = "Re-exports (Live weight million tonnes)", y = "Processing Country") +
-  facet_wrap(~habitat_method, scales = "free_y")
-
-```

### 2023-05-15 prep_data.R

@@ -16,7 +16,7 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20230420"
+outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20230428"
 outdir <- "outputs"
 
 #-------------------------------------------------------------------------------
@@ -90,68 +90,68 @@ write.csv(artis_fm, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
 # Write out the bilater non-food trade data to use in supply calculations
 write.csv(artis_nonfood, file.path(outdir, "artis_nonfood.csv"), row.names = FALSE)
 
-hs_versions <- c("96", "02", "07", "12", "17")
-datadir <- "../ARTIS/qa/model_inputs_20221129_NEW"
-
-hs_clade_match <- data.frame()
-for (i in 1:length(hs_versions)) {
-  curr_hs <- hs_versions[i]
-  print(curr_hs)
-  
-  # Determine most specific clade of each HS code (but if clade is not reported in production data (i.e., hs_taxa_match$SciName), return NA)
-  # To match to clade, even if not reported in production data, set match_to_prod to FALSE
-  curr_hs_clade_match <- match_hs_to_clade(hs_taxa_match = read.csv(file.path(datadir, paste("hs-taxa-match_HS", curr_hs, ".csv", sep = ""))) %>%
-                                             select(-c(sciname_habitat, code_habitat)),
-                                           prod_taxa_classification = read.csv(file.path(datadir, "clean_fao_taxa.csv")),
-                                           match_to_prod = FALSE) %>%
-    # pad HS codes with zeroes
-    mutate(Code = as.character(Code)) %>%
-    mutate(Code = if_else(str_detect(Code, "^30"), true = str_replace(Code, pattern = "^30", replacement = "030"),
-                          if_else(str_detect(Code, "^511"), true = str_replace(Code, pattern = "^511", replacement = "0511"),
-                                  false = Code)))# %>%
-  # # Filter down to just the codes in cc_m (recreated above from country_est)
-  # filter(Code %in% cc_m)
-  
-  curr_hs_clade_match <- curr_hs_clade_match %>%
[…]

### 2023-05-16 prep_data.R
new file paths based on newest Zorro run

@@ -33,9 +33,11 @@ con <- dbConnect(RPostgres::Postgres(),
 # Check that connection is established by checking which tables are present
 dbListTables(con)
 
-# ARTIS dataframe
-artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
-  select(-record_id) 
+# # ARTIS dataframe
+# artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
+#   select(-record_id) 
+
+artis <- read.csv("/Volumes/jgephart/ARTIS/Outputs/S_net/snet_20230412/custom_ts/mid_custom_ts.csv")
 
 artis_fm <- artis %>% 
   filter(hs6 == "230120") %>% 
@@ -89,69 +91,6 @@ write.csv(artis_fm, file.path(outdir, "artis_fm.csv"), row.names = FALSE)
 
 # Write out the bilater non-food trade data to use in supply calculations
 write.csv(artis_nonfood, file.path(outdir, "artis_nonfood.csv"), row.names = FALSE)
-
-# hs_versions <- c("96", "02", "07", "12", "17")
-# datadir <- "../ARTIS/qa/model_inputs_20221129_NEW"
-# 
-# hs_clade_match <- data.frame()
-# for (i in 1:length(hs_versions)) {
-#   curr_hs <- hs_versions[i]
-#   print(curr_hs)
-#   
-#   # Determine most specific clade of each HS code (but if clade is not reported in production data (i.e., hs_taxa_match$SciName), return NA)
-#   # To match to clade, even if not reported in production data, set match_to_prod to FALSE
-#   curr_hs_clade_match <- match_hs_to_clade(hs_taxa_match = read.csv(file.path(datadir, paste("hs-taxa-match_HS", curr_hs, ".csv", sep = ""))) %>%
-#                                              select(-c(sciname_habitat, code_habitat)),
-#                                            prod_taxa_classification = read.csv(file.path(datadir, "clean_fao_taxa.csv")),
-#                                            match_to_prod = FALSE) %>%
-#     # pad HS codes with zeroes
-#     mutate(Code = as.character(Code)) %>%
-#     mutate(Code = if_else(str_detect(Code, "^30"), true = str_replace(Code, pattern = "^30", replacement = "030"),
-#                           if_else(str_detect(Code, "^511"), true = str_replace(Code, pattern = "^511", replacement = "0511"),
[…]

### 2024-02-29 prep_data.R
cleaning manuscript code files

@@ -16,8 +16,7 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "/Volumes/jgephart/ARTIS/Outputs/ms_seafood_globalization/20230428"
-outdir <- "outputs"
+outdir <- "qa/outputs_20240221"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -33,11 +32,19 @@ con <- dbConnect(RPostgres::Postgres(),
 # Check that connection is established by checking which tables are present
 dbListTables(con)
 
-# # ARTIS dataframe
-# artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
-#   select(-record_id) 
+# ARTIS dataframe
+artis_query <- 'SELECT * FROM snet'
+artis <- dbGetQuery(con, artis_query) %>%
+  select(-record_id) %>%
+  filter(!is.na(live_weight_t) & !is.na(product_weight_t)) %>%
+  mutate(hs6 = as.character(hs6)) %>%
+  mutate(hs6 = case_when(
+    str_length(hs6) == 5 ~ paste("0", hs6, sep = ""),
+    TRUE ~ hs6
+  ))
 
-artis <- read.csv("/Volumes/jgephart/ARTIS/Outputs/S_net/snet_20230412/custom_ts/mid_custom_ts.csv")
+code_max_resolved <- dbGetQuery(con, 'SELECT * FROM code_max_resolved_taxa') %>%
+  select(-record_id)
 
 artis_fm <- artis %>% 
   filter(hs6 == "230120") %>% 
@@ -74,16 +81,22 @@ code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa"
 sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
   select(-record_id)
 
+pop <- dbGetQuery(con, "SELECT * FROM population") %>%
+  select(-record_id)
+
+consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
+  select(-record_id)
+
 # Close database connection
 dbDisconnect(con)
 rm(list = c("con"))
 
-# Population data
-pop <- read.csv("/Volumes/jgephart/ARTIS/Outputs/clean_metadata/fao_annual_pop.csv") %>% 
+# Population data with region data
+pop <- pop %>%
   left_join(country_metadata, by = c("iso3c")) %>%
[…]

### 2024-02-29 prep_data.R
Add script for draft of new figure

@@ -16,7 +16,7 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "qa/outputs_20240221"
+outdir <- "outputs"
 
 #-------------------------------------------------------------------------------
 # Initial database pulls
@@ -54,12 +54,12 @@ artis_fm <- artis %>%
             product_weight_t = sum(product_weight_t, na.rm = TRUE))
 
 artis_nonfood <- artis %>% 
-  # Keep only FMFO and ornamental species
-  filter(hs6 %in% c("230120", "051191", "030110", "030111", "030119"))
+  # Keep only FMFO, and ornamental species
+  filter(hs6 %in% c("230120",  "030110", "030111", "030119"))
 
 artis <- artis %>% 
   # Remove FMFO and ornamental species
-  filter(!(hs6 %in% c("230120", "051191", "030110", "030111", "030119")))
+  filter(!(hs6 %in% c("230120",  "030110", "030111", "030119")))
 
 # Production dataframe
 prod <- dbGetQuery(con, "SELECT * FROM production") %>%

### 2024-02-29 top_flows_top_species_figure.R

@@ -1,0 +1,248 @@
+# Draft figure of species-level information
+library(tidytext)
+
+con <- dbConnect(RPostgres::Postgres(),
+                 dbname=Sys.getenv("DB_NAME"),
+                 host="localhost",
+                 port="5432",
+                 user=Sys.getenv("DB_USERNAME"),
+                 password=Sys.getenv("DB_PASSWORD"))
+
+artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
+  select(-record_id)
+
+# Country metadata
+country_metadata <- dbGetQuery(con, "SELECT * FROM countries") %>%
+  select(-record_id)
+
+# Sciname metadata
+sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
+  select(-record_id)
+
+# Code max resolved taxa
+code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
+  select(-record_id) 
+
+# Consumption
+consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
+  select(-record_id)
+
+# Population 
+pop <- dbGetQuery(con, "SELECT * FROM population") %>%
+  select(-record_id)
+
+# Close database connection
+dbDisconnect(con)
+rm(list = c("con"))
+
+# Set colors
+region6_palette <- c(
+  "#741A32", # maroon
+  "#B34232", # burnt orange
+  "#D38F35", #  orange
+  "#D4B95F", # khaki
+  "#4FA2A2", # teal
+  "#114F59" # dark teal
+)
+
+artis_palette <- colorRampPalette(region6_palette)
+
+# FIX IT: Check on consumer NEI 
+consumption_scaled <- consumption %>%
+  left_join(pop, by = c("year", "consumer_iso3c" = "iso3c")) %>%
+  mutate(consumption_kg_percap = 1000*consumption_live_t/pop) %>%
+  group_by(year, consumer_iso3c) %>%
+  mutate(total_consumption_kg_percap = sum(consumption_kg_percap)) %>% 
+  ungroup() %>%
+  mutate(scale_factor = consumption_kg_percap/total_consumption_kg_percap) %>%
+  mutate(total_consumption_kg_percap = case_when(
+    total_consumption_kg_percap > 100 ~ 100, 
+    TRUE ~ total_consumption_kg_percap
+    )) %>%
[…]

### 2024-02-29 ms_figures.Rmd

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
-
+knitr::opts_chunk$set(
+	echo = FALSE,
+	message = FALSE,
+	warning = FALSE
+)
 library(tidyverse)
@@
 # Set directories
-datadir <- "outputs"
-outdir <- "outputs"
+datadir <- "qa/outputs_20240221"
+outdir <- "qa/outputs_20240221"
 ```
@@
   theme(legend.position = "bottom",
-        axis.title.y = element_text(size = 16),
+        axis.title.y = element_text(size = 12),
         axis.text.x = element_text(size = 12),
         axis.text.y = element_text(size = 12),
-        legend.text = element_text(size = 9),
+        legend.title = element_text(size = 7),
+        legend.text = element_text(size = 6)
         )
@@
   scale_color_manual(values = habitat_source_colors) +
-  labs(y = "% Production\nExported", x = "", color = "Source") +
+  labs(y = "% Production\nexported", x = "", color = "Source") +
   theme_bw() +
   theme(
-    text = element_text(size = 16)
+    text = element_text(size = 10),
+    legend.title = element_text(size = 7),
+    legend.text = element_text(size = 6)
   )
@@
 
-## Figure 1 c: Degree or measure of degree distribution by production method and environment
+## Figure 1c: Degree or measure of degree distribution by production method and environment
 
@@
   ggplot(aes(x = year, y = all_in_degree, color = habitat_method)) +
-  geom_smooth() +
+  geom_smooth(alpha = 0.2) +
   scale_color_manual(values = habitat_source_colors) +
@@
   theme(
-    text = element_text(size = 16),
-    axis.title.x = element_blank()
+    text = element_text(size = 10),
+    axis.title.x = element_blank(),
+    legend.title = element_text(size = 7),
+    legend.text = element_text(size = 6)
   )
@@
   theme(
-    text = element_text(size = 16)
+    text = element_text(size = 10),
+    legend.title = element_text(size = 7),
+    legend.text = element_text(size = 6)
   )
@@
 
-```{r fig1, fig.height = 4, fig.width = 4}
+```{r fig1, fig.height = 6, fig.width = 6}
 ggarrange(fig1a,
@@
         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        strip.background = element_blank()) +
+        strip.background = element_blank(),
+        plot.margin = unit(c(t=0, r=1, b=0, l=0.5), "cm")) +
   facet_wrap(~region, nrow = 1)
 
-fig2a
-```
-
-```{r fig2a_old, eval=FALSE}
-# Imports by Region
-fig2a <- artis_region %>% 
-  ggplot(aes(x = year, y = live_weight_t/1000000, fill = exporter_region)) +
-  geom_area() +
-  labs(x = "", y = "", fill = "Exporter") +
-  scale_fill_manual(values = region_colors) +
-  facet_wrap(~importer_region, nrow = 1) +
-  theme_bw() +
-  theme(legend.position = "none",
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        axis.text.x = element_blank(),
-        plot.margin=unit(c(0.5,0,0,0), "cm"))
+fig2a <- ggdraw(fig2a) +
+  draw_label("Net import", x = 0.97, y = 0.55, angle = 270, size = 9) +
+  draw_label("Net export", x = 0.97, y = 0.3, angle = 270, size = 9)
 
@@
   scale_fill_manual(values = habitat_source_colors) +
-  labs(x = "", y = "Trade (million t, live weight)", color = "Ave. annual change in export \n(1000 t live weight)",
+  scale_y_continuous(position = "right") +
+  labs(x = "", y = "", color = "Ave. annual change in export \n(1000 t live weight)",
        fill = "Source") +
@@
         strip.background = element_blank(),
-        plot.margin = margin(t = 8, r = 8, l = 2, unit = "mm")) +
+        plot.margin = margin(t = 6, r = 7, l = 6, unit = "mm")) +
   facet_grid(rows = vars(exporter_region), cols = vars(importer_region), 
-             labeller = label_wrap_gen(width = 5), scales = "free")
+             labeller = label_wrap_gen(width = 5), scales = "free", switch = "y")
 
 fig2b <- ggdraw(fig2b) +
-  draw_label("Exporting region", x = 0.975, y = 0.5, angle = 270) +
-  draw_label("Importing region", x = 0.5, y = 0.975)
+  draw_label("Exporting region", x = 0.02, y = 0.5, angle = 90, size = 10) +
+  draw_label("Importing region", x = 0.5, y = 0.965, size = 10) +
+  draw_label("Trade (million t, live weight)", x = 0.96, y = 0.5, angle = 270, size = 10)
 
@@
 
-```{r fig2c_old, eval=FALSE}
-# Figure 2c
-# Exports by Region
-fig2c <- artis_region %>% 
-  ggplot(aes(x = year, y = live_weight_t/1000000, fill = importer_region)) +
-  geom_area() +
-  labs(x = "", y = "", fill = "Importer") +
-  scale_fill_manual(values = region_colors) +
-  facet_wrap(~exporter_region, ncol = 1) +
-  theme_bw() +
-  theme(legend.position = "none",
-        strip.background = element_blank(),
-        strip.text.x = element_blank(),
-        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
-        plot.margin=unit(c(0,0.5,0,0), "cm"))
-
-fig2c
-```
-
-```{r fig2_old, eval=FALSE}
-# Overall Figure 2
-ggarrange(ggarrange(fig2a, NULL,
-                    ncol = 2, nrow = 1,
-                    widths = c(4.5, 1.25), labels = c("a", "")),
-          ggarrange(fig2b, 
-                    ggarrange(NULL, fig2c,
-                              ncol = 1, heights = c(0.3, 4)), 
-                    ncol = 2, nrow = 1,
-                    widths = c(4.5, 1), labels = c("b", "c")),
-          ncol = 1, nrow = 2, 
-          heights = c(1, 4.5))
-
-ggsave(file.path(outdir, "fig2.png"))
-```
-
-```{r fig2, fig.height=4, fig.width=3}
+```{r fig2, fig.height=8, fig.width=6.5}
 # Overall Figure 2
@@
           fig2b, labels = c("", "b"),
-          ncol = 1, heights = c(1.75, 4))
+          ncol = 1, heights = c(3, 4))
 
@@
   scale_color_manual(values = habitat_source_colors) +
+  guides(color = guide_legend(nrow = 1)) +
   theme_bw() +
@@
   scale_fill_manual(values = habitat_source_colors) +
+  guides(fill = guide_legend(nrow = 1)) +
   theme_bw() +
@@
 fig3c <- ggplot() +
-  # geom_ribbon(data = fig3c_data %>%
-  #               filter(supply_source == "domestic"),
-  #             aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
-  # geom_ribbon(data = fig3c_data %>%
-  #               filter(supply_source == "foreign"),
-  #             aes(ymin = supply_min_percent, ymax = supply_max_percent, x = year), fill = "grey70") +
   geom_line(data = fig3c_data, aes(x = year, y = supply_percent, color = supply_source)) +
@@
 
-```{r fig3, fig.height = 4, fig.width = 5}
+```{r fig3, fig.height = 6, fig.width = 7}
 fig3 <- ggarrange(
@@
 
-```{r regional_sankey, fig.height = 2, fig.width = 2}
+```{r regional_sankey, fig.height = 3, fig.width = 3}
 regional_artis <- read.csv(file.path(datadir, "regional_artis_by_source.csv"))
-# plot_regional_sankey_method_habitat(regional_artis, 2019, 2019)
 
@@
     scale_fill_manual(values = region_colors) +
-    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    geom_sankey_label(size = 4, color = "white", fill = "gray40") +
     theme_sankey(base_size = 18) +
@@
     plot.title = element_text(hjust = 0.5),
-    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    plot.margin = margin(c(t=-0.5, r=-10, b=-1, l=-18), unit = "mm")
     )
+
+marine_cap_p <- ggdraw(marine_cap_p) +
+  draw_label("Marine capture", x = 0.5, y = 0.97, size = 9)
 
@@
 
-```{r marine_aqua, fig.height = 2, fig.width = 2}
+```{r marine_aqua, fig.height = 3, fig.width = 3}
 marine_aqua_regions <- regional_artis %>%
@@
     plot.title = element_text(hjust = 0.5),
-    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    plot.margin = margin(c(t=-0.5, r=-10, b=-1, l=-18), unit = "mm")
     )
+
+marine_aqua_p <- ggdraw(marine_aqua_p) +
+  draw_label("Marine aquaculture", x = 0.5, y = 0.97, size = 9)
 
@@
 
-```{r inland_cap, fig.height = 2, fig.width = 2}
+```{r inland_cap, fig.height = 3, fig.width = 3}
 inland_cap_regions <- regional_artis %>%
@@
     plot.title = element_text(hjust = 0.5),
-    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    plot.margin = margin(c(t=-0.5, r=-18, b=-1, l=-18), unit = "mm")
     )
+
+inland_cap_p <- ggdraw(inland_cap_p) +
+  draw_label("Inland capture", x = 0.5, y = 0.97, size = 9)
 
@@
 
-```{r inland_aqua, fig.height = 2, fig.width = 2}
+```{r inland_aqua, fig.height = 3, fig.width = 3}
 inland_aqua_regions <- regional_artis %>%
@@
     scale_fill_manual(values = region_colors) +
-    geom_sankey_label(size = 3, color = "white", fill = "gray40") +
+    geom_sankey_label(size = 4, color = "white", fill = "gray40") +
     theme_sankey(base_size = 18) +
-  labs(x = NULL) +
-    #labs(title = "Marine", x = NULL, y = "CAPTURE") +
+  labs(x = NULL, title = NULL) +
   theme(
@@
     axis.text.x = element_blank(),
-    plot.title = element_text(hjust = 0.5),
-    plot.margin = margin(c(0,0,0,0), unit = "mm")
+    plot.title = element_text(hjust = 0.5, size = 9),
+    plot.margin = margin(c(t=-0.5, r=-18, b=-1, l=-18), unit = "mm")
     )
+
+inland_aqua_p <- ggdraw(inland_aqua_p) +
+  draw_label("Inland aquaculture", x = 0.5, y = 0.97, size = 9)
 
@@
 ```
-```{r regional_sankey_by_source, fig.height = 2.5, fig.width = 3}
+```{r regional_sankey_by_source, fig.height = 8, fig.width = 10}
 ggarrange(marine_cap_p, inland_cap_p,
@@
           labels = c("a", "b", "c", "d"))
-          # labels = c("Marine Capture", "Inland Capture",
-          #            "Marine Aquaculture", "Inland Aquaculture"))
 
@@
 
-```{r si_top_exporters, fig.height = 5, fig.width = 4}
+```{r si_top_exporters, fig.height = 7, fig.width = 6}
 
-top_exporters_old <- read.csv(file.path(datadir, "top_exporters_old.csv")) %>%
+top_exporters_old <- read.csv(file.path(datadir, "si_top_exporters_old_fill.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
                                             "marine aquaculture", "inland aquaculture", "unknown")))
-top_exporters_recent <- read.csv(file.path(datadir, "top_exporters_recent.csv")) %>%
+top_exporters_recent <- read.csv(file.path(datadir, "si_top_exporters_recent_fill.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
 exporters_old_plot <- top_exporters_old %>%
-  ggplot(aes(x = live_weight_t/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
+  pivot_longer(cols = c("domestic", "foreign"), names_to = "dom_source", values_to = "live_weight_t_2") %>%
+  ggplot(aes(x = live_weight_t_2/1000000, y = reorder(exporter_iso3c, ranking), fill = dom_source)) +
   geom_bar(stat = "identity") +
@@
   scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
+  labs(y = "", x = "Live weight (million t)", fill = "Export Source") +
   xlim(c(0,6)) +
@@
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.text = element_text(size = 8)
+  )
 
 exporters_recent_plot <- top_exporters_recent %>%
-  ggplot(aes(x = live_weight_t/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
+  pivot_longer(cols = c("domestic", "foreign"), names_to = "dom_source", values_to = "live_weight_t_2") %>%
+  ggplot(aes(x = live_weight_t_2/1000000, y = reorder(exporter_iso3c, ranking), fill = dom_source)) +
   geom_bar(stat = "identity") +
@@
   scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
+  labs(y = "", x = "Live weight (million t)", fill = "Export Source") +
   facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.text = element_text(size = 8)
+  )
 
@@
   ncol = 2,
-  labels = c("'96-'00", "'16-'20")
+  labels = c("'96-'00", "'16-'20"),
+  common.legend = TRUE
 )
@@
 
-```{r si_top_importers, fig.height = 5, fig.width = 4}
+```{r si_top_importers, fig.height = 7, fig.width = 6}
 
-top_importers_old <- read.csv(file.path(datadir, "top_importers_old.csv")) %>%
+top_importers_old <- read.csv(file.path(datadir, "si_top_importers_old_fill.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
                                             "marine aquaculture", "inland aquaculture", "unknown")))
-top_importers_recent <- read.csv(file.path(datadir, "top_importers_recent.csv")) %>%
+top_importers_recent <- read.csv(file.path(datadir, "si_top_importers_recent_fill.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
 importers_old_plot <- top_importers_old %>%
-  ggplot(aes(x = live_weight_t/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method)) +
+  pivot_longer(cols = c("domestic", "foreign"), names_to = "dom_source", values_to = "live_weight_t_2") %>%
+  ggplot(aes(x = live_weight_t_2/1000000, y = reorder(importer_iso3c, ranking), fill = dom_source)) +
   geom_bar(stat = "identity") +
@@
   xlim(c(0, 5.5)) +
-  labs(y = "", x = "Live weight (million t)") +
+  labs(y = "", x = "Live weight (million t)", fill = "Import Source") +
   facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.text = element_text(size = 8)
+  )
 
 importers_recent_plot <- top_importers_recent %>%
-  ggplot(aes(x = live_weight_t/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method)) +
+  pivot_longer(cols = c("domestic", "foreign"), names_to = "dom_source", values_to = "live_weight_t_2") %>%
+  ggplot(aes(x = live_weight_t_2/1000000, y = reorder(importer_iso3c, ranking), fill = dom_source)) +
   geom_bar(stat = "identity") +
@@
   xlim(c(0, 5.5)) +
-  labs(y = "", x = "Live weight (million t)") +
+  labs(y = "", x = "Live weight (million t)", fill = "Import Source") +
   facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.text = element_text(size = 8)
+  )
 
@@
 
-```{r si_regional_changes, fig.height = 4, fig.width = 4}
+```{r si_regional_changes, fig.height = 6, fig.width = 6}
 
@@
 
-```{r si_export_decreases_increases, fig.height = 4, fig.width = 4.5}
+```{r si_export_decreases_increases, fig.height = 6, fig.width = 5}
 top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv")) %>%
@@
   geom_segment(aes(x=0, xend=change/1000000, y= reorder(exporter_iso3c, ranking), yend=exporter_iso3c)) +
-  geom_point(size=4) +
+  geom_point(size=2) +
   scale_color_manual(values = habitat_source_colors) +
@@
   theme_minimal() +
-  theme(legend.position = "none")
+  theme(
+    legend.position = "none",
+    axis.text = element_text(size = 8)
+  )
 
@@
 
-```{r si_import_decreases_increases, fig.height = 4, fig.width = 4.5}
+```{r si_import_decreases_increases, fig.height = 6, fig.width = 5}
 top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv")) %>%
@@
   geom_segment(aes(x=0, xend=change/1000000, y= reorder(importer_iso3c, ranking), yend=importer_iso3c)) +
-  geom_point(size=4) +
+  geom_point(size=2) +
   scale_color_manual(values = habitat_source_colors) +
@@
 
-```{r regional_avg_annual_change, fig.height = 4, fig.width = 5}
+```{r regional_avg_annual_change, fig.height = 8, fig.width = 7}
 regional_avg_annual_change <- read.csv(file.path(datadir, "regional_avg_annual_change.csv"))
@@
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 8)
+  )
 
@@
 
-```{r ldc_net_exports_habitat_method}
+```{r ldc_net_exports_habitat_method, fig.height = 4, fig.width = 7}
 ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv")) %>%
@@
   theme_minimal() +
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 12)
+  )
 
@@
 
-```{r custom_artis_ts}
+```{r custom_artis_ts, fig.height = 4, fig.width = 6}
 
@@
   ) +
-  labs(x = "Year", y = "Live Weight (tonnes)", color = "HS Version") +
+  labs(x = "Year", y = "Product Weight (tonnes)", color = "HS Version") +
   theme_bw() +
-  theme(legend.position = "bottom")
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 10)
+  )
 
@@
     legend.position = "bottom",
+    axis.text = element_text(size = 10),
     legend.box.margin=margin(t = -25)
@@
     legend.position = "bottom",
+    axis.text = element_text(size = 10),
     panel.spacing = unit(5, "mm"),
@@
 
-```{r dom_source_overview, fig.height = 5, fig.width = 7}
+```{r dom_source_overview, fig.height = 5, fig.width = 8}
  ggarrange(
@@
 ```
-
-```{r top_processing_countries_by_source, fig.height = 4, fig.width = 6}
-
-top_processing_by_source <- read.csv(file.path(datadir, "top_processing_by_source.csv"))
-
-top_processing_by_source %>%
-  filter(year == 2019) %>%
-  # Set factor levels
-  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
-                                                            "marine aquaculture", "inland aquaculture",
-                                                            "unknown" ))) %>%
-  ggplot(aes(x = live_weight_t / 1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method)) +
-  geom_bar(stat = "identity") +
-  scale_fill_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  theme_bw() +
-  labs(x = "Live weight million tonnes", y = "", fill = "Source") +
-  facet_wrap(~habitat_method, scales = "free_y") +
-  theme(legend.position = "none")
-
-ggsave(file.path(outdir, "foreign_exports_2019_by_source.png"))
-```

### 2024-02-29 ms_results.Rmd

@@
 # Directory
-datadir <- "outputs"
+datadir <- "qa/outputs_20240221"
 
@@
 d) Number of countries comprising 75% of global exports by production method and environment with the global total in the black line.
-e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](outputs/fig1.png)
+e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](qa/outputs_20240221/fig1.png)
 
@@
 Values represent million tonnes in live weight equivalents.
-Note the scale difference across rows.](outputs/fig2.png)
+Note the scale difference across rows.](qa/outputs_20240221/fig2.png)
 
@@
 d) Regional aquatic food domestic versus foreign sourcing over time.
-](outputs/fig3.png)
+](qa/outputs_20240221/fig3.png)

### 2024-02-29 top_flows_top_species_figure.R
move comment

@@ -47,7 +47,7 @@ region6_palette <- c(
 
 artis_palette <- colorRampPalette(region6_palette)
 
-# FIX IT: Check on consumer NEI 
+
 consumption_scaled <- consumption %>%
   left_join(pop, by = c("year", "consumer_iso3c" = "iso3c")) %>%
   mutate(consumption_kg_percap = 1000*consumption_live_t/pop) %>%
@@ -68,6 +68,7 @@ consumption_scaled <- consumption %>%
   left_join(country_metadata %>%
               select(iso3c, "source_region" = "owid_region"), 
             by = c("source_country_iso3c" = "iso3c")) %>%
+  # FIX IT: Check on consumer NEI 
   filter(consumer_region != "Other nei", source_region != "Other nei") %>%
   # Add common names associated with sciname_hs_modified names
   left_join(sciname_metadata %>% 

### 2024-05-01 prep_data.R
edits for manuscript archive

@@ -16,8 +16,8 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "outputs"
-
+# outdir <- "qa/outputs_20240315"
+outdir <- "qa/manuscript_archive/manuscript_inputs"
 #-------------------------------------------------------------------------------
 # Initial database pulls
 
@@ -33,8 +33,12 @@ con <- dbConnect(RPostgres::Postgres(),
 dbListTables(con)
 
 # ARTIS dataframe
+artis_query <- "SELECT * FROM snet WHERE 
+((hs_version = 'HS96' AND year >= 1996 AND year <= 2003) OR
+(hs_version = 'HS02' AND year >= 2004 AND year <= 2009) OR
+(hs_version = 'HS07' AND year >= 2010 AND year <= 2012) OR
+(hs_version = 'HS12' AND year >= 2013 AND year <= 2020))"
 
-artis_query <- 'SELECT * FROM snet'
 artis <- dbGetQuery(con, artis_query) %>%
   select(-record_id) %>%
   filter(!is.na(live_weight_t) & !is.na(product_weight_t)) %>%
@@ -43,9 +47,6 @@ artis <- dbGetQuery(con, artis_query) %>%
     str_length(hs6) == 5 ~ paste("0", hs6, sep = ""),
     TRUE ~ hs6
   ))
-
-code_max_resolved <- dbGetQuery(con, 'SELECT * FROM code_max_resolved_taxa') %>%
-  select(-record_id)
 
 artis_fm <- artis %>% 
   filter(hs6 == "230120") %>% 
@@ -85,7 +86,11 @@ sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
 pop <- dbGetQuery(con, "SELECT * FROM population") %>%
   select(-record_id)
 
-consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
+consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption WHERE 
+((hs_version = 'HS96' AND year >= 1996 AND year <= 2003) OR
+(hs_version = 'HS02' AND year >= 2004 AND year <= 2009) OR
+(hs_version = 'HS07' AND year >= 2010 AND year <= 2012) OR
+(hs_version = 'HS12' AND year >= 2013 AND year <= 2020))") %>%
   select(-record_id)
 
 # Close database connection
[…]

### 2024-05-01 ms_results.Rmd
edits on ms results for manuscript archive

@@
 # Directory
-datadir <- "qa/outputs_20240221"
+datadir <- "qa/manuscript_archive/manuscript_inputs"
 
@@
 
-regional_supply_per_cap <- read.csv(file.path(datadir, "fig3b_data.csv"))
+regional_supply_per_cap <- read.csv(file.path(datadir, "fig_4b_data.csv"))
 
@@
 
-foreign_consumption_region <- read.csv(file.path(datadir, "fig3d_data.csv"))
+foreign_consumption_region <- read.csv(file.path(datadir, "fig_4d_data.csv"))
 

### 2024-05-01 ms_figures.Rmd
ms figures edits for manuscript edits

@@
 library(ggpubr)
+library(rnaturalearth)
+library(CoordinateCleaner)
 ```
@@
 # Set directories
-datadir <- "qa/outputs_20240221"
-outdir <- "qa/outputs_20240221"
+datadir <- "qa/manuscript_archive/manuscript_inputs"
+outdir <- "qa/manuscript_archives/figures"
 ```
@@
 
+region6_palette <- c(
+  "#741A32", # maroon
+  "#B34232", # burnt orange
+  "#D38F35", #  orange
+  "#D4B95F", # khaki
+  "#4FA2A2", # teal
+  "#114F59" # dark teal
+)
+
 ```
@@
                                             "marine aquaculture", "inland aquaculture", "unknown"))) %>%
-  ggplot(aes(x = year, y = all_in_degree, color = habitat_method)) +
+  ggplot(aes(x = year, y = all_out_degree, color = habitat_method)) +
   geom_smooth(alpha = 0.2) +
@@
 ggsave(file.path(outdir, "fig1.png"))
-```
-
-# Figure 2: Trade patterns
-
-```{r fig2_data}
-artis_region_method <- read.csv(file.path(datadir, "artis_region_method.csv"))
-artis_region <- read.csv(file.path(datadir, "artis_region.csv"))
-```
-
-```{r fig2a}
-
+# ggsave(file.path(outdir, "fig1.pdf"), device = "pdf", height = )
+```
+
+# Figure 2: Foreign consumption trends
+```{r foreign_consumption_setup}
+marine_cap_foreign_consumption <- read.csv(file.path(datadir, "marine_capture_foreign_consumption.csv"))
+
+marine_aqua_foreign_consumption <- read.csv(file.path(datadir, "marine_aquaculture_foreign_consumption.csv"))
+
+inland_cap_foreign_consumption <- read.csv(file.path(datadir, "inland_capture_foreign_consumption.csv"))
+
+inland_aqua_foreign_consumption <- read.csv(file.path(datadir, "inland_aquaculture_foreign_consumption.csv"))
+
+marine_cap_sciname <- read.csv(file.path(datadir, "marine_capture_top_scinames_foreign_consumption.csv"))
+
+marine_aqua_sciname <- read.csv(file.path(datadir, "marine_aquaculture_top_scinames_foreign_consumption.csv"))
+
+inland_cap_sciname <- read.csv(file.path(datadir, "inland_capture_top_scinames_foreign_consumption.csv"))
+
+inland_aqua_sciname <- read.csv(file.path(datadir, "inland_aquaculture_top_scinames_foreign_consumption.csv"))
+
+# Load world map data
+world <- ne_countries(scale = "medium", returnclass = "sf") %>%
+  filter(iso_a3 != "ATA")
+
+consumer_map_df <- marine_cap_foreign_consumption %>%
+  mutate(habitat_method = "Marine Capture") %>%
+  bind_rows(marine_aqua_foreign_consumption %>%
+              mutate(habitat_method = "Marine Aquaculture")) %>%
+  bind_rows(inland_cap_foreign_consumption %>%
+              mutate(habitat_method = "Inland Capture")) %>%
+  bind_rows(inland_aqua_foreign_consumption %>%
+              mutate(habitat_method = "Inland Aquaculture")) %>%
+  group_by(consumer_iso3c, habitat_method) %>%
+  summarize(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
+  ungroup() %>%
+  left_join(
+    world,
+    by = c("consumer_iso3c"="iso_a3")
+  )
+
+country_centroids <- countryref %>%
+  filter(str_count(as.character(adm1_code))==3) %>%
+  group_by(iso3) %>%
+  summarise(centroid.lon = mean(centroid.lon),
+            centroid.lat = mean(centroid.lat)) %>%
+  ungroup() %>%
+  mutate(
+    centroid.lon = case_when(
+      iso3 == "KIR" ~ 172.9717,
+      TRUE ~ centroid.lon
+    ),
+    centroid.lat = case_when(
+      iso3 == "KIR" ~ 1.4518,
+      TRUE ~ centroid.lat
+    )
+  )
+
+trade_map <- ggplot(world) +
+  geom_sf(size = 0.1, color = "white", fill = "grey", alpha = 0.5)
+
+# Maximum and minimum limits for maps and stacked bar charts
+consumption_live_magnitude <- 1e6
+
+map_min <- 0
+map_max <- max(consumer_map_df$consumption_live_t, na.rm = TRUE) / consumption_live_magnitude
+arrows_min <- 0
+arrows_max <- max(
+  c(marine_cap_foreign_consumption$consumption_live_t,
+    marine_aqua_foreign_consumption$consumption_live_t,
+    inland_cap_foreign_consumption$consumption_live_t,
+    inland_aqua_foreign_consumption$consumption_live_t),
+  na.rm = TRUE
+) / consumption_live_magnitude
+
+sciname_df <- marine_cap_sciname %>%
+  mutate(habitat_method = "Marine Capture") %>%
+  bind_rows(marine_aqua_sciname %>%
+              mutate(habitat_method = "Marine Aquaculture")) %>%
+  bind_rows(inland_cap_sciname %>%
+              mutate(habitat_method = "Inland Capture")) %>%
+  bind_rows(inland_aqua_sciname %>%
+              mutate(habitat_method = "Inland Aquaculture")) %>%
+  group_by(common_name, flow, habitat_method) %>%
+  summarize(consumption_live_t = sum(consumption_live_t, na.rm = TRUE)) %>%
+  ungroup()
+
+stacked_bar_min <- 0
+stacked_bar_max <- max(
+  sciname_df$consumption_live_t,
+  na.rm = TRUE
+) / consumption_live_magnitude
+
+```
+
+```{r marine_cap_foreign_consumption}
+
+marine_cap_map <- trade_map +
+  geom_sf(data = consumer_map_df %>%
+            filter(habitat_method == "Marine Capture"), 
+          aes(fill = consumption_live_t/consumption_live_magnitude, geometry = geometry), color = "white", size = 0.1) +
+  scale_fill_gradient(low = "#86ADA7", high = "#0F2D59", limits=c(map_min, map_max)) +
+  theme_bw() + 
+  theme(
+    legend.position = "none",
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_blank(),
+    panel.spacing = unit(2, "mm"),
+    panel.border = element_blank(),
+    plot.background = element_blank(),
+    plot.margin = unit(c(0,0,-5,0),"mm")
+  )
+
+marine_cap_arrows <- marine_cap_foreign_consumption %>%
+  slice_max(n = 10, order_by = consumption_live_t) %>%
+  # Join with lat/long data for centroids
+  left_join(country_centroids, by = c("source_country_iso3c" = "iso3")) %>%
+  left_join(country_centroids, by = c("consumer_iso3c" = "iso3")) %>%
+  mutate(centroid.lat.x = 1.05 * centroid.lat.x,
+         centroid.lat.y = 0.95 * centroid.lat.y)
+
+marine_cap_map <- marine_cap_map +
+  geom_curve(data = marine_cap_arrows, 
+                 aes(x = centroid.lon.x, y = centroid.lat.x, 
+                     xend = centroid.lon.y, yend = centroid.lat.y,
+                     color = consumption_live_t/consumption_live_magnitude),
+                 size = 1.2,
+                 alpha = 0.75,
+                 curvature = -0.35, arrow = arrow(length = unit(3, "mm"), angle = 20)) +
+      scale_colour_gradient(low = "#F7AF75", high = "#E24027", limits = c(arrows_min, arrows_max))
+
+marine_cap_map
+
+marine_cap_stacked_bar <- marine_cap_sciname %>%
+  mutate(common_name = case_when(
+    common_name == "alaska pollock(=walleye poll.)" ~ "alaska pollock",
+    TRUE ~ common_name
+  )) %>%
+  mutate(common_name = gsub(" ", "\n", common_name)) %>%
+  mutate(common_name = reorder(common_name, desc(total))) %>%
+  ggplot(aes(x = consumption_live_t/consumption_live_magnitude, y = flow, fill = region)) +
+  geom_bar(stat = "identity", position = "stack", width = 0.8) +
+  scale_fill_manual(values = region6_palette) +
+  labs(x = "", y = "") +
+  lims(x = c(stacked_bar_min, stacked_bar_max)) +
+  facet_grid(vars(common_name), switch = "y") +
+  theme_minimal() +
+  theme(
+    strip.text.y.left = element_text(size = 8),
+    strip.placement = "outside",
+    legend.position = "none",
+    panel.spacing = unit(2, "mm"),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    axis.text.y = element_text(size = 8),
+    plot.margin = unit(c(0, 5, -5, -10), "mm")
+  )
+
+marine_cap_stacked_bar
+
+marine_cap <- ggarrange(
+  marine_cap_map, marine_cap_stacked_bar,
+  nrow = 1, ncol = 2,
+  widths = c(1.5, 1)
+)
+
+marine_cap <- annotate_figure(
+  marine_cap, top = text_grob("Marine Capture")#, vjust = 1),
+)
+
+marine_cap
+
+# ggsave(file.path(outdir, "marine_cap_map_bar.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
+```
+
+```{r marine_aqua_foreign_consumption}
+marine_aqua_map <- trade_map +
+  geom_sf(data = consumer_map_df %>%
+            filter(habitat_method == "Marine Aquaculture"), 
+          aes(fill = consumption_live_t/consumption_live_magnitude, geometry = geometry), color = "white", size = 0.1) +
+  scale_fill_gradient(low = "#86ADA7", high = "#0F2D59", limits = c(map_min, map_max)) +
+  theme_bw() + 
+  theme(
+    legend.position = "none",
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_blank(),
+    panel.spacing = unit(2, "mm"),
+    panel.border = element_blank(),
+    plot.background = element_blank(),
+    plot.margin = unit(c(0,0,-5,0),"mm")
+  )
+
+marine_aqua_arrows <- marine_aqua_foreign_consumption %>%
+  slice_max(n = 10, order_by = consumption_live_t) %>%
+  # Join with lat/long data for centroids
+  left_join(country_centroids, by = c("source_country_iso3c" = "iso3")) %>%
+  left_join(country_centroids, by = c("consumer_iso3c" = "iso3")) %>%
+  mutate(centroid.lat.x = 1.05 * centroid.lat.x,
+         centroid.lat.y = 0.95 * centroid.lat.y)
+
+marine_aqua_map <- marine_aqua_map +
+  geom_curve(data = marine_aqua_arrows, 
+                 aes(x = centroid.lon.x, y = centroid.lat.x, 
+                     xend = centroid.lon.y, yend = centroid.lat.y,
+                     color = consumption_live_t/consumption_live_magnitude),
+                 size = 1.2,
+                 alpha = 0.75,
+                 curvature = -0.35, arrow = arrow(length = unit(3, "mm"), angle = 20)) +
+      scale_colour_gradient(low = "#F7AF75", high = "#E24027", limits = c(arrows_min, arrows_max))
+
+marine_aqua_map
+
+marine_aqua_stacked_bar <- marine_aqua_sciname %>%
+  mutate(common_name = gsub(" ", "\n", common_name))%>%
+  mutate(common_name = reorder(common_name, desc(total))) %>%
+  ggplot(aes(x = consumption_live_t/consumption_live_magnitude, y = flow, fill = region)) +
+  geom_bar(stat = "identity", position = "stack", width = 0.8) +
+  scale_fill_manual(values = region6_palette) +
+  labs(x = "", y = "") +
+  lims(x = c(stacked_bar_min, stacked_bar_max)) +
+  facet_grid(vars(common_name), switch = "y") +
+  theme_minimal() +
+  theme(
+    strip.text.y.left = element_text(size = 8),
+    strip.placement = "right",
+    legend.position = "none",
+    panel.spacing = unit(2, "mm"),
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    plot.margin = unit(c(0, 5, -5, -10), "mm")
+  )
+
+marine_aqua_stacked_bar
+
+marine_aqua <- ggarrange(
+  marine_aqua_map, marine_aqua_stacked_bar,
+  nrow = 1, ncol = 2,
+  widths = c(1.5, 1)
+)
+
+marine_aqua <- annotate_figure(
+  marine_aqua, top = text_grob("Marine Aquaculture")#, vjust = 1),
+)
+
+marine_aqua
+
+# ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
+```
+
+```{r inland_cap_foreign_consumption}
+
+inland_cap_map <- trade_map +
+  geom_sf(data = consumer_map_df %>%
+            filter(habitat_method == "Inland Capture"), 
+          aes(fill = consumption_live_t/consumption_live_magnitude, geometry = geometry), color = "white", size = 0.1) +
+  scale_fill_gradient(low = "#86ADA7", high = "#0F2D59", limits = c(map_min, map_max)) +
+  theme_bw() + 
+  theme(
+    legend.position = "none",
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_blank(),
+    panel.spacing = unit(2, "mm"),
+    panel.border = element_blank(),
+    plot.background = element_blank(),
+    plot.margin = unit(c(0,0,-5,-10),"mm")
+  )
+
+inland_cap_arrows <- inland_cap_foreign_consumption %>%
+  slice_max(n = 10, order_by = consumption_live_t) %>%
+  # Join with lat/long data for centroids
+  left_join(country_centroids, by = c("source_country_iso3c" = "iso3")) %>%
+  left_join(country_centroids, by = c("consumer_iso3c" = "iso3")) %>%
+  mutate(centroid.lat.x = 1.05 * centroid.lat.x,
+         centroid.lat.y = 0.95 * centroid.lat.y)
+
+inland_cap_map <- inland_cap_map +
+  geom_curve(data = inland_cap_arrows, 
+                 aes(x = centroid.lon.x, y = centroid.lat.x, 
+                     xend = centroid.lon.y, yend = centroid.lat.y,
+                     color = consumption_live_t/consumption_live_magnitude),
+                 size = 1.2,
+                 alpha = 0.75,
+                 curvature = -0.35, arrow = arrow(length = unit(3, "mm"), angle = 20)) +
+      scale_colour_gradient(low = "#F7AF75", high = "#E24027", limits = c(arrows_min, arrows_max))
+
+inland_cap_map
+
+inland_cap_stacked_bar <- inland_cap_sciname %>%
+  mutate(common_name = gsub(" nei", "", common_name)) %>%
+  mutate(common_name = case_when(
+    common_name == "ray-finned fishes" ~ "\nray-finned fishes",
+    common_name == "freshwater siluroids" ~ "\nfreshwater siluroids",
+    common_name == "nile perch" ~ "\nnile perch",
+    TRUE ~ common_name
+  )) %>%
+  mutate(common_name = gsub(" ", "\n", common_name)) %>%
+  mutate(common_name = reorder(common_name, desc(total))) %>%
+  ggplot(aes(x = consumption_live_t/consumption_live_magnitude, y = flow, fill = region)) +
+  geom_bar(stat = "identity", position = "stack", width = 0.8) +
+  scale_fill_manual(values = region6_palette) +
+  labs(x = "", y = "") +
+  lims(x = c(stacked_bar_min, stacked_bar_max)) +
+  facet_grid(vars(common_name), switch = "y") +
+  theme_minimal() +
+  theme(
+    strip.text.y.left = element_text(size = 8),
+    strip.placement = "outside",
+    legend.position = "none",
+    axis.ticks.x = element_blank(),
+    axis.text.x = element_blank(),
+    panel.spacing = unit(2, "mm"),
+    plot.margin = margin(r = 5, l=-10, b=-5, unit="mm")# unit(c(, 5, 0, -10), "mm")
+  )
+
+inland_cap_stacked_bar
+
+inland_cap <- ggarrange(
+  inland_cap_map, inland_cap_stacked_bar,
+  nrow = 1, ncol = 2,
+  widths = c(1.5, 1)
+)
+
+inland_cap <- annotate_figure(
+  inland_cap, top = text_grob("Inland Capture")#, vjust = 1),
+)
+
+inland_cap
+
+# ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
+```
+
+```{r inland_aqua_foreign_consumption}
+inland_aqua_map <- trade_map +
+  geom_sf(data = consumer_map_df %>%
+            filter(habitat_method == "Inland Aquaculture"), 
+          aes(fill = consumption_live_t/consumption_live_magnitude, geometry = geometry), color = "white", size = 0.1) +
+  scale_fill_gradient(low = "#86ADA7", high = "#0F2D59", limits = c(map_min, map_max)) +
+  theme_bw() + 
+  theme(
+    legend.position = c(0.5, 0.05),
+    legend.direction = "horizontal",
+    legend.box = "vertical",
+    legend.spacing.y = unit(1, "mm"),
+    legend.spacing.x = unit(5, "mm"),
+    legend.title = element_text(size = 10),
+    legend.box.margin = margin(b = 3, t = 10, r = 20, unit = "mm"),
+    legend.margin = margin(t = 0, unit = "mm"),
+    axis.title = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text = element_blank(),
+    panel.grid.major = element_blank(),
+    panel.background = element_blank(),
+    panel.spacing = unit(2, "mm"),
+    panel.border = element_blank(),
+    plot.background = element_blank(),
+    legend.key.height = unit(2.5, "mm"),
+    plot.margin = margin(t=-10, unit="mm")
+  )
+
+inland_aqua_arrows <- inland_aqua_foreign_consumption %>%
+  slice_max(n = 10, order_by = consumption_live_t) %>%
+  # Join with lat/long data for centroids
+  left_join(country_centroids, by = c("source_country_iso3c" = "iso3")) %>%
+  left_join(country_centroids, by = c("consumer_iso3c" = "iso3")) %>%
+  mutate(centroid.lat.x = 1.05 * centroid.lat.x,
+         centroid.lat.y = 0.95 * centroid.lat.y)
+
+inland_aqua_map <- inland_aqua_map +
+  geom_curve(data = inland_aqua_arrows, 
+                 aes(x = centroid.lon.x, y = centroid.lat.x, 
+                     xend = centroid.lon.y, yend = centroid.lat.y,
+                     color = consumption_live_t/consumption_live_magnitude),
+                 size = 1.2,
+                 alpha = 0.75,
+                 curvature = -0.35, arrow = arrow(length = unit(3, "mm"), angle = 20)) +
+      scale_colour_gradient(low = "#F7AF75", high = "#E24027", limits = c(arrows_min, arrows_max)) +
+  labs(fill = "Consumption (mil. t.)",
+       colour = "Export (mil. t.)           ")
+
+inland_aqua_map
+
+inland_aqua_stacked_bar <- inland_aqua_sciname %>%
+  mutate(common_name = gsub(" nei", "", common_name)) %>%
+  mutate(common_name = case_when(
+    common_name == "ray-finned fishes" ~ "\nray-finned fishes",
+    common_name == "striped catfish" ~ "\nstriped catfish",
+    common_name == "freshwater siluroids" ~ "\nfreshwater siluroids",
+    common_name == "nile tilapia" ~ "\nnile tilapia",
+    TRUE ~ common_name
+  )) %>%
+  mutate(common_name = gsub(" ", "\n", common_name)) %>%
+  mutate(common_name = reorder(common_name, desc(total))) %>%
+  mutate(region = case_when(
+    region == "North America" ~ "N. America",
+    region == "South America" ~ "S. America",
+    TRUE ~ region
+  )) %>%
+  ggplot(aes(x = consumption_live_t/consumption_live_magnitude, y = flow, fill = region)) +
+  geom_bar(stat = "identity", position = "stack", width = 0.8) +
+  scale_fill_manual(values = region6_palette) +
+  labs(x = "", y = "", fill = "") +
+  lims(x = c(stacked_bar_min, stacked_bar_max)) +
+  facet_grid(vars(common_name), switch = "y") +
+  theme_minimal() +
+  theme(
+    strip.placement = "outside",
+    strip.text.y.left = element_text(size = 8),
+    legend.position = "bottom",
+    legend.box = "horizontal",
+    legend.text = element_text(size = 8),
+    legend.key.size = unit(3, "mm"),
+    legend.box.margin = margin(l = -25, unit = "mm"),
+    legend.margin = margin(t = -7, b = 0, unit = "mm"),
+    panel.spacing = unit(2, "mm"),
+    plot.margin = unit(c(0, 5, 0, -10), "mm")
+  )
+
+inland_aqua_stacked_bar
+
+inland_aqua <- ggarrange(
+  inland_aqua_map, inland_aqua_stacked_bar,
+  nrow = 1, ncol = 2,
+  widths = c(1.5, 1))
+
+inland_aqua <- annotate_figure(
+  inland_aqua, top = text_grob("Inland Aquaculture")
+)
+
+inland_aqua
+
+ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
+```
+
+```{r foreign_consumption, fig.height = 15, fig.width = 8}
+ggarrange(marine_cap,
+          marine_aqua,
+          inland_cap,
+          inland_aqua,
+          ncol = 1,
+          heights = c(1, 1, 1, 1.1),
+          common.legend = TRUE)
+
+ggsave(file.path(outdir, "fig2.png"), width = 9, height = 13, dpi = 500, unit = "in", device = "png")
+```
+
+# Figure 3: Trade patterns
+
+```{r fig3_data}
+artis_region_method <- read.csv(file.path(datadir, "fig_3_data.csv"))
+```
+
+```{r fig3a}
 region_net_import <- artis_region_method %>% 
@@
 
-fig2a <- ggplot() +
+fig3a <- ggplot() +
   geom_area(data = artis_region_method %>% 
@@
 
-fig2a <- ggdraw(fig2a) +
+fig3a <- ggdraw(fig3a) +
   draw_label("Net import", x = 0.97, y = 0.55, angle = 270, size = 9) +
@@
 
-fig2a
-```
-
-```{r fig2b}
+fig3a
+```
+
+```{r fig3b}
 # Matrix of plots showcasing inter-regional trade trends
-fig2b <- artis_region_method %>% 
+fig3b <- artis_region_method %>% 
     mutate(habitat_method = factor(habitat_method, 
@@
 
-fig2b <- ggdraw(fig2b) +
+fig3b <- ggdraw(fig3b) +
   draw_label("Exporting region", x = 0.02, y = 0.5, angle = 90, size = 10) +
@@
 
-fig2b
-```
-
-```{r fig2, fig.height=8, fig.width=6.5}
-# Overall Figure 2
-ggarrange(ggarrange(fig2a, NULL,
+fig3b
+```
+
+```{r fig3, fig.height=8, fig.width=6.5}
+# Overall Figure 3
+ggarrange(ggarrange(fig3a, NULL,
                     ncol = 2, nrow = 1,
                     widths = c(4.5, 0.2), labels = c("a", "")), 
-          fig2b, labels = c("", "b"),
+          fig3b, labels = c("", "b"),
           ncol = 1, heights = c(3, 4))
 
-ggsave(file.path(outdir, "fig2.png"))
-```
-
-# Figure 3
-```{r fig3a}
+ggsave(file.path(outdir, "fig3.png"))
+```
+
+# Figure 3: Foreign consumption maps and top species traded
+
+# Figure 4
+```{r fig4a}
 # Global per capita supply
-fig3a_data <- read.csv(file.path(datadir, "fig3a_data.csv"))
-
-fig3a <- fig3a_data %>%
+fig4a_data <- read.csv(file.path(datadir, "fig_4a_data.csv"))
+
+fig4a <- fig4a_data %>%
   # Set factor levels
@@
 
-fig3a
-```
-
-```{r fig3b}
+fig4a
+```
+
+```{r fig4b}
 # Per capita supply by region
-fig3b_data <- read.csv(file.path(datadir, "fig3b_data.csv"))
-
-fig3b <- fig3b_data %>%
+fig4b_data <- read.csv(file.path(datadir, "fig_4b_data.csv"))
+
+fig4b <- fig4b_data %>%
    # Set factor levels
@@
 
-fig3b
-```
-
-```{r fig3c}
+fig4b
+```
+
+```{r fig4c}
 # Global percent of supply by domestic / foreign exports
-fig3c_data <- read.csv(file.path(datadir, "fig3c_data.csv"))
-
-fig3c <- ggplot() +
-  geom_line(data = fig3c_data, aes(x = year, y = supply_percent, color = supply_source)) +
+fig4c_data <- read.csv(file.path(datadir, "fig_4c_data.csv"))
+
+fig4c <- ggplot() +
+  geom_line(data = fig4c_data, aes(x = year, y = supply_percent, color = supply_source)) +
   scale_color_manual(values = c("#F37542", "#0F2D59")) +
@@
 
-fig3c
-```
-
-```{r fig3d}
-fig3d_data <- read.csv(file.path(datadir, "fig3d_data.csv"))
-
-fig3d <- fig3d_data %>%
+fig4c
+```
+
+```{r fig4d}
+fig4d_data <- read.csv(file.path(datadir, "fig_4d_data.csv"))
+
+fig4d <- fig4d_data %>%
   ggplot(aes(x = year, y = supply_percent, color = supply_source)) +
@@
 
-fig3d
-```
-
-```{r fig3, fig.height = 6, fig.width = 7}
-fig3 <- ggarrange(
-  ggarrange(fig3a, fig3b, 
+fig4d
+```
+
+```{r fig4, fig.height = 6, fig.width = 7}
+fig4 <- ggarrange(
+  ggarrange(fig4a, fig4b, 
             common.legend = TRUE, nrow = 1, widths = c(1, 2),
             labels = c("a", "b")),
-  ggarrange(fig3c, fig3d,
+  ggarrange(fig4c, fig4d,
             common.legend = TRUE, nrow = 1, widths = c(1, 2),
@@
 
-fig3
-
-ggsave(file.path(outdir, "fig3.png"))
+fig4
+
+ggsave(file.path(outdir, "fig4.png"))
 ```
@@
 ```{r regional_sankey, fig.height = 3, fig.width = 3}
-regional_artis <- read.csv(file.path(datadir, "regional_artis_by_source.csv"))
+regional_artis <- read.csv(file.path(datadir, "si_fig_1_data.csv"))
 
@@
 
-ggsave(file.path(outdir, "regional_sankey_by_source.png"))
-```
-
-```{r si_top_exporters, fig.height = 7, fig.width = 6}
-
-top_exporters_old <- read.csv(file.path(datadir, "si_top_exporters_old_fill.csv")) %>%
+ggsave(file.path(outdir, "si_fig1.png"))
+```
+
+```{r si_fig_2_top_exporters, fig.height = 7, fig.width = 6}
+
+top_exporters_old <- read.csv(file.path(datadir, "si_fig_2_exporters_old.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
                                             "marine aquaculture", "inland aquaculture", "unknown")))
-top_exporters_recent <- read.csv(file.path(datadir, "si_top_exporters_recent_fill.csv")) %>%
+top_exporters_recent <- read.csv(file.path(datadir, "si_fig_2_exporters_recent.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
 
-ggsave(file.path(outdir, "si_top_exporters.png"))
-```
-
-```{r si_top_importers, fig.height = 7, fig.width = 6}
-
-top_importers_old <- read.csv(file.path(datadir, "si_top_importers_old_fill.csv")) %>%
+ggsave(file.path(outdir, "si_fig2.png"))
+```
+
+```{r si_fig3_top_importers, fig.height = 7, fig.width = 6}
+
+top_importers_old <- read.csv(file.path(datadir, "si_fig_3_importers_old.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
                                             "marine aquaculture", "inland aquaculture", "unknown")))
-top_importers_recent <- read.csv(file.path(datadir, "si_top_importers_recent_fill.csv")) %>%
+top_importers_recent <- read.csv(file.path(datadir, "si_fig_3_importers_recent.csv")) %>%
   mutate(habitat_method = factor(habitat_method, 
@@
 
-ggsave(file.path(outdir, "si_top_importers.png"))
-```
-
-```{r si_regional_changes, fig.height = 6, fig.width = 6}
-
-# Largest regional export changes
-top_regional_export_change <- read.csv(file.path(datadir, "top_regional_export_change.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-
-regional_export_change_plot <- top_regional_export_change %>%
-  ggplot(aes(x = change/1000000, y = reorder(exporter_region, ranking), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=reorder(exporter_region, ranking), yend=exporter_region)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-# Largest regional import changes
-top_regional_import_change <- read.csv(file.path(datadir, "top_regional_import_change.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-
-regional_import_change_plot <- top_regional_import_change %>%
-  ggplot(aes(x = change/1000000, y = reorder(importer_region, change), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y=reorder(importer_region, change), yend=importer_region)) +
-  geom_point(size=4) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(legend.position = "none")
-
-ggarrange(
-  regional_export_change_plot,
-  regional_import_change_plot,
-  ncol = 2,
-  labels = c("Export", "Import")
-)
-
-ggsave(file.path(outdir, "regional_changes.png"))
-```
-
-```{r si_export_decreases_increases, fig.height = 6, fig.width = 5}
-top_export_increases <- read.csv(file.path(datadir, "top_export_increases.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-top_export_decreases <- read.csv(file.path(datadir, "top_export_decreases.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-
-exporter_increases_plot <- top_export_increases %>%
-  ggplot(aes(x = change/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y= reorder(exporter_iso3c, ranking), yend=exporter_iso3c)) +
-  geom_point(size=2) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(
-    legend.position = "none",
-    axis.text = element_text(size = 8)
-  )
-
-exporter_decreases_plot <- top_export_decreases %>%
-  ggplot(aes(x = change/1000000, y = reorder(exporter_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y= reorder(exporter_iso3c, ranking), yend=exporter_iso3c)) +
-  geom_point(size=2) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(
-    legend.position = "none",
-    axis.text = element_text(size = 8)
-  )
-
-ggarrange(
-  exporter_decreases_plot, exporter_increases_plot, nrow = 1
-)
-
-ggsave(file.path(outdir, "si_export_changes.png"))
-```
-
-```{r si_import_decreases_increases, fig.height = 6, fig.width = 5}
-top_import_increases <- read.csv(file.path(datadir, "top_import_increases.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-top_import_decreases <- read.csv(file.path(datadir, "top_import_decreases.csv")) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown")))
-
-importer_increases_plot <- top_import_increases %>%
-  ggplot(aes(x = change/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y= reorder(importer_iso3c, ranking), yend=importer_iso3c)) +
-  geom_point(size=2) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(
-    legend.position = "none",
-    axis.text = element_text(size = 8)
-  )
-
-importer_decreases_plot <- top_import_decreases %>%
-  ggplot(aes(x = change/1000000, y = reorder(importer_iso3c, ranking), fill = habitat_method, color = habitat_method)) +
-  geom_segment(aes(x=0, xend=change/1000000, y= reorder(importer_iso3c, ranking), yend=importer_iso3c)) +
-  geom_point(size=2) +
-  scale_color_manual(values = habitat_source_colors) +
-  scale_y_reordered() +
-  labs(y = "", x = "Live weight (million t)") +
-  facet_wrap(~habitat_method, ncol = 1, scales = "free_y") +
-  theme_minimal() +
-  theme(
-    legend.position = "none",
-    axis.text = element_text(size = 8)
-  )
-
-ggarrange(
-  importer_decreases_plot, importer_increases_plot, nrow = 1
-)
-
-ggsave(file.path(outdir, "si_import_changes.png"))
-```
-
-```{r regional_avg_annual_change, fig.height = 8, fig.width = 7}
-regional_avg_annual_change <- read.csv(file.path(datadir, "regional_avg_annual_change.csv"))
+ggsave(file.path(outdir, "si_fig3.png"))
+```
+
+```{r si_fig4_regional_avg_annual_change, fig.height = 8, fig.width = 7}
+regional_avg_annual_change <- read.csv(file.path(datadir, "si_fig_4_data.csv"))
 
@@
 
-ggsave(file.path(outdir, "regional_avg_annual_change.png"))
-```
-
-```{r ldc_net_exports_habitat_method, fig.height = 4, fig.width = 7}
-ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "ldc_net_habitat_method_summary.csv")) %>%
-  select(year, habitat_method, net_export)
-
-ldc_net_total <- ldc_net_habitat_method_summary %>%
-  group_by(year) %>%
-  summarise(net_export = sum(net_export)) %>%
-  mutate(habitat_method = "total")
-
-ldc_net_habitat_method_summary <- ldc_net_habitat_method_summary %>%
-  bind_rows(ldc_net_total) %>%
-  mutate(habitat_method = factor(habitat_method, 
-                                 levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "total")))
-
-ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
-  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
-  geom_line(size = 1) +
-  scale_color_manual(values = habitat_source_colors) +
-  labs(x = "", y = "Net export (million t)", color = "") +
-  theme_minimal() +
-  theme(
-    legend.position = "bottom",
-    axis.text = element_text(size = 12)
-  )
-
-ldc_net_exports_by_source_plot
-
-ggsave(file.path(outdir, "ldc_net_exports_by_source.png"))
-```
-
-```{r custom_artis_ts, fig.height = 4, fig.width = 6}
-
-artis_ts <- read.csv(file.path(datadir, "artis_ts.csv"))
-custom_ts <- read.csv(file.path(datadir, "custom_ts.csv"))
-
-ggplot() +
-  geom_line(data = artis_ts, aes(x = year, y = product_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
-  geom_line(data = custom_ts, aes(x = year, y = product_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
-  scale_color_manual(
-    values = c("Custom ARTIS Time series" = "black", "HS96" = "#741A32", "HS02" = "#B34232", "HS07" = "#D38F35", "HS12" = "red", "HS17" = "#114F59")
-  ) +
-  labs(x = "Year", y = "Product Weight (tonnes)", color = "HS Version") +
-  theme_bw() +
-  theme(
-    legend.position = "bottom",
-    axis.text = element_text(size = 10)
-  )
-
-ggsave(file.path(outdir, "custom_artis_ts.png"))
-```
-
-```{r dom_source_ts, fig.height = 4, fig.width = 6}
-dom_source_ts <- read.csv(file.path(datadir, "dom_source_ts.csv"))
+ggsave(file.path(outdir, "si_fig4.png"))
+```
+
+```{r si_fig5a, fig.height = 4, fig.width = 6}
+dom_source_ts <- read.csv(file.path(datadir, "si_fig_5a_data.csv"))
 
@@
 
-```{r dom_source_by_habitat_ts, fig.height = 4, fig.width = 7}
-dom_source_by_habitat_method <- read.csv(file.path(datadir, "dom_source_by_habitat_method.csv"))
+```{r si_fig5b, fig.height = 4, fig.width = 7}
+dom_source_by_habitat_method <- read.csv(file.path(datadir, "si_fig_5b_data.csv"))
 
@@
 
-```{r dom_source_overview, fig.height = 5, fig.width = 8}
+```{r si_fig5, fig.height = 5, fig.width = 8}
  ggarrange(
@@
 
-ggsave(file.path(outdir, "dom_source_2019_overview.png"))
-```
+ggsave(file.path(outdir, "si_fig5.png"))
+```
+
+```{r si_fig6_ldc_net_exports_habitat_method, fig.height = 4, fig.width = 7}
+ldc_net_habitat_method_summary <- read.csv(file.path(datadir, "si_fig_6_data.csv")) %>%
+  select(year, habitat_method, net_export)
+
+ldc_net_total <- ldc_net_habitat_method_summary %>%
+  group_by(year) %>%
+  summarise(net_export = sum(net_export)) %>%
+  mutate(habitat_method = "total")
+
+ldc_net_habitat_method_summary <- ldc_net_habitat_method_summary %>%
+  bind_rows(ldc_net_total) %>%
+  mutate(habitat_method = factor(habitat_method, 
+                                 levels = c("marine capture", "inland capture", 
+                                            "marine aquaculture", "inland aquaculture", "total")))
+
+ldc_net_exports_by_source_plot <- ldc_net_habitat_method_summary %>%
+  ggplot(aes(x = year, y = net_export/1000000, color = habitat_method)) +
+  geom_line(size = 1) +
+  scale_color_manual(values = habitat_source_colors) +
+  labs(x = "", y = "Net export (million t)", color = "") +
+  theme_minimal() +
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 12)
+  )
+
+ldc_net_exports_by_source_plot
+
+ggsave(file.path(outdir, "si_fig6.png"))
+```
+
+```{r si_fig7_custom_artis_ts, fig.height = 4, fig.width = 6}
+artis_ts <- read.csv(file.path(datadir, "si_fig_7_artis_ts.csv")) %>%
+  mutate(hs_version = factor(hs_version,
+                             levels = c("HS96", "HS02", "HS07", "HS12", "HS17")))
+custom_ts <- read.csv(file.path(datadir, "si_fig_7_custom_ts.csv"))
+
+ggplot() +
+  geom_line(data = artis_ts, aes(x = year, y = product_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
+  geom_line(data = custom_ts, aes(x = year, y = product_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
+  scale_color_manual(
+    values = c("Custom ARTIS Time series" = "black", "HS96" = "#741A32", "HS02" = "#B34232", "HS07" = "#D38F35", "HS12" = "red", "HS17" = "#114F59")
+  ) +
+  labs(x = "Year", y = "Product Weight (tonnes)", color = "HS Version") +
+  theme_bw() +
+  theme(
+    legend.position = "bottom",
+    axis.text = element_text(size = 10)
+  )
+
+ggsave(file.path(outdir, "si_fig7.png"))
+```

### 2024-05-02 ms_figures.Rmd
ms figures labels fixed based on pdf journal size requirements

@@
 datadir <- "qa/manuscript_archive/manuscript_inputs"
-outdir <- "qa/manuscript_archives/figures"
+outdir <- "qa/manuscript_archive/figures"
 ```
@@
   theme(legend.position = "bottom",
+        text = element_text(size = 10),
         axis.title.y = element_text(size = 12),
@@
 
-ggsave(file.path(outdir, "fig1.png"))
-# ggsave(file.path(outdir, "fig1.pdf"), device = "pdf", height = )
+# ggsave(file.path(outdir, "fig1.png"))
+ggsave(file.path(outdir, "fig1.pdf"), device = "pdf", height = 150, width = 180, units = "mm")
 ```
@@
     common_name == "alaska pollock(=walleye poll.)" ~ "alaska pollock",
+    common_name == "ray-finned fishes" ~ "ray finned fishes",
     TRUE ~ common_name
@@
     axis.text.x = element_blank(),
-    axis.text.y = element_text(size = 8),
+    # axis.text.y = element_text(size = 8),
     plot.margin = unit(c(0, 5, -5, -10), "mm")
@@
 marine_cap
-
-# ggsave(file.path(outdir, "marine_cap_map_bar.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
 ```
@@
 marine_aqua
-
-# ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
 ```
@@
   mutate(common_name = case_when(
-    common_name == "ray-finned fishes" ~ "\nray-finned fishes",
+    common_name == "ray-finned fishes" ~ "ray finned fishes",
     common_name == "freshwater siluroids" ~ "\nfreshwater siluroids",
@@
   mutate(common_name = case_when(
-    common_name == "ray-finned fishes" ~ "\nray-finned fishes",
+    common_name == "ray-finned fishes" ~ "ray finned fishes",
     common_name == "striped catfish" ~ "\nstriped catfish",
@@
 
-ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
+# ggsave(file.path(outdir, "test.png"), width = 6, height = 3, dpi = 300, units = "in", device = "png")
 ```
@@
 
-ggsave(file.path(outdir, "fig2.png"), width = 9, height = 13, dpi = 500, unit = "in", device = "png")
+# ggsave(file.path(outdir, "fig2.png"), width = 9, height = 13, dpi = 500, unit = "in", device = "png")
+ggsave(file.path(outdir, "fig2.pdf"), width = 180, height = 320, unit = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "fig3.png"))
+# ggsave(file.path(outdir, "fig3.png"))
+ggsave(file.path(outdir, "fig3.pdf"), height = 180, width = 180, unit = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "fig4.png"))
+# ggsave(file.path(outdir, "fig4.png"))
+ggsave(file.path(outdir, "fig4.pdf"), height = 180, width = 180, unit = "mm", device = "pdf")
 ```
@@
 ```{r regional_sankey, fig.height = 3, fig.width = 3}
-regional_artis <- read.csv(file.path(datadir, "si_fig_1_data.csv"))
+regional_artis <- read.csv(file.path(datadir, "si_fig_1_data.csv")) %>%
+  mutate(exporter_region = case_when(
+    exporter_region == "North America" ~ "N.America",
+    exporter_region == "South America" ~ "S.America",
+    TRUE ~ exporter_region
+  )) %>%
+  mutate(importer_region = case_when(
+    importer_region == "North America" ~ "N.America",
+    importer_region == "South America" ~ "S.America",
+    TRUE ~ importer_region
+  ))
 
@@
                               levels = c("Europe", "Oceania", "Asia",
-                                         "South America", "North America",
+                                         "S.America", "N.America",
                                          "Africa"))
@@
     plot.title = element_text(hjust = 0.5),
-    plot.margin = margin(c(t=-0.5, r=-10, b=-1, l=-18), unit = "mm")
+    plot.margin = margin(c(t=-0.5, r=-7, b=-1, l=-25), unit = "mm")
     )
@@
 
-ggsave(file.path(outdir, "si_fig1.png"))
+# ggsave(file.path(outdir, "si_fig1.png"))
+ggsave(file.path(outdir, "si_fig1.pdf"), height = 200, width = 180, unit = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "si_fig2.png"))
+# ggsave(file.path(outdir, "si_fig2.png"))
+ggsave(file.path(outdir, "si_fig2.pdf"), height = 200, width = 180, units = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "si_fig3.png"))
+# ggsave(file.path(outdir, "si_fig3.png"))
+ggsave(file.path(outdir, "si_fig3.pdf"), height = 200, width = 180, units = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "si_fig4.png"))
+# ggsave(file.path(outdir, "si_fig4.png"))
+ggsave(file.path(outdir, "si_fig4.pdf"), height = 200, width = 180, unit = "mm", device = "pdf")
 ```
@@
 dom_source_by_habitat_method_p <- dom_source_by_habitat_method %>%
+  mutate(habitat_method = gsub(" ", "\n", habitat_method)) %>%
   # Set factor levels
-  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
-                                                            "marine aquaculture", "inland aquaculture",
+  mutate(habitat_method = factor(habitat_method, levels = c("marine\ncapture", "inland\ncapture",
+                                                            "marine\naquaculture", "inland\naquaculture",
                                                             "unknown" ))) %>%
@@
 
-ggsave(file.path(outdir, "si_fig5.png"))
+# ggsave(file.path(outdir, "si_fig5.png"))
+ggsave(file.path(outdir, "si_fig5.pdf"), height = 125, width = 180, unit = "mm", device = "pdf")
 ```
@@
 
-ggsave(file.path(outdir, "si_fig6.png"))
+# ggsave(file.path(outdir, "si_fig6.png"))
+ggsave(file.path(outdir, "si_fig6.pdf"), height = 80, width = 180, unit = "mm", device = "pdf")
 ```
@@
 ggplot() +
-  geom_line(data = artis_ts, aes(x = year, y = product_weight_t, color = hs_version), linetype = "dashed", linewidth = 1) +
-  geom_line(data = custom_ts, aes(x = year, y = product_weight_t, color = hs_version), linewidth = 0.5, linetype = "solid") +
+  geom_line(data = artis_ts, aes(x = year, y = product_weight_t/1e6, color = hs_version), linetype = "dashed", linewidth = 1) +
+  geom_line(data = custom_ts, aes(x = year, y = product_weight_t/1e6, color = hs_version), linewidth = 0.5, linetype = "solid") +
   scale_color_manual(
@@
   ) +
-  labs(x = "Year", y = "Product Weight (tonnes)", color = "HS Version") +
+  labs(x = "Year", y = "Product Weight (million tonnes)", color = "HS Version") +
   theme_bw() +
@@
 
-ggsave(file.path(outdir, "si_fig7.png"))
-```
+# ggsave(file.path(outdir, "si_fig7.png"))
+ggsave(file.path(outdir, "si_fig7.pdf"), height = 80, width = 180, unit = "mm", device = "pdf")
+```

### 2024-05-02 ms_figures.Rmd
fixed height of fig 3 for fig3b y axis
figures fit A4 page restrictions

@@
   ggplot(aes(x = consumption_live_t/consumption_live_magnitude, y = flow, fill = region)) +
-  geom_bar(stat = "identity", position = "stack", width = 0.8) +
+  geom_bar(stat = "identity", position = "stack", width = 0.7) +
   scale_fill_manual(values = region6_palette) +
@@
     legend.position = "none",
-    panel.spacing = unit(2, "mm"),
+    panel.spacing = unit(0.5, "mm"),
     axis.ticks.x = element_blank(),
@@
     axis.text.x = element_blank(),
-    panel.spacing = unit(2, "mm"),
+    panel.spacing = unit(0.5, "mm"),
     plot.margin = margin(r = 5, l=-10, b=-5, unit="mm")# unit(c(, 5, 0, -10), "mm")
@@
     legend.margin = margin(t = -7, b = 0, unit = "mm"),
-    panel.spacing = unit(2, "mm"),
+    panel.spacing = unit(0.5, "mm"),
     plot.margin = unit(c(0, 5, 0, -10), "mm")
@@
 # ggsave(file.path(outdir, "fig2.png"), width = 9, height = 13, dpi = 500, unit = "in", device = "png")
-ggsave(file.path(outdir, "fig2.pdf"), width = 180, height = 320, unit = "mm", device = "pdf")
+ggsave(file.path(outdir, "fig2.pdf"), width = 180, height = 295, unit = "mm", device = "pdf")
 ```
@@
 # ggsave(file.path(outdir, "fig3.png"))
-ggsave(file.path(outdir, "fig3.pdf"), height = 180, width = 180, unit = "mm", device = "pdf")
+ggsave(file.path(outdir, "fig3.pdf"), height = 210, width = 180, unit = "mm", device = "pdf")
 ```

### 2024-05-02 top_flows_top_species_figure.R
cleaning up unused code

@@ -1,249 +1,0 @@
-# Draft figure of species-level information
-library(tidytext)
-
-con <- dbConnect(RPostgres::Postgres(),
-                 dbname=Sys.getenv("DB_NAME"),
-                 host="localhost",
-                 port="5432",
-                 user=Sys.getenv("DB_USERNAME"),
-                 password=Sys.getenv("DB_PASSWORD"))
-
-artis <- dbGetQuery(con, "SELECT * FROM snet") %>%
-  select(-record_id)
-
-# Country metadata
-country_metadata <- dbGetQuery(con, "SELECT * FROM countries") %>%
-  select(-record_id)
-
-# Sciname metadata
-sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
-  select(-record_id)
-
-# Code max resolved taxa
-code_max_resolved_taxa <- dbGetQuery(con, "SELECT * FROM code_max_resolved_taxa") %>%
-  select(-record_id) 
-
-# Consumption
-consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption") %>%
-  select(-record_id)
-
-# Population 
-pop <- dbGetQuery(con, "SELECT * FROM population") %>%
-  select(-record_id)
-
-# Close database connection
-dbDisconnect(con)
-rm(list = c("con"))
-
-# Set colors
-region6_palette <- c(
-  "#741A32", # maroon
-  "#B34232", # burnt orange
-  "#D38F35", #  orange
-  "#D4B95F", # khaki
-  "#4FA2A2", # teal
-  "#114F59" # dark teal
-)
-
-artis_palette <- colorRampPalette(region6_palette)
-
-
-consumption_scaled <- consumption %>%
-  left_join(pop, by = c("year", "consumer_iso3c" = "iso3c")) %>%
-  mutate(consumption_kg_percap = 1000*consumption_live_t/pop) %>%
-  group_by(year, consumer_iso3c) %>%
-  mutate(total_consumption_kg_percap = sum(consumption_kg_percap)) %>% 
-  ungroup() %>%
-  mutate(scale_factor = consumption_kg_percap/total_consumption_kg_percap) %>%
-  mutate(total_consumption_kg_percap = case_when(
-    total_consumption_kg_percap > 100 ~ 100, 
-    TRUE ~ total_consumption_kg_percap
-    )) %>%
-  mutate(consumption_kg_percap= total_consumption_kg_percap*scale_factor) %>%
[…]

### 2024-05-03 ms_figures.Rmd
fig 1c interpretation of number of average export partners

@@
                                             "marine aquaculture", "inland aquaculture", "unknown"))) %>%
-  ggplot(aes(x = year, y = all_out_degree, color = habitat_method)) +
+  ggplot(aes(x = year, y = all_in_degree, color = habitat_method)) +
   geom_smooth(alpha = 0.2) +

### 2024-05-03 prep_data.R
comment added no need to run this script to replicate results or figures

@@ -1,4 +1,8 @@
 
+################################################################################
+######## NO NEED TO RUN THIS SCRIPT TO REPLICATE RESULTS OR FIGURES ############
+################################################################################
+
 library(tidyverse)
 library(countrycode)
 library(circlize)
@@ -16,7 +20,6 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-# outdir <- "qa/outputs_20240315"
 outdir <- "qa/manuscript_archive/manuscript_inputs"
 #-------------------------------------------------------------------------------
 # Initial database pulls

### 2024-05-03 ms_results.Rmd
replicable data directory
image paths updated

@@
 # Directory
-datadir <- "qa/manuscript_archive/manuscript_inputs"
-
+datadir <- "data"
 global_export_by_source <- read.csv(file.path(datadir, "global_export_by_source.csv"))
@@
 d) Number of countries comprising 75% of global exports by production method and environment with the global total in the black line.
-e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](qa/outputs_20240221/fig1.png)
+e) Number of countries comprising 75% of global imports by production method and environment with the global total in the black line.](figures/fig1.pdf)
 
@@
 Values represent million tonnes in live weight equivalents.
-Note the scale difference across rows.](qa/outputs_20240221/fig2.png)
+Note the scale difference across rows.](figures/fig2.pdf)
 
@@
 d) Regional aquatic food domestic versus foreign sourcing over time.
-](qa/outputs_20240221/fig3.png)
+](figures/fig3.pdf)

### 2024-07-24 ms_figures.Rmd
replicable data and figure directories
update file path and confirm no results changed

@@
 
-```{r load_data}
+```{r data_setup}
 # Set directories
-datadir <- "qa/manuscript_archive/manuscript_inputs"
-outdir <- "qa/manuscript_archive/figures"
+datadir <- "data"
+outdir <- "figures"
+
+# Create figures directory
+if (!dir.exists(outdir)) {
+  dir.create(outdir)
+}
 ```
@@
 ```{r fig1de}
-habitat_source_colors_v2 <- c("#F7AF75", # light orange = inland aquaculture
+habitat_source_colors_v2 <- c("#1B708F", # blue = marine capture
                           "#86ADA7", # light blue = freshwater capture
                           "#F37542", # medium orange = marine aquaculture
-                              "#1B708F", # blue = marine capture
+                          "#F7AF75", # light orange = inland aquaculture
                               "black" # unknown
@@
   geom_line(data = import_n_countries_concentration_by_source %>% 
+              filter(habitat_method != "unknown") %>%
               mutate(habitat_method = factor(habitat_method, 
                                  levels = c("marine capture", "inland capture", 
-                                            "marine aquaculture", "inland aquaculture", "unknown"))), 
+                                            "marine aquaculture", "inland aquaculture"))), 
       aes(x = year, y = n, color = habitat_method)) +
@@
     world,
-    by = c("consumer_iso3c"="iso_a3")
+    by = c("consumer_iso3c"="iso_a3_eh")
   )

### 2024-07-24 prep_data.R

@@ -20,7 +20,7 @@ library(tidytext)
 #-------------------------------------------------------------------------------
 # Directory listing
 
-outdir <- "qa/manuscript_archive/manuscript_inputs"
+outdir <- "data"
 #-------------------------------------------------------------------------------
 # Initial database pulls
 
@@ -89,7 +89,7 @@ sciname_metadata <- dbGetQuery(con, "SELECT * FROM sciname") %>%
 pop <- dbGetQuery(con, "SELECT * FROM population") %>%
   select(-record_id)
 
-consumption <- dbGetQuery(con, "SELECT * FROM complete_consumption WHERE 
+consumption <- dbGetQuery(con, "SELECT * FROM consumption WHERE 
 ((hs_version = 'HS96' AND year >= 1996 AND year <= 2003) OR
 (hs_version = 'HS02' AND year >= 2004 AND year <= 2009) OR
 (hs_version = 'HS07' AND year >= 2010 AND year <= 2012) OR
